---
title: "Bottom-up effects of an herbicide and an antibiotic on *Potamopyrgus antipodarum*"
format: 
  html:
    self-contained: true
    toc: true
    style: "floating"
---

## Autoren noch anpassen

Sophie Oster, Verena C. Schreiner, Tobias Schmitt, Patrick Fink, Markus Weitere, Ibrahim Araoye, Ralf Schulz, Mirco Bundschuh

R script written by Sophie Oster\
checked by Verena C. Schreiner

\
RPTU Kaiserslautern-Landau\
Fortstrasse 7\
76829 Landau in der Pfalz\
GERMANY

This study was conducted with four chemical treatments, as explanatory variables:

-   treatment (categorical)
    -   uncontaminated control (negative control, NC)
    -   herbicide propyzamide (Pro)
    -   antibiotic ciprofloxacin (Cip)
    -   mixture of propyzamide and ciprofloxacin (Mix)


# 1 R version 4.3.2

R version used for this script: 4.3.2, use the following code chunk to download the correct version.

```{r}
# Check the current R version
current_r_version <- as.character(getRversion())

# Specify the required R version
required_r_version <- "4.3.2"

if (current_r_version != required_r_version) {
  # Install 'rig' if not already installed
  if (!requireNamespace("rig", quietly = TRUE)) {
    install.packages("rig")
  }
  
  # Load the 'rig' package
  library(rig)
  
  # Check if the required R version is installed
  installed_versions <- rig::rig_ls()
  if (!(required_r_version %in% installed_versions$version)) {
    # Install the required R version using 'rig'
    rig::rig_add(required_r_version)
  }
  
  # Set the required R version as default
  rig::rig_default(required_r_version)
  
  # Notify the user about the version change
  stop(paste("Please restart R using version", required_r_version, "and rerun the script."))
} else {
  message("You are already using the correct R version:", required_r_version)
}

```

# 2 Packages in a reproducible environment

```{r, message = FALSE, warning = FALSE}
# Restore the environment:
# renv::restore()

library(stats)
library(rstatix)
library(dplyr)
library(tidyverse)
library(tidyr)
library(broom)
library(dunn.test)
library(patchwork)
library(vegan)
library(data.table)
library(ggplot2)
library(ggrepel)
library(ggforce)
library(concaveman)
library(gridExtra)
library(stringr)
library(shades)
library(ggpubr)
library(base)
library(ggokabeito)
library(multcomp)
library(car)
library(asbio)
library(plotrix)
library(lmtest)
library(boot)
library(moments)
library(lme4)
library(DAAG)
library(performance)
library(emmeans)
library(purrr)
library(DT)
library(Hmisc)

if (!requireNamespace("devtools", quietly = TRUE)) {
  install.packages("devtools")
}

library(devtools)

if (!requireNamespace("pairwiseAdonis", quietly = TRUE)) {
  devtools::install_github("pmartinezarbizu/pairwiseAdonis/pairwiseAdonis")
}

library(pairwiseAdonis)

# renv::snapshot()
```

# 3 Biofilm endpoints

## Chlorophyll and ash free dry weight of biofilm after colonisation + chronic contamination of 14 days

-   Unit chlorophyll a, b & c: µg chlorophyll / cm²

-   Unit afdw: mg / cm²

```{r}
# insert data
biofilm_endpoints <-
  read.table(file.path("data\\biofilm_endpoints.txt"),
             header = TRUE, sep = "\t")

View(biofilm_endpoints)
```

## 3.1 Chlorophyll

Caclulate mean chla, chlb + chlc of each treatment:

```{r}
mean_chlorophylls <- biofilm_endpoints %>%
  dplyr::group_by(treatment) %>%
  dplyr::summarize(
    mean_chlorophyll_a = mean(chla, na.rm = TRUE),
    mean_chlorophyll_b = mean(chlb, na.rm = TRUE),
    mean_chlorophyll_c = mean(chlc, na.rm = TRUE)
  )

print(mean_chlorophylls)
```


#### Histogram chlorophyll a, b + c

```{r}
# chl a
hist(biofilm_endpoints $ chla) 
skewness(biofilm_endpoints $ chla, na.rm = TRUE)
shapiro.test(biofilm_endpoints $ chla)

# chl b
hist(biofilm_endpoints $ chlb) 
skewness(biofilm_endpoints $ chlb, na.rm = TRUE) 
shapiro.test(biofilm_endpoints $ chlb)

# chl c
hist(biofilm_endpoints $ chlc) 
skewness(biofilm_endpoints $ chlc, na.rm = TRUE) 
shapiro.test(biofilm_endpoints $ chlc)
```

All not normal distributed, right skewed but skewness \< 1.5 therefore no data transformation for model

#### Model chla

Linear model with treatment as explanatory variable.

```{r}
# treatment as factor
biofilm_endpoints $ week <- factor(biofilm_endpoints $ week)
biofilm_endpoints $ treatment <- factor(biofilm_endpoints $ treatment)

model_chla <- lm(chla ~ week + treatment + week:treatment, data = biofilm_endpoints)

par(mfrow = c(2, 2))

plot(model_chla) # point cloud looks evenly distributed --> normal distribution

qreference(residuals(model_chla)) 

drop1(model_chla, test = 'F')

```

Now check differences between treatments within weeks.

```{r}
chla_inter <- lsmeans(model_chla, ~ treatment | week )
chla_inter_dunn1 <- contrast(chla_inter, 'pairwise')
(chla_inter_dunn_sum1 <- summary(chla_inter_dunn1, infer = TRUE, adjust = 'mvt'))
```

No significant differences of the treatments within week 1 and 2. Only Cip-Mix and Cip-Pro significant in week 3.

Now check for differences between treatments.

```{r}
anova(model_chla)
# week variable highly significant 

chla_inter2 <- lsmeans(model_chla, ~ week |  treatment)
chla_inter_dunn2 <- contrast(chla_inter2, 'pairwise')
(chla_inter_dunn_sum2 <- summary(chla_inter_dunn2, infer = TRUE, adjust = 'mvt'))

```

-   NC\
    -   Week 1 & 2 similar\
    -   Week 2 & 3 similar\
    -   difference between week 1 & 3
-   Cip\
    -   Week 1 & 2 similar\
    -   Week 1 & 3 similar\
    -   difference between week 2 & 3
-   Pro\
    -   Week 1 & 2 similar\
    -   differences between week 1 & 3 and 2 & 3
-   Mix\
    -   all weeks different

Week 1 & 2 quite similar, expect of Mix treatment. Week 3 quite different to week 1 & 2 despite of Cip treatment.

#### Model chlb

```{r}
model_chlb <- lm(chlb ~ week + treatment + week:treatment, data = biofilm_endpoints)

par(mfrow = c(2, 2))

plot(model_chlb) # point cloud looks evenly distributed --> normal distribution

qreference(residuals(model_chlb)) # normal distribution assumed

drop1(model_chlb, test = 'F')
```

No significant differences. Still check for differences between treatments.

```{r}
chlb_inter <- lsmeans(model_chlb, ~ treatment | week )
chlb_inter_dunn1 <- contrast(chlb_inter, 'pairwise')
(chlb_inter_dunn_sum1 <- summary(chlb_inter_dunn1, infer = TRUE, adjust = 'mvt'))
```

No differences between treatments for chlb.

Now check for differences between weeks.

```{r}
anova(model_chlb)
# week variable highly significant 

chlb_inter2 <- lsmeans(model_chlb, ~ week |  treatment)
chlb_inter_dunn3 <- contrast(chlb_inter2, 'pairwise')
(chlb_inter_dunn_sum3 <- summary(chlb_inter_dunn3, infer = TRUE, adjust = 'mvt'))
```

Similar results/trends as for chla. Week 1 & 2 quite similar; differences in:

-   Pro treatment: week 2 & 3

-   Mix treatment: week 1 & 2; 2 & 3

#### Model chlc

```{r}
model_chlc <- lm(chlc ~ week + treatment + week:treatment, data = biofilm_endpoints)

par(mfrow = c(2, 2))

plot(model_chlc) # point cloud looks evenly distributed --> normal distribution

qreference(residuals(model_chlc)) # normal distribution assumed

drop1(model_chlc, test = 'F')
```

No significant differences. Still check for differences between treatments.

```{r}
chlc_inter <- lsmeans(model_chlc, ~ treatment | week )
chlc_inter_dunn1 <- contrast(chlc_inter, 'pairwise')
(chlc_inter_dunn_sum1 <- summary(chlc_inter_dunn1, infer = TRUE, adjust = 'mvt'))
```

No differences between treatments for chlorophyll c. Now check for differences between weeks:

```{r}
anova(model_chlc)

# week variable significant

chlc_inter2 <- lsmeans(model_chlc, ~ week |  treatment)
chlc_inter_dunn3 <- contrast(chlc_inter2, 'pairwise')
(chlc_inter_dunn_sum3 <- summary(chlc_inter_dunn3, infer = TRUE, adjust = 'mvt'))
```

-   NC\
    -   all weeks similar
-   Cip\
    -   all weeks similar
-   Pro\
    -   all weeks similar
-   Mix\
    -   Week 1 & 2, and 2 & 3 different

#### Boxplots chlorophyll a, b + c

```{r}
# Chlorophyll a
chla_plot <- ggplot(biofilm_endpoints, 
                    aes(x = factor(week), y = chla, 
                        fill = factor(treatment, levels = c("NC", "Cip", "Pro", "Mix")))) +
  geom_boxplot() +
  scale_fill_manual(values = c("NC" = "#009E73", "Cip" = "#56B4E9", "Pro" = "#E69F00", "Mix" = "#F0E442")) +
  theme_bw() +
  labs(x = NULL , y = expression("Chl a (µg/cm"^"2"*")"), fill = "Treatment") +
  scale_x_discrete(labels = c("Week 1", "Week 2", "Week 3"))

# Chlorophyll b
chlb_plot <- ggplot(biofilm_endpoints, 
                    aes(x = factor(week), y = chlb, 
                        fill = factor(treatment, levels = c("NC", "Cip", "Pro", "Mix")))) +
  geom_boxplot() +
  scale_fill_manual(values = c("NC" = "#009E73", "Cip" = "#56B4E9", "Pro" = "#E69F00", "Mix" = "#F0E442")) +
  theme_bw() +
  labs(x = NULL , y = expression("Chl b (µg/cm"^"2"*")"), fill = "Treatment") +
  scale_x_discrete(labels = c("Week 1", "Week 2", "Week 3"))

# Chlorophyll c
chlc_plot <- ggplot(biofilm_endpoints, 
                    aes(x = factor(week), y = chlc, 
                        fill = factor(treatment, levels = c("NC", "Cip", "Pro", "Mix")))) +
  geom_boxplot() +
  scale_fill_manual(values = c("NC" = "#009E73", "Cip" = "#56B4E9", "Pro" = "#E69F00", "Mix" = "#F0E442")) +
  theme_bw() +
  labs(x = NULL , y = expression("Chl c (µg/cm"^"2"*")"), fill = "Treatment") +
  scale_x_discrete(labels = c("Week 1", "Week 2", "Week 3"))

# combine and save with AFDW plot
```

## 3.2 AFDW after colonisation

Calculate mean AFDW of each treatment with distinction of the 3 weeks.

```{r}
mean_afdw_after_colonisation <- biofilm_endpoints %>%
  group_by(week, treatment) %>%
  summarize(mean_afdw_after_colonisation = mean(afdw_after_colonisation))
```

Now calculate mean AFDW of each treatment without taking into account the different weekly colonizations:

```{r}
mean_afdw_after_colonisation_treatments <- biofilm_endpoints %>%
  group_by(treatment) %>%
  summarize(mean_afdw_after_colonisation_treatments = mean(afdw_after_colonisation))

print(mean_afdw_after_colonisation_treatments)
```

#### Histogram AFDW after colonisation

```{r}
hist(biofilm_endpoints $ afdw_after_colonisation) 
skewness(biofilm_endpoints $ afdw_after_colonisation, na.rm = TRUE) 
shapiro.test(biofilm_endpoints $ afdw_after_colonisation)
```

Not normal distributed, right skewed but skewness \<1.5 therefore no data transformation for model.

#### Model AFDW after colonisation

Linear model with treatment and week and their interaction (week:treatment) as explanatory variables.

```{r}
model_afdw_after_colonisation <- lm(afdw_after_colonisation ~ week + treatment + week:treatment, data = biofilm_endpoints)

par(mfrow = c(2, 2))

plot(model_afdw_after_colonisation) # not perfectly even distributed, but ok

qreference(residuals(model_afdw_after_colonisation)) # normal distribution assumed

drop1(model_afdw_after_colonisation, test = 'F')
```

Significant difference for week:treatment. Check differences between treatments within weeks.

```{r}
afdw_after_colonisation_inter <- lsmeans(model_afdw_after_colonisation, ~ treatment | week )

afdw_after_colonisation_inter_dunn1 <- contrast(afdw_after_colonisation_inter, 'pairwise')

(afdw_after_colonisation_inter_dunn_sum1 <- summary(afdw_after_colonisation_inter_dunn1, infer = TRUE, adjust = 'mvt'))
```

Significant differences in week 3 Cip-NC, Cip-Pro, Mix-NC and Mix-Pro.

Now check for differences within weeks.

```{r}
anova(model_afdw_after_colonisation)
# week variable highly significant 

afdw_after_colonisation_inter2 <- lsmeans(model_afdw_after_colonisation, ~ week |  treatment)

afdw_after_colonisation_inter_dunn2 <- contrast(afdw_after_colonisation_inter2, 'pairwise')

(afdw_after_colonisation_inter_dunn_sum2 <- summary(afdw_after_colonisation_inter_dunn2, infer = TRUE, adjust = 'mvt'))
```

-   Cip

    -   Difference between week 1 and 3

-   Mix

    -   Week 1 & 2 similar; week 1 & 3 and 2 & 3 different.

-   NC

    -   all weeks different from each other

-   Pro

    -   Week 1 & 2 similar; week 1 & 3 and 2 & 3 different.

#### Boxplot AFDW after colonisation

```{r}
afdw_after_colonisation_boxplot <- ggplot(data = biofilm_endpoints, 
               mapping = aes(x = factor(week), 
                             y = afdw_after_colonisation, 
                             fill = factor(treatment, levels = c("NC", "Cip", "Pro", "Mix")))) +
  geom_boxplot() +
  scale_fill_manual(values = c("NC" = "#009E73", "Cip" = "#56B4E9", "Pro" = "#E69F00", "Mix" = "#F0E442")) + 
  scale_color_okabe_ito() + 
  theme_bw() +
  labs(x = NULL , y = expression("AFDW (mg/cm"^"2"*")"), fill = "Treatment") +
  scale_x_discrete(labels = c("Week 1", "Week 2", "Week 3")) 

afdw_after_colonisation_boxplot 

# Now combine with chlorophyll plots
combined_plot <- chla_plot / chlb_plot / chlc_plot / afdw_after_colonisation_boxplot + 
  plot_layout(guides = "collect") & 
  theme(legend.position = "bottom")

# save the plot
ggsave(
  "output/biofilm_endpoints.png",
  plot = combined_plot,
  width = 25,
  height = 30,
  units = "cm",
  dpi = 600
)

```

# 4 Dryweight and growth snails

Unit of dry weight = mg

```{r}
# insert data (Culture snails from in-house breeding)
dryweight_snails <-
  read.table(file.path("data\\dryweight_snails.txt"),
             header = TRUE, sep = "\t")

View(dryweight_snails)
# 3 snails from the 20 baseline (culture) snails were lost during the analyses.
```

#### Calculate mean dryweight of all treatments and the culture snails

```{r}
mean_dryweight_snails <- dryweight_snails %>%
  group_by(treatment) %>%
  dplyr::summarize(mean_dryweight_snails = mean(dryweight_mg),
            sd_dryweight_snails = sd(dryweight_mg))
  

print(mean_dryweight_snails)

```

Snails from feeding assay (fed with biofilm) on average heavier than snails from culture (fed with Tetraphyll flakes).

Calculate growth:

```{r}
growth_snails <- dryweight_snails %>%
  dplyr::mutate(growth = ((dryweight_mg - median(dryweight_mg[treatment == "Culture"])) / 21) * 1000) 
# unit: growth_µg_dw_per_day
# multiply by 1000 to obtain a growth unit of µg dryweight per day

# calculate the growth in percentage in comparison to culture snails
percental_growth <- growth_snails %>%
  group_by(treatment) %>%
  dplyr::summarize(mean_dryweight = mean(dryweight_mg)) %>%
  mutate(percental_growth = (mean_dryweight / mean_dryweight[treatment == "Culture"]) * 100 - 100)

percental_growth

```

#### Histogram snail growth

```{r}
# filter culture snails
filtered_growth <- growth_snails %>%
  filter(!treatment %in% c("Culture"))

hist(filtered_growth $ growth) 
skewness(filtered_growth $ growth, na.rm = TRUE) 
shapiro.test(filtered_growth $ growth)
```

Growth normal distributed.

#### Model growth snails

```{r}
# explanatory variable treatment as factor
filtered_growth $ treatment <- factor(filtered_growth $ treatment)
  
model_growth_snails <- lm(growth ~ treatment, data = filtered_growth)

plot(model_growth_snails) # point cloud looks evenly distributed --> normal distribution

qreference(residuals(model_growth_snails)) # normal distribution

anova(model_growth_snails)
```

Explanatory variable "treatment" highly significant. Now check differences between treatments:

```{r}
growth_snails_inter <- lsmeans(model_growth_snails, ~ treatment)

growth_snails_inter_dunn1 <- contrast(growth_snails_inter, 'pairwise')

growth_snails_inter_dunn_sum1 <- summary(growth_snails_inter_dunn1, infer = TRUE, adjust = 'mvt')

growth_snails_inter_dunn1

```

-   Significant difference between growth of Cip and Mix snails compared to NC; almost significant difference (p=0.0564) between Pro treatment and NC

#### Boxplot growth snails

```{r}

# Plot
growth_snails_boxplot <- 
  ggplot(data = filtered_growth %>%
                 filter(growth > 0),
   mapping = aes(x = factor(treatment, levels = c("NC", "Cip", "Pro", "Mix")), y =    growth)) +
  # Mean 
  stat_summary(fun = mean, geom = "point", aes(fill = treatment),
               shape = 21, size = 5, color = "grey20", stroke = 0.5) +

  # 95% CI 
  # Black outline (thicker, underneath)
stat_summary(fun.data = mean_cl_normal, geom = "errorbar",
             aes(group = treatment),  # don't map color here
             width = 0.2, linewidth = 1, color = "grey30") +

# Colored CI on top
stat_summary(fun.data = mean_cl_normal, geom = "errorbar",
             aes(color = treatment),
             width = 0.2, linewidth = 0.8) +

#  stat_summary(fun.data = mean_cl_normal, geom = "errorbar", aes(color = treatment   ), width = 0.2, linewidth = 0.7) +

  # Jittered 
  geom_point(aes(color = treatment),
             position = position_jitter(width = 0.1, height = 0),
             alpha = 0.7, size = 2) +

  # Color scales
  scale_fill_manual(values = c("NC" = "#009E73", "Cip" = "#56B4E9", "Pro" = "#E69F00", "Mix" = "#F0E442")) +
  scale_color_manual(values = c("NC" = "#009E73", "Cip" = "#56B4E9", "Pro" = "#E69F00", "Mix" = "#F0E442")) +
  guides(fill = "none") +

  # Theme and labels
  #theme_bw() +
  labs(x = "Treatment", y = "Growth (µg dryweight / day)", color = "") +
  scale_y_continuous(limits = c(0,30), breaks = seq(0,30, by = 5)) +
  scale_x_discrete(labels = c("NC", "Cip", "Pro", "Mix")) +
  #labs(title = expression(paste("Growth of ", italic("Potamopyrgus antipodarum"), " after 21-day feeding assay"))) +

  theme_minimal(base_size = 11) +
  theme(
      legend.position = "none",
      panel.background = element_rect(fill = "grey100"),
      axis.title = element_text(
      size = 11,
      face = "bold",
      colour = "grey30"
    ),
    panel.border = element_rect(fill = NA, colour = "grey30"),
   # legend.key = element_blank(),
   # legend.title = element_text(
     # size = 11,
     # face = "bold",
      #colour = "grey30"
   # ),
    #legend.text = element_text(size = 10, colour = "grey30"),
    panel.grid.major = element_line(colour = "grey95", size = 0.4),
    panel.grid.minor = element_line(colour = "grey95", size = 0.2)
  ) +

  # Significance annotations
  annotate("text", x = 1.5, y = max(filtered_growth$growth) + 3,
           label = "p=0.022*", size = 3, color = "grey30") +
  geom_segment(aes(x = 1, xend = 2, y = max(growth) + 2, yend = max(growth) + 2),
               lineend = "round", linewidth = 0.5, color = "grey30") +
  annotate("text", x = 2.5, y = max(filtered_growth$growth) + 7,
           label = "p=0.005**", size = 3, color = "grey30") +
  geom_segment(aes(x = 1, xend = 4, y = max(growth) + 6, yend = max(growth) + 6),
               lineend = "round", linewidth = 0.5, color = "grey30") +
  annotate("text", x = 2, y = max(filtered_growth$growth) + 5,
           label = "p=0.056", size = 3, color = "grey30") +
  geom_segment(aes(x = 1, xend = 3, y = max(growth) + 4, yend = max(growth) + 4),
               lineend = "round", linewidth = 0.5, color = "grey30")


growth_snails_boxplot
```

```{r}
# save the plot
ggsave(
  "output/growth_snails_boxplot.png",
  width = 16.5,
  height = 12,
  units = "cm",
  dpi = 600
)
```

# 5 Fatty acid composition snails

Statistics for fatty acid composition of snails in absolute values (µg fatty acid / mg dry weight snail).

```{r}
# insert data
FA_Snails_absolute <-
  read.table(file.path("data\\FA_Snails_absolute.txt"),
             header = TRUE, sep = "\t")

View(FA_Snails_absolute)
# Mix S_29 no data, fatty acid extraction did not work
```

#### Total SAFA, MUFA & PUFA content

1\) Total SAFA content

```{r}
# total SAFA, MUFA, PUFA content; sa = "snail absolute" 
# total SAFA content 
# filter saturated fatty acids in new table 
SAFA_sa <- FA_Snails_absolute %>%
filter(fa %in% c("C12:0", "C13:0", "C14:0", "C15:0", "C16:0", "C18:0", "C20:0"))

# Sum SAFA per replicate and treatment
total_SAFA_per_snail <- SAFA_sa %>% 
  group_by(replicate, treatment) %>% 
  summarize(sum_SAFA = sum(fa_conc), .groups = "drop") 
print(total_SAFA_per_snail) 

# Calculate mean and standard deviation per treatment
mean_SAFA_sa <- total_SAFA_per_snail %>%
  group_by(treatment) %>%
  summarize(
    mean_SAFA = mean(sum_SAFA, na.rm = TRUE),
    sd_SAFA = sd(sum_SAFA, na.rm = TRUE),
    .groups = "drop"
  )

# Extract mean of control group for percental difference
control_mean_SAFA_sa <- mean_SAFA_sa %>%
  filter(treatment == "NC") %>%
  pull(mean_SAFA)

# Calculate percental difference from control
mean_SAFA_sa <- mean_SAFA_sa %>%
  mutate(percent_diff_from_NC = 100 * (mean_SAFA - control_mean_SAFA_sa) / control_mean_SAFA_sa)

print(mean_SAFA_sa)
```

2\) Total MUFA content

```{r}
# filter mono-unsaturated fatty acids in new table
MUFA_sa <- FA_Snails_absolute %>%
  filter(fa %in% c("C16:1", "C18:1", "C20:1", "C22:1"))

# Sum MUFA per replicate and treatment
total_MUFA_per_snail <- MUFA_sa %>% 
  group_by(replicate, treatment) %>% 
  summarize(sum_MUFA = sum(fa_conc), .groups = "drop") 
print(total_MUFA_per_snail) 

# Calculate mean and standard deviation per treatment
mean_MUFA_sa <- total_MUFA_per_snail %>%
  group_by(treatment) %>%
  summarize(
    mean_MUFA = mean(sum_MUFA, na.rm = TRUE),
    sd_MUFA = sd(sum_MUFA, na.rm = TRUE),
    .groups = "drop"
  )

# Extract mean of control group for percental difference
control_mean_MUFA_sa <- mean_MUFA_sa %>%
  filter(treatment == "NC") %>%
  pull(mean_MUFA)

# Calculate percental difference from control
mean_MUFA_sa <- mean_MUFA_sa %>%
  mutate(percent_diff_from_NC = 100 * (mean_MUFA - control_mean_MUFA_sa) / control_mean_MUFA_sa)

print(mean_MUFA_sa)
```

3\) Total PUFA content

```{r}
# filter poly-unsaturated fatty acids in new table
PUFA_sa <- FA_Snails_absolute %>%
  filter(fa %in% c("C18:2(n-6c)", "C18:3(n-3)", "C18:3(n-6)", "C18:4(n-3)", "C20:2(n-6)", "C20:3(n-6)", "C20:4(n-6)", "C20:5(n-3)", "C22:5(n-3)", "C22:6(n-3)"))
  
# Sum PUFA per replicate and treatment
total_PUFA_per_snail <- PUFA_sa %>% 
  group_by(replicate, treatment) %>% 
  summarize(sum_PUFA = sum(fa_conc), .groups = "drop") 
print(total_PUFA_per_snail) 

# Calculate mean and standard deviation per treatment
mean_PUFA_sa <- total_PUFA_per_snail %>%
  group_by(treatment) %>%
  summarize(
    mean_PUFA = mean(sum_PUFA, na.rm = TRUE),
    sd_PUFA = sd(sum_PUFA, na.rm = TRUE),
    .groups = "drop"
  )

# Extract mean of control group for percental difference
control_mean_PUFA_sa <- mean_PUFA_sa %>%
  filter(treatment == "NC") %>%
  pull(mean_PUFA)

# Calculate percental difference from control
mean_PUFA_sa <- mean_PUFA_sa %>%
  mutate(percent_diff_from_NC = 100 * (mean_PUFA - control_mean_PUFA_sa) / control_mean_PUFA_sa)

print(mean_PUFA_sa)
```

4\) Total FA per snail

```{r}
# total FA 
total_FA_snails <- bind_rows(total_SAFA_per_snail, total_MUFA_per_snail, total_PUFA_per_snail)

total_FA_per_snail <- total_FA_snails %>%
  group_by(replicate, treatment) %>%
  summarize(sum_FA = sum(sum_SAFA, sum_MUFA, sum_PUFA, na.rm = TRUE))
print(total_FA_per_snail)

# Calculate mean and standard deviation per treatment
mean_FA_sa <- total_FA_per_snail %>%
  group_by(treatment) %>%
  summarize(
    mean_FA = mean(sum_FA, na.rm = TRUE),
    sd_FA = sd(sum_FA, na.rm = TRUE),
    .groups = "drop"
  )

# Extract mean of control group for percental difference
control_mean_FA_sa <- mean_FA_sa %>%
  filter(treatment == "NC") %>%
  pull(mean_FA)

# Calculate percental difference from control
mean_FA_sa <- mean_FA_sa %>%
  mutate(percent_diff_from_NC = 100 * (mean_FA - control_mean_FA_sa) / control_mean_FA_sa)

print(mean_FA_sa)
```

#### Histogram FA snails

```{r}
hist(FA_Snails_absolute $ fa_conc) 
skewness(FA_Snails_absolute $ fa_conc, na.rm = TRUE) 
shapiro.test(FA_Snails_absolute $ fa_conc)
```

Fatty acid concentration not normal distributed and right skewed.

#### NMDS FA snails

Change table to wide format:

```{r}
# change table to wide format (sqrt for square root transformation in next step)
FA_Snails_absolute_wide_sqrt <-
  pivot_wider(FA_Snails_absolute,
              names_from = fa,
              values_from = fa_conc)

# square root transformation, to reduce discriminatory power of dominant FA groups
FA_Snails_absolute_wide_sqrt[, 3:23] <- sqrt(FA_Snails_absolute_wide_sqrt[, 3:23])

# Only keep columns/FAs with more than 5 non-zero values (fatty acids that have just been found in 5 replicates will be removed)
FA_Snails_absolute_wide_sqrt <-
  FA_Snails_absolute_wide_sqrt %>%
  dplyr::select(
    1:2,
    where(~sum(. !=0, na.rm = TRUE)>5)
  )
```

Perform metaMDS:

```{r}
# Remove first and second column for metaMDS
FA_Snails_absolute_wide_sqrt <- FA_Snails_absolute_wide_sqrt[, -c(1, 2)]

# Rename (- C, - () ) fatty acids for later plot
FA_Snails_absolute_wide_sqrt <- FA_Snails_absolute_wide_sqrt %>%
  rename(
    `12:0`     = `C12:0`,   
    `13:0`     = `C13:0`,
    `14:0`     = `C14:0`,
    `15:0`     = `C15:0`,
    `16:0`     = `C16:0`,
    `16:1`     = `C16:1`,
    `18:0`     = `C18:0`,
    `18:1`     = `C18:1`,
    `18:2n-6c` = `C18:2(n-6c)`,
    `18:3n-6`  = `C18:3(n-6)`,
    `18:3n-3`  = `C18:3(n-3)`,
    `18:4n-3`  = `C18:4(n-3)`,
    `20:0`     = `C20:0`,
    `20:1`     = `C20:1`,
    `20:2n-6`  = `C20:2(n-6)`,
    `20:3n-6`  = `C20:3(n-6)`,
    `20:4n-6`  = `C20:4(n-6)`,
    `20:5n-3`  = `C20:5(n-3)`,
    `22:1`     = `C22:1`,
    `22:5n-3`  = `C22:5(n-3)`,
    `22:6n-3`  = `C22:6(n-3)`
)


NMDS_Snails_absolute <-
  metaMDS(
    FA_Snails_absolute_wide_sqrt,
    autotransform = FALSE,
    k = 2 ,
    distance = "bray",
    na.rm = TRUE
  )

NMDS_Snails_absolute$stress
```

Create fa vectors

```{r}
# separation in SAFA; MUFA & PUFA to build vectors in NMDS
# "sa" = snail absolute
FA_variables.SAFAs_sa <-
  envfit(
    NMDS_Snails_absolute,
    subset(
      FA_Snails_absolute_wide_sqrt,
      select = c(
        "12:0",
        "13:0",
        "14:0",
        "15:0",
        "16:0",
        "18:0",
        "20:0"
      )
    )
    ,
    permutations = 999,
    na.rm = TRUE
  )

FA_variables.MUFAs_sa <-
  envfit(
    NMDS_Snails_absolute,
    subset(
      FA_Snails_absolute_wide_sqrt,
      select = c(
        "16:1",
        "18:1",
        "20:1",
        "22:1"
      )
    )
    ,
    permutations = 999,
    na.rm = TRUE
  )

FA_variables.PUFAs_sa <-
  envfit(
    NMDS_Snails_absolute,
    subset(
      FA_Snails_absolute_wide_sqrt,
      select = c(
        "18:2n-6c",
        "18:3n-6",
        "18:3n-3",
        "18:4n-3",
        "20:2n-6",
        "20:3n-6",
        "20:4n-6",
        "20:5n-3",
        "22:5n-3",
        "22:6n-3"
      )
    )
    ,
    permutations = 999,
    na.rm = TRUE
  )

# Extract NMDS scores 
data.scores_snail_absolute = as.data.frame(scores(NMDS_Snails_absolute)$sites)

Treatments <-
  FA_Snails_absolute $ treatment[FA_Snails_absolute $ fa == "C16:0"]

# add treatment column to NMDS scores
data.scores_snail_absolute$Treatment = Treatments

# vectors for SAFA, MUFA and PUFA
SAFAs_coord_cont_sa = as.data.frame(scores(FA_variables.SAFAs_sa, "vectors")) * ordiArrowMul(FA_variables.SAFAs_sa)
MUFAs_coord_cont_sa = as.data.frame(scores(FA_variables.MUFAs_sa, "vectors")) * ordiArrowMul(FA_variables.MUFAs_sa)
PUFAs_coord_cont_sa = as.data.frame(scores(FA_variables.PUFAs_sa, "vectors")) * ordiArrowMul(FA_variables.PUFAs_sa)
```

Plot NMDS:

```{r, warning=FALSE}
# For the plot: reorder treatment variable in data.scores_snail_absolute
data.scores_snail_absolute$Treatment <- factor(
  data.scores_snail_absolute$Treatment,
  levels = c("NC", "Cip", "Pro", "Mix")
)

# Compute convex hull for each Treatment
polygon_data_snails <- data.scores_snail_absolute %>%
  group_by(Treatment) %>%
  summarise(hull_points = list(data.frame(
    NMDS1 = NMDS1[chull(NMDS1, NMDS2)],
    NMDS2 = NMDS2[chull(NMDS1, NMDS2)]
  ))) %>%
  unnest(hull_points)

# Function to rescale coordinates
rescale_vectors <- function(df, scale_factor = 1) {
  df$length <- sqrt(df$NMDS1^2 + df$NMDS2^2)
  max_length <- max(df$length)
  df$NMDS1 <- df$NMDS1 / max_length * scale_factor
  df$NMDS2 <- df$NMDS2 / max_length * scale_factor
  df$length <- NULL
  return(df)
}

# Apply to each group with same scale_factor (e.g. 0.2)
PUFAs_coord_cont_sa <- rescale_vectors(PUFAs_coord_cont_sa, 0.4)
MUFAs_coord_cont_sa <- rescale_vectors(MUFAs_coord_cont_sa, 0.4)
SAFAs_coord_cont_sa <- rescale_vectors(SAFAs_coord_cont_sa, 0.4)

# plot
absol_snail_with_vectors <- ggplot(data = data.scores_snail_absolute, aes(x = NMDS1, y = NMDS2)) +
    geom_polygon(
    data = polygon_data_snails,
    aes(x = NMDS1, y = NMDS2, group = Treatment, fill = Treatment),
    alpha = 0.2,
    show.legend = FALSE
    ) +
  geom_point(
    aes(shape = Treatment, color = Treatment),
    size = 5,
    alpha = 1,
    show.legend = TRUE
  ) +
  scale_shape_manual(values = c("NC" = 16, "Pro" = 17, "Cip" = 18, "Mix" = 15)) +
  scale_color_manual(values = c("NC" = "#009E73", "Pro" = "#E69F00", "Cip" = "#56B4E9", "Mix" = "#F0E442")) +
  scale_fill_manual(values = c("NC" = "#009E73", "Pro" = "#E69F00", "Cip" = "#56B4E9", "Mix" = "#F0E442")) + 
  geom_segment(aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2), 
               data = PUFAs_coord_cont_sa, linewidth = 0.5, alpha = 0.2, colour = "black") +
  geom_segment(aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2), 
               data = MUFAs_coord_cont_sa, linewidth = 0.5, alpha = 0.2, colour = "black") +
  geom_segment(aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2), 
               data = SAFAs_coord_cont_sa, linewidth = 0.5, alpha = 0.2, colour = "black") +
  geom_text_repel(
    data = PUFAs_coord_cont_sa, 
    aes(x = NMDS1, y = NMDS2, label = row.names(PUFAs_coord_cont_sa)), 
    colour = "black",
    size = 2.5) +
  geom_text_repel(
    data = MUFAs_coord_cont_sa, 
    aes(x = NMDS1, y = NMDS2, label = row.names(MUFAs_coord_cont_sa)), 
    colour = "#996600",
    size = 2.5) +
  geom_text_repel(
    data = SAFAs_coord_cont_sa, 
    aes(x = NMDS1, y = NMDS2, label = row.names(SAFAs_coord_cont_sa)), 
    colour = "#0000FF",
    size = 2.5) +
  theme_bw() +
  theme(
    axis.title = element_text(
      size = 11,
      face = "bold",
      colour = "grey30"
    ),
    panel.background = element_blank(),
    panel.border = element_rect(fill = NA, colour = "grey30"),
    legend.key = element_blank(),
    legend.title = element_text(
      size = 11,
      face = "bold",
      colour = "grey30"
    ),
    legend.text = element_text(size = 10, colour = "grey30"),
    panel.grid.major = element_line(colour = "grey95", size = 0.4),
    panel.grid.minor = element_line(colour = "grey95", size = 0.2)
  ) 

absol_snail_with_vectors
# position of FA names manually adjusted in vector graphic
```

Save the plot as .png:

```{r}
ggsave(
  "output/NMDS_snails_absolute_vectors.png",
  width = 16.5,
  height = 12,
  units = "cm",
  dpi = 600
)
```

Save the plot as vector graphic (minor adaptions for better readability with Inkscape):

```{r}
ggsave(
  "output/NMDS_snails_absolute_vectors.svg",
  device = "svg",
  width = 16.5,
  height = 16,
  units = "cm",
  dpi = 600
)
```

Now plot NMDS without fatty acid vectors:

```{r}
# plot
absol_snail <- ggplot(data = data.scores_snail_absolute, aes(x = NMDS1, y = NMDS2)) +
    geom_polygon(
    data = polygon_data_snails,
    aes(x = NMDS1, y = NMDS2, group = Treatment, fill = Treatment),
    alpha = 0.2,
    show.legend = FALSE
  ) +
  geom_point(
    aes(shape = Treatment, color = Treatment),
    size = 5,
    alpha = 1,
    show.legend = TRUE
  ) +
  scale_shape_manual(values = c("NC" = 16, "Pro" = 17, "Cip" = 18, "Mix" = 15)) +
  scale_color_manual(values = c("NC" = "#009E73", "Pro" = "#E69F00", "Cip" = "#56B4E9", "Mix" = "#F0E442")) +
  scale_color_manual(values = c("NC" = "#009E73", "Pro" = "#E69F00", "Cip" = "#56B4E9", "Mix" = "#F0E442")) +  
  scale_fill_manual(values = c("NC" = "#009E73", "Pro" = "#E69F00", "Cip" = "#56B4E9", "Mix" = "#F0E442")) + 
  theme_bw() +
  theme(
    axis.title = element_text(
      size = 11,
      face = "bold",
      colour = "grey30"
    ),
    panel.background = element_blank(),
    panel.border = element_rect(fill = NA, colour = "grey30"),
    legend.key = element_blank(),
    legend.title = element_text(
      size = 11,
      face = "bold",
      colour = "grey30"
    ),
    legend.text = element_text(size = 10, colour = "grey30"),
    panel.grid.major = element_line(colour = "grey95", size = 0.4),
    panel.grid.minor = element_line(colour = "grey95", size = 0.2)
  ) 

absol_snail
```

Save the plot as .png:

```{r}
ggsave(
  "output/NMDS_snails_absolute.png",
  width = 16.5,
  height = 12,
  units = "cm",
  dpi = 600
)
```

Save the plot as vector graphic:

```{r}
ggsave(
  "output/NMDS_snails_absolute.svg",
  device = "svg",
  width = 16.5,
  height = 16,
  units = "cm",
  dpi = 600
)
```

#### PERMANOVA FA snails

Create distance matrix with fatty acids:

```{r}
FA_Snails_dist <- vegdist(FA_Snails_absolute_wide_sqrt, method = 'bray')
```

Separate explanatory variables from fatty acid information:

```{r}
# change initial table to wide format again
FA_Snails_absolute_wide_permanova <-
  pivot_wider(FA_Snails_absolute,
              names_from = fa,
              values_from = fa_conc)

# square root transformation again, to reduce discriminatory power of dominant fa groups
FA_Snails_absolute_wide_permanova[, 3:28] <- sqrt(FA_Snails_absolute_wide_permanova[, 3:28])

# Only keep columns/FAs with more than 5 non-zero values (fatty acids that have just been found in 5 replicates will be removed)
FA_Snails_absolute_wide_permanova <-
  FA_Snails_absolute_wide_permanova %>%
  dplyr::select(
    1:2,
    where(~sum(. !=0, na.rm = TRUE)>5)
  )

# delete fa columns
FA_Snails_absolute_wide_permanova <- FA_Snails_absolute_wide_permanova[, c(1, 2)]
FA_Snails_absolute_wide_permanova $ treatment <- as.factor(FA_Snails_absolute_wide_permanova $ treatment)
```

Perform PERMANOVA:

```{r}
perm_FA_Snails <- adonis2(FA_Snails_dist ~ treatment, data = FA_Snails_absolute_wide_permanova, permutations = 999)
print(perm_FA_Snails)
```

Pairwise comparisons (by: Martinez Arbizu, P. (2020). pairwiseAdonis: Pairwise multilevel comparison using adonis. R package version 0.4):

```{r}
pairwise_results <- pairwise.adonis2(
  FA_Snails_dist ~ treatment,
  data = FA_Snails_absolute_wide_permanova,
  permutations = 999
)

# View the results
print(pairwise_results)
```

All treatments except of Pro_vs_Mix significant differences.

#### SIMPER FA snails

```{r}
# Perform SIMPER
simper_FA_Snails <-
  with(
    FA_Snails_absolute_wide_permanova,
    simper(FA_Snails_absolute_wide_sqrt, treatment,  permutations = 999)
  )

summary(simper_FA_Snails)

str(simper_FA_Snails)

# Generate list of SIMPER tables where contributors explain >70% dissimilarity
simper_top70_tables <- map(simper_FA_Snails, function(x) {
  if (!is.null(x$p)) {
    # Create a data frame for this comparison
    df <- data.frame(
      fatty_acid = names(x$average),
      average = x$average,
      sd = x$sd,
      ratio = x$ratio,
      cusum = x$cusum,
      p = x$p,
      stringsAsFactors = FALSE
    ) %>%
      arrange(desc(average))  # ensure correct ordering for cusum cutoff

    # Filter for cumulative contribution <= 70%
    df %>% filter(cusum <= 0.7)
  } else {
    NULL
  }
})

# Name the list elements with comparison names
names(simper_top70_tables) <- names(simper_FA_Snails)

# save csv
walk2(
  simper_top70_tables,
  names(simper_top70_tables),
  ~
    write.csv(.x, file = file.path(
      "output", paste0("SIMPER_Snails_top70_", .y, ".csv")
    ), row.names = FALSE)
)

# Combine results into a data frame
simper_results_snails <- do.call(rbind, lapply(simper_FA_Snails, function(x) {
  if (!is.null(x$p)) {
    data.frame(
      fatty_acid = names(x$p),  
      p = as.numeric(x$p),      
      average = as.numeric(x$average),  
      sd = as.numeric(x$sd),
      ratio = as.numeric(x$ratio),
      cusum = as.numeric(x$cusum),
      stringsAsFactors = FALSE
    )
  } else {
    NULL  
  }
}))

# Show only significant results
simper_results_snails %>%
  dplyr::filter(p <= 0.05) %>%
  dplyr::select(fatty_acid, p, average, sd, ratio, cusum)  

# How often do they occur
simper_results_snails %>%
  dplyr::filter(p <= 0.05) %>%
  dplyr::count(fatty_acid) %>%
  dplyr::arrange(desc(n)) %>%
  print()
```

# 6 Fatty acid composition biofilm

Statistics for fatty acid composition of the biofilm in absolute values (µg fatty acid / mg dry weight biofilm).

```{r}
# insert data
FA_Biofilm_absolute <-
  read.table(file.path("data\\FA_Biofilm_absolute.txt"),
             header = TRUE,
             sep = "\t")

View(FA_Biofilm_absolute)
```

#### Total SAFA, MUFA & PUFA content

1\) Total SAFA content

```{r}
# total SAFA, MUFA, PUFA content; ba = "biofilm absolute"
# total SAFA content
# filter saturated fatty acids in new table
SAFA_ba <- FA_Biofilm_absolute %>%
    filter(fa %in% c("C12:0", "C13:0", "C14:0", "C15:0", "C16:0", "C18:0", "C20:0", "C22:0", "C24:0"))

# Sum SAFA per replicate and treatment
total_SAFA_per_replicate <- SAFA_ba %>%
  group_by(replicate, treatment) %>%
  summarize(sum_SAFA = sum(fa_conc), .groups = "drop")
print(total_SAFA_per_replicate)

# Calculate mean and standard deviation per treatment
mean_SAFA_ba <- total_SAFA_per_replicate %>%
  group_by(treatment) %>%
  summarize(
    mean_SAFA = mean(sum_SAFA, na.rm = TRUE),
    sd_SAFA = sd(sum_SAFA, na.rm = TRUE),
    .groups = "drop"
  )

# Extract mean of control group for percental difference
control_mean_SAFA_ba <- mean_SAFA_ba %>%
  filter(treatment == "NC") %>%
  pull(mean_SAFA)

# Calculate percental difference from control
mean_SAFA_ba <- mean_SAFA_ba %>%
  mutate(percent_diff_from_NC = 100 * (mean_SAFA - control_mean_SAFA_ba) / control_mean_SAFA_ba)

print(mean_SAFA_ba)

```

2\) Total MUFA content

```{r}
# filter mono-unsaturated fatty acids in new table
MUFA_ba <- FA_Biofilm_absolute %>%
  filter(fa %in% c("C16:1", "C17:1", "C18:1", "C20:1","C22:1"))

# Sum MUFA per replicate and treatment
total_MUFA_per_replicate <- MUFA_ba %>%
  group_by(replicate, treatment) %>%
  summarize(sum_MUFA = sum(fa_conc), .groups = "drop")
print(total_MUFA_per_replicate)

# Calculate mean and standard deviation per treatment
mean_MUFA_ba <- total_MUFA_per_replicate %>%
  group_by(treatment) %>%
  summarize(
    mean_MUFA = mean(sum_MUFA, na.rm = TRUE),
    sd_MUFA = sd(sum_MUFA, na.rm = TRUE),
    .groups = "drop"
  )

# Extract mean of control group for percental difference
control_mean_MUFA_ba <- mean_MUFA_ba %>%
  filter(treatment == "NC") %>%
  pull(mean_MUFA)

# Calculate percental difference from control
mean_MUFA_ba <- mean_MUFA_ba %>%
  mutate(percent_diff_from_NC = 100 * (mean_MUFA - control_mean_MUFA_ba) / control_mean_MUFA_ba)

print(mean_MUFA_ba)
```

3\) Total PUFA content

```{r}
# filter poly-unsaturated fatty acids in new table
PUFA_ba <- FA_Biofilm_absolute %>%
  filter(fa %in% c("C18:2(n-6c)", "C18:3(n-3)", "C18:3(n-6)", "C18:4(n-3)", "C20:2(n-6)", "C20:3(n-6)", "C20:4(n-6)", "C20:5(n-3)", "C22:6(n-3)"))
  
# Sum PUFA per replicate and treatment
total_PUFA_per_replicate <- PUFA_ba %>%
  group_by(replicate, treatment) %>%
  summarize(sum_PUFA = sum(fa_conc), .groups = "drop")
print(total_PUFA_per_replicate)

# Calculate mean and standard deviation per treatment
mean_PUFA_ba <- total_PUFA_per_replicate %>%
  group_by(treatment) %>%
  summarize(
    mean_PUFA = mean(sum_PUFA, na.rm = TRUE),
    sd_PUFA = sd(sum_PUFA, na.rm = TRUE),
    .groups = "drop"
  )

# Extract mean of control group for percental difference
control_mean_PUFA_ba <- mean_PUFA_ba %>%
  filter(treatment == "NC") %>%
  pull(mean_PUFA)

# Calculate percental difference from control
mean_PUFA_ba <- mean_PUFA_ba %>%
  mutate(percent_diff_from_NC = 100 * (mean_PUFA - control_mean_PUFA_ba) / control_mean_PUFA_ba)

print(mean_PUFA_ba)
```

4\) Total FA content

```{r}
# total FA 
total_FA_biofilm <- bind_rows(total_SAFA_per_replicate, total_MUFA_per_replicate, total_PUFA_per_replicate)

total_FA_per_replicate <- total_FA_biofilm %>%
  group_by(replicate, treatment) %>%
  summarize(sum_FA = sum(sum_SAFA, sum_MUFA, sum_PUFA, na.rm = TRUE))
print(total_FA_per_replicate)

# Calculate mean and standard deviation per treatment
mean_FA_ba <- total_FA_per_replicate %>%
  group_by(treatment) %>%
  summarize(
    mean_FA = mean(sum_FA, na.rm = TRUE),
    sd_FA = sd(sum_FA, na.rm = TRUE),
    .groups = "drop"
  )

# Extract mean of control group for percental difference
control_mean_FA_ba <- mean_FA_ba %>%
  filter(treatment == "NC") %>%
  pull(mean_FA)

# Calculate percental difference from control
mean_FA_ba <- mean_FA_ba %>%
  mutate(percent_diff_from_NC = 100 * (mean_FA - control_mean_FA_ba) / control_mean_FA_ba)

print(mean_FA_ba)
```

#### Histogram FA biofilm

```{r}
hist(FA_Biofilm_absolute $ fa_conc) 
skewness(FA_Biofilm_absolute $ fa_conc, na.rm = TRUE) 
shapiro.test(FA_Biofilm_absolute $ fa_conc)
```

Biofilm fatty acid concentration not normal distributed and right skewed.

#### NMDS FA biofilm

Change table to wide format:

```{r}
# add prefix (week) 1,2 or 3 to treatments for later subgrouping
FA_Biofilm_absolute <- FA_Biofilm_absolute %>%
  mutate(treatment = if_else(week == "1", paste("1_", treatment, sep = ""), treatment))

FA_Biofilm_absolute <- FA_Biofilm_absolute %>%
  mutate(treatment = if_else(week == "2", paste("2_", treatment, sep = ""), treatment))

FA_Biofilm_absolute <- FA_Biofilm_absolute %>%
  mutate(treatment = if_else(week == "3", paste("3_", treatment, sep = ""), treatment))

# change table to wide format
FA_Biofilm_absolute_wide_sqrt <-
  pivot_wider(FA_Biofilm_absolute,
              names_from = fa,
              values_from = fa_conc)

# square root transformation, to reduce discriminatory power of dominant fa groups
FA_Biofilm_absolute_wide_sqrt[, 4:37] <- sqrt(FA_Biofilm_absolute_wide_sqrt[, 4:37])

View(FA_Biofilm_absolute_wide_sqrt)

# Only keep columns/FAs with more than 5 non-zero values (fatty acids that have just been found in 5 replicates will be removed)
FA_Biofilm_absolute_wide_sqrt <-
  FA_Biofilm_absolute_wide_sqrt %>%
  dplyr::select(
    1:3,
    where(~sum(. !=0, na.rm = TRUE)>5)
  )
```

Perform metaMDS:

```{r}
# Rename and ease col-names for later NMDS plot
FA_Biofilm_absolute_wide_sqrt <- FA_Biofilm_absolute_wide_sqrt %>%
  rename(
    `12:0`     = `C12:0`,   
    `13:0`     = `C13:0`,
    `14:0`     = `C14:0`,
    `15:0`     = `C15:0`,
    `16:0`     = `C16:0`,
    `16:1`     = `C16:1`,
    `17:1`     = `C17:1`,
    `18:0`     = `C18:0`,
    `18:1`     = `C18:1`,
    `18:2n-6c` = `C18:2(n-6c)`,
    `18:3n-6`  = `C18:3(n-6)`,
    `18:3n-3`  = `C18:3(n-3)`,
    `18:4n-3`  = `C18:4(n-3)`,
    `20:0`     = `C20:0`,
    `20:1`     = `C20:1`,
    `20:2n-6`  = `C20:2(n-6)`,
    `20:4n-6`  = `C20:4(n-6)`,
    `20:5n-3`  = `C20:5(n-3)`,
    `22:0`     = `C22:0`,
    `22:1`     = `C22:1`,
    `22:6n-3`  = `C22:6(n-3)`,
    `24:0`     = `C24:0`
  )

# Remove columns 1:3 for metaMDS
FA_Biofilm_absolute_wide_sqrt <- FA_Biofilm_absolute_wide_sqrt[, -c(1, 2, 3)]

NMDS_Biofilm_absolute <-
  metaMDS(
    FA_Biofilm_absolute_wide_sqrt,
    autotransform = FALSE,
    k = 2 ,
    distance = "bray",
    na.rm = TRUE
  )

NMDS_Biofilm_absolute$stress
```

Create fa vectors

```{r}
# separation in SAFA; MUFA & PUFA to build vectors in NMDS
# "ba" = biofilm absolute
FA_variables.SAFAs_ba <-
  envfit(
    NMDS_Biofilm_absolute,
    subset(
      FA_Biofilm_absolute_wide_sqrt,
      select = c(
        "12:0",
        "13:0",
        "14:0",
        "15:0",
        "16:0",
        "18:0",
        "20:0",
        "22:0",
        "24:0"
      )
    )
    ,
    permutations = 999,
    na.rm = TRUE
  )

FA_variables.MUFAs_ba <-
  envfit(
    NMDS_Biofilm_absolute,
    subset(
      FA_Biofilm_absolute_wide_sqrt,
      select = c(
        "16:1",
        "17:1",
        "18:1",
        "20:1",
        "22:1"
      )
    )
    ,
    permutations = 999,
    na.rm = TRUE
  )

FA_variables.PUFAs_ba <-
  envfit(
    NMDS_Biofilm_absolute,
    subset(
      FA_Biofilm_absolute_wide_sqrt,
      select = c(
        "18:2n-6c",
        "18:3n-6",
        "18:3n-3",
        "18:4n-3",
        "20:2n-6",
        "20:4n-6",
        "20:5n-3",
        "22:6n-3"
      )
    )
    ,
    permutations = 999,
    na.rm = TRUE
  )

# Extract NMDS scores
data.scores_biofilm_absolute = as.data.frame(scores(NMDS_Biofilm_absolute)$sites)

Treatments <-
  FA_Biofilm_absolute$treatment[FA_Biofilm_absolute$fa == "C16:0"]

# add treatment column to NMDS scores
data.scores_biofilm_absolute$Treatment = Treatments

# vectors for SAFA, MUFA and PUFA
SAFAs_coord_cont_ba = as.data.frame(scores(FA_variables.SAFAs_ba, "vectors")) * ordiArrowMul(FA_variables.SAFAs_ba)
MUFAs_coord_cont_ba = as.data.frame(scores(FA_variables.MUFAs_ba, "vectors")) * ordiArrowMul(FA_variables.MUFAs_ba)
PUFAs_coord_cont_ba = as.data.frame(scores(FA_variables.PUFAs_ba, "vectors")) * ordiArrowMul(FA_variables.PUFAs_ba)

# create weeks as subgroups for plotting
data.scores_biofilm_absolute <- data.scores_biofilm_absolute %>%
  mutate(time.point = case_when(
    str_detect(Treatment, "^1_") ~ "Week1",
    str_detect(Treatment, "^2_") ~ "Week2",
    str_detect(Treatment, "^3_") ~ "Week3",
    TRUE ~ "Other"
  ))

```

```{r}
# first delete 1_, 2_ & 3_ prefix in treatments
data.scores_biofilm_absolute $ Treatment <- sub("1_", "", data.scores_biofilm_absolute $ Treatment)
data.scores_biofilm_absolute $ Treatment <- sub("2_", "", data.scores_biofilm_absolute $ Treatment)
data.scores_biofilm_absolute $ Treatment <- sub("3_", "", data.scores_biofilm_absolute $ Treatment)
```

Plot NMDS:

```{r}
# Reorder treatment variable for the plot
data.scores_biofilm_absolute$Treatment <- factor(
  data.scores_biofilm_absolute$Treatment,
  levels = c("NC", "Cip", "Pro", "Mix")
)

# Compute convex hull for each Treatment
polygon_data_biofilm_treatment <- data.scores_biofilm_absolute %>%
  group_by(Treatment) %>%
  summarise(hull_points = list(data.frame(
    NMDS1 = NMDS1[chull(NMDS1, NMDS2)],
    NMDS2 = NMDS2[chull(NMDS1, NMDS2)]
  ))) %>%
  unnest(hull_points)

# plot
absol_biofilm_with_vectors = ggplot(data = data.scores_biofilm_absolute, aes(x = NMDS1, y = NMDS2)) +
    geom_polygon(
    data = polygon_data_biofilm_treatment,
    aes(x = NMDS1, y = NMDS2, group = Treatment, fill = Treatment),
    alpha = 0.2,
    show.legend = FALSE  
  ) +
  geom_point( aes(shape = time.point, color = Treatment),
    size = 5,
    alpha = 1
  ) +
  scale_shape_manual(values = c(15, 16, 17)) +
  scale_color_manual(values = c("NC" = "#009E73", "Pro" = "#E69F00", "Cip" = "#56B4E9", "Mix" = "#F0E442")) +
  scale_fill_manual(values = c("NC" = "#009E73", "Pro" = "#E69F00", "Cip" = "#56B4E9", "Mix" = "#F0E442")) +
  labs(shape = "Week", color = "Treatment") +
  geom_segment(data = PUFAs_coord_cont_ba, aes(x = 0, y = 0, xend = NMDS1 * 0.3, yend = NMDS2 * 0.3), linewidth = 0.5, alpha = 0.2, colour = "black") +
  geom_segment(data = MUFAs_coord_cont_ba, aes(x = 0, y = 0, xend = NMDS1 * 0.2, yend = NMDS2 * 0.2), linewidth = 0.5, alpha = 0.2, colour = "black") +
  geom_segment(data = SAFAs_coord_cont_ba, aes(x = 0, y = 0, xend = NMDS1 * 0.3, yend = NMDS2 * 0.3), linewidth = 0.5, alpha = 0.2, colour = "black") +
  geom_text_repel(
    data = PUFAs_coord_cont_ba, 
    aes(x = NMDS1 * 0.3, y = NMDS2 * 0.3, label = row.names(PUFAs_coord_cont_ba)), 
    colour = "black",
    size = 3) +
  geom_text_repel(
    data = MUFAs_coord_cont_ba, 
    aes(x = NMDS1 * 0.2, y = NMDS2 * 0.2, label = row.names(MUFAs_coord_cont_ba)), 
    colour = "#996600",
    size = 3) +
  geom_text_repel(
    data = SAFAs_coord_cont_ba, 
    aes(x = NMDS1 * 0.3, y = NMDS2 * 0.3, label = row.names(SAFAs_coord_cont_ba)), 
    colour = "#0000FF",
    size = 3) +
  theme_bw() +
  theme(
    axis.title = element_text(
      size = 11,
      face = "bold",
      colour = "grey30"
    ),
    panel.background = element_blank(),
    panel.border = element_rect(fill = NA, colour = "grey30"),
    legend.key = element_blank(),
    legend.title = element_text(
      size = 11,
      face = "bold",
      colour = "grey30"
    ),
    legend.text = element_text(size = 10, colour = "grey30"),
    plot.margin = margin(
      t = 0.2,
      b = 0.2,
      l = 0.2,
      r = 0.2,
      unit = "cm"
    ),
    panel.grid.major = element_line(colour = "grey95", size = 0.4),
    panel.grid.minor = element_line(colour = "grey95", size = 0.2)
  ) 
  
absol_biofilm_with_vectors
# position of FA names manually adjusted in vector graphic with Inkscape
```

Save the plot as .png:

```{r}
ggsave(
  "output/NMDS_Biofilm_absolute.png",
  width = 16.5,
  height = 12,
  units = "cm",
  dpi = 600
)
```

Save the plot as vector graphic:

```{r}
ggsave(
  "output/NMDS_Biofilm_absolute.svg",
  device = "svg",
  width = 16.5,
  height = 16,
  units = "cm",
  dpi = 600
)
```

Now plot again without vectors and treatment polygon:

```{r}
absol_biofilm_treatment_polygon = ggplot(data = data.scores_biofilm_absolute, aes(x = NMDS1, y = NMDS2)) +
    geom_polygon(
    data = polygon_data_biofilm_treatment,
    aes(x = NMDS1, y = NMDS2, group = Treatment, fill = Treatment),
    alpha = 0.2,
    show.legend = FALSE  
  ) +
  geom_point( aes(shape = time.point, color = Treatment),
    size = 5,
    alpha = 1
  ) +
  scale_shape_manual(values = c(15, 16, 17)) +
  scale_color_manual(values = c("NC" = "#009E73", "Pro" = "#E69F00", "Cip" = "#56B4E9", "Mix" = "#F0E442")) +
  scale_fill_manual(values = c("NC" = "#009E73", "Pro" = "#E69F00", "Cip" = "#56B4E9", "Mix" = "#F0E442")) +
  labs(shape = "Week", color = "Treatment") +
  theme_bw() +
  theme(
    axis.title = element_text(
      size = 11,
      face = "bold",
      colour = "grey30"
    ),
    panel.background = element_blank(),
    panel.border = element_rect(fill = NA, colour = "grey30"),
    legend.key = element_blank(),
    legend.title = element_text(
      size = 11,
      face = "bold",
      colour = "grey30"
    ),
    legend.text = element_text(size = 10, colour = "grey30"),
    plot.margin = margin(
      t = 0.2,
      b = 0.2,
      l = 0.2,
      r = 0.2,
      unit = "cm"
    ),
    panel.grid.major = element_line(colour = "grey95", size = 0.4),
    panel.grid.minor = element_line(colour = "grey95", size = 0.2)
  ) 
  
  
absol_biofilm_treatment_polygon
```

Save the plot as .png:

```{r}
ggsave(
  "output/NMDS_Biofilm_treatment_polygon.png",
  width = 16.5,
  height = 12,
  units = "cm",
  dpi = 600
)
```

Now plot with week polygon:

```{r}
# Compute convex hull for each week
polygon_data_biofilm_week <- data.scores_biofilm_absolute %>%
  group_by(time.point) %>%
  summarise(hull_points = list(data.frame(
    NMDS1 = NMDS1[chull(NMDS1, NMDS2)],
    NMDS2 = NMDS2[chull(NMDS1, NMDS2)]
  ))) %>%
  unnest(hull_points)

# Plot
absol_biofilm_week_polygon = ggplot(data = data.scores_biofilm_absolute, aes(x = NMDS1, y = NMDS2)) +
    geom_polygon(
    data = polygon_data_biofilm_week,
    aes(x = NMDS1, y = NMDS2, group = time.point, fill = time.point),
    alpha = 0.2,
    show.legend = FALSE  
  ) +
  geom_point( aes(shape = time.point, color = Treatment),
    size = 5,
    alpha = 1
  ) +
  scale_shape_manual(values = c(15, 16, 17)) +
  scale_color_manual(values = c("NC" = "#009E73", "Pro" = "#E69F00", "Cip" = "#56B4E9", "Mix" = "#F0E442")) +
  scale_fill_manual(values = c("NC" = "#009E73", "Pro" = "#E69F00", "Cip" = "#56B4E9", "Mix" = "#F0E442")) +
  labs(shape = "Week", color = "Treatment") +
  theme_bw() +
  theme(
    axis.title = element_text(
      size = 11,
      face = "bold",
      colour = "grey30"
    ),
    panel.background = element_blank(),
    panel.border = element_rect(fill = NA, colour = "grey30"),
    legend.key = element_blank(),
    legend.title = element_text(
      size = 11,
      face = "bold",
      colour = "grey30"
    ),
    legend.text = element_text(size = 10, colour = "grey30"),
    plot.margin = margin(
      t = 0.2,
      b = 0.2,
      l = 0.2,
      r = 0.2,
      unit = "cm"
    ),
    panel.grid.major = element_line(colour = "grey95", size = 0.4),
    panel.grid.minor = element_line(colour = "grey95", size = 0.2)
  ) 
  
  
absol_biofilm_week_polygon
```

Save the plot as .png:

```{r}
ggsave(
  "output/NMDS_Biofilm_week_polygon.png",
  width = 16.5,
  height = 12,
  units = "cm",
  dpi = 600
)
```

#### PERMANOVA FA biofilm

Create distance matrix with fatty acids:

```{r}
FA_biofilm_dist <- vegdist(FA_Biofilm_absolute_wide_sqrt, method = 'bray')
```

Separate explanatory variables from fatty acid information:

```{r}
FA_biofilm_absolute_wide_permanova <-
  pivot_wider(FA_Biofilm_absolute,
              names_from = fa,
              values_from = fa_conc)

# square root transformation again, to reduce discriminatory power of dominant fa groups
FA_biofilm_absolute_wide_permanova[, 4:37] <- sqrt(FA_biofilm_absolute_wide_permanova[, 4:37])

# delete fa columns
FA_biofilm_absolute_wide_permanova <-
  FA_biofilm_absolute_wide_permanova[, c(1, 2, 3)]

FA_biofilm_absolute_wide_permanova$week <-
  as.factor(FA_biofilm_absolute_wide_permanova$week)

FA_biofilm_absolute_wide_permanova$treatment <-
  as.factor(FA_biofilm_absolute_wide_permanova$treatment)
```

Perform PERMANOVA for treatments:

```{r}
# first remove 1_, 2_, 3_ and prefixes
FA_biofilm_absolute_wide_permanova$treatment <- gsub("^[1-3]_","", FA_biofilm_absolute_wide_permanova$treatment)

# perform PERMANOVA for treatments
perm_FA_biofilm_treatment <- adonis2(FA_biofilm_dist ~ treatment, data = FA_biofilm_absolute_wide_permanova, permutations = 999)
print(perm_FA_biofilm_treatment)

```

No significant differences between biofilm treatments. Anyway pairwise comparisons (by: Martinez Arbizu, P. (2020). pairwiseAdonis: Pairwise multilevel comparison using adonis. R package version 0.4):

```{r}
pairwise_results_biofilm_treatment <- pairwise.adonis2(
  FA_biofilm_dist ~ treatment,
  data = FA_biofilm_absolute_wide_permanova,
  permutations = 99
)

# View the results
print(pairwise_results_biofilm_treatment)
```

No significant differences. Now PERMANOVA for week variable:

```{r}
perm_FA_biofilm_week <- adonis2(FA_biofilm_dist ~ week, data = FA_biofilm_absolute_wide_permanova, permutations = 999)
print(perm_FA_biofilm_week)
```

Now pairwise tests:

```{r}
pairwise_results_biofilm_week <- pairwise.adonis2(
  FA_biofilm_dist ~ week,
  data = FA_biofilm_absolute_wide_permanova,
  permutations = 99
)

# View the results
print(pairwise_results_biofilm_week)
```

All weeks different, as already indicated by NMDS.

#### SIMPER FA biofilm

First SIMPER for treatments:

```{r}
# create extra table with treatments
biofilm_treatments_table <- FA_biofilm_absolute_wide_permanova

View(biofilm_treatments_table)

# Perform SIMPER
simper_FA_biofilm_treatments <-
  with(
    biofilm_treatments_table,
    simper(FA_Biofilm_absolute_wide_sqrt, treatment,  permutations = 999)
  )

summary(simper_FA_biofilm_treatments)
```

```{r}
# extract FAs which explain >70% of dissimilarities
simper_top70_tables_biofilm_treat <- map(simper_FA_biofilm_treatments, function(x) {
  if (!is.null(x$p)) {
    # Create a data frame for this comparison
    df <- data.frame(
      fatty_acid = names(x$average),
      average = x$average,
      sd = x$sd,
      ratio = x$ratio,
      cusum = x$cusum,
      p = x$p,
      stringsAsFactors = FALSE
    ) %>%
      arrange(desc(average))  

    # Filter for cumulative contribution <= 70%
    df %>% filter(cusum <= 0.7)
  } else {
    NULL
  }
})

# Name the list elements with comparison names
names(simper_top70_tables_biofilm_treat) <- names(simper_FA_biofilm_treatments)

# save csv
walk2(
  simper_top70_tables_biofilm_treat,
  names(simper_top70_tables_biofilm_treat),
  ~
    write.csv(.x, file = file.path(
      "output", paste0("SIMPER_Biofilm_Treatment_top70_", .y, ".csv")
    ), row.names = FALSE)
)
```

```{r}
# Combine results into a data frame
simper_results_biofilm_treatments <- do.call(rbind, lapply(simper_FA_biofilm_treatments, function(x) {
  # Check if the element contains valid data
  if (!is.null(x$p)) {
    data.frame(
      fatty_acid = names(x$p),  
      p = as.numeric(x$p),      
      average = as.numeric(x$average),  
      sd = as.numeric(x$sd),
      ratio = as.numeric(x$ratio),
      cusum = as.numeric(x$cusum),
      stringsAsFactors = FALSE
    )
  } else {
    NULL  # Skip if no p-values
  }
}))

# Show only significant results
simper_results_biofilm_treatments %>%
  dplyr::filter(p <= 0.05) %>%
  dplyr::select(fatty_acid, p, average, sd, ratio, cusum)  
```

```{r}
# How often do they occur
simper_results_biofilm_treatments %>%
  dplyr::filter(p <= 0.05) %>%
  dplyr::count(fatty_acid) %>%
  dplyr::arrange(desc(n)) %>%
  print()
```

Now perform SIMPER for weeks:

```{r}
simper_FA_biofilm_week <- with(biofilm_treatments_table, simper(FA_Biofilm_absolute_wide_sqrt, week,  permutations = 999))

summary(simper_FA_biofilm_week)
```

```{r}
# extract FAs which explain >70% of dissimilarities
simper_top70_tables_biofilm_week <- map(simper_FA_biofilm_week, function(x) {
  if (!is.null(x$p)) {
    # Create a data frame for this comparison
    df <- data.frame(
      fatty_acid = names(x$average),
      average = x$average,
      sd = x$sd,
      ratio = x$ratio,
      cusum = x$cusum,
      p = x$p,
      stringsAsFactors = FALSE
    ) %>%
      arrange(desc(average))  

    # Filter for cumulative contribution <= 70%
    df %>% filter(cusum <= 0.7)
  } else {
    NULL
  }
})

# Name the list elements with comparison names
names(simper_top70_tables_biofilm_week) <- names(simper_FA_biofilm_week)

# save csv
walk2(
  simper_top70_tables_biofilm_week,
  names(simper_top70_tables_biofilm_week),
  ~
    write.csv(.x, file = file.path(
      "output", paste0("SIMPER_Biofilm_Week_top70_", .y, ".csv")
    ), row.names = FALSE)
)
```

```{r}
# Combine results into a data frame
simper_results_biofilm_week <- do.call(rbind, lapply(simper_FA_biofilm_week, function(x) {
  # Check if the element contains valid data
  if (!is.null(x$p)) {
    data.frame(
      fatty_acid = names(x$p),  
      p = as.numeric(x$p),      
      average = as.numeric(x$average),  
      sd = as.numeric(x$sd),
      ratio = as.numeric(x$ratio),
      cusum = as.numeric(x$cusum),
      stringsAsFactors = FALSE
    )
  } else {
    NULL  # Skip if no p-values
  }
}))

# Show only significant results
simper_results_biofilm_week %>%
  dplyr::filter(p <= 0.05) %>%
  dplyr::select(fatty_acid, p, average, sd, ratio, cusum)  
```

```{r}
# How often do they occur
simper_results_biofilm_week %>%
  dplyr::filter(p <= 0.05) %>%
  dplyr::count(fatty_acid) %>%
  dplyr::arrange(desc(n)) %>%
  print()
```

# 7 Temperature and light conditions

## 7.1 Biofilm colonisation and contamination

```{r}
# insert data
colonisation_temp_light <-
  read.table(file.path("data\\colonisation_temp_light.txt"),
             header = TRUE, sep = "\t")

View(colonisation_temp_light)

# average temperature
mean(colonisation_temp_light $ temperature_colonisation)
# 18.28751 °C

# mean light intensity
# filter only daytime
average_light_intensity_colonisation <- colonisation_temp_light %>%
  dplyr::filter(light_intensity_colonisation > 0) %>%
  summarize(average_light_intensity_colonisation = mean(light_intensity_colonisation))

print(average_light_intensity_colonisation)
# 14019.26 lux

# temperature plot
# add a small amount of random noise to the y-values
colonisation_temp_light $ temperature_colonisation <- jitter(colonisation_temp_light $ temperature_colonisation)

t2 <- ggplot(
  data = colonisation_temp_light,
  mapping = aes( x = day, y = temperature_colonisation, color = time_of_day)) +
  geom_point(alpha = 0.5) +  # Individual measurements with transparency
  geom_smooth() +
  labs(
    title = "Temperature profile during biofilm colonisation",
    x = "Day", y = "Temperature (°C)"
  ) +
  theme_bw() +
  scale_x_continuous(breaks = seq(0, 77, by = 5)) +
  scale_color_continuous(
    name = "Time of Day",  
    breaks = c(0, 5, 10, 15, 20),
    labels = c("08:00 pm", "03:00 pm", "10:00 am", "05:00 am", "0:00 am"),
    limits = c(0, 22),
  )

t2 

# save the plot
ggsave(
  "output/temperature_colonisation.png",
  width = 16.5,
  height = 12,
  units = "cm",
  dpi = 600
)

# light intensity plot
# Filter the data to include only rows where light_intensity > 0
filtered_data2 <- colonisation_temp_light %>% 
  filter(light_intensity_colonisation > 0)

# add a small amount of random noise to the y-values
filtered_data2 $ light_intensity_colonisation <- jitter(filtered_data2 $ light_intensity_colonisation)

l2 <- ggplot(
  data = filtered_data2,
  mapping = aes( x = day, y = light_intensity_colonisation)) +
  geom_point(size = 1) +  # Individual measurements with transparency
  geom_smooth(color = "#E69F00", linewidth = 1) +
  labs(
    title = "Light intensity profile during biofilm colonisation",
    x = "Day", y = "Light intensity (lux)"
  ) +
  theme_bw() +
  scale_x_continuous(breaks = seq(0, 77, by = 5)) 

l2 

# save the plot
ggsave(
  "output/light_intensity_colonisation.png",
  width = 16.5,
  height = 12,
  units = "cm",
  dpi = 600
)
```

## 7.2 Feeding assay

```{r}
# insert data
feeding_assay_temp_light <-
  read.table(file.path("data\\feeding_assay_temp_light.txt"),
             header = TRUE, 
             sep = "\t"
  )
             
View(feeding_assay_temp_light)

# average temperature
mean(feeding_assay_temp_light $ temperature)
# 13.27927 °C

# mean light intensity
# filter only daytime
average_light_intensity_feeding <- feeding_assay_temp_light %>%
  dplyr::filter(light_intensity > 0) %>%
  summarize(average_light_intensity_feeding = mean(light_intensity))

print(average_light_intensity_feeding)
# 330.3552 lux

# temperature plot
# add a small amount of random noise to the y-values
feeding_assay_temp_light $ temperature <- jitter(feeding_assay_temp_light $ temperature)

t1 <- ggplot(
  data = feeding_assay_temp_light,
  mapping = aes( x = day, y = temperature, color = time_of_day)) +
  geom_point(alpha = 0.5) +  # Individual measurements with transparency
  geom_smooth() +
  labs(
    title = "Temperature profile during 21-day feeding assay",
    x = "Day", y = "Temperature (°C)"
  ) +
  theme_bw() +
  scale_x_continuous(breaks = 0:21) +
  scale_color_continuous(
    name = "Time of Day",  
    breaks = c(0, 5, 10, 15, 20),
    labels = c("08:00 pm", "03:00 pm", "10:00 am", "05:00 am", "0:00 am"),
    limits = c(0, 22),
  )

t1 

# save the plot
ggsave(
  "output/temperature_feeding_assay.png",
  width = 16.5,
  height = 12,
  units = "cm",
  dpi = 600
)

# light intensity plot
# Filter the data to include only rows where light_intensity > 0
filtered_data1 <- feeding_assay_temp_light %>% 
  filter(light_intensity > 0)

# add a small amount of random noise to the y-values
filtered_data1 $ light_intensity <- jitter(filtered_data1 $ light_intensity)

l1 <- ggplot(
  data = filtered_data1,
  mapping = aes( x = day, y = light_intensity)) +
  geom_point(size = 1) +  # Individual measurements with transparency
  geom_smooth(color = "#E69F00", linewidth = 1) +
  labs(
    title = "Light intensity profile during 21-day feeding assay",
    x = "Day", y = "Light intensity (lux)"
  ) +
  theme_bw() +
  scale_x_continuous(breaks = 0:21) 

l1 

# save the plot
ggsave(
  "output/light_intensity_feeding_assay.png",
  width = 16.5,
  height = 12,
  units = "cm",
  dpi = 600
)
```
