---
title: "Bottom-up effects of an herbicide and an antibiotic on *Potamopyrgus antipodarum*"
format: 
  html:
    toc: true
    style: "floating"
---

Sophie Oster, Verena C. Schreiner, Tobias Schmitt, Patrick Fink, Markus Weitere, Ibrahim Arayoe, Ralf Schulz, Mirco Bundschuh

R script written by Sophie Oster\
checked by Verena C. Schreiner\
RPTU Kaiserslautern-Landau\
Fortstrasse 7\
76829 Landau in der Pfalz\
GERMANY

Corresponding authors email address:\
1) Sophie Oster: oster\@rptu.de\
2) Mirco Bundschuh: mirco.bundschuh\@rptu.de

This study was conducted with four treatments and three time points explanatory variables

-   treatment (categorical)
    -   uncontaminated control (negative control, NC)
    -   herbicide propyzamide (Pro)
    -   antibiotic ciprofloxacin (Cip)
    -   mixture of propyzamide and ciprofloxacin (Mix)
-   time.point (categorical)
    -   week 1
    -   week 2
    -   week 3

# 1 Packages in a reproducible environment

```{r, message = FALSE, warning = FALSE}
# renv::init() 
# renv::snapshot()

library(stats)
library(rstatix)
library(dplyr)
library(tidyverse)
library(tidyr)
library(broom)
library(dunn.test)
library(patchwork)
library(vegan)
library(data.table)
library(ggplot2)
library(ggrepel)
library(ggforce)
library(concaveman)
library(gridExtra)
library(stringr)
library(shades)
library(ggpubr)
library(base)
library(ggokabeito)
library(multcomp)
library(car)
library(asbio)
library(plotrix)
library(lmtest)
library(boot)
library(moments)
library(lme4)
library(DAAG)
library(performance)
library(emmeans)
library(purrr)
library(DT)
```

# 2 Set working directory

```{r}
prj <- getwd()
setwd(prj)
```

# 3 Biofilm endpoints

## Chlorophyll and ash free dry weight of biofilm after colonisation + chronic contamination of 14 days

-   Unit chlorophyll a, b & c: µg chlorophyll / cm²

-   Unit afdw after colonisation: mg / cm²

```{r}
# insert data
biofilm_endpoints <-
  read.table(file.path(prj, "data\\biofilm_endpoints.txt"),
             header = TRUE, sep = "\t")

View(biofilm_endpoints)
```

## 3.1 Chlorophyll

#### Caclulate mean chla, chlb + chlc of each treatment, to compare in manuscript

```{r, message = FALSE, warning = FALSE}
#chl a
mean_chlorophyll_a <- biofilm_endpoints %>%
  group_by(week, treatment) %>%
  summarize(mean_chlorophyll_a = mean(chla))

print(mean_chlorophyll_a)

# chl b
mean_chlorophyll_b <- biofilm_endpoints %>%
  group_by(week, treatment) %>%
  summarize(mean_chlorophyll_b = mean(chlb))

print(mean_chlorophyll_b)

# chl c
mean_chlorophyll_c <- biofilm_endpoints %>%
  group_by(week, treatment) %>%
  summarize(mean_chlorophyll_c = mean(chlc))

print(mean_chlorophyll_c)
```

#### Histogram chlorophyll a, b + c

```{r}
# chl a
a <- hist(biofilm_endpoints $ chla) 
skewness(biofilm_endpoints $ chla, na.rm = TRUE) 

# chl b
b <- hist(biofilm_endpoints $ chlb) 
skewness(biofilm_endpoints $ chlb, na.rm = TRUE) 

# chl c
c <- hist(biofilm_endpoints $ chlc) 
skewness(biofilm_endpoints $ chlc, na.rm = TRUE) 
```

All not normal distributed and right skewed -\> log10 transformation in model

#### Model chla

Linear model with treatment and week and their interaction (week:treatment) as explanatory variables.

```{r}
# week as factor
biofilm_endpoints $ week <- factor(biofilm_endpoints $ week)

model_chla <- lm(log10(chla) ~ week + treatment + week:treatment, data = biofilm_endpoints)

plot(model_chla) # point cloud looks evenly distributed --> normal distribution

qreference(residuals(model_chla)) # normal distribution assumed

drop1(model_chla, test = 'F')

```

Interaction week:treatment significant (p=0.02965). Now check differences between treatments within weeks.

```{r}
chla_inter <- lsmeans(model_chla, ~ treatment | week )
chla_inter_dunn1 <- contrast(chla_inter, 'pairwise')
(chla_inter_dunn_sum1 <- summary(chla_inter_dunn1, infer = TRUE, adjust = 'mvt'))
```

No significant differences of the treatments within week 1 and 3. Only NC-Mix significant in week 2 (p=0.003).

Now check for differences between weeks.

```{r}
anova(model_chla)
# week variable highly significant (p=7.689e-08)

chla_inter2 <- lsmeans(model_chla, ~ week |  treatment)
chla_inter_dunn2 <- contrast(chla_inter2, 'pairwise')
(chla_inter_dunn_sum2 <- summary(chla_inter_dunn2, infer = TRUE, adjust = 'mvt'))

```

-   NC\
    -   Week 1 & 2 similar\
    -   Week 2 & 3 similar\
    -   difference between week 1 & 3
-   Cip\
    -   Week 1 & 2 similar\
    -   Week 1 & 3 similar\
    -   difference between week 2 & 3
-   Pro\
    -   Week 1 & 2 similar\
    -   differences between week 1 & 3 and 2 & 3
-   Mix\
    -   all weeks different

Week 1 & 2 quite similar, expect of Mix treatment. Week 3 quite different to week 1 & 2 despite of Cip tretment.

#### Model chlb

```{r}
model_chlb <- lm(log10(chlb) ~ week + treatment + week:treatment, data = biofilm_endpoints)

plot(model_chlb) # point cloud looks evenly distributed --> normal distribution

qreference(residuals(model_chlb)) # normal distribution assumed

drop1(model_chlb, test = 'F')
```

No significant differences. Still check for differences between treatments.

```{r}
chlb_inter <- lsmeans(model_chlb, ~ treatment | week )
chlb_inter_dunn1 <- contrast(chlb_inter, 'pairwise')
(chlb_inter_dunn_sum1 <- summary(chlb_inter_dunn1, infer = TRUE, adjust = 'mvt'))
```

No differences between treatments for chlb.

Now check for differences between weeks.

```{r}
anova(model_chlb)
# week variable highly significant (p=5.259e-05)

chlb_inter2 <- lsmeans(model_chlb, ~ week |  treatment)
chlb_inter_dunn3 <- contrast(chlb_inter2, 'pairwise')
(chlb_inter_dunn_sum3 <- summary(chlb_inter_dunn3, infer = TRUE, adjust = 'mvt'))
```

Similar results/trends as for chla. Week 1 & 2 quite similar; differences in:

-   Pro treatment: week 2 & 3

-   Mix treatment: week 1 & 2; 2 & 3

#### Model chlc

```{r}
model_chlc <- lm(log10(chlc) ~ week + treatment + week:treatment, data = biofilm_endpoints)

plot(model_chlc) # point cloud looks evenly distributed --> normal distribution

qreference(residuals(model_chlc)) # normal distribution assumed

drop1(model_chlc, test = 'F')
```

No significant differences. Still check for differences between treatments.

```{r}
chlc_inter <- lsmeans(model_chlc, ~ treatment | week )
chlc_inter_dunn1 <- contrast(chlc_inter, 'pairwise')
(chlc_inter_dunn_sum1 <- summary(chlc_inter_dunn1, infer = TRUE, adjust = 'mvt'))
```

No differences between treatments for chlorophyll c.

#### Boxplots chlorophyll a, b + c

```{r}
# chlorophyll a
chla <- ggplot(data = biofilm_endpoints, 
               mapping = aes(x = factor(week), 
                             y = chla, 
                             fill = factor(treatment, levels = c("NC", "Cip", "Pro", "Mix")))) +
  geom_boxplot() +
  scale_fill_manual(values = c("NC" = "#009E73", "Cip" = "#56B4E9", "Pro" = "#E69F00", "Mix" = "#F0E442")) + 
  scale_color_okabe_ito() + 
  theme_bw() +
  labs(x = NULL , y = expression("Chlorophyll a (µg/cm"^"2"*")"), fill = "Treatment") +
  scale_x_discrete(labels = c("Week 1", "Week 2", "Week 3"))

chla 

# save the plot
ggsave(chla, 
       filename = "chla.png", 
       path = 'C:\\Users\\Oster Test\\OneDrive\\Desktop\\Promotion\\Bioassays\\Schneckenversuch 2023\\Bottom_up_effects_biofilm_grazer\\output')


# chlorophyll b
chlb <- ggplot(data = biofilm_endpoints, 
               mapping = aes(x = factor(week), 
                             y = chlb, 
                             fill = factor(treatment, levels = c("NC", "Cip", "Pro", "Mix")))) +
  geom_boxplot() +
  scale_fill_manual(values = c("NC" = "#009E73", "Cip" = "#56B4E9", "Pro" = "#E69F00", "Mix" = "#F0E442")) + 
  scale_color_okabe_ito() + 
  theme_bw() +
  labs(x = NULL , y = expression("Chlorophyll b (µg/cm"^"2"*")"), fill = "Treatment") +
  scale_x_discrete(labels = c("Week 1", "Week 2", "Week 3"))

chlb 

# save the plot
ggsave(chlb, 
       filename = "chlb.png", 
       path = 'C:\\Users\\Oster Test\\OneDrive\\Desktop\\Promotion\\Bioassays\\Schneckenversuch 2023\\Bottom_up_effects_biofilm_grazer\\output')


# chlorophyll c
chlc <- ggplot(data = biofilm_endpoints, 
               mapping = aes(x = factor(week), 
                             y = chlc, 
                             fill = factor(treatment, levels = c("NC", "Cip", "Pro", "Mix")))) +
  geom_boxplot() +
  scale_fill_manual(values = c("NC" = "#009E73", "Cip" = "#56B4E9", "Pro" = "#E69F00", "Mix" = "#F0E442")) + 
  scale_color_okabe_ito() + 
  theme_bw() +
  labs(x = NULL , y = expression("Chlorophyll c (µg/cm"^"2"*")"), fill = "Treatment") +
  scale_x_discrete(labels = c("Week 1", "Week 2", "Week 3"))

chlc 

# save the plot
ggsave(chlc, 
       filename = "chlc.png", 
       path = 'C:\\Users\\Oster Test\\OneDrive\\Desktop\\Promotion\\Bioassays\\Schneckenversuch 2023\\Bottom_up_effects_biofilm_grazer\\output')
```

## 3.2 AFDW after colonisation

#### Calculate mean afdw of each treatment for manuscript.

```{r}
mean_afdw_after_colonisation <- biofilm_endpoints %>%
  group_by(week, treatment) %>%
  summarize(mean_afdw_after_colonisation = mean(afdw_after_colonisation))

print(mean_afdw_after_colonisation)
```

#### Histogram afdw after colonisation

```{r}
hist(biofilm_endpoints $ afdw_after_colonisation) 
skewness(biofilm_endpoints $ afdw_after_colonisation, na.rm = TRUE) 
```

Not normal distributed and right skewed. Log10 transformation for model

#### Model afdw after colonisation

Linear model with treatment and week and their interaction (week:treatment) as explanatory variables.

```{r}
model_afdw_after_colonisation <- lm(log10(afdw_after_colonisation) ~ week + treatment + week:treatment, data = biofilm_endpoints)

plot(model_afdw_after_colonisation) # not perfectly even distributed, but ok

qreference(residuals(model_afdw_after_colonisation)) # normal distribution assumed

drop1(model_afdw_after_colonisation, test = 'F')
```

Significant difference for week:treatment (p=0.009063). Check differences between treatments within weeks.

```{r}
afdw_after_colonisation_inter <- lsmeans(model_afdw_after_colonisation, ~ treatment | week )

afdw_after_colonisation_inter_dunn1 <- contrast(afdw_after_colonisation_inter, 'pairwise')

(afdw_after_colonisation_inter_dunn_sum1 <- summary(afdw_after_colonisation_inter_dunn1, infer = TRUE, adjust = 'mvt'))
```

Significant differences between week 2 Mix-NC and NC-Pro; and week 3 Cip-NC and Cip-Pro.

Now check for differences within weeks.

```{r}
anova(model_afdw_after_colonisation)
# week variable highly significant (p=2.541e-07)

afdw_after_colonisation_inter2 <- lsmeans(model_afdw_after_colonisation, ~ week |  treatment)

afdw_after_colonisation_inter_dunn2 <- contrast(afdw_after_colonisation_inter2, 'pairwise')

(afdw_after_colonisation_inter_dunn_sum2 <- summary(afdw_after_colonisation_inter_dunn2, infer = TRUE, adjust = 'mvt'))
```

-   Cip

    -   Difference between week 1 and 3

-   Mix

    -   Week 1 & 2 similar; week 1 & 3 and 2 & 3 different.

-   NC

    -   all weeks different from each other

-   Pro

    -   Week 1 & 2 similar; week 1 & 3 and 2 & 3 different.

#### Boxplot afdw after colonisation

```{r}
afdw_after_colonisation_boxplot <- ggplot(data = biofilm_endpoints, 
               mapping = aes(x = factor(week), 
                             y = afdw_after_colonisation, 
                             fill = factor(treatment, levels = c("NC", "Cip", "Pro", "Mix")))) +
  geom_boxplot() +
  scale_fill_manual(values = c("NC" = "#009E73", "Cip" = "#56B4E9", "Pro" = "#E69F00", "Mix" = "#F0E442")) + 
  scale_color_okabe_ito() + 
  theme_bw() +
  labs(x = NULL , y = expression("Ash free dry weight (mg/cm"^"2"*")"), fill = "Treatment") +
  scale_x_discrete(labels = c("Week 1", "Week 2", "Week 3"))

afdw_after_colonisation_boxplot 

# @ VS: Hier auch log Skala, um die Daten besser zu sehen?

# save the plot
ggsave(afdw_after_colonisation_boxplot, 
       filename = "afdw_after_colonisation.png", 
       path = 'C:\\Users\\Oster Test\\OneDrive\\Desktop\\Promotion\\Bioassays\\Schneckenversuch 2023\\Bottom_up_effects_biofilm_grazer\\output')

```

# 4 Feeding rate snails and ash free dry weight after feeding

-   NA's due to missing snail and zero values of feeding rate due to no measurable feeding in respective week

-   Unit feeding rate: µg / day and unit ash free dry weight (afdw): mg / cm²

-   Column "used_for_FA_analysis" highlights the ten snails used for FA analyses for each treatment

```{r}
# insert data
afdw_feeding_rate <-
  read.table(file.path(prj, "data\\feeding_assay_afdw_feeding_rate.txt"),
             header = TRUE, sep = "\t")

View(afdw_feeding_rate)
```

## 4.1 Feeding rate snails

#### Calculate mean feeding rate for each week

```{r}
mean_feeding_rate <- afdw_feeding_rate %>%
  group_by(week, treatment) %>%
  summarize(mean_feeding_rate = mean(feeding_rate, na.rm = TRUE))

print(mean_feeding_rate)
```

#### Histogram feeding rate

```{r}
hist(afdw_feeding_rate $ feeding_rate) 
skewness(afdw_feeding_rate $ feeding_rate, na.rm = TRUE) 
```

Feeding rate variable not normal distributed and right skewed -\> log10 transformation for model

#### Model feeding rate

```{r}
# week and beaker as factor
afdw_feeding_rate $ week <- factor(afdw_feeding_rate $ week)

afdw_feeding_rate $ beaker <- factor(afdw_feeding_rate $ beaker)

# @ VS: muss die explanatory variable "treatment" auch ein faktor sein? (Frage gilt auch für alle anderen Modelle)

# log10 transformation; adding 0.001 to variable for samples with feeding rate = 0
afdw_feeding_rate <- afdw_feeding_rate %>%
  mutate(log_feeding_rate = log10(feeding_rate + 0.001))

# response variable: log_feeding_rate
# fixed effects: week + treatment + week:treatment
# random effect with constant intercept: (1|beaker)
# first remove NAs
afdw_feeding_rate <- na.omit(afdw_feeding_rate)

feeding_rate_lmer <- lmer(log_feeding_rate ~ week + treatment + week:treatment + (1|beaker), data = afdw_feeding_rate)
```

Warning "boundary (singular) fit" because random effects of beaker are very small.

```{r}
isSingular(feeding_rate_lmer)
# isSingular = TRUE

# look at fitted model
ranef(feeding_rate_lmer)
```

No systematic effect by "beaker", so simplify model:

```{r}
model_feeding_rate <- lm(log_feeding_rate ~ week + treatment + week:treatment, data = afdw_feeding_rate)

plot(model_feeding_rate) # point cloud looks evenly distributed --> normal distribution

qreference(residuals(model_feeding_rate)) # normality not perfect, but ok

drop1(model_feeding_rate, test = 'Chisq')
# interaction of timepoint (week) and treatment highly significant
# now check at which timepoints is a difference between treatments
anova(model_feeding_rate)

# conduct Tukey HSD-test for pairwise comparisons
feeding_rate_tukey <- lsmeans(model_feeding_rate, ~ treatment | week)
feeding_rate_tukey

feeding_rate_tukey1 <- contrast(feeding_rate_tukey, 'pairwise')
feeding_rate_tukey1

feeding_rate_summary <- summary(feeding_rate_tukey1, infer = TRUE, adjust = 'mvt')
feeding_rate_summary

```

Treatments within week 2 & 3 similar. Within week 2 differences between Cip-Pro and Mix-Pro. Now check differences between weeks.

```{r}
feeding_rate_inter1 <- lsmeans(model_feeding_rate, ~ week |  treatment)

feeding_rate_inter_dunn1 <- contrast(feeding_rate_inter1, 'pairwise')

(feeding_rate_inter_dunn_sum1 <- summary(feeding_rate_inter_dunn1, infer = TRUE, adjust = 'mvt'))
```

Weeks for all treatments similiar, only difference for treatment Pro week 1 & 2 and 1 & 3.

#### Boxplot feeding rate after 7/14/21 days

```{r}
afdw_after_feeding_boxplot <- ggplot(data = afdw_feeding_rate, 
       mapping = aes(x = factor(week), 
                     y = afdw, 
                     fill = factor(treatment, levels = c("NC", "Cip", "Pro", "Mix")))) +
  geom_boxplot() +
  scale_fill_manual(values = c("NC" = "#009E73", "Cip" = "#56B4E9", "Pro" = "#E69F00", "Mix" = "#F0E442")) + 
  scale_color_okabe_ito() + 
  theme_bw() +
  labs(x = NULL , fill = "Treatment") +
  scale_x_discrete(labels = c("Week 1", "Week 2", "Week 3")) +
  scale_y_continuous(trans = "log10") +
  labs(y = expression("Feeding rate (µg/day)" ~ log[10]))

afdw_after_feeding_boxplot

# save the plot
ggsave(afdw_after_feeding_boxplot, 
       filename = "afdw_after_feeding_boxplot.png", 
       path = 'C:\\Users\\Oster Test\\OneDrive\\Desktop\\Promotion\\Bioassays\\Schneckenversuch 2023\\Bottom_up_effects_biofilm_grazer\\output')


```

## 4.2 AFDW after feeding

#### Calculate mean of afdw after feeding for each week

```{r}
mean_afdw_after_feeding <- afdw_feeding_rate %>%
  group_by(week, treatment) %>%
  summarize(mean_afdw_after_feeding = mean(afdw))

print(mean_afdw_after_feeding)
```

#### Histogram afdw after feeding

```{r}
hist(afdw_feeding_rate $ afdw) 
skewness(afdw_feeding_rate $ afdw, na.rm = TRUE) 
```

AFDW after feeding not normal distributed. Log10 transformation for model.

#### Model afdw after feeding

```{r}
# log10 transformation; adding 0.001 
afdw_feeding_rate <- afdw_feeding_rate %>%
  mutate(log_afdw = log10(afdw + 0.001))
# @ VS: Die Addition von 0.001 auch bei afdw machen, obwohl es keine 0-Werte gibt?

# response variable: log_afdw
# fixed effects: week + treatment + week:treatment
# random effect with constant intercept: (1|beaker)
afdw_after_feeding_lmer <- lmer(log_afdw ~ week + treatment + week:treatment + (1|beaker), data = afdw_feeding_rate)

afdw_after_feeding_lmer %>% r2
# Conditional R2: explained portion of the variance of the fixed and random effect together
# Marginal R2: declared share of the variance of the fixed components alone

plot(afdw_after_feeding_lmer) # point cloud looks evenly distributed --> normal distribution

qreference(residuals(afdw_after_feeding_lmer)) # normality

drop1(afdw_after_feeding_lmer, test = 'Chisq')
# interaction of timepoint (week) and treatment highly significant
# now check at which timepoints is a difference between treatments
anova(afdw_after_feeding_lmer)

# conduct Tukey HSD-test for pairwise comparisons
afdw_tukey <- lsmeans(afdw_after_feeding_lmer, ~ treatment | week)
afdw_tukey

afdw_tukey1 <- contrast(afdw_tukey, 'pairwise')
afdw_tukey

afdw_summary <- summary(afdw_tukey1, infer = TRUE, adjust = 'mvt')
afdw_summary
```

Here treatments quite different within weeks. Compare weeks:

```{r}
afdw_after_feeding_inter1 <- lsmeans(afdw_after_feeding_lmer, ~ week |  treatment)

afdw_after_feeding_dunn1 <- contrast(afdw_after_feeding_inter1, 'pairwise')

(afdw_after_feeding_dunn_sum1 <- summary(afdw_after_feeding_dunn1, infer = TRUE, adjust = 'mvt'))
```

Week 1 & 2 for treatments Mix & Pro similar. Everything else shows differences.

#### Boxplot afdw after feeding of 7/14/21 days

```{r}
afdw_after_feeding_boxplot <- ggplot(data = afdw_feeding_rate, 
       mapping = aes(x = factor(week), 
                     y = afdw, 
                     fill = factor(treatment, levels = c("NC", "Cip", "Pro", "Mix")))) +
  geom_boxplot() +
  scale_fill_manual(values = c("NC" = "#009E73", "Cip" = "#56B4E9", "Pro" = "#E69F00", "Mix" = "#F0E442")) + 
  scale_color_okabe_ito() + 
  theme_bw() +
  labs(x = NULL , fill = "Treatment") +
  scale_x_discrete(labels = c("Week 1", "Week 2", "Week 3")) +
  scale_y_continuous(trans = "log10") +
  labs(y = expression("AFDW (mg/cm"^"2"*")" ~ log[10]))

afdw_after_feeding_boxplot

# save the plot
ggsave(afdw_after_feeding_boxplot, 
       filename = "afdw_after_feeding_boxplot.png", 
       path = 'C:\\Users\\Oster Test\\OneDrive\\Desktop\\Promotion\\Bioassays\\Schneckenversuch 2023\\Bottom_up_effects_biofilm_grazer\\output')
```

# 5 Dryweight and growth snails

\@ VS: Ich hatte 20 Schnecken aus der Zucht eingewogen, um das Wachstum zu berechnen. Ich habe 3 Schnecken verloren, und die restlichen 7 (in der alten Tabelle, die du gesehen hast, waren es 10 Culture Schnecken) in der neuen Tabelle ergänzt. Die Schnecken wurden alle trocken gewogen. Also das ist das Gewicht nach Gefriertrocknung und ohne Schale, die habe ich bei allen Schnecken immer entfernt, da da teilweise Biofilm drauf gewachsen ist. Ich würde den Endpunkt an sich gerne drinlassen, du hast aber natürlich Recht, dass die Standardabweichung sehr groß ist, aber das stelle ich in den Boxplots ja dar.. was denkst du? Du hast da mehr Erfahrung, wenn du davon abrätst, dann nehme ich ihn raus. Ich habe die Schneckenzucht auch noch, also könnte ich noch mehr Culture snails einwiegen? Der Versuch aber dann 1 Jahr her.. Wahrscheinlich nicht mehr vergleichbar?

```{r}
# insert data
dryweight_snails <-
  read.table(file.path(prj, "data\\dryweight_snails.txt"),
             header = TRUE, sep = "\t")

View(dryweight_snails)
# 3 snails from the 20 baseline (culture) snails were lost during the analyses.
```

#### Calculate mean dryweight of all treatments and the culture snails

```{r}
mean_dryweight_snails <- dryweight_snails %>%
              group_by(treatment) %>%
              summarize(mean_dryweight_snails = mean(dryweight_mg))

print(mean_dryweight_snails)
```

Snails from feeding assay (fed with biofilm) heavier than snails from culture (fed with Tetraphyll flakes).

Calculate growth:

```{r}
growth_snails <- dryweight_snails %>%
  dplyr::mutate(growth_µg_dw_per_day = ((dryweight_mg - median(dryweight_mg[treatment == "Culture"])) / 7) * 1000) 
# multiply by 1000 to obtain a growth unit of µg dryweight per day

# calculate the growth in percentage in comparison to culture snails
percental_growth <- growth_snails %>%
  group_by(treatment) %>%
  summarize(mean_dryweight = mean(dryweight_mg)) %>%
  mutate(percental_growth = (mean_dryweight / mean_dryweight[treatment == "Culture"]) * 100 - 100)

percental_growth

```

#### Histogram snail growth

```{r}
# filter culture snails
filtered_growth <- growth_snails %>%
  filter(!treatment %in% c("Culture"))

hist(filtered_growth $ growth_µg_dw_per_day) 
skewness(filtered_growth $ growth_µg_dw_per_day, na.rm = TRUE) 
shapiro.test(filtered_growth $ growth_µg_dw_per_day)
```

Growth normal distributed.

#### Model growth snails

```{r}
model_growth_snails <- lm(growth_µg_dw_per_day ~ treatment, data = filtered_growth)

plot(model_growth_snails) # point cloud looks evenly distributed --> normal distribution

qreference(residuals(model_growth_snails)) # normal distribution

anova(model_growth_snails)
```

Explanatory variable "treatment" highly significant. Now check differences between treatments:

```{r}
growth_snails_inter <- lsmeans(model_growth_snails, ~ treatment)

growth_snails_inter_dunn1 <- contrast(growth_snails_inter, 'pairwise')

growth_snails_inter_dunn_sum1 <- summary(growth_snails_inter_dunn1, infer = TRUE, adjust = 'mvt')

growth_snails_inter_dunn1

```

-   Significant difference between growth of Cip and Mix snails compared to NC; almost significant difference (p=0.0564) between Pro treatment and NC

#### Boxplot growth snails

```{r}
# plot
growth_snails_boxplot <- ggplot(data = filtered_growth, 
                                mapping = aes(x = factor(treatment, levels = c("NC", "Cip", "Pro", "Mix")), 
                                              y = growth_µg_dw_per_day, 
                                              fill = factor(treatment, levels = c("NC", "Cip", "Pro", "Mix")))) +
  geom_boxplot() +
  scale_fill_manual(values = c("NC" = "#009E73", "Cip" = "#56B4E9", "Pro" = "#E69F00", "Mix" = "#F0E442"),
                    breaks = c("NC", "Cip", "Pro", "Mix")) + 
  scale_color_okabe_ito() + 
  theme_bw() +
  labs(x = NULL , y = "Growth (µg dryweight / day)", fill = "Treatment") +
  scale_x_discrete(labels = c("NC", "Cip", "Pro", "Mix")) +
  labs(title = expression(paste("Growth of ", italic("Potamopyrgus antipodarum"), " after 21-day feeding assay"))) +
  annotate("text", x = 1.5, y = max(filtered_growth$growth_µg_dw_per_day) + 2.5, 
           label = "*", size = 6) +
  geom_segment(aes(x = 1, xend = 2, y = max(growth_µg_dw_per_day) + 2, yend = max(growth_µg_dw_per_day) + 2),
               lineend = "round", linewidth = 0.8) +
  annotate("text", x = 2.5, y = max(filtered_growth$growth_µg_dw_per_day) + 6, 
           label = "**", size = 6) +
  geom_segment(aes(x = 1, xend = 4, y = max(growth_µg_dw_per_day) + 5, yend = max(growth_µg_dw_per_day) + 5),
               lineend = "round", linewidth = 0.8) 


growth_snails_boxplot

# save the plot
ggsave(growth_snails_boxplot, 
       filename = "growth_snails_boxplot.png", 
       path = 'C:\\Users\\Oster Test\\OneDrive\\Desktop\\Promotion\\Bioassays\\Schneckenversuch 2023\\Bottom_up_effects_biofilm_grazer\\output')
```

# 6 Fatty acid composition snails

```{r}
# insert data
FA_Snails_absolute <-
  read.table(file.path(prj, "data\\FA_Snails_absolute.txt"),
             header = TRUE, sep = "\t")

View(FA_Snails_absolute)

# remove NAs, measurement < LOQ
FA_Snails_absolute <- na.omit(FA_Snails_absolute)
```

In FA_Snails_absolute.txt: Snail 13, 14 and 29 (S_13, S_14, S_29) removed due to strongly deviating measured values compared to other snails.

Following fatty acids were not considered for FA content in snails due to too less values: C10:0, C11:0, C14:1, C15:1, C17:1, C20:3(n-3), C20:3(n-6), C21:0, C22:0 and C24:1.

Therefore, delete rows with respective fatty acids:

```{r}
FA_Snails_absolute <- FA_Snails_absolute %>%
  filter(!fa %in% c("C10:0", "C11:0", "C14:1", "C15:1", "C17:1", "C20:3(n-3)", "C20:3(n-6)", "C21:0", "C22:0", "C24:1"))
```

#### Histogram FA Snails

```{r}
hist(FA_Snails_absolute $ fa_conc) 
skewness(FA_Snails_absolute $ fa_conc, na.rm = TRUE) 
shapiro.test(FA_Snails_absolute $ fa_conc)
```

Fatty acid concentration not normal distributed and right skewed.

#### Model FA Snails

```{r}
# log10 transformation; adding 0.001 
FA_Snails_absolute <- FA_Snails_absolute %>%
  mutate(log_fa_conc = log10(fa_conc + 0.001))

# response variable: log_fa_conc
# fixed effects: treatment + fa + treatment:fa
# random effect with constant intercept: (1|replicate); first change to factor
FA_Snails_absolute $ replicate <- as.factor(FA_Snails_absolute $ replicate)

FA_snails_lmer <- lmer(log_fa_conc ~ treatment + fa + treatment:fa + (1|replicate), data = FA_Snails_absolute)
```

```{r}
plot(FA_snails_lmer) 
# @ VS: das sieht nicht wirklich gleichmäßig verteilt aus, oder? Was würdest du jetzt machen?

qreference(residuals(FA_snails_lmer)) 
# @ VS: das sieht schon wieder okay aus 

anova(FA_snails_lmer)
```

```{r}
FA_snails_inter <- lsmeans(FA_snails_lmer, ~ treatment | fa)
FA_snails_inter

FA_snails_inter_dunn1 <- contrast(FA_snails_inter, 'pairwise')

FA_snails_inter_dunn_sum1 <- summary(FA_snails_inter_dunn1, infer = TRUE, adjust = 'mvt')

FA_snails_inter_dunn_sum1

```

Except the following remarks, fatty acid concentration of the 4 treatments quite similar.

-   C12:0 : Ciprofloxacin treatment to Mix and NC (and almost to Pro) significantly different

-   C13:0 : Ciprofloxacin to all other treatments significantly different. Mix and Pro also different

-   C14:0 : Cip-Mix and Mix-NC different

-   C18:3(n-6) : Cip-NC and almost Mix-NC different

-   C22:1 : NC different to all chemical treatments. Mix and Pro also different.

#### Total SAFA, MUFA & PUFA content

1\) Total SAFA content

```{r}
# total SAFA, MUFA, PUFA content; sa = "snail absolute"
# total SAFA content
# filter saturated fatty acids in new table
SAFA_sa <- FA_Snails_absolute %>%
    filter(fa %in% c("C08:0", "C12:0", "C13:0", "C14:0", "C15:0", "C16:0", "C17:0", "C18:0", "C20:0", "C24:0"))

# calculate the total SAFA content per snail
total_SAFA_per_snail <- SAFA_sa %>%
  group_by(replicate, treatment) %>%
  summarize(sum_SAFA = sum(fa_conc))
print(total_SAFA_per_snail)

# mean
mean_SAFA_sa <- total_SAFA_per_snail %>%
  group_by(treatment) %>%
  summarize(mean_SAFA = mean(sum_SAFA, na.rm = TRUE))
print(mean_SAFA_sa)
```

2\) Total MUFA content

```{r}
# filter mono-unsaturated fatty acids in new table
MUFA_sa <- FA_Snails_absolute %>%
  filter(fa %in% c("C16:1", "C18:1", "C20:1", "C22:1"))

# total MUFA content per snail
total_MUFA_per_snail <- MUFA_sa %>%
  group_by(replicate, treatment) %>%
  summarize(sum_MUFA = sum(fa_conc))
print(total_MUFA_per_snail)

# mean
mean_MUFA_sa <- total_MUFA_per_snail %>%
  group_by(treatment) %>%
  summarize(mean_MUFA = mean(sum_MUFA, na.rm = TRUE))
print(mean_MUFA_sa)
```

3\) Total PUFA content

```{r}
# filter poly-unsaturated fatty acids in new table
PUFA_sa <- FA_Snails_absolute %>%
  filter(fa %in% c("C18:2(n-6c)", "C18:2(n-6t)", "C18:3(n-3)", "C18:3(n-6)", "C18:4(n-3)", "C20:2(n-6)", "C20:3(n-6)", "C20:4(n-6)", "C20:5(n-3)", "C22:2", "C22:5(n-3)", "C22:6(n-3)"))
  
# calculate total PUFA content per snail
total_PUFA_per_snail <- PUFA_sa %>%
  group_by(replicate, treatment) %>%
  summarize(sum_PUFA = sum(fa_conc))
print(total_PUFA_per_snail)

# mean
mean_PUFA_sa <- total_PUFA_per_snail %>%
  group_by(treatment) %>%
  summarize(mean_PUFA = mean(sum_PUFA, na.rm = TRUE))
print(mean_PUFA_sa)
```

4\) Total FA per snail

```{r}
# total FA 
total_FA_snails <- bind_rows(total_SAFA_per_snail, total_MUFA_per_snail, total_PUFA_per_snail)

total_FA_per_snail <- total_FA_snails %>%
  group_by(replicate, treatment) %>%
  summarize(sum_FA = sum(sum_SAFA, sum_MUFA, sum_PUFA, na.rm = TRUE))
print(total_FA_per_snail)

# mean
mean_total_FA_sa <- total_FA_per_snail %>%
  group_by(treatment) %>%
  summarize(mean_FA = mean(sum_FA, na.rm = TRUE))
print(mean_total_FA_sa)
```

#### NMDS FA snails

Change table to wide format:

```{r}
# change table to wide format
FA_Snails_absolute_filtered <- FA_Snails_absolute %>%
   dplyr::select(-log_fa_conc)


FA_Snails_absolute_wide <-
  pivot_wider(FA_Snails_absolute_filtered,
              names_from = fa,
              values_from = fa_conc)

# square root transformation, to reduce discriminatory power of dominant fa groups
FA_Snails_absolute_wide <- sqrt(FA_Snails_absolute_wide[, -c(1, 2)])
View(FA_Snails_absolute_wide)

#set NA to 0 
FA_Snails_absolute_wide[is.na(FA_Snails_absolute_wide)] <- 0
View(FA_Snails_absolute_wide)
```

Perform metaMDS:

```{r}
# Add treatment column and rename (- C, - () ) fatty acids for later plot
colnames(FA_Snails_absolute_wide) <-
  c(
    "8:0",
    "12:0",
    "13:0",
    "14:0",
    "15:0",
    "16:0",
    "16:1",
    "17:0",
    "18:0",
    "18:1",
    "18:2n-6t",
    "18:2n-6c",
    "18:3n-6",
    "18:3n-3",
    "18:4n-3",
    "20:0",
    "20:1",
    "20:2n-6",
    "20:4n-6",
    "20:5n-3",
    "22:1",
    "22:2",
    "22:5n-3",
    "22:6n-3",
    "24:0"
  )


NMDS_Snails_absolute <-
  metaMDS(
    FA_Snails_absolute_wide,
    autotransform = FALSE,
    k = 2 ,
    distance = "bray",
    na.rm = TRUE
  )

NMDS_Snails_absolute$stress # stress = 0.098
```

Create fa vectors

```{r}
# separation in SAFA; MUFA & PUFA to build vectors in NMDS
# "sa" = snail absolute
FA_variables.SAFAs_sa <-
  envfit(
    NMDS_Snails_absolute,
    subset(
      FA_Snails_absolute_wide,
      select = c(
        "8:0",
        "12:0",
        "13:0",
        "14:0",
        "15:0",
        "16:0",
        "17:0",
        "18:0",
        "20:0",
        "24:0"
      )
    )
    ,
    permutations = 999,
    na.rm = TRUE
  )

FA_variables.MUFAs_sa <-
  envfit(
    NMDS_Snails_absolute,
    subset(
      FA_Snails_absolute_wide,
      select = c(
        "16:1",
        "18:1",
        "20:1",
        "22:1"
      )
    )
    ,
    permutations = 999,
    na.rm = TRUE
  )

FA_variables.PUFAs_sa <-
  envfit(
    NMDS_Snails_absolute,
    subset(
      FA_Snails_absolute_wide,
      select = c(
        "18:2n-6t",
        "18:2n-6c",
        "18:3n-6",
        "18:3n-3",
        "18:4n-3",
        "20:2n-6",
        "20:4n-6",
        "20:5n-3",
        "22:5n-3",
        "22:6n-3"
      )
    )
    ,
    permutations = 999,
    na.rm = TRUE
  )

# Extract NMDS scores 
data.scores_snail_absolute = as.data.frame(scores(NMDS_Snails_absolute)$sites)

Treatments <-
  FA_Snails_absolute $ treatment[FA_Snails_absolute $ fa == "C16:0"]

# add treatment column to NMDS scores
data.scores_snail_absolute$Treatment = Treatments

# vectors for SAFA, MUFA and PUFA
SAFAs_coord_cont_sa = as.data.frame(scores(FA_variables.SAFAs_sa, "vectors")) * ordiArrowMul(FA_variables.SAFAs_sa)
MUFAs_coord_cont_sa = as.data.frame(scores(FA_variables.MUFAs_sa, "vectors")) * ordiArrowMul(FA_variables.MUFAs_sa)
PUFAs_coord_cont_sa = as.data.frame(scores(FA_variables.PUFAs_sa, "vectors")) * ordiArrowMul(FA_variables.PUFAs_sa)

```

Plot NMDS:

@VS: ich hab viel herumprobiert mit anderen Plots, base R plots ausprobiert und mit ggplot ordination plots gemacht. Ich finde diese Darstellung allerdings weiterhin am besten.

```{r}
# plot
absol_Snail <- ggplot(data = data.scores_snail_absolute, aes(x = NMDS1, y = NMDS2)) +
  geom_point(
    aes(shape = Treatment, color = Treatment),
    size = 5,
    alpha = 1,
    show.legend = TRUE
  ) +
  scale_shape_manual(values = c("NC" = 16, "Pro" = 17, "Cip" = 18, "Mix" = 15)) +
  scale_color_manual(values = c("NC" = "#009E73", "Pro" = "#E69F00", "Cip" = "#56B4E9", "Mix" = "#F0E442")) +
  stat_conf_ellipse(
    aes(colour = Treatment, fill = Treatment),
    alpha = 0.3,
    geom = "polygon",
    show.legend = FALSE
  ) +
  scale_color_manual(values = c("NC" = "#009E73", "Pro" = "#E69F00", "Cip" = "#56B4E9", "Mix" = "#F0E442")) +  
  scale_fill_manual(values = c("NC" = "#009E73", "Pro" = "#E69F00", "Cip" = "#56B4E9", "Mix" = "#F0E442")) + 
  geom_segment(aes(x = 0, y = 0, xend = NMDS1 * 0.2, yend = NMDS2 * 0.2), 
               data = PUFAs_coord_cont_sa, linewidth = 0.5, alpha = 0.2, colour = "black") +
    geom_segment(aes(x = 0, y = 0, xend = NMDS1 * 0.2, yend = NMDS2 * 0.2), 
               data = MUFAs_coord_cont_sa, linewidth = 0.5, alpha = 0.2, colour = "black") +
      geom_segment(aes(x = 0, y = 0, xend = NMDS1 * 0.1, yend = NMDS2 * 0.1), 
               data = SAFAs_coord_cont_sa, linewidth = 0.5, alpha = 0.2, colour = "black") +
  geom_text_repel(
    data = PUFAs_coord_cont_sa, 
    aes(x = NMDS1 * 0.2, y = NMDS2 * 0.2, label = row.names(PUFAs_coord_cont_sa)), 
    colour = "grey30",
    size = 3) +
    geom_text_repel(
    data = MUFAs_coord_cont_sa, 
    aes(x = NMDS1 * 0.2, y = NMDS2 * 0.2, label = row.names(MUFAs_coord_cont_sa)), 
    colour = "grey30",
    size = 3) +
      geom_text_repel(
    data = SAFAs_coord_cont_sa, 
    aes(x = NMDS1 * 0.1, y = NMDS2 * 0.1, label = row.names(SAFAs_coord_cont_sa)), 
    colour = "grey30",
    size = 3) +
  theme_bw() +
  theme(
    axis.title = element_text(
      size = 11,
      face = "bold",
      colour = "grey30"
    ),
    panel.background = element_blank(),
    panel.border = element_rect(fill = NA, colour = "grey30"),
    legend.key = element_blank(),
    legend.title = element_text(
      size = 11,
      face = "bold",
      colour = "grey30"
    ),
    legend.text = element_text(size = 10, colour = "grey30"),
    plot.margin = margin(
      t = 0.2,
      b = 0.2,
      l = 0.2,
      r = 0.2,
      unit = "cm"
    ),
    panel.grid.major = element_line(colour = "grey95", size = 0.4),
    panel.grid.minor = element_line(colour = "grey95", size = 0.2)
  ) +
  annotate(
    "text",
    x = 0.35,
    y = -0.19,
    label = paste("stress:", NMDS_Snails_absolute$stress %>% round(2)),
    size = 3.5
  ) 


absol_Snail
# position of FA names manually adjusted in vector graphic

# save the plot
ggsave(absol_Snail, 
       filename = "NMDS_snails_absolute.png", 
       path = 'C:\\Users\\Oster Test\\OneDrive\\Desktop\\Promotion\\Bioassays\\Schneckenversuch 2023\\Bottom_up_effects_biofilm_grazer\\output')
```

#### Permanova FA Snails

Create distance matrix with fatty acids

```{r}
FA_Snails_dist <- vegdist(FA_Snails_absolute_wide, method = 'bray')
```

Separate explanatory variables from fatty acid information

```{r}
expl_FA_Snails <-
  pivot_wider(FA_Snails_absolute_filtered,
              names_from = fa,
              values_from = fa_conc)

# delete fa columns
expl_FA_Snails <- expl_FA_Snails[, c(1, 2)]
expl_FA_Snails$treatment <- as.factor(expl_FA_Snails$treatment)
```

Make Permanova

```{r}
perm_FA_Snails <- adonis2(FA_Snails_dist ~ treatment, data = expl_FA_Snails, permutations = 999)
```

Make anova

```{r}
rda_FA_Snails_1 <- rda(decostand(FA_Snails_dist, "hellinger")
                                   ~ treatment, data = expl_FA_Snails)

summary(rda_FA_Snails_1, display = NULL)

anova_result <- anova(rda_FA_Snails_1, by = "margin")

```

Explanatory variable "treatment" significant, now perform pairwise comparisons

```{r}
# @ VS: hier komme ich nicht weiter, ich hab multiconstrained und normale post-hoc tests probiert.. wie würdest du hier weitermachen?

```

# 7 Fatty acid composition biofilm

```{r}
# insert data
FA_Biofilm_absolute <-
  read.table(file.path(prj, "data\\FA_Biofilm_absolute.txt"),
             header = TRUE, sep = "\t")

View(FA_Biofilm_absolute)

# remove NAs, measurement < LOQ
FA_Biofilm_absolute <- na.omit(FA_Biofilm_absolute)
```

Following fatty acids were not considered for biofilm analysis due to too less values: C11:0, C14:1, C15:1, C21:0, C22:1 and C24:1.

Therefore, delete rows with respective fatty acids:

```{r}
FA_Biofilm_absolute <- FA_Biofilm_absolute %>%
  filter(!fa %in% c("C11:0", "C14:1", "C15:1", "C21:0", "C22:1", "C24:1"))
```

#### Histogram FA Biofilm

```{r}
hist(FA_Biofilm_absolute $ fa_conc) 
skewness(FA_Biofilm_absolute $ fa_conc, na.rm = TRUE) 
shapiro.test(FA_Biofilm_absolute $ fa_conc)
```

Fatty acid concentration not normal distributed and right skewed.

#### Model FA Biofilm

```{r}
# log10 transformation; adding 0.001 
FA_Biofilm_absolute <- FA_Biofilm_absolute %>%
  mutate(log_fa_conc = log10(fa_conc + 0.001))

# response variable: log_fa_conc
# fixed effects: treatment + fa +  + week + treatment:fa + week:fa + treatment:week
# random effect with constant intercept: (1|replicate); first change to factor
FA_Biofilm_absolute $ replicate <- as.factor(FA_Biofilm_absolute $ replicate)
FA_Biofilm_absolute $ week <- as.factor(FA_Biofilm_absolute $ week)

FA_Biofilm_lmer <- lmer(log_fa_conc ~ treatment + fa + + week + week:fa + treatment:week + treatment:fa + (1|replicate), data = FA_Biofilm_absolute)
# @ VS: ist das richtig so? alle 3 explanatory variables und jeweils alle interaktionen zu testen?
```

```{r}
plot(FA_Biofilm_lmer) 
# @ VS: das sieht nicht wirklich gleichmäßig verteilt aus, oder? Was würdest du jetzt machen?

qreference(residuals(FA_Biofilm_lmer)) 
# @ VS: das sieht schon wieder okay aus 

anova(FA_Biofilm_lmer)

drop1(FA_Biofilm_lmer, test = 'Chisq')
```

```{r}
FA_Biofilm_inter <- lsmeans(FA_Biofilm_lmer, ~ treatment | fa)
FA_Biofilm_inter

FA_Biofilm_inter_dunn1 <- contrast(FA_Biofilm_inter, 'pairwise')

FA_Biofilm_inter_dunn_sum1 <- summary(FA_Biofilm_inter_dunn1, infer = TRUE, adjust = 'mvt')

FA_Biofilm_inter_dunn_sum1
# @ VS hier kommen bei mir "Inf" Werte raus, irgendwie klappt das so nicht, ich habe nicht herausgefunden, was ich ändern muss
```

#### Total SAFA, MUFA & PUFA content

1\) Total SAFA content

```{r}
# total SAFA, MUFA, PUFA content; ba = "biofilm absolute"
# total SAFA content
# filter saturated fatty acids in new table
SAFA_ba <- FA_Biofilm_absolute %>%
    filter(fa %in% c("C08:0", "C10:0", "C12:0", "C13:0", "C14:0", "C15:0", "C16:0", "C17:0", "C18:0", "C20:0", "C22:0", "C24:0"))

# calculate the total SAFA content per replicate
total_SAFA_per_replicate <- SAFA_ba %>%
  group_by(replicate, treatment) %>%
  summarize(sum_SAFA = sum(fa_conc))
print(total_SAFA_per_replicate)

# mean
mean_SAFA_ba <- total_SAFA_per_replicate %>%
  group_by(treatment) %>%
  summarize(mean_SAFA = mean(sum_SAFA, na.rm = TRUE))
print(mean_SAFA_ba)
```

2\) Total MUFA content

```{r}
# filter mono-unsaturated fatty acids in new table
MUFA_ba <- FA_Biofilm_absolute %>%
  filter(fa %in% c("C16:1", "C17:1", "C18:1", "C20:1"))

# total MUFA content per replicate
total_MUFA_per_replicate <- MUFA_ba %>%
  group_by(replicate, treatment) %>%
  summarize(sum_MUFA = sum(fa_conc))
print(total_MUFA_per_replicate)

# mean
mean_MUFA_ba <- total_MUFA_per_replicate %>%
  group_by(treatment) %>%
  summarize(mean_MUFA = mean(sum_MUFA, na.rm = TRUE))
print(mean_MUFA_ba)
```

3\) Total PUFA content

```{r}
# filter poly-unsaturated fatty acids in new table
PUFA_ba <- FA_Biofilm_absolute %>%
  filter(fa %in% c("C18:2(n-6c)", "C18:2(n-6t)", "C18:3(n-3)", "C18:3(n-6)", "C18:4(n-3)", "C20:2(n-6)", "C20:3(n-6)", "C20:3(n-3)", "C20:4(n-6)", "C20:5(n-3)", "C22:2", "C22:5(n-3)", "C22:6(n-3)"))
  
# calculate total PUFA content per replicate
total_PUFA_per_replicate <- PUFA_ba %>%
  group_by(replicate, treatment) %>%
  summarize(sum_PUFA = sum(fa_conc))
print(total_PUFA_per_replicate)

# mean
mean_PUFA_ba <- total_PUFA_per_replicate %>%
  group_by(treatment) %>%
  summarize(mean_PUFA = mean(sum_PUFA, na.rm = TRUE))
print(mean_PUFA_ba)
```

4\) Total FA content

```{r}
# total FA 
total_FA_biofilm <- bind_rows(total_SAFA_per_replicate, total_MUFA_per_replicate, total_PUFA_per_replicate)

total_FA_per_replicate <- total_FA_biofilm %>%
  group_by(replicate, treatment) %>%
  summarize(sum_FA = sum(sum_SAFA, sum_MUFA, sum_PUFA, na.rm = TRUE))
print(total_FA_per_replicate)

# mean
mean_total_FA_ba <- total_FA_per_replicate %>%
  group_by(treatment) %>%
  summarize(mean_FA = mean(sum_FA, na.rm = TRUE))
print(mean_total_FA_ba)
```

#### NMDS FA Biofilm

Change table to wide format:

```{r}
# add prefix (week) 1,2 or 3 to treatments for later subgrouping
FA_Biofilm_absolute <- FA_Biofilm_absolute %>%
  mutate(treatment = if_else(week == "1", paste("1_", treatment, sep = ""), treatment))

FA_Biofilm_absolute <- FA_Biofilm_absolute %>%
  mutate(treatment = if_else(week == "2", paste("2_", treatment, sep = ""), treatment))

FA_Biofilm_absolute <- FA_Biofilm_absolute %>%
  mutate(treatment = if_else(week == "3", paste("3_", treatment, sep = ""), treatment))

# change table to wide format
FA_Biofilm_absolute_filtered <- FA_Biofilm_absolute %>%
   dplyr::select(-log_fa_conc)


FA_Biofilm_absolute_wide <-
  pivot_wider(FA_Biofilm_absolute_filtered,
              names_from = fa,
              values_from = fa_conc)

# square root transformation, to reduce discriminatory power of dominant fa groups
FA_Biofilm_absolute_wide <- sqrt(FA_Biofilm_absolute_wide[, -c(1, 2, 3)])
View(FA_Biofilm_absolute_wide)

#set NA to 0 
FA_Biofilm_absolute_wide[is.na(FA_Biofilm_absolute_wide)] <- 0
View(FA_Biofilm_absolute_wide)
```

Perform metaMDS:

```{r}
#add treatment column and ease colnames for later NMDS plot
colnames(FA_Biofilm_absolute_wide) <-
  c(
    "8:0",
    "10:0",
    "12:0",
    "13:0",
    "14:0",
    "15:0",
    "16:0",
    "16:1",
    "17:0",
    "17:1",
    "18:0",
    "18:1",
    "18:2n-6t",
    "18:2n-6c",
    "18:3n-6",
    "18:3n-3",
    "18:4n-3",
    "20:0",
    "20:1",
    "20:2n-6",
    "20:3n-6",
    "20:4n-6",
    "20:5n-3",
    "22:0",
    "22:5n-3",
    "22:6n-3",
    "24:0"
  )

NMDS_Biofilm_absolute <-
  metaMDS(
    FA_Biofilm_absolute_wide,
    autotransform = FALSE,
    k = 2 ,
    distance = "bray",
    na.rm = TRUE
  )

NMDS_Biofilm_absolute$stress
```

Create fa vectors

```{r}
# separation in SAFA; MUFA & PUFA to build vectors in NMDS
# "ba" = biofilm absolute
FA_variables.SAFAs_ba <-
  envfit(
    NMDS_Biofilm_absolute,
    subset(
      FA_Biofilm_absolute_wide,
      select = c(
        "8:0",
        "10:0",
        "12:0",
        "13:0",
        "14:0",
        "15:0",
        "16:0",
        "17:0",
        "18:0",
        "20:0",
        "22:0",
        "24:0"
      )
    )
    ,
    permutations = 999,
    na.rm = TRUE
  )

FA_variables.MUFAs_ba <-
  envfit(
    NMDS_Biofilm_absolute,
    subset(
      FA_Biofilm_absolute_wide,
      select = c(
        "16:1",
        "17:1",
        "18:1",
        "20:1"
      )
    )
    ,
    permutations = 999,
    na.rm = TRUE
  )

FA_variables.PUFAs_ba <-
  envfit(
    NMDS_Biofilm_absolute,
    subset(
      FA_Biofilm_absolute_wide,
      select = c(
        "18:2n-6t",
        "18:2n-6c",
        "18:3n-6",
        "18:3n-3",
        "18:4n-3",
        "20:2n-6",
        "20:3n-6",
        "20:4n-6",
        "20:5n-3",
        "22:5n-3",
        "22:6n-3"
      )
    )
    ,
    permutations = 999,
    na.rm = TRUE
  )

# Extract NMDS scores
data.scores_biofilm_absolute = as.data.frame(scores(NMDS_Biofilm_absolute)$sites)

Treatments <-
  FA_Biofilm_absolute$treatment[FA_Biofilm_absolute$fa == "C16:0"]

# add treatment column to NMDS scores
data.scores_biofilm_absolute$Treatment = Treatments

# vectors for SAFA, MUFA and PUFA
SAFAs_coord_cont_ba = as.data.frame(scores(FA_variables.SAFAs_ba, "vectors")) * ordiArrowMul(FA_variables.SAFAs_ba)
MUFAs_coord_cont_ba = as.data.frame(scores(FA_variables.MUFAs_ba, "vectors")) * ordiArrowMul(FA_variables.MUFAs_ba)
PUFAs_coord_cont_ba = as.data.frame(scores(FA_variables.PUFAs_ba, "vectors")) * ordiArrowMul(FA_variables.PUFAs_ba)

# create weeks as subgroups for plotting
data.scores_biofilm_absolute <- data.scores_biofilm_absolute %>%
  mutate(time.point = case_when(
    str_detect(Treatment, "^1_") ~ "Week1",
    str_detect(Treatment, "^2_") ~ "Week2",
    str_detect(Treatment, "^3_") ~ "Week3",
    TRUE ~ "Other"
  ))

```

```{r}
# first delete 1_, 2_ & 3_ prefix in treatments
data.scores_biofilm_absolute $ Treatment <- sub("1_", "", data.scores_biofilm_absolute $ Treatment)
data.scores_biofilm_absolute $ Treatment <- sub("2_", "", data.scores_biofilm_absolute $ Treatment)
data.scores_biofilm_absolute $ Treatment <- sub("3_", "", data.scores_biofilm_absolute $ Treatment)
```

Plot NMDS:

```{r}
absol_Biofilm = ggplot(data = data.scores_biofilm_absolute, aes(x = NMDS1, y = NMDS2)) +
  geom_point( aes(shape = time.point, color = Treatment),
    size = 5,
    alpha = 1
  ) +
  scale_shape_manual(values = c(15, 16, 17)) +
  scale_color_manual(values = c("NC" = "#009E73", "Pro" = "#E69F00", "Cip" = "#56B4E9", "Mix" = "#F0E442")) +
  labs(shape = "Week", color = "Treatment") +
  geom_segment(data = PUFAs_coord_cont_ba, aes(x = 0, y = 0, xend = NMDS1 * 0.3, yend = NMDS2 * 0.3), linewidth = 0.5, alpha = 0.2, colour = "black") +
  geom_segment(data = MUFAs_coord_cont_ba, aes(x = 0, y = 0, xend = NMDS1 * 0.2, yend = NMDS2 * 0.2), linewidth = 0.5, alpha = 0.2, colour = "black") +
  geom_segment(data = SAFAs_coord_cont_ba, aes(x = 0, y = 0, xend = NMDS1 * 0.3, yend = NMDS2 * 0.3), linewidth = 0.5, alpha = 0.2, colour = "black") +
  geom_text_repel(
    data = PUFAs_coord_cont_ba, 
    aes(x = NMDS1 * 0.3, y = NMDS2 * 0.3, label = row.names(PUFAs_coord_cont_ba)), 
    colour = "grey30",
    size = 3) +
  geom_text_repel(
    data = MUFAs_coord_cont_ba, 
    aes(x = NMDS1 * 0.2, y = NMDS2 * 0.2, label = row.names(MUFAs_coord_cont_ba)), 
    colour = "grey30",
    size = 3) +
  geom_text_repel(
    data = SAFAs_coord_cont_ba, 
    aes(x = NMDS1 * 0.3, y = NMDS2 * 0.3, label = row.names(SAFAs_coord_cont_ba)), 
    colour = "grey30",
    size = 3) +
  theme_bw() +
  theme(
    axis.title = element_text(
      size = 11,
      face = "bold",
      colour = "grey30"
    ),
    panel.background = element_blank(),
    panel.border = element_rect(fill = NA, colour = "grey30"),
    #axis.ticks = element_blank(),
    #axis.text = element_blank(),
    legend.key = element_blank(),
    legend.title = element_text(
      size = 11,
      face = "bold",
      colour = "grey30"
    ),
    legend.text = element_text(size = 10, colour = "grey30"),
    plot.margin = margin(
      t = 0.2,
      b = 0.2,
      l = 0.2,
      r = 0.2,
      unit = "cm"
    ),
    panel.grid.major = element_line(colour = "grey95", size = 0.4),
    panel.grid.minor = element_line(colour = "grey95", size = 0.2)
  ) +
  annotate(
    "text",
    x = 0.32,
    y = -0.3,
    label = paste("stress:", NMDS_Biofilm_absolute$stress %>% round(2)),
    size = 3
  )
  
  
absol_Biofilm





# save the plot
ggsave(absol_Biofilm, 
       filename = "NMDS_Biofilm_absolute.png", 
       path = 'C:\\Users\\Oster Test\\OneDrive\\Desktop\\Promotion\\Bioassays\\Schneckenversuch 2023\\Bottom_up_effects_biofilm_grazer\\output')
```

#### Permanova FA Biofilm

```{r}
# @ VS: Hier muss die Woche als explanatory variable wieder dazu, oder? Dann muss ich die Tabellen wahrscheinlich anders formen, als für die FA der Schnecken?
```

# 8 Temperature and light conditions

@ VS: Dieses Skript ist nur zum plotten der Temperatur und Lichtverhältnisse dar. Mirco möchte diese Plots gerne in den Anhang.

## 8.1 Biofilm colonisation and contamination

```{r}
# insert data
colonisation_temp_light <-
  read.table(file.path(prj, "data\\colonisation_temp_light.txt"),
             header = TRUE, sep = "\t")

View(colonisation_temp_light)

# average temperature
mean(colonisation_temp_light $ temperature_colonisation)
# 18.28751 °C

# mean light intensity
# filter only daytime
average_light_intensity_colonisation <- colonisation_temp_light %>%
  dplyr::filter(light_intensity_colonisation > 0) %>%
  summarize(average_light_intensity_colonisation = mean(light_intensity_colonisation))

print(average_light_intensity_colonisation)
# 14019.26 lux

# temperature plot
# add a small amount of random noise to the y-values
colonisation_temp_light $ temperature_colonisation <- jitter(colonisation_temp_light $ temperature_colonisation)

t2 <- ggplot(
  data = colonisation_temp_light,
  mapping = aes( x = day, y = temperature_colonisation, color = time_of_day)) +
  geom_point(alpha = 0.5) +  # Individual measurements with transparency
  geom_smooth() +
  labs(
    title = "Temperature profile during biofilm colonisation",
    x = "Day", y = "Temperature (°C)"
  ) +
  theme_bw() +
  scale_x_continuous(breaks = seq(0, 77, by = 5)) +
  scale_color_continuous(
    name = "Time of Day",  
    breaks = c(0, 5, 10, 15, 20),
    labels = c("0:00 am", "05:00 am", "10:00 am", "03:00 pm", "08:00 pm"),
    limits = c(0, 22),
  )

t2 

# save the plot
ggsave(t2, 
       filename = "temperature_colonisation.png", 
       path = 'C:\\Users\\Oster Test\\OneDrive\\Desktop\\Promotion\\Bioassays\\Schneckenversuch 2023\\Bottom_up_effects_biofilm_grazer\\output')

# light intensity plot
# Filter the data to include only rows where light_intensity > 0
filtered_data2 <- colonisation_temp_light %>% 
  filter(light_intensity_colonisation > 0)

# add a small amount of random noise to the y-values
filtered_data2 $ light_intensity_colonisation <- jitter(filtered_data2 $ light_intensity_colonisation)

l2 <- ggplot(
  data = filtered_data2,
  mapping = aes( x = day, y = light_intensity_colonisation)) +
  geom_point(size = 1) +  # Individual measurements with transparency
  geom_smooth(color = "#E69F00", linewidth = 1) +
  labs(
    title = "Light intensity profile during biofilm colonisation",
    x = "Day", y = "Light intensity (lux)"
  ) +
  theme_bw() +
  scale_x_continuous(breaks = seq(0, 77, by = 5)) 


l2 

# save the plot
ggsave(l2, 
       filename = "light_intensity_colonisation.png", 
       path = 'C:\\Users\\Oster Test\\OneDrive\\Desktop\\Promotion\\Bioassays\\Schneckenversuch 2023\\Bottom_up_effects_biofilm_grazer\\output')


```

## 8.2 Feeding assay

```{r}
# insert data
feeding_assay_temp_light <-
  read.table(file.path(prj, "data\\feeding_assay_temp_light.txt"),
             header = TRUE, 
             sep = "\t"
  )
             
View(feeding_assay_temp_light)

# average temperature
mean(feeding_assay_temp_light $ temperature)
# 13.27927 °C

# mean light intensity
# filter only daytime
average_light_intensity_feeding <- feeding_assay_temp_light %>%
  dplyr::filter(light_intensity > 0) %>%
  summarize(average_light_intensity_feeding = mean(light_intensity))

print(average_light_intensity_feeding)
# 330.3552 lux

# temperature plot
# add a small amount of random noise to the y-values
feeding_assay_temp_light $ temperature <- jitter(feeding_assay_temp_light $ temperature)

t1 <- ggplot(
  data = feeding_assay_temp_light,
  mapping = aes( x = day, y = temperature, color = time_of_day)) +
  geom_point(alpha = 0.5) +  # Individual measurements with transparency
  geom_smooth() +
  labs(
    title = "Temperature profile during 21-day feeding assay",
    x = "Day", y = "Temperature (°C)"
  ) +
  theme_bw() +
  scale_x_continuous(breaks = 0:21) +
  scale_color_continuous(
    name = "Time of Day",  
    breaks = c(0, 5, 10, 15, 20),
    labels = c("0:00 am", "05:00 am", "10:00 am", "03:00 pm", "08:00 pm"),
    limits = c(0, 22),
  )

t1 

# save the plot
ggsave(t1, 
       filename = "temperature_feeding_assay.png", 
       path = 'C:\\Users\\Oster Test\\OneDrive\\Desktop\\Promotion\\Bioassays\\Schneckenversuch 2023\\Bottom_up_effects_biofilm_grazer\\output')


# light intensity plot
# Filter the data to include only rows where light_intensity > 0
filtered_data1 <- feeding_assay_temp_light %>% 
  filter(light_intensity > 0)

# add a small amount of random noise to the y-values
filtered_data1 $ light_intensity <- jitter(filtered_data1 $ light_intensity)

l1 <- ggplot(
  data = filtered_data1,
  mapping = aes( x = day, y = light_intensity)) +
  geom_point(size = 1) +  # Individual measurements with transparency
  geom_smooth(color = "#E69F00", linewidth = 1) +
  labs(
    title = "Light intensity profile during 21-day feeding assay",
    x = "Day", y = "Light intensity (lux)"
  ) +
  theme_bw() +
  scale_x_continuous(breaks = 0:21) 


l1 

# save the plot
ggsave(l1, 
       filename = "light_intensity_feeding_assay.png", 
       path = 'C:\\Users\\Oster Test\\OneDrive\\Desktop\\Promotion\\Bioassays\\Schneckenversuch 2023\\Bottom_up_effects_biofilm_grazer\\output')


```
