---
title: "Xenobiotics modify vertical energy transfer from photoautotrophic biofilms to grazers"
format: 
  html:
    self-contained: true
    toc: true
    toc-depth: 2
    number-sections: true
    style: "floating"
editor_options: 
  chunk_output_type: console
---

Sophie Oster, Patrick Fink, Tobias Schmitt, Ralf Schulz, Verena C. Schreiner & Mirco Bundschuh

R script written by Sophie Oster\
checked by Verena C. Schreiner

\
RPTU Kaiserslautern-Landau\
Fortstrasse 7\
76829 Landau in der Pfalz\
GERMANY

This study was conducted with four chemical treatments, as explanatory variables:

-   treatment (categorical)
    -   uncontaminated control (negative control, NC)
    -   herbicide propyzamide 10 ug / L (Pro)
    -   antibiotic ciprofloxacin 100 ug / L (Cip)
    -   mixture of 10 ug / L  propyzamide and 100 ug / L ciprofloxacin (Mix)


# R version 4.3.2

R version used for this script: 4.3.2, use the following code chunk to download the correct version.

```{r}
# Check the current R version
current_r_version <- as.character(getRversion())

# Specify the required R version
required_r_version <- "4.3.2"

if (current_r_version != required_r_version) {
  
  sysname <- Sys.info()[["sysname"]]
  machine <- tolower(Sys.info()[["machine"]])
  
  # Choose the correct official CRAN download
  download_url <- switch(
    sysname,
    "Windows" = sprintf(
      "https://cran.r-project.org/bin/windows/base/old/%s/R-%s-win.exe",
      required_r_version, required_r_version
    ),
    "Darwin" = if (grepl("arm64|aarch64", machine)) {
      sprintf(
        "https://cran.r-project.org/bin/macosx/big-sur-arm64/base/R-%s-arm64.pkg",
        required_r_version
      )
    } else {
      sprintf(
        "https://cran.r-project.org/bin/macosx/big-sur-x86_64/base/R-%s-x86_64.pkg",
        required_r_version
      )
    },
    "Linux" = sprintf(
      "https://cran.r-project.org/src/base/R-4/R-%s.tar.gz",
      required_r_version
    ),
    stop("Unsupported operating system: ", sysname)
  )
  
  destfile <- file.path(tempdir(), basename(download_url))
  
  message("This project requires R ", required_r_version,
          ". Downloading it from CRAN now...")
  
  download.file(download_url, destfile = destfile, mode = "wb")
  
  message("Downloaded to: ", destfile)
  
  if (sysname == "Windows") {
    shell.exec(normalizePath(destfile))
  } else if (sysname == "Darwin") {
    system2("open", shQuote(destfile), wait = FALSE)
  } else if (sysname == "Linux") {
    message(
      "Linux detected: the R source tarball was downloaded.\n",
      "Please install R ", required_r_version,
      " using your distribution's preferred method, or build from source."
    )
    utils::browseURL("https://cran.r-project.org/doc/manuals/r-patched/R-admin.html")
  }
  
  stop(
    "Please install R ", required_r_version,
    ", restart R with that version, and rerun the script."
  )
  
} else {
  message("You are already using the correct R version: ", required_r_version)
}
```

# Packages in a reproducible environment

```{r, message = FALSE, warning = FALSE}
# Restore the environment:
# renv::restore()

library(rstatix)
library(dplyr)
library(tidyverse)
library(tidyr)
library(broom)
library(dunn.test)
library(patchwork)
library(vegan)
library(data.table)
library(ggplot2)
library(ggrepel)
library(ggforce)
library(concaveman)
library(gridExtra)
library(stringr)
library(shades)
library(ggpubr)
library(ggokabeito)
library(multcomp)
library(car)
library(asbio)
library(plotrix)
library(lmtest)
library(boot)
library(moments)
library(lme4)
library(DAAG)
library(performance)
library(emmeans)
library(purrr)
library(DT)
library(Hmisc)
library(svglite)
library(mgcv)       
library(DHARMa)   
library(marginaleffects)
library(writexl)
library(ARTool)

if (!requireNamespace("devtools", quietly = TRUE)) {
  install.packages("devtools")
}

library(devtools)

if (!requireNamespace("pairwiseAdonis", quietly = TRUE)) {
  devtools::install_github("pmartinezarbizu/pairwiseAdonis/pairwiseAdonis")
}

library(pairwiseAdonis)
```

To examine the interaction between both xenobiotics, we further analysed their combined effects using linear models based on presence or absence of each xenobiotic. When a significant interaction was detected, we applied the post-estimation inference framework proposed by Madge Pimentel and colleagues (2025). This approach assesses interaction strength relative to additive, multiplicative, and dominance null models, and tests deviations from zero using simulation-based confidence intervals.  

Load these functions:
```{r}
# Madge Pimentel, I., Albini, D., Beermann, A. J., Leese, F., Macaulay, S. J., Matthaei, C. D., ... & Schäfer, R. B. (2025). Hypothesis‐Driven Research on Multiple Stressors: An Analytical Framework for Stressor Interactions. Ecology and evolution, 15(8), e71959.

# hypothesis function
hypothesis.function <- function(df.cf, stressor1, stressor2,
                                strategy = c("average", "rowwise"),
                                by = NULL,
                                null = c("ad", "mu", "do", "all")) {
  strategy <- match.arg(strategy)
  null <- match.arg(null)

  df <- df.cf |>
    dplyr::arrange(dataset, rowidcf)

  C.df <- dplyr::filter(df, dataset == "C")
  S1.df <- dplyr::filter(df, dataset == "S1")
  S2.df <- dplyr::filter(df, dataset == "S2")
  S12.df <- dplyr::filter(df, dataset == "S12")

  change_x1 <- as.numeric(S12.df[[stressor1]]) - as.numeric(C.df[[stressor1]])
  change_x2 <- as.numeric(S12.df[[stressor2]]) - as.numeric(C.df[[stressor2]])

  C <- C.df$estimate
  S1 <- S1.df$estimate
  S2 <- S2.df$estimate
  S12 <- S12.df$estimate

  if (null == "ad") {
    row.est <- (S12 - S1 - S2 + C) / (change_x1 * change_x2)
  } else if (null == "mu") {
    row.est <- (log(S12) - log(S1) - log(S2) + log(C)) / (change_x1 * change_x2)
  } else if (null == "do") {
    row.est <- (pmin(S12, C) - pmin(S1, S2)) / (change_x1 * change_x2)
  }

  if (null != "all") {
    rowwise <- C.df |>
      dplyr::select(-estimate, -dataset) |>
      dplyr::mutate(
        estimate = row.est,
        term = null,
        change_x1 = change_x1,
        change_x2 = change_x2,
        .before = 1
      )
  }

  if (null == "all") {
    row.est.ad <- (S12 - S1 - S2 + C) / (change_x1 * change_x2)
    row.est.mu <- (log(S12) - log(S1) - log(S2) + log(C)) / (change_x1 * change_x2)
    row.est.do <- (pmin(S12, C) - pmin(S1, S2)) / (change_x1 * change_x2)

    rowwise <- C.df |>
      dplyr::select(-estimate, -dataset) |>
      dplyr::mutate(
        ad = row.est.ad,
        mu = row.est.mu,
        do = row.est.do,
        change_x1 = change_x1,
        change_x2 = change_x2
      ) |>
      tidyr::pivot_longer(cols = c("ad", "mu", "do"),
                          names_to = "term", values_to = "estimate")
  }

  if (strategy == "average") {
    grouping <- "term"
    if (!is.null(by)) grouping <- c(grouping, by)
    out <- rowwise |>
      dplyr::group_by(dplyr::across(all_of(grouping))) |>
      dplyr::summarise(estimate = mean(estimate), .groups = "drop")
  } else {
    out <- rowwise
  }

  return(out)
} 

# nd generator
nd.generator <- function(df.ref, stressor1, stressor2, diff = NULL){
    df.prep <- df.ref |>
        dplyr::mutate(dataset = "C", rowidcf = dplyr::row_number())

    if (is.null(diff)) {
        df.out <- df.prep |>
            tibble::add_case(
                df.prep |> dplyr::mutate(
                    !!stressor1 := factor(levels(.data[[stressor1]])[
                        as.numeric(.data[[stressor1]]) + 1],
                        levels = levels(.data[[stressor1]])
                    ),
                    dataset = "S1"
                )
            ) |>
            tibble::add_case(
                df.prep |> dplyr::mutate(
                    !!stressor2 := factor(levels(.data[[stressor2]])[
                        as.numeric(.data[[stressor2]]) + 1],
                        levels = levels(.data[[stressor2]])
                    ),
                    dataset = "S2"
                )
            ) |>
            tibble::add_case(
                df.prep |> dplyr::mutate(
                    !!stressor1 := factor(levels(.data[[stressor1]])[
                        as.numeric(.data[[stressor1]]) + 1],
                        levels = levels(.data[[stressor1]])
                    ),
                    !!stressor2 := factor(levels(.data[[stressor2]])[
                        as.numeric(.data[[stressor2]]) + 1],
                        levels = levels(.data[[stressor2]])
                    ),
                    dataset = "S12"
                )
            )
    } else if (length(diff) == 1) {
        if (is.factor(df.prep[[stressor1]])) {
            df.out <- df.prep |>
                tibble::add_case(df.prep |> dplyr::mutate(
                    !!stressor1 := factor(levels(.data[[stressor1]])[
                        as.numeric(.data[[stressor1]]) + 1],
                        levels = levels(.data[[stressor1]])
                    ),
                    dataset = "S1"
                )) |>
                tibble::add_case(df.prep |> dplyr::mutate(
                    !!stressor2 := .data[[stressor2]] + diff,
                    dataset = "S2"
                )) |>
                tibble::add_case(df.prep |> dplyr::mutate(
                    !!stressor1 := factor(levels(.data[[stressor1]])[
                        as.numeric(.data[[stressor1]]) + 1],
                        levels = levels(.data[[stressor1]])
                    ),
                    !!stressor2 := .data[[stressor2]] + diff,
                    dataset = "S12"
                ))
        } else {
            df.out <- df.prep |>
                tibble::add_case(df.prep |> dplyr::mutate(
                    !!stressor1 := .data[[stressor1]] + diff,
                    dataset = "S1"
                )) |>
                tibble::add_case(df.prep |> dplyr::mutate(
                    !!stressor2 := factor(levels(.data[[stressor2]])[
                        as.numeric(.data[[stressor2]]) + 1],
                        levels = levels(.data[[stressor2]])
                    ),
                    dataset = "S2"
                )) |>
                tibble::add_case(df.prep |> dplyr::mutate(
                    !!stressor1 := .data[[stressor1]] + diff,
                    !!stressor2 := factor(levels(.data[[stressor2]])[
                        as.numeric(.data[[stressor2]]) + 1],
                        levels = levels(.data[[stressor2]])
                    ),
                    dataset = "S12"
                ))
        }
    } else if (length(diff) == 2) {
        df.out <- df.prep |>
            tibble::add_case(dplyr::mutate(df.prep,
                !!stressor1 := .data[[stressor1]] + diff[1],
                dataset = "S1")) |>
            tibble::add_case(dplyr::mutate(df.prep,
                !!stressor2 := .data[[stressor2]] + diff[2],
                dataset = "S2")) |>
            tibble::add_case(dplyr::mutate(df.prep,
                !!stressor1 := .data[[stressor1]] + diff[1],
                !!stressor2 := .data[[stressor2]] + diff[2],
                dataset = "S12"))
    }

    if ("rowid" %in% names(df.out)) {
        df.out <- dplyr::select(df.out, -rowid)
    }

    return(df.out)
}
```


# Biofilm endpoints

Chlorophyll content and ash free dry weight (AFDW) of biofilms

-   Unit chlorophyll a, b & c: µg chlorophyll / cm²

-   Unit afdw: mg / cm²

```{r}
# insert data
biofilm_endpoints <-
  read.table(file.path("data\\biofilm_endpoints.txt"),
             header = TRUE, sep = "\t")

View(biofilm_endpoints)
```

##  Chlorophyll a, b and c content

Caclulate mean, se and 95% CI chla, chlb + chlc of each treatment:

```{r}
# chla
mean_chla <- biofilm_endpoints %>%
  group_by(treatment) %>%
  dplyr::summarize(
    mean_chla = mean(chla, na.rm = TRUE),
    n = sum(!is.na(chla)),
    se = sd(chla, na.rm = TRUE) / sqrt(n),
    t_crit = qt(0.975, df = n - 1),
    ci_lower = mean_chla - t_crit * se,
    ci_upper = mean_chla + t_crit * se,
    .groups = "drop"
  ) %>%
  dplyr::select(treatment, mean_chla, ci_lower, ci_upper, se)

print(mean_chla)

#chlb
mean_chlb <- biofilm_endpoints %>%
  group_by(treatment) %>%
  dplyr::summarize(
    mean_chlb = mean(chlb, na.rm = TRUE),
    n = sum(!is.na(chlb)),
    se = sd(chlb, na.rm = TRUE) / sqrt(n),
    t_crit = qt(0.975, df = n - 1),
    ci_lower = mean_chlb - t_crit * se,
    ci_upper = mean_chlb + t_crit * se,
    .groups = "drop"
  ) %>%
  dplyr::select(treatment, mean_chlb, ci_lower, ci_upper, se)

print(mean_chlb)

#chlc
mean_chlc <- biofilm_endpoints %>%
  group_by(treatment) %>%
  dplyr::summarize(
    mean_chlc = mean(chlc, na.rm = TRUE),
    n = sum(!is.na(chlc)),
    se = sd(chlc, na.rm = TRUE) / sqrt(n),
    t_crit = qt(0.975, df = n - 1),
    ci_lower = mean_chlc - t_crit * se,
    ci_upper = mean_chlc + t_crit * se,
    .groups = "drop"
  ) %>%
  dplyr::select(treatment, mean_chlc, ci_lower, ci_upper, se)

print(mean_chlc)
```


### Histogram chlorophyll a, b + c

```{r}
# chl a
hist(biofilm_endpoints $ chla) 
skewness(biofilm_endpoints $ chla, na.rm = TRUE)
shapiro.test(biofilm_endpoints $ chla)

# chl b
hist(biofilm_endpoints $ chlb) 
skewness(biofilm_endpoints $ chlb, na.rm = TRUE) 
shapiro.test(biofilm_endpoints $ chlb)

# chl c
hist(biofilm_endpoints $ chlc) 
skewness(biofilm_endpoints $ chlc, na.rm = TRUE) 
shapiro.test(biofilm_endpoints $ chlc)
```

All not normal distributed, right skewed but skewness max. 1.6 therefore no data transformation for model

### Model chla

Linear model with xenobiotic treatment as explanatory variable.
```{r}
# treatment as factor
biofilm_endpoints $ treatment <- factor(biofilm_endpoints $ treatment)

model_chla <- lm(chla ~ treatment, data = biofilm_endpoints)
par(mfrow = c(2, 2))
plot(model_chla) 
qreference(residuals(model_chla)) # not normal distributed

pairwise.wilcox.test(biofilm_endpoints$chla,
                     biofilm_endpoints$treatment,
                     p.adjust.method = "holm")

# add presence / absence
# interaction effect model with presence / absence of single xenobiotics
biofilm_endpoints$Cip <- ifelse(biofilm_endpoints$treatment %in% c("Cip", "Mix"),
                                1, 0)

biofilm_endpoints$Cip <- factor(biofilm_endpoints$Cip, levels = c(0, 1))

biofilm_endpoints$Pro <- ifelse(biofilm_endpoints$treatment %in% c("Pro", "Mix"),
                                1, 0)

biofilm_endpoints$Pro <- factor(biofilm_endpoints$Pro, levels = c(0, 1))

# sanity check
table(biofilm_endpoints$treatment,
      biofilm_endpoints$Cip,
      biofilm_endpoints$Pro)

chla_mixture <- lm(chla ~ Cip * Pro, data = biofilm_endpoints)
par(mfrow = c(2, 2))
plot(chla_mixture) # not normally distributed
qreference(residuals(chla_mixture)) 

kruskal_chla_interaction <- biofilm_endpoints %>%
  dplyr::mutate(treatment = interaction(Cip, Pro, sep = "_")) %>%
  kruskal.test(chla ~ treatment, data = .)

kruskal_chla_interaction
# no significant interaction
```
No significant differences between chemical treatments in chlorophyll a content.

### Model chlb

```{r}
model_chlb <- lm(chlb ~ treatment, data = biofilm_endpoints)
par(mfrow = c(2, 2))
plot(model_chlb)
qreference(residuals(model_chlb)) # normality acceptable
drop1(model_chlb, test = 'F')

chlb_inter <- lsmeans(model_chlb, ~ treatment)
chlb_inter_dunn1 <- contrast(chlb_inter, 'pairwise')
(chlb_inter_dunn_sum1 <- summary(chlb_inter_dunn1, infer = TRUE, adjust = 'mvt'))
```

No significant differences between chemical treatments in chlorophyll b content.
Still assess potential interaction:
```{r}
chlb_mixture <- lm(chlb ~ Cip * Pro, data = biofilm_endpoints)
par(mfrow = c(2, 2))
plot(chlb_mixture) # normality acceptable
qreference(residuals(chlb_mixture)) 
drop1(chlb_mixture, test = 'F')
# no significant interaction
```


### Model chlc

```{r}
model_chlc <- lm(chlc ~ treatment, data = biofilm_endpoints)
par(mfrow = c(2, 2))
plot(model_chlc) # normality acceptable
qreference(residuals(model_chlc))
drop1(model_chlc, test = 'F')
```
No significant differences between chemical treatments in chlorophyll c content.
Still check for interactions of Cip * Pro:
```{r}
chlc_mixture <- lm(chlc ~ Cip * Pro, data = biofilm_endpoints)
par(mfrow = c(2, 2))
plot(chlc_mixture) # normality acceptable
qreference(residuals(chlc_mixture)) 
drop1(chlc_mixture, test = 'F')
# no significant interaction
```

### Graphs chlorophyll a, b + c

```{r, Warning = FALSE}
# chl a
chla_plot <-
  ggplot(data = biofilm_endpoints,
         mapping = aes(x = factor(
           treatment, levels = c("NC", "Pro", "Cip", "Mix")
         ), y = chla)) +
  # Mean
  stat_summary(
    fun = mean,
    geom = "point",
    aes(fill = treatment),
    shape = 21,
    size = 5,
    color = "grey20",
    stroke = 0.5
  ) +
  
  # 95% CI
  # Black outline (thicker, underneath)
  stat_summary(
    fun.data = mean_cl_normal,
    geom = "errorbar",
    aes(group = treatment),
    width = 0.2,
    linewidth = 1,
    color = "grey30"
  ) +
  
  # Colored CI on top
  stat_summary(
    fun.data = mean_cl_normal,
    geom = "errorbar",
    aes(color = treatment),
    width = 0.2,
    linewidth = 0.8
  ) +
  
  # Jittered
  geom_point(
    aes(color = treatment),
    position = position_jitter(width = 0.25, height = 0),
    alpha = 0.8,
    size = 2
  ) +
  
  # Color scales
  scale_fill_manual(values = c(
    "NC" = "#009E73",
    "Pro" = "#E69F00",
    "Cip" = "#56B4E9",
    "Mix" = "#999999"
  )) +
  scale_color_manual(values = c(
    "NC" = "#009E73",
    "Pro" = "#E69F00",
    "Cip" = "#56B4E9",
    "Mix" = "#999999"
  )) +
  guides(fill = "none") +
  
  # Theme and labels
  labs(x = NULL,
       y = expression(bold("Chl a (µg/cm" ^ "2" * ")")),
       color = "") +
  scale_x_discrete(labels = NULL) +
  theme_minimal(base_size = 11) +
  theme(
    legend.position = "none",
    panel.background = element_rect(fill = "grey100"),
    axis.title.y = element_text(
      size = 13,
      face = "bold",
      colour = "grey30"
  ),
    axis.text.x = element_text(size = 12),  
    axis.text.y = element_text(size = 12),
    panel.border = element_rect(fill = NA, colour = "grey30"),
    panel.grid.major = element_line(colour = "grey85", size = 0.4),
    panel.grid.minor = element_line(colour = "grey95", size = 0.2)
  ) 

# Chlorophyll b
chlb_plot <-
  ggplot(data = biofilm_endpoints,
         mapping = aes(x = factor(
           treatment, levels = c("NC", "Pro", "Cip", "Mix")
         ), y = chlb)) +
  # Mean
  stat_summary(
    fun = mean,
    geom = "point",
    aes(fill = treatment),
    shape = 21,
    size = 5,
    color = "grey20",
    stroke = 0.5
  ) +
  
  # 95% CI
  # Black outline (thicker, underneath)
  stat_summary(
    fun.data = mean_cl_normal,
    geom = "errorbar",
    aes(group = treatment),
    width = 0.2,
    linewidth = 1,
    color = "grey30"
  ) +
  
  # Colored CI on top
  stat_summary(
    fun.data = mean_cl_normal,
    geom = "errorbar",
    aes(color = treatment),
    width = 0.2,
    linewidth = 0.8
  ) +
  
  # Jittered
  geom_point(
    aes(color = treatment),
    position = position_jitter(width = 0.25, height = 0),
    alpha = 0.8,
    size = 2
  ) +
  
  # Color scales
  scale_fill_manual(values = c(
    "NC" = "#009E73",
    "Pro" = "#E69F00",
    "Cip" = "#56B4E9",
    "Mix" = "#999999"
  )) +
  scale_color_manual(values = c(
    "NC" = "#009E73",
    "Pro" = "#E69F00",
    "Cip" = "#56B4E9",
    "Mix" = "#999999"
  )) +
  guides(fill = "none") +
  
  # Theme and labels
  labs(x = NULL,
       y = expression(bold("Chl b (µg/cm" ^ "2" * ")")),
       color = "") +
  scale_x_discrete(labels = NULL) +
  theme_minimal(base_size = 11) +
  theme(
    legend.position = "none",
    panel.background = element_rect(fill = "grey100"),
    axis.title = element_text(
      size = 13,
      face = "bold",
      colour = "grey30"
    ),
    axis.text.x = element_text(size = 12),  
    axis.text.y = element_text(size = 12),
    panel.border = element_rect(fill = NA, colour = "grey30"),
    panel.grid.major = element_line(colour = "grey85", size = 0.4),
    panel.grid.minor = element_line(colour = "grey95", size = 0.2)
  ) 

# Chlorophyll c
chlc_plot <-
  ggplot(data = biofilm_endpoints,
         mapping = aes(x = factor(
           treatment, levels = c("NC", "Pro", "Cip", "Mix")
         ), y = chlc)) +
  # Mean
  stat_summary(
    fun = mean,
    geom = "point",
    aes(fill = treatment),
    shape = 21,
    size = 5,
    color = "grey20",
    stroke = 0.5
  ) +
  
  # 95% CI
  # Black outline (thicker, underneath)
  stat_summary(
    fun.data = mean_cl_normal,
    geom = "errorbar",
    aes(group = treatment),
    width = 0.2,
    linewidth = 1,
    color = "grey30"
  ) +
  
  # Colored CI on top
  stat_summary(
    fun.data = mean_cl_normal,
    geom = "errorbar",
    aes(color = treatment),
    width = 0.2,
    linewidth = 0.8
  ) +
  
  # Jittered
  geom_point(
    aes(color = treatment),
    position = position_jitter(width = 0.25, height = 0),
    alpha = 0.8,
    size = 2
  ) +
  
  # Color scales
  scale_fill_manual(values = c(
    "NC" = "#009E73",
    "Pro" = "#E69F00",
    "Cip" = "#56B4E9",
    "Mix" = "#999999"
  )) +
  scale_color_manual(values = c(
    "NC" = "#009E73",
    "Pro" = "#E69F00",
    "Cip" = "#56B4E9",
    "Mix" = "#999999"
  )) +
  guides(fill = "none") +
  
  # Theme and labels
  labs(x = "Treatment",
       y = expression(bold("Chl c (µg/cm" ^ "2" * ")")),
       color = "") +
  scale_x_discrete(labels = c("Control", "Propyzamide", "Ciprofloxacin", "Mixture")) +
  theme_minimal(base_size = 11) +
  theme(
    legend.position = "none",
    panel.background = element_rect(fill = "grey100"),
    axis.title = element_text(
      size = 13,
      face = "bold",
      colour = "grey30"
    ),
    axis.text.x = element_text(size = 11),  
    axis.text.y = element_text(size = 12),
    panel.border = element_rect(fill = NA, colour = "grey30"),
    panel.grid.major = element_line(colour = "grey85", size = 0.4),
    panel.grid.minor = element_line(colour = "grey95", size = 0.2)
  ) 

# combine and save with AFDW plot later
```

## AFDW 

Calculate mean, se and 95% CI AFDW of each treatment.

```{r}
mean_afdw <- biofilm_endpoints %>%
  group_by(treatment) %>%
  dplyr::summarize(
    mean_afdw = mean(afdw, na.rm = TRUE),
    n = sum(!is.na(afdw)),
    se = sd(afdw, na.rm = TRUE) / sqrt(n),
    t_crit = qt(0.975, df = n - 1),
    ci_lower = mean_afdw - t_crit * se,
    ci_upper = mean_afdw + t_crit * se,
    .groups = "drop"
  ) %>%
  dplyr::select(treatment, mean_afdw, ci_lower, ci_upper, se)

print(mean_afdw)
```

### Histogram AFDW after colonisation

```{r}
hist(biofilm_endpoints $ afdw) 
skewness(biofilm_endpoints $ afdw, na.rm = TRUE) 
shapiro.test(biofilm_endpoints $ afdw)
```

Not normal distributed, right skewed but skewness \<1.5 therefore no data transformation for model.

### Model AFDW 

```{r, warning=FALSE}
model_afdw <- lm(afdw ~ treatment, data = biofilm_endpoints)
par(mfrow = c(2, 2))
plot(model_afdw) # rather s shape
qreference(residuals(model_afdw))

pairwise.wilcox.test(biofilm_endpoints$afdw,
                     biofilm_endpoints$treatment,
                     p.adjust.method = "holm")
```

No significant differences for treatment variable of total biofilm biomass.

### Boxplot AFDW 

```{r, Warning = FALSE}
afdw_boxplot <-
  ggplot(data = biofilm_endpoints,
         mapping = aes(x = factor(
           treatment, levels = c("NC", "Pro", "Cip", "Mix")
         ), y = afdw)) +
  # Mean
  stat_summary(
    fun = mean,
    geom = "point",
    aes(fill = treatment),
    shape = 21,
    size = 5,
    color = "grey20",
    stroke = 0.5
  ) +
  
  # 95% CI
  # Black outline (thicker, underneath)
  stat_summary(
    fun.data = mean_cl_normal,
    geom = "errorbar",
    aes(group = treatment),
    width = 0.2,
    linewidth = 1,
    color = "grey30"
  ) +
  
  # Colored CI on top
  stat_summary(
    fun.data = mean_cl_normal,
    geom = "errorbar",
    aes(color = treatment),
    width = 0.2,
    linewidth = 0.8
  ) +
  
  # Jittered
  geom_point(
    aes(color = treatment),
    position = position_jitter(width = 0.25, height = 0),
    alpha = 0.8,
    size = 2
  ) +
  
  # Color scales
  scale_fill_manual(values = c(
    "NC" = "#009E73",
    "Pro" = "#E69F00",
    "Cip" = "#56B4E9",
    "Mix" = "#999999"
  )) +
  scale_color_manual(values = c(
    "NC" = "#009E73",
    "Pro" = "#E69F00",
    "Cip" = "#56B4E9",
    "Mix" = "#999999"
  )) +
  guides(fill = "none") +
  
  # Theme and labels
  labs(x = "Treatment",
       y = expression(bold("AFDW (mg/cm"^"2"*")")),
       color = "") +
  scale_x_discrete(labels = c("Control", "Propyzamide", "Ciprofloxacin", "Mixture")) +
  theme_minimal(base_size = 11) +
  theme(
    legend.position = "none",
    panel.background = element_rect(fill = "grey100"),
    axis.title = element_text(
      size = 13,
      face = "bold",
      colour = "grey30"
    ),
    axis.text.x = element_text(size = 11),  
    axis.text.y = element_text(size = 12),
    panel.border = element_rect(fill = NA, colour = "grey30"),
    panel.grid.major = element_line(colour = "grey85", size = 0.4),
    panel.grid.minor = element_line(colour = "grey95", size = 0.2)
  ) 


afdw_boxplot 

# Now combine with chlorophyll plots
combined_plot <- chla_plot / chlb_plot / chlc_plot / afdw_boxplot + 
  plot_layout(ncol = 2) 

# save the plot
ggsave(
  "output/biofilm_endpoints.png",
  plot = combined_plot,
  width = 25,
  height = 30,
  units = "cm",
  dpi = 600
)

```

# Fatty acid composition biofilm

Statistics for fatty acid composition of the biofilm in absolute values (µg fatty acid / mg dry weight biofilm).

```{r}
# insert data
FA_Biofilm_absolute <-
  read.table(file.path("data\\FA_Biofilm_absolute.txt"),
             header = TRUE,
             sep = "\t")

View(FA_Biofilm_absolute)
```

## Absolute SAFA, MUFA & PUFA concentrations

1\) Total SAFA content

```{r, warning=FALSE}
# total SAFA, MUFA, PUFA content; ba = "biofilm absolute"
# total SAFA content
# filter saturated fatty acids in new table
SAFA_ba <- FA_Biofilm_absolute %>%
    filter(fa %in% c("C08:0", "C10:0", "C11:0","C12:0", "C13:0", "C14:0", "C15:0", "C16:0", "C18:0", "C20:0", "C21:0", "C22:0", "C24:0"))

# Sum SAFA per replicate and treatment
total_SAFA_per_replicate <- SAFA_ba %>%
  dplyr::group_by(replicate, treatment) %>%
  dplyr::summarize(sum_SAFA = sum(fa_conc), .groups = "drop")
print(total_SAFA_per_replicate)

# Statistical differences
hist(total_SAFA_per_replicate $ sum_SAFA) # normal distributed
shapiro.test(total_SAFA_per_replicate $ sum_SAFA) # normal distributed
leveneTest(sum_SAFA ~ treatment, data = total_SAFA_per_replicate) # homogenous
anova_SAFA <- aov(sum_SAFA ~ treatment, data = total_SAFA_per_replicate)
summary(anova_SAFA)
TukeyHSD(anova_SAFA)

# Calculate mean, se and 95% CI per treatment
mean_SAFA_ba <- total_SAFA_per_replicate %>%
  dplyr::group_by(treatment) %>%
  dplyr::summarize(
    mean_SAFA = mean(sum_SAFA, na.rm = TRUE),
    n = sum(!is.na(sum_SAFA)),
    se = sd(sum_SAFA, na.rm = TRUE) / sqrt(n),
    t_crit = qt(0.975, df = n - 1),
    ci_lower = mean_SAFA - t_crit * se,
    ci_upper = mean_SAFA + t_crit * se,
    .groups = "drop"
  ) %>%
  dplyr::select(treatment, mean_SAFA, ci_lower, ci_upper, se)

# Extract mean of control group for percental difference
control_mean_SAFA_ba <- mean_SAFA_ba %>%
  filter(treatment == "NC") %>%
  pull(mean_SAFA)

# Calculate percental difference from control
mean_SAFA_ba <- mean_SAFA_ba %>%
  mutate(percent_diff_from_NC = 100 * (mean_SAFA - control_mean_SAFA_ba) / control_mean_SAFA_ba)

print(mean_SAFA_ba)
```

2\) Total MUFA content

```{r, warning=FALSE}
# filter mono-unsaturated fatty acids in new table
MUFA_ba <- FA_Biofilm_absolute %>%
  filter(fa %in% c("C14:1", "C15:1","C16:1", "C17:1", "C18:1", "C20:1","C22:1", "C24:1"))

# Sum MUFA per replicate and treatment
total_MUFA_per_replicate <- MUFA_ba %>%
  dplyr::group_by(replicate, treatment) %>%
  dplyr::summarize(sum_MUFA = sum(fa_conc), .groups = "drop")
print(total_MUFA_per_replicate)

# Statistical differences
hist(total_MUFA_per_replicate $ sum_MUFA) # normal distributed
shapiro.test(total_MUFA_per_replicate $ sum_MUFA) # normal distributed
leveneTest(sum_MUFA ~ treatment, data = total_MUFA_per_replicate) # homogenous
anova_MUFA <- aov(sum_MUFA ~ treatment, data = total_MUFA_per_replicate)
summary(anova_MUFA)
TukeyHSD(anova_MUFA)

# Calculate mean, se and 95% CI per treatment
mean_MUFA_ba <- total_MUFA_per_replicate %>%
  dplyr::group_by(treatment) %>%
  dplyr::summarize(
    mean_MUFA = mean(sum_MUFA, na.rm = TRUE),
    n = sum(!is.na(sum_MUFA)),
    se = sd(sum_MUFA, na.rm = TRUE) / sqrt(n),
    t_crit = qt(0.975, df = n - 1),
    ci_lower = mean_MUFA - t_crit * se,
    ci_upper = mean_MUFA + t_crit * se,
    .groups = "drop"
  ) %>%
  dplyr::select(treatment, mean_MUFA, ci_lower, ci_upper, se)

# Extract mean of control group for percental difference
control_mean_MUFA_ba <- mean_MUFA_ba %>%
  filter(treatment == "NC") %>%
  pull(mean_MUFA)

# Calculate percental difference from control
mean_MUFA_ba <- mean_MUFA_ba %>%
  mutate(percent_diff_from_NC = 100 * (mean_MUFA - control_mean_MUFA_ba) / control_mean_MUFA_ba)

print(mean_MUFA_ba)
```

3\) Total PUFA content

```{r, warning=FALSE}
# filter poly-unsaturated fatty acids in new table
PUFA_ba <- FA_Biofilm_absolute %>%
  filter(fa %in% c("C18:2(n-6c)", "C18:2(n-6t)", "C18:3(n-3)", "C18:3(n-6)", "C18:4(n-3)", "C20:2(n-6)", "C20:3(n-3)", "C20:3(n-6)", "C20:4(n-6)", "C20:5(n-3)", "C22:2", "C22:5(n-3)", "C22:6(n-3)"))
  
# Sum PUFA per replicate and treatment
total_PUFA_per_replicate <- PUFA_ba %>%
  dplyr::group_by(replicate, treatment) %>%
  dplyr::summarize(sum_PUFA = sum(fa_conc), .groups = "drop")
print(total_PUFA_per_replicate)

# Statistical differences
hist(total_PUFA_per_replicate $ sum_PUFA) # normal distributed
shapiro.test(total_PUFA_per_replicate $ sum_PUFA) # normal distributed
leveneTest(sum_PUFA ~ treatment, data = total_PUFA_per_replicate) # homogenous
anova_PUFA <- aov(sum_PUFA ~ treatment, data = total_PUFA_per_replicate)
summary(anova_PUFA)
TukeyHSD(anova_PUFA)

# Calculate mean, se and 95% CI per treatment
mean_PUFA_ba <- total_PUFA_per_replicate %>%
  dplyr::group_by(treatment) %>%
  dplyr::summarize(
    mean_PUFA = mean(sum_PUFA, na.rm = TRUE),
    n = sum(!is.na(sum_PUFA)),
    se = sd(sum_PUFA, na.rm = TRUE) / sqrt(n),
    t_crit = qt(0.975, df = n - 1),
    ci_lower = mean_PUFA - t_crit * se,
    ci_upper = mean_PUFA + t_crit * se,
    .groups = "drop"
  ) %>%
  dplyr::select(treatment, mean_PUFA, ci_lower, ci_upper, se)

# Extract mean of control group for percental difference
control_mean_PUFA_ba <- mean_PUFA_ba %>%
  filter(treatment == "NC") %>%
  pull(mean_PUFA)

# Calculate percental difference from control
mean_PUFA_ba <- mean_PUFA_ba %>%
  mutate(percent_diff_from_NC = 100 * (mean_PUFA - control_mean_PUFA_ba) / control_mean_PUFA_ba)

print(mean_PUFA_ba)
```

SAFA, MUFA & PUFA plot relative to control:
```{r, warning=FALSE}
prep_class_from_replicates <- function(df_rep, value_col, class_name){
  df_rep %>%
    group_by(treatment) %>%
    summarise(
      mean = mean(.data[[value_col]], na.rm = TRUE),
      n    = sum(!is.na(.data[[value_col]])),
      se   = sd(.data[[value_col]], na.rm = TRUE) / sqrt(n),
      tcrit = qt(0.975, df = n - 1),
      lwr  = mean - tcrit * se,
      upr  = mean + tcrit * se,
      .groups = "drop"
    ) %>%
    dplyr::mutate(Class = class_name) %>%
    dplyr::select(treatment, Class, mean, lwr, upr)
}

biofilm_class_abs <- bind_rows(
  prep_class_from_replicates(total_SAFA_per_replicate, "sum_SAFA", "SAFA"),
  prep_class_from_replicates(total_MUFA_per_replicate, "sum_MUFA", "MUFA"),
  prep_class_from_replicates(total_PUFA_per_replicate, "sum_PUFA", "PUFA")
) %>%
  mutate(
    Class = factor(Class, levels = c("SAFA","PUFA","MUFA")),
    treatment = factor(treatment, levels = c("NC", "Cip","Pro","Mix"))
  )

biofilm_class_abs

# relative values
biofilm_class_rel <- biofilm_class_abs %>%
  group_by(Class) %>%
  mutate(
    ctrl_mean = mean[treatment == "NC"][1],
    mean = mean / ctrl_mean,
    lwr  = lwr  / ctrl_mean,
    upr  = upr  / ctrl_mean
  ) %>%
  ungroup() %>%
  dplyr::select(-ctrl_mean)


plot_df <- biofilm_class_rel  

col_xeno <- c(
  "Pro" = "#E69F00",
  "Cip" = "#56B4E9",
  "Mix" = "#999999"
)

plot_df <- plot_df %>%
  mutate(
    show = ifelse(treatment == "NC", 0, 1),  
    treatment = factor(treatment, levels = c("NC","Pro","Cip","Mix")),
    Class = factor(Class, levels = c("PUFA", "MUFA", "SAFA")),
    lwr = ifelse(lwr < 0, 0, lwr)
  )

# plot
fa_biofilm_relative_changes <- ggplot(plot_df, aes(x = mean, y = Class)) +

  geom_vline(xintercept = 1, linewidth = 1, colour = "grey30") +

  geom_errorbarh(
    aes(xmin = lwr, xmax = upr, group = treatment, alpha = show),
    height = 0,
    linewidth = 1,
    colour = "grey30",
    position = position_dodge(width = 0.45),
    show.legend = FALSE
  ) +

  geom_errorbarh(
    aes(xmin = lwr, xmax = upr, colour = treatment, alpha = show),
    height = 0,
    linewidth = 0.8,
    position = position_dodge(width = 0.45),
    show.legend = FALSE
  ) +

  geom_point(
    aes(colour = treatment, alpha = show),
    size = 4,
    position = position_dodge(width = 0.45),
    show.legend = FALSE
  ) +

  scale_colour_manual(
    values = c(
    "NC" = "#009E73",
    "Pro" = "#E69F00",
    "Cip" = "#56B4E9",
    "Mix" = "#999999"
    ),
    labels = c(
      "NC"  = "Control",
      "Cip" = "Ciprofloxacin",
      "Pro" = "Propyzamide",
      "Mix" = "Mixture"
    ),
    name = "Treatment"
  ) +

  scale_alpha_identity() +

  guides(
    colour = guide_legend(override.aes = list(alpha = 1))
  ) +

  labs(
    x = "Fold change relative to control",
    y = NULL
  ) +

  theme_minimal(base_size = 11) +
  theme(
    legend.position = "right",
    legend.title = element_text(size = 12, face = "bold", colour = "grey30"),
    legend.text  = element_text(size = 11, colour = "grey30"),
    panel.background = element_rect(fill = "grey100"),
    axis.title = element_text(size = 13, face = "bold", colour = "grey30"),
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(size = 13, face = "bold", colour = "grey30"),
    panel.border = element_rect(fill = NA, colour = "grey30"),
    panel.grid.major = element_line(colour = "grey95", size = 0.4),
    panel.grid.minor = element_line(colour = "grey95", size = 0.2)
  )

fa_biofilm_relative_changes

# save the plot
ggsave(
  "output/fa_biofilm_relative_changes.png",
  plot = fa_biofilm_relative_changes,
  width = 16.5,
  height = 12,
  units = "cm",
  dpi = 600
)
```


4\) Total FA content

```{r, warning=FALSE}
# total FA 
total_FA_biofilm <- bind_rows(total_SAFA_per_replicate, total_MUFA_per_replicate, total_PUFA_per_replicate)

total_FA_per_replicate <- total_FA_biofilm %>%
  dplyr::group_by(replicate, treatment) %>%
  dplyr::summarize(sum_FA = sum(sum_SAFA, sum_MUFA, sum_PUFA, na.rm = TRUE))
print(total_FA_per_replicate)

# statistical differences will be calculated by PERMANOVA

# Calculate mean, se and 95% CI per treatment
mean_FA_ba <- total_FA_per_replicate %>%
  dplyr::group_by(treatment) %>%
  dplyr::summarize(
    mean_FA = mean(sum_FA, na.rm = TRUE),
    n = sum(!is.na(sum_FA)),
    se = sd(sum_FA, na.rm = TRUE) / sqrt(n),
    t_crit = qt(0.975, df = n - 1),
    ci_lower = mean_FA - t_crit * se,
    ci_upper = mean_FA + t_crit * se,
    .groups = "drop"
  ) %>%
  dplyr::select(treatment, mean_FA, ci_lower, ci_upper, se)

# Extract mean of control group for percental difference
control_mean_FA_ba <- mean_FA_ba %>%
  filter(treatment == "NC") %>%
  pull(mean_FA)

# Calculate percental difference from control
mean_FA_ba <- mean_FA_ba %>%
  mutate(percent_diff_from_NC = 100 * (mean_FA - control_mean_FA_ba) / control_mean_FA_ba)

print(mean_FA_ba)
```

## Histogram FA biofilm

```{r}
hist(FA_Biofilm_absolute $ fa_conc) 
skewness(FA_Biofilm_absolute $ fa_conc, na.rm = TRUE) 
shapiro.test(FA_Biofilm_absolute $ fa_conc)
```

Biofilm fatty acid concentration not normal distributed and right skewed.

## NMDS FA biofilm

Change table to wide format:

```{r}
# change table to wide format
FA_Biofilm_absolute_wide_sqrt <-
  pivot_wider(FA_Biofilm_absolute,
              names_from = fa,
              values_from = fa_conc)

# square root transformation, to reduce discriminatory power of dominant fa groups
FA_Biofilm_absolute_wide_sqrt[, 4:37] <- sqrt(FA_Biofilm_absolute_wide_sqrt[, 4:37])

View(FA_Biofilm_absolute_wide_sqrt)

# Only keep columns/FAs with more than 5 non-zero values (fatty acids that have just been found in 5 replicates will be removed)
FA_Biofilm_absolute_wide_sqrt <-
  FA_Biofilm_absolute_wide_sqrt %>%
  dplyr::select(
    1:3,
    where(~sum(. !=0, na.rm = TRUE)>5)
  )
```

Perform metaMDS:

```{r}
# Rename and ease col-names for later NMDS plot
FA_Biofilm_absolute_wide_sqrt <- FA_Biofilm_absolute_wide_sqrt %>%
  rename(
    `12:0`     = `C12:0`,   
    `13:0`     = `C13:0`,
    `14:0`     = `C14:0`,
    `15:0`     = `C15:0`,
    `16:0`     = `C16:0`,
    `16:1`     = `C16:1`,
    `17:1`     = `C17:1`,
    `18:0`     = `C18:0`,
    `18:1`     = `C18:1`,
    `18:2n-6c` = `C18:2(n-6c)`,
    `18:3n-6`  = `C18:3(n-6)`,
    `18:3n-3`  = `C18:3(n-3)`,
    `18:4n-3`  = `C18:4(n-3)`,
    `20:0`     = `C20:0`,
    `20:1`     = `C20:1`,
    `20:2n-6`  = `C20:2(n-6)`,
    `20:4n-6`  = `C20:4(n-6)`,
    `20:5n-3`  = `C20:5(n-3)`,
    `22:0`     = `C22:0`,
    `22:1`     = `C22:1`,
    `22:6n-3`  = `C22:6(n-3)`,
    `24:0`     = `C24:0`
  )

# Remove columns 1:3 for metaMDS
FA_Biofilm_absolute_wide_sqrt <- FA_Biofilm_absolute_wide_sqrt[, -c(1, 2, 3)]

NMDS_Biofilm_absolute <-
  metaMDS(
    FA_Biofilm_absolute_wide_sqrt,
    autotransform = FALSE,
    k = 2 ,
    distance = "bray",
    na.rm = TRUE
  )

NMDS_Biofilm_absolute$stress
```

Create fa vectors:

```{r}
# separation in SAFA; MUFA & PUFA to build vectors in NMDS
# "ba" = biofilm absolute
FA_variables.SAFAs_ba <-
  envfit(
    NMDS_Biofilm_absolute,
    subset(
      FA_Biofilm_absolute_wide_sqrt,
      select = c(
        "12:0",
        "13:0",
        "14:0",
        "15:0",
        "16:0",
        "18:0",
        "20:0",
        "22:0",
        "24:0"
      )
    )
    ,
    permutations = 999,
    na.rm = TRUE
  )

FA_variables.MUFAs_ba <-
  envfit(
    NMDS_Biofilm_absolute,
    subset(
      FA_Biofilm_absolute_wide_sqrt,
      select = c(
        "16:1",
        "17:1",
        "18:1",
        "20:1",
        "22:1"
      )
    )
    ,
    permutations = 999,
    na.rm = TRUE
  )

FA_variables.PUFAs_ba <-
  envfit(
    NMDS_Biofilm_absolute,
    subset(
      FA_Biofilm_absolute_wide_sqrt,
      select = c(
        "18:2n-6c",
        "18:3n-6",
        "18:3n-3",
        "18:4n-3",
        "20:2n-6",
        "20:4n-6",
        "20:5n-3",
        "22:6n-3"
      )
    )
    ,
    permutations = 999,
    na.rm = TRUE
  )

# Extract NMDS scores
data.scores_biofilm_absolute = as.data.frame(scores(NMDS_Biofilm_absolute)$sites)

Treatments <-
  FA_Biofilm_absolute$treatment[FA_Biofilm_absolute$fa == "C16:0"]

# add treatment column to NMDS scores
data.scores_biofilm_absolute$Treatment = Treatments

# vectors for SAFA, MUFA and PUFA
SAFAs_coord_cont_ba = as.data.frame(scores(FA_variables.SAFAs_ba, "vectors")) * ordiArrowMul(FA_variables.SAFAs_ba)
MUFAs_coord_cont_ba = as.data.frame(scores(FA_variables.MUFAs_ba, "vectors")) * ordiArrowMul(FA_variables.MUFAs_ba)
PUFAs_coord_cont_ba = as.data.frame(scores(FA_variables.PUFAs_ba, "vectors")) * ordiArrowMul(FA_variables.PUFAs_ba)

```

Plot NMDS:

```{r, warning=FALSE}
# Reorder treatment variable for the plot
data.scores_biofilm_absolute$Treatment <- factor(
  data.scores_biofilm_absolute$Treatment,
  levels = c("NC", "Pro", "Cip", "Mix"),
  labels  = c("Control", "Propyzamide", "Ciprofloxacin", "Mixture")
)

# Function to rescale coordinates
rescale_vectors <- function(df, scale_factor = 1) {
  df$length <- sqrt(df$NMDS1^2 + df$NMDS2^2)
  max_length <- max(df$length)
  df$NMDS1 <- df$NMDS1 / max_length * scale_factor
  df$NMDS2 <- df$NMDS2 / max_length * scale_factor
  df$length <- NULL
  return(df)
}

# Apply to each group
PUFAs_coord_cont_ba <- rescale_vectors(PUFAs_coord_cont_ba, 0.3)
MUFAs_coord_cont_ba <- rescale_vectors(MUFAs_coord_cont_ba, 0.3)
SAFAs_coord_cont_ba <- rescale_vectors(SAFAs_coord_cont_ba, 0.3)


# Compute convex hull for each Treatment
polygon_data_biofilm_treatment <- data.scores_biofilm_absolute %>%
  group_by(Treatment) %>%
  summarise(hull_points = list(data.frame(
    NMDS1 = NMDS1[chull(NMDS1, NMDS2)],
    NMDS2 = NMDS2[chull(NMDS1, NMDS2)]
  ))) %>%
  unnest(hull_points)

# plot
absol_biofilm_with_vectors = ggplot(data = data.scores_biofilm_absolute, aes(x = NMDS1, y = NMDS2)) +
  geom_polygon(
    data = polygon_data_biofilm_treatment,
    aes(
      x = NMDS1,
      y = NMDS2,
      group = Treatment,
      fill = Treatment
    ),
    alpha = 0.2,
    show.legend = FALSE
  ) +
  geom_point(aes(shape = Treatment, color = Treatment),
             size = 5,
             alpha = 1,
             show.legend = TRUE) +
  scale_shape_manual(values = c(
    "Control" = 16,
    "Propyzamide" = 17,
    "Ciprofloxacin" = 18,
    "Mixture" = 15
  )) +
  scale_color_manual(values = c(
    "Control" = "#009E73",
    "Propyzamide" = "#E69F00",
    "Ciprofloxacin" = "#56B4E9",
    "Mixture" = "#999999"
  )) +
  scale_fill_manual(values = c(
    "Control" = "#009E73",
    "Propyzamide" = "#E69F00",
    "Ciprofloxacin" = "#56B4E9",
    "Mixture" = "#999999"
  ))  +
  labs(color = "Treatment") +
  geom_segment(
    data = PUFAs_coord_cont_ba,
    aes(
      x = 0,
      y = 0,
      xend = NMDS1,
      yend = NMDS2
    ),
    linewidth = 0.5,
    alpha = 0.2,
    colour = "black"
  ) +
  geom_segment(
    data = MUFAs_coord_cont_ba,
    aes(
      x = 0,
      y = 0,
      xend = NMDS1,
      yend = NMDS2
    ),
    linewidth = 0.5,
    alpha = 0.2,
    colour = "black"
  ) +
  geom_segment(
    data = SAFAs_coord_cont_ba,
    aes(
      x = 0,
      y = 0,
      xend = NMDS1,
      yend = NMDS2
    ),
    linewidth = 0.5,
    alpha = 0.2,
    colour = "black"
  ) +
  geom_text_repel(
    data = PUFAs_coord_cont_ba,
    aes(
      x = NMDS1,
      y = NMDS2,
      label = row.names(PUFAs_coord_cont_ba)
    ),
    colour = "black",
    size = 2.5
  ) +
  geom_text_repel(
    data = MUFAs_coord_cont_ba,
    aes(
      x = NMDS1,
      y = NMDS2,
      label = row.names(MUFAs_coord_cont_ba)
    ),
    colour = "#996600",
    size = 2.5
  ) +
  geom_text_repel(
    data = SAFAs_coord_cont_ba,
    aes(
      x = NMDS1,
      y = NMDS2,
      label = row.names(SAFAs_coord_cont_ba)
    ),
    colour = "#0000FF",
    size = 2.5
  ) +
  theme_bw() +
  theme(
    axis.title = element_text(
      size = 13,
      face = "bold",
      colour = "grey30"
    ),
    panel.background = element_blank(),
    panel.border = element_rect(fill = NA, colour = "grey30"),
    legend.key = element_blank(),
    legend.title = element_text(
      size = 11,
      face = "bold",
      colour = "grey30"
    ),
    legend.text = element_text(size = 10, colour = "grey30"),
    plot.margin = margin(
      t = 0.2,
      b = 0.2,
      l = 0.2,
      r = 0.2,
      unit = "cm"
    ),
    panel.grid.major = element_line(colour = "grey95", size = 0.4),
    panel.grid.minor = element_line(colour = "grey95", size = 0.2)
  )

absol_biofilm_with_vectors
# position of FA names for publication were manually adjusted in vector graphic with the software Inkscape
```

Save the plot as .png:

```{r}
ggsave(
  "output/NMDS_Biofilm_absolute.png",
  width = 16.5,
  height = 12,
  units = "cm",
  dpi = 600
)
```

Save the plot as vector graphic:

```{r}
ggsave(
  "output/NMDS_Biofilm_absolute.svg",
  device = "svg",
  width = 16.5,
  height = 16,
  units = "cm",
  dpi = 600
)
```

## PERMANOVA FA biofilm

Create distance matrix with fatty acids:

```{r}
FA_biofilm_dist <- vegdist(FA_Biofilm_absolute_wide_sqrt, method = 'bray')
```

Separate explanatory variables from fatty acid information:

```{r}
FA_biofilm_absolute_wide_permanova <-
  pivot_wider(FA_Biofilm_absolute,
              names_from = fa,
              values_from = fa_conc)

# square root transformation again, to reduce discriminatory power of dominant fa groups
FA_biofilm_absolute_wide_permanova[, 4:37] <- sqrt(FA_biofilm_absolute_wide_permanova[, 4:37])

# delete fa columns
FA_biofilm_absolute_wide_permanova <-
  FA_biofilm_absolute_wide_permanova[, c(1, 2, 3)]

FA_biofilm_absolute_wide_permanova$treatment <-
  as.factor(FA_biofilm_absolute_wide_permanova$treatment)
```

Perform PERMANOVA for treatments:

```{r}
# perform PERMANOVA for treatments
perm_FA_biofilm_treatment <- adonis2(FA_biofilm_dist ~ treatment, data = FA_biofilm_absolute_wide_permanova, permutations = 999)
print(perm_FA_biofilm_treatment)

```

No significant differences between biofilm treatments. Anyway pairwise comparisons (by: Martinez Arbizu, P. (2020). pairwiseAdonis: Pairwise multilevel comparison using adonis. R package version 0.4):

```{r}
pairwise_results_biofilm_treatment <- pairwise.adonis2(
  FA_biofilm_dist ~ treatment,
  data = FA_biofilm_absolute_wide_permanova,
  permutations = 99
)

# View the results
print(pairwise_results_biofilm_treatment)

# Cip*Pro interaction?
# add presence / absence columns
FA_biofilm_absolute_wide_permanova$Cip <- ifelse(FA_biofilm_absolute_wide_permanova$treatment %in% c("Cip", "Mix"),
                                                 1,
                                                 0)

FA_biofilm_absolute_wide_permanova$Cip <- 
  factor(FA_biofilm_absolute_wide_permanova$Cip, levels = c(0, 1))

FA_biofilm_absolute_wide_permanova$Pro <- ifelse(FA_biofilm_absolute_wide_permanova$treatment %in% c("Pro", "Mix"),
                                                 1,
                                                 0)

FA_biofilm_absolute_wide_permanova$Pro <- 
  factor(FA_biofilm_absolute_wide_permanova$Pro, levels = c(0, 1))

# sanity check
table(FA_biofilm_absolute_wide_permanova$treatment,
      FA_biofilm_absolute_wide_permanova$Cip)

table(FA_biofilm_absolute_wide_permanova$treatment,
      FA_biofilm_absolute_wide_permanova$Pro)

# permanova
perm_FA_biofilm_Cip_Pro <- adonis2(FA_biofilm_dist ~ Cip * Pro,
        data = FA_biofilm_absolute_wide_permanova,
        permutations = 999)

print(perm_FA_biofilm_Cip_Pro)
# no interaction
```

## SIMPER FA biofilm

```{r}
# create extra table with treatments
biofilm_treatments_table <- FA_biofilm_absolute_wide_permanova

View(biofilm_treatments_table)

# Perform SIMPER
simper_FA_biofilm_treatments <-
  with(
    biofilm_treatments_table,
    simper(FA_Biofilm_absolute_wide_sqrt, treatment,  permutations = 999)
  )

summary(simper_FA_biofilm_treatments)
```

```{r}
simper_sum <- summary(simper_FA_biofilm_treatments)

# save results
out_dir <- "output"
dir.create(out_dir, showWarnings = FALSE)

safe_name <- function(x) gsub("[^A-Za-z0-9_]+", "_", x)

iwalk(simper_sum, function(x, nm) {
  if (is.null(x)) return(invisible(NULL))

  # SIMPER summary as a data.frame:
  # columns: average, sd, ratio, ava, avb, cumsum, p
  # rownames are the fatty acid IDs
  df <- as.data.frame(x)

  # keep rows up to the first that reaches/exceeds 70%
  idx <- which(df$cumsum >= 0.70)[1]
  if (is.na(idx)) idx <- nrow(df)
  out <- df[seq_len(idx), , drop = FALSE]

  # just keep 3 decimals
  out_write <- out
  num_cols <- vapply(out_write, is.numeric, logical(1))
  out_write[, num_cols] <- lapply(out_write[, num_cols, drop = FALSE],
                                  function(col) formatC(col, format = "f", digits = 3))

  file <- file.path(out_dir, paste0(safe_name(nm), "_top70.tsv"))
  # write row names so the TSV mirrors the SIMPER structure
  write.table(out_write, file, sep = "\t", quote = FALSE, col.names = NA)
})
```

```{r}
# Combine results into a data frame
simper_results_biofilm_treatments <- do.call(rbind, lapply(simper_FA_biofilm_treatments, function(x) {
  # Check if the element contains valid data
  if (!is.null(x$p)) {
    data.frame(
      fatty_acid = names(x$p),  
      p = as.numeric(x$p),      
      average = as.numeric(x$average),  
      sd = as.numeric(x$sd),
      ratio = as.numeric(x$ratio),
      cusum = as.numeric(x$cusum),
      stringsAsFactors = FALSE
    )
  } else {
    NULL  # Skip if no p-values
  }
}))

# Show only significant results
simper_results_biofilm_treatments %>%
  dplyr::filter(p <= 0.05) %>%
  dplyr::select(fatty_acid, p, average, sd, ratio, cusum)  
```

```{r}
# How often do they occur
simper_results_biofilm_treatments %>%
  dplyr::filter(p <= 0.05) %>%
  dplyr::count(fatty_acid) %>%
  dplyr::arrange(desc(n)) %>%
  print()
```

# Dryweight snails

Unit of soft body dry weight = mg

```{r}
# insert data (Culture snails from in-house breeding)
dryweight_snails <-
  read.table(file.path("data\\dryweight_snails.txt"),
             header = TRUE, sep = "\t")

View(dryweight_snails)
# 3 snails from the 20 baseline (culture) snails were lost during the analyses.
```

## Calculate mean dryweight, se and 95% CI of all treatment snails and the culture snails

```{r}
mean_dryweight_snails <- dryweight_snails %>%
  group_by(treatment) %>%
  dplyr::summarize(
    mean_dryweight_snails = mean(dryweight_mg, na.rm = TRUE),
    n = sum(!is.na(dryweight_mg)),
    se = sd(dryweight_mg, na.rm = TRUE) / sqrt(n),
    t_crit = qt(0.975, df = n - 1),
    ci_lower = mean_dryweight_snails - t_crit * se,
    ci_upper = mean_dryweight_snails + t_crit * se,
    .groups = "drop"
  ) %>%
  dplyr::select(treatment, mean_dryweight_snails, ci_lower, ci_upper, se)
  
print(mean_dryweight_snails)
```

Snails from feeding assay (fed with biofilm) on average heavier than snails from culture (fed with Tetraphyll flakes).

```{r}
#culture/baseline snails
culture_snails <- dryweight_snails %>%
  dplyr::filter(c(treatment == "Culture"))

hist(culture_snails $ dryweight_mg) 
skewness(culture_snails $ dryweight_mg, na.rm = TRUE) 
shapiro.test(culture_snails $ dryweight_mg)
# culture / baseline snails normal distributed

# calculate differences in final dry weight in comparison to culture snails
percentage_differences <- dryweight_snails %>%
  group_by(treatment) %>%
  dplyr::summarize(mean_dryweight = mean(dryweight_mg)) %>%
  mutate(percentage_differences = (mean_dryweight / mean_dryweight[treatment == "Culture"]) * 100 - 100)

percentage_differences
```

## Histogram

```{r}
# filter culture snails
filtered_dry_weight <- dryweight_snails %>%
  filter(!treatment %in% c("Culture"))

hist(filtered_dry_weight $ dryweight_mg) 
skewness(filtered_dry_weight $ dryweight_mg, na.rm = TRUE) 
shapiro.test(filtered_dry_weight $ dryweight_mg)
```

Soft tissue dry weights normal distributed.

## Model dry weight snails

```{r}
# explanatory variable treatment as factor
filtered_dry_weight $ treatment <- factor(filtered_dry_weight $ treatment)
  
model_snails <- lm(dryweight_mg ~ treatment, data = filtered_dry_weight)
par(mfrow = c(2, 2))
plot(model_snails) # normal distribution assumed
qreference(residuals(model_snails)) 
drop1(model_snails, test = 'F')
```

Explanatory variable "treatment" highly significant. 

```{r}
dw_snails_inter <- lsmeans(model_snails, ~ treatment)

dw_snails_inter_dunn1 <- contrast(dw_snails_inter, 'pairwise')

dw_snails_inter_dunn_sum1 <- summary(dw_snails_inter_dunn1, infer = TRUE, adjust = 'mvt')

dw_snails_inter_dunn1
```

Significant difference between dry weight of Cip and Mix snails compared to NC; almost significant difference between Pro treatment and NC.

Test for interaction effects of Cip and Pro:
```{r}
# interaction effect model with presence / absence of single xenobiotics
filtered_dry_weight$Cip <- ifelse(filtered_dry_weight$treatment %in% c("Cip", "Mix"),
                                1, 0)

filtered_dry_weight$Cip <- factor(filtered_dry_weight$Cip, levels = c(0, 1))

filtered_dry_weight$Pro <- ifelse(filtered_dry_weight$treatment %in% c("Pro", "Mix"),
                                1, 0)

filtered_dry_weight$Pro <- factor(filtered_dry_weight$Pro, levels = c(0, 1))

# sanity check
table(filtered_dry_weight$treatment,
      filtered_dry_weight$Cip,
      filtered_dry_weight$Pro)

# interaction model
model_dryweight_interact <- lm(dryweight_mg ~ Cip * Pro, data = filtered_dry_weight)
par(mfrow = c(2, 2))
plot(model_dryweight_interact) # point cloud looks evenly distributed --> normal distribution
qreference(residuals(model_dryweight_interact)) 
drop1(model_dryweight_interact, test = 'F')
# Cip:Pro interaction not significant
```

```{r}
# df.ref only control values
df.ref.dryweight<- filtered_dry_weight |>
  dplyr::filter(Cip == 0, Pro == 0) 

nd_dryweight_snails <- nd.generator(df.ref.dryweight, stressor1 = "Cip", stressor2 = "Pro", diff = NULL)

#sanity check  
table(nd_dryweight_snails$dataset, nd_dryweight_snails$Cip)
table(nd_dryweight_snails$dataset, nd_dryweight_snails$Pro)

#check for AD, MU (all values >0) and OD
hyp_all_dryweight_snails <- predictions(
  model_dryweight_interact,
  newdata = nd_dryweight_snails,
  hypothesis = function(x) {
    hypothesis.function(
      x,
      stressor1 = "Cip",
      stressor2 = "Pro",
      null = "all",        # <- AD, MU, DO together
      strategy = "average"
    )
  }
) |>
  inferences(method = "simulation")

hyp_all_dryweight_snails

hyp_all_dryweight_snails$term <- dplyr::recode(hyp_all_dryweight_snails$term,
                                    "ad" = "AD",
                                    "do" = "DO",
                                    "mu" = "MU")

ggplot(hyp_all_dryweight_snails, aes(x = term, y = estimate)) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.2) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  labs(x = "Null model", y = "Interaction estimate")
```
Both ciprofloxacin and propyzamide significantly affected snail dry weight in the additive model, whereas the interaction term was not significant. Post-estimation inference following Pimentel et al. (2025) showed that the additive, multiplicative, and dominance null models could not be rejected, indicating that the observed mixture response did not deviate significantly from any of these expectations. Together, these results suggest that the effects of both stressors on snail dry weight were largely independent and did not exhibit a detectable interaction.

## Boxplot dry weight snails

```{r, warning = FALSE}
# Plot
snails_boxplot <-
  ggplot(data = filtered_dry_weight,
         mapping = aes(x = factor(
           treatment, levels = c("NC", "Pro", "Cip", "Mix")
         ), y = dryweight_mg)) +
  # Mean
  stat_summary(
    fun = mean,
    geom = "point",
    aes(fill = treatment),
    shape = 21,
    size = 5,
    color = "grey20",
    stroke = 0.5
  ) +
  
  # 95% CI
  # Black outline (thicker, underneath)
  stat_summary(
    fun.data = mean_cl_normal,
    geom = "errorbar",
    aes(group = treatment),
    width = 0.2,
    linewidth = 1,
    color = "grey30"
  ) +
  
  # Colored CI on top
  stat_summary(
    fun.data = mean_cl_normal,
    geom = "errorbar",
    aes(color = treatment),
    width = 0.2,
    linewidth = 0.8
  ) +
  
  # Jittered
geom_point(
  aes(color = treatment),
  position = position_jitter(width = 0.25, height = 0),
  alpha = 0.8,
  size = 2
) +
  
  # Color scales
  scale_fill_manual(values = c(
    "NC" = "#009E73",
    "Pro" = "#E69F00",
    "Cip" = "#56B4E9",
    "Mix" = "#999999"
  )) +
  scale_color_manual(values = c(
    "NC" = "#009E73",
    "Pro" = "#E69F00",
    "Cip" = "#56B4E9",
    "Mix" = "#999999"
  )) +
  guides(fill = "none") +
  
  # Theme and labels
  labs(x = "Treatment", y = "Grazer Dry Weight (mg)", color = "") +
  scale_x_discrete(labels = c("Control", "Propyzamide", "Ciprofloxacin", "Mixture")) +
  theme_minimal(base_size = 11) +
  theme(
    legend.position = "none",
    panel.background = element_rect(fill = "grey100"),
    axis.title = element_text(
      size = 13,
      face = "bold",
      colour = "grey30"
    ),
    axis.text.x = element_text(size = 12),  
    axis.text.y = element_text(size = 12),
    panel.border = element_rect(fill = NA, colour = "grey30"),
    panel.grid.major = element_line(colour = "grey85", size = 0.4),
    panel.grid.minor = element_line(colour = "grey95", size = 0.2)
  ) +
  # Significance annotations
  annotate(
    "text",
    x = 1.5,
    y = max(filtered_dry_weight$dryweight_mg) + 0.07,
    label = "p=0.056",
    size = 4,
    color = "grey30"
  ) +
  geom_segment(
    aes(
      x = 1,
      xend = 2,
      y = max(dryweight_mg) + 0.05,
      yend = max(dryweight_mg) + 0.05
    ),
    lineend = "round",
    linewidth = 0.5,
    color = "grey30"
  ) +
  # Vertical tick on the left 
  geom_segment(
    aes(
      x = 1,
      xend = 1,
      y = max(dryweight_mg) + 0.04,
      yend = max(dryweight_mg) + 0.05
    ),
    color = "grey30",
    linewidth = 0.5
  ) +
  # # Vertical tick on the right 
  geom_segment(
    aes(
      x = 2,
      xend = 2,
      y = max(dryweight_mg) + 0.04,
      yend = max(dryweight_mg) + 0.05
    ),
    color = "grey30",
    linewidth = 0.5
  ) +
  annotate(
    "text",
    x = 2.5,
    y = max(filtered_dry_weight$dryweight_mg) + 0.17,
    label = "p=0.005**",
    size = 4,
    color = "grey30"
  ) +
  geom_segment(
    aes(
      x = 1,
      xend = 4,
      y = max(dryweight_mg) + 0.15,
      yend = max(dryweight_mg) + 0.15
    ),
    lineend = "round",
    linewidth = 0.5,
    color = "grey30"
  ) +
  geom_segment(
    aes(
      x = 1,
      xend = 1,
      y = max(dryweight_mg) + 0.14,
      yend = max(dryweight_mg) + 0.15
    ),
    color = "grey30",
    linewidth = 0.5
  ) +
  geom_segment(
    aes(
      x = 4,
      xend = 4,
      y = max(dryweight_mg) + 0.14,
      yend = max(dryweight_mg) + 0.15
    ),
    color = "grey30",
    linewidth = 0.5
  ) +
  annotate(
    "text",
    x = 2,
    y = max(filtered_dry_weight$dryweight_mg) + 0.12,
    label = "p=0.022*",
    size = 4,
    color = "grey30"
  ) +
  geom_segment(
    aes(
      x = 1,
      xend = 3,
      y = max(dryweight_mg) + 0.1,
      yend = max(dryweight_mg) + 0.1
    ),
    lineend = "round",
    linewidth = 0.5,
    color = "grey30"
  ) +
  geom_segment(
    aes(
      x = 1,
      xend = 1,
      y = max(dryweight_mg) + 0.09,
      yend = max(dryweight_mg) + 0.1
    ),
    color = "grey30",
    linewidth = 0.5
  ) +
  geom_segment(
    aes(
      x = 3,
      xend = 3,
      y = max(dryweight_mg) + 0.09,
      yend = max(dryweight_mg) + 0.1
    ),
    color = "grey30",
    linewidth = 0.5
  )


snails_boxplot
```

```{r, warning = FALSE}
# save the plot
ggsave(
  "output/snails_boxplot.png",
  width = 16.5,
  height = 12,
  units = "cm",
  dpi = 600
)
```

# Feeding rate snails

-   NA's due to missing snail and zero values of feeding rate due to no measurable feeding in respective week

-   Unit feeding rate: µg / day 

-   Column "used_for_FA_analysis" highlights the ten snails used for FA analyses for each treatment

```{r}
# insert data
afdw_feeding_rate <-
  read.table(file.path("data\\feeding_assay_afdw_feeding_rate.txt"),
             header = TRUE, sep = "\t")

View(afdw_feeding_rate)
```

## Feeding rate snails

#### Histogram feeding rate
```{r}
hist(afdw_feeding_rate $ feeding_rate) 
skewness(afdw_feeding_rate $ feeding_rate, na.rm = TRUE) # skewness > 3
shapiro.test(afdw_feeding_rate $ feeding_rate) # not normal distributed
# following visual observation of absolute data & histogram, check for outliers
```

Outlier check:
```{r}
#Run function one time, after that you can use the shortcut outlierKD (your subset, variable)
outlierKD_feeding_rate <- function(dt, id_col = "treatment", start_col = 6, end_col = 6) {
  # Initialize an empty list to store results for each column
  outlier_results <- list()
  
  
  for (i in start_col:end_col) {
    col_name <- colnames(dt)[i]  # Get the column name
    
    # Extract the column values
    var_name <- dt[[col_name]]
    
    # Check if the column is numeric
    if (is.numeric(var_name)) {
      na1 <- sum(is.na(var_name))
      m1 <- mean(var_name, na.rm = TRUE)
      
      # Identify outliers using boxplot stats
      outlier <- boxplot.stats(var_name)$out  
      
      # Find rows with outliers
      outlier_rows <- which(var_name %in% outlier)
      
      # If no outliers, skip the rest of the loop
      if (length(outlier_rows) == 0) next
      
      # Get replicate values for the outliers
      id_values <- dt[[id_col]][outlier_rows]
      
      # Mean of outliers
      mo <- mean(outlier)
      
      # Collect the outlier information for this column
      outlier_results[[col_name]] <- data.frame(
        Column = col_name,
        Row = outlier_rows,
        Replicate = id_values,
        Outlier = outlier
      )
      
      cat("\nOutliers for column:", col_name, "\n")
      cat("Outliers identified:", length(outlier), "\n")
      cat("Proportion (%) of outliers:", 
          round(length(outlier) / sum(!is.na(var_name)) * 100, 1), "\n")
      cat("Mean of the outliers:", round(mo, 2), "\n")
      cat("Mean with outliers:", round(m1, 2), "\n")
    }
  }
  
  # Combine all outlier information into one data frame
  if (length(outlier_results) > 0) {
    combined_outliers <- do.call(rbind, outlier_results)
  } else {
    combined_outliers <- data.frame()  # If no outliers were found, return an empty data frame
  }
  
  # Return combined data frame with outliers information
  return(combined_outliers)
}

# Call the function and store the results
outlier_info_combined_feeding_rate <- outlierKD_feeding_rate(afdw_feeding_rate, id_col = "treatment", start_col = 5, end_col = 5)

# Print the combined outlier information
print(outlier_info_combined_feeding_rate)
```

```{r}
# Assuming outlier_info_combined is the combined outlier data frame
outlier_info_combined_feeding_rate <- outlier_info_combined_feeding_rate  # If not already assigned

# Count the number of outliers for each replicate
outlier_counts_feeding_rate <- table(outlier_info_combined_feeding_rate$Replicate)

# Convert to a data frame for better readability
outlier_counts_feeding_rate_df <- as.data.frame(outlier_counts_feeding_rate)
colnames(outlier_counts_feeding_rate_df) <- c("Replicate", "Outlier_Count")

View(outlier_counts_feeding_rate_df)
```

Remove outliers:
```{r}
# Extract the row numbers of the outliers
outlier_rows_feeding_rate <- outlier_info_combined_feeding_rate$Row

# Remove the outlier rows from the dataset
afdw_feeding_rate_cleaned_1 <- afdw_feeding_rate[-outlier_rows_feeding_rate, ]
# 38 outliers identified and removed

# remove NAs
afdw_feeding_rate_cleaned_1 <- na.omit(afdw_feeding_rate_cleaned_1)
# 30 NAs removed
```

Histogram feeding rate again
```{r}
hist(afdw_feeding_rate_cleaned_1 $ feeding_rate) 
skewness(afdw_feeding_rate_cleaned_1 $ feeding_rate, na.rm = TRUE) 
shapiro.test(afdw_feeding_rate_cleaned_1 $ feeding_rate)
```
Feeding rate still not normal distributed, right skewed but skewness < 1.5, so no data transformation for model. 

Calculate mean feeding rate + se + 95 CI for each treatment
```{r}
mean_feeding_rate <- afdw_feeding_rate_cleaned_1 %>%
  group_by(treatment) %>%
  dplyr::summarize(
    mean_feeding = mean(feeding_rate, na.rm = TRUE),
    n = sum(!is.na(feeding_rate)),
    se = sd(feeding_rate, na.rm = TRUE) / sqrt(n),
    t_crit = qt(0.975, df = n - 1),
    ci_lower = mean_feeding - t_crit * se,
    ci_upper = mean_feeding + t_crit * se,
    .groups = "drop"
  ) %>%
  dplyr::select(treatment, mean_feeding, ci_lower, ci_upper, se)

print(mean_feeding_rate)
```

#### Model feeding rate
```{r}
# treatment and beaker as factor
afdw_feeding_rate_cleaned_1 $ treatment <- factor(afdw_feeding_rate_cleaned_1 $ treatment)
afdw_feeding_rate_cleaned_1 $ beaker <- factor(afdw_feeding_rate_cleaned_1 $ beaker)

# response variable: feeding_rate
# fixed effects: treatment
# random effect with constant intercept: (1|beaker)
feeding_rate_lmer <- lmer(feeding_rate ~ treatment + (1|beaker), data = afdw_feeding_rate_cleaned_1)
par(mfrow = c(2, 2))
plot(feeding_rate_lmer) 
qreference(residuals(feeding_rate_lmer)) # rather S shape

pairwise.wilcox.test(afdw_feeding_rate_cleaned_1$feeding_rate,
                     afdw_feeding_rate_cleaned_1$treatment,
                     p.adjust.method = "holm")
# trend NC:Mix p = 0.059
```

Mix vs. Control trend p = 0.059   

Ciprofloxacin x propyzamide interaction
```{r}
afdw_feeding_rate_cleaned_1$Cip <- ifelse(afdw_feeding_rate_cleaned_1$treatment %in% c("Cip", "Mix"),
                                1, 0)

afdw_feeding_rate_cleaned_1$Cip <- factor(afdw_feeding_rate_cleaned_1$Cip, levels = c(0, 1))

afdw_feeding_rate_cleaned_1$Pro <- ifelse(afdw_feeding_rate_cleaned_1$treatment %in% c("Pro", "Mix"),
                                1, 0)

afdw_feeding_rate_cleaned_1$Pro <- factor(afdw_feeding_rate_cleaned_1$Pro, levels = c(0, 1))

# sanity check
table(afdw_feeding_rate_cleaned_1$treatment,
      afdw_feeding_rate_cleaned_1$Cip,
      afdw_feeding_rate_cleaned_1$Pro)

# interaction model
model_feeding_rate_interact <- lm(feeding_rate ~ Cip * Pro, data = afdw_feeding_rate_cleaned_1)
par(mfrow = c(2, 2))
plot(model_feeding_rate_interact) # not normal distributed, S shape
qreference(residuals(model_feeding_rate_interact))

kruskal_feeding_rate_interaction <- afdw_feeding_rate_cleaned_1 %>%
  dplyr::mutate(treatment = interaction(Cip, Pro, sep = "_")) %>%
  kruskal.test(feeding_rate ~ treatment, data = .)

kruskal_feeding_rate_interaction
```


#### Boxplot feeding rate
```{r}
# Plot
feeding_rate_boxplot <-
  ggplot(data = afdw_feeding_rate_cleaned_1,
         mapping = aes(x = factor(
           treatment, levels = c("NC", "Pro", "Cip", "Mix")
         ), y = feeding_rate)) +
  # Mean
  stat_summary(
    fun = mean,
    geom = "point",
    aes(fill = treatment),
    shape = 21,
    size = 5,
    color = "grey20",
    stroke = 0.5
  ) +
  
  # 95% CI
  # Black outline (thicker, underneath)
  stat_summary(
    fun.data = mean_cl_normal,
    geom = "errorbar",
    aes(group = treatment),
    width = 0.2,
    linewidth = 1,
    color = "grey30"
  ) +
  
  # Colored CI on top
  stat_summary(
    fun.data = mean_cl_normal,
    geom = "errorbar",
    aes(color = treatment),
    width = 0.2,
    linewidth = 0.8
  ) +
  
  # Jittered
geom_point(
  aes(color = treatment),
  position = position_jitter(width = 0.25, height = 0),
  alpha = 0.8,
  size = 2
) +
  
  # Color scales
  scale_fill_manual(values = c(
    "NC" = "#009E73",
    "Pro" = "#E69F00",
    "Cip" = "#56B4E9",
    "Mix" = "#999999"
  )) +
  scale_color_manual(values = c(
    "NC" = "#009E73",
    "Pro" = "#E69F00",
    "Cip" = "#56B4E9",
    "Mix" = "#999999"
  )) +
  guides(fill = "none") +
  
  # Theme and labels
  labs(x = "Treatment", y = "Grazer Feeding Rate (µg / day)", color = "") +
  scale_x_discrete(labels = c("Control", "Propyzamide", "Ciprofloxacin", "Mixture")) +
  theme_minimal(base_size = 11) +
  theme(
    legend.position = "none",
    panel.background = element_rect(fill = "grey100"),
    axis.title = element_text(
      size = 13,
      face = "bold",
      colour = "grey30"
    ),
    axis.text.x = element_text(size = 12),  
    axis.text.y = element_text(size = 12),
    panel.border = element_rect(fill = NA, colour = "grey30"),
    panel.grid.major = element_line(colour = "grey85", size = 0.4),
    panel.grid.minor = element_line(colour = "grey95", size = 0.2)
  ) 

feeding_rate_boxplot
```

```{r, warning = FALSE}
# save the plot
ggsave(
  "output/feeding_rate.png",
  width = 16.5,
  height = 12,
  units = "cm",
  dpi = 600
)
```

# Fatty acid composition snails

Statistics for fatty acid composition of snails in absolute values (µg fatty acid / mg dry weight snail (soft body)).

```{r}
# insert data
FA_Snails_absolute <-
  read.table(file.path("data\\FA_Snails_absolute.txt"),
             header = TRUE, sep = "\t")

View(FA_Snails_absolute)
# Mix S_29 no data, fatty acid extraction did not work
```

## Absolute SAFA, MUFA & PUFA concentrations

1\) Total SAFA content

```{r, warning=FALSE}
# total SAFA, MUFA, PUFA content; sa = "snail absolute" 
# total SAFA content 
# filter saturated fatty acids in new table 
SAFA_sa <- FA_Snails_absolute %>%
filter(fa %in% c("C08:0", "C10:0", "C11:0", "C12:0", "C13:0", "C14:0", "C15:0", "C16:0", "C18:0", "C20:0", "C21:0", "C22:0", "C24:0"))

# Sum SAFA per replicate and treatment
total_SAFA_per_snail <- SAFA_sa %>% 
  dplyr::group_by(replicate, treatment) %>% 
  dplyr::summarize(sum_SAFA = sum(fa_conc), .groups = "drop") 
print(total_SAFA_per_snail) 

# Statistical differences
hist(total_SAFA_per_snail $ sum_SAFA) # normal distributed
shapiro.test(total_SAFA_per_snail $ sum_SAFA) # normal distributed
leveneTest(sum_SAFA ~ treatment, data = total_SAFA_per_snail) # homogenous
anova_SAFA <- aov(sum_SAFA ~ treatment, data = total_SAFA_per_snail)
summary(anova_SAFA)
TukeyHSD(anova_SAFA)

# Calculate mean, se and 95% CI per treatment
mean_SAFA_sa <- total_SAFA_per_snail %>%
  dplyr::group_by(treatment) %>%
  dplyr::summarize(
    mean_SAFA = mean(sum_SAFA, na.rm = TRUE),
    n = sum(!is.na(sum_SAFA)),
    se = sd(sum_SAFA, na.rm = TRUE) / sqrt(n),
    t_crit = qt(0.975, df = n - 1),
    ci_lower = mean_SAFA - t_crit * se,
    ci_upper = mean_SAFA + t_crit * se,
    .groups = "drop"
  ) %>%
  dplyr::select(treatment, mean_SAFA, ci_lower, ci_upper, se)

# Extract mean of control group for percental difference
control_mean_SAFA_sa <- mean_SAFA_sa %>%
  filter(treatment == "NC") %>%
  pull(mean_SAFA)

# Calculate percental difference from control
mean_SAFA_sa <- mean_SAFA_sa %>%
  mutate(percent_diff_from_NC = 100 * (mean_SAFA - control_mean_SAFA_sa) / control_mean_SAFA_sa)

print(mean_SAFA_sa)
```

2\) Total MUFA content

```{r, warning=FALSE}
# filter mono-unsaturated fatty acids in new table
MUFA_sa <- FA_Snails_absolute %>%
  filter(fa %in% c("C14:1", "C15:1", "C16:1", "C17:1", "C18:1", "C20:1", "C22:1", "C24:1"))

# Sum MUFA per replicate and treatment
total_MUFA_per_snail <- MUFA_sa %>% 
  dplyr::group_by(replicate, treatment) %>% 
  dplyr::summarize(sum_MUFA = sum(fa_conc), .groups = "drop") 
print(total_MUFA_per_snail) 

# Statistical differences
hist(total_MUFA_per_snail $ sum_MUFA) 
shapiro.test(total_MUFA_per_snail $ sum_MUFA) 
leveneTest(sum_MUFA ~ treatment, data = total_MUFA_per_snail)
pairwise.wilcox.test(total_MUFA_per_snail$sum_MUFA,
                     total_MUFA_per_snail$treatment,
                     p.adjust.method = "BH")

# Calculate mean, se and 95% CI per treatment
mean_MUFA_sa <- total_MUFA_per_snail %>%
  dplyr::group_by(treatment) %>%
  dplyr::summarize(
    mean_MUFA = mean(sum_MUFA, na.rm = TRUE),
    n = sum(!is.na(sum_MUFA)),
    se = sd(sum_MUFA, na.rm = TRUE) / sqrt(n),
    t_crit = qt(0.975, df = n - 1),
    ci_lower = mean_MUFA - t_crit * se,
    ci_upper = mean_MUFA + t_crit * se,
    .groups = "drop"
  ) %>%
  dplyr::select(treatment, mean_MUFA, ci_lower, ci_upper, se)

# Extract mean of control group for percental difference
control_mean_MUFA_sa <- mean_MUFA_sa %>%
  filter(treatment == "NC") %>%
  pull(mean_MUFA)

# Calculate percental difference from control
mean_MUFA_sa <- mean_MUFA_sa %>%
  mutate(percent_diff_from_NC = 100 * (mean_MUFA - control_mean_MUFA_sa) / control_mean_MUFA_sa)

print(mean_MUFA_sa)
```

3\) Total PUFA content

```{r, warning=FALSE}
# filter poly-unsaturated fatty acids in new table
PUFA_sa <- FA_Snails_absolute %>%
  filter(fa %in% c("C18:2(n-6c)", "C18:2(n-6t)", "C18:3(n-3)", "C18:3(n-6)", "C18:4(n-3)", "C20:2(n-6)", "C20:3(n-6)", "C20:3(n-3)", "C20:4(n-6)", "C20:5(n-3)", "C22:2", "C22:5(n-3)", "C22:6(n-3)"))
  
# Sum PUFA per replicate and treatment
total_PUFA_per_snail <- PUFA_sa %>% 
  dplyr::group_by(replicate, treatment) %>% 
  dplyr::summarize(sum_PUFA = sum(fa_conc), .groups = "drop") 
print(total_PUFA_per_snail) 

# statistics
hist(total_PUFA_per_snail $ sum_PUFA)
shapiro.test(total_PUFA_per_snail $ sum_PUFA) 
leveneTest(sum_PUFA ~ treatment, data = total_PUFA_per_snail)
anova_PUFA <- aov(sum_PUFA ~ treatment, data = total_PUFA_per_snail)
summary(anova_PUFA)
TukeyHSD(anova_PUFA)

# Calculate mean, se and 95% CI per treatment
mean_PUFA_sa <- total_PUFA_per_snail %>%
  dplyr::group_by(treatment) %>%
  dplyr::summarize(
    mean_PUFA = mean(sum_PUFA, na.rm = TRUE),
    n = sum(!is.na(sum_PUFA)),
    se = sd(sum_PUFA, na.rm = TRUE) / sqrt(n),
    t_crit = qt(0.975, df = n - 1),
    ci_lower = mean_PUFA - t_crit * se,
    ci_upper = mean_PUFA + t_crit * se,
    .groups = "drop"
  ) %>%
  dplyr::select(treatment, mean_PUFA, ci_lower, ci_upper, se)

# Extract mean of control group for percental difference
control_mean_PUFA_sa <- mean_PUFA_sa %>%
  filter(treatment == "NC") %>%
  pull(mean_PUFA)

# Calculate percental difference from control
mean_PUFA_sa <- mean_PUFA_sa %>%
  mutate(percent_diff_from_NC = 100 * (mean_PUFA - control_mean_PUFA_sa) / control_mean_PUFA_sa)

print(mean_PUFA_sa)
```

SAFA, MUFA & PUFA relative differences to control plot:
```{r, warning=FALSE}
snail_class_abs <- bind_rows(
  prep_class_from_replicates(total_SAFA_per_snail, "sum_SAFA", "SAFA"),
  prep_class_from_replicates(total_MUFA_per_snail, "sum_MUFA", "MUFA"),
  prep_class_from_replicates(total_PUFA_per_snail, "sum_PUFA", "PUFA")
) %>%
  mutate(
    Class = factor(Class, levels = c("SAFA","PUFA","MUFA")),
    treatment = factor(treatment, levels = c("NC","Pro","Cip","Mix"))
  )

snail_class_abs

# relative values
snail_class_rel <- snail_class_abs %>%
  group_by(Class) %>%
  mutate(
    ctrl_mean = mean[treatment == "NC"][1],
    mean = mean / ctrl_mean,
    lwr  = lwr  / ctrl_mean,
    upr  = upr  / ctrl_mean
  ) %>%
  ungroup() %>%
  dplyr::select(-ctrl_mean)


plot_df_snails <- snail_class_rel  

plot_df_snails <- plot_df_snails %>%
  mutate(
    show = ifelse(treatment == "NC", 0, 1),  
    treatment = factor(treatment, levels = c("NC","Pro","Cip","Mix")),
    Class = factor(Class, levels = c("PUFA", "MUFA", "SAFA")),
    lwr = ifelse(lwr < 0, 0, lwr)
  )

# plot
fa_snails_relative_changes <- ggplot(plot_df_snails, aes(x = mean, y = Class)) +
  geom_vline(xintercept = 1,
             linewidth = 1,
             colour = "grey30") +
  
  # Black CI outline
  geom_errorbarh(
    aes(xmin = lwr, xmax = upr, group = treatment, alpha = show),
    height = 0,
    linewidth = 1,
    colour = "grey30",
    position = position_dodge(width = 0.45)
  ) +
  
  # Colored CI overlay
  geom_errorbarh(
    aes(xmin = lwr, xmax = upr, colour = treatment, alpha = show),
    height = 0,
    linewidth = 0.8,
    position = position_dodge(width = 0.45)
  ) +
  
  geom_point(
    aes(colour = treatment, alpha = show),
    size = 4,
    position = position_dodge(width = 0.45)
  ) +
  
  scale_colour_manual(
    values = c(
    "NC" = "#009E73",
    "Pro" = "#E69F00",
    "Cip" = "#56B4E9",
    "Mix" = "#999999"
    ),
    labels = c(
      "NC"  = "Control",
      "Pro" = "Propyzamide",
      "Cip" = "Ciprofloxacin",
      "Mix" = "Mixture"
    ),
    name = "Treatment"
  ) +
  scale_alpha_identity() +
  guides(
    colour = guide_legend(override.aes = list(alpha = 1))
  ) +

  labs(
    x = "Fold change relative to control",
    y = NULL
  ) +

  theme_minimal(base_size = 11) +
  theme(
    legend.position = "none",
    legend.title = element_text(size = 12, face = "bold", colour = "grey30"),
    legend.text  = element_text(size = 11, colour = "grey30"),
    panel.background = element_rect(fill = "grey100"),
    axis.title = element_text(size = 13, face = "bold", colour = "grey30"),
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(size = 13, face = "bold", colour = "grey30"),
    panel.border = element_rect(fill = NA, colour = "grey30"),
    panel.grid.major = element_line(colour = "grey95", size = 0.4),
    panel.grid.minor = element_line(colour = "grey95", size = 0.2)
  )

fa_snails_relative_changes

# save the plot
ggsave(
  "output/fa_snails_relative_changes.png",
  plot = fa_snails_relative_changes,
  width = 16.5,
  height = 12,
  units = "cm",
  dpi = 600
)
```

4\) Total FA per snail
```{r}
# total FA 
total_FA_snails <- bind_rows(total_SAFA_per_snail, total_MUFA_per_snail, total_PUFA_per_snail)

total_FA_per_snail <- total_FA_snails %>%
  dplyr::group_by(replicate, treatment) %>%
  dplyr::summarize(sum_FA = sum(sum_SAFA, sum_MUFA, sum_PUFA, na.rm = TRUE))
print(total_FA_per_snail)

# Calculate mean, se and 95% CI per treatment
mean_FA_sa <- total_FA_per_snail %>%
  dplyr::group_by(treatment) %>%
  dplyr::summarize(
    mean_FA = mean(sum_FA, na.rm = TRUE),
    n = sum(!is.na(sum_FA)),
    se = sd(sum_FA, na.rm = TRUE) / sqrt(n),
    t_crit = qt(0.975, df = n - 1),
    ci_lower = mean_FA - t_crit * se,
    ci_upper = mean_FA + t_crit * se,
    .groups = "drop"
  ) %>%
  dplyr::select(treatment, mean_FA, ci_lower, ci_upper, se)

# Extract mean of control group for percental difference
control_mean_FA_sa <- mean_FA_sa %>%
  filter(treatment == "NC") %>%
  pull(mean_FA)

# Calculate percental difference from control
mean_FA_sa <- mean_FA_sa %>%
  mutate(percent_diff_from_NC = 100 * (mean_FA - control_mean_FA_sa) / control_mean_FA_sa)

print(mean_FA_sa)

# linear model with total FA per snail
# treatment as factor
total_FA_per_snail$treatment <- as.factor(total_FA_per_snail$treatment)

model_FA_snails <- lm(sum_FA ~ treatment, data = total_FA_per_snail)
par(mfrow = c(2, 2))
plot(model_FA_snails) # normality acceptable
qreference(residuals(model_FA_snails)) 
drop1(model_FA_snails, test = 'F')
# treatment highly significant 

# conduct Tukey HSD-test for pairwise comparisons
FA_snails_tukey <- lsmeans(model_FA_snails, ~ treatment)
FA_snails_tukey

FA_snails_tukey1 <- contrast(FA_snails_tukey, 'pairwise')
FA_snails_tukey

FA_snails_summary <- summary(FA_snails_tukey1, infer = TRUE, adjust = 'mvt')
FA_snails_summary
# Cip vs. Mix p = 0.0005
# Cip vs. Pro p = 0.0095
# Mix vs. Control p = 0.001
# Control vs. Pro p = 0.02
```

Look into interaction of Cip and Pro on total FA of snails:
```{r}
# interaction effect model with presence / absence of single xenobiotics
total_FA_per_snail$Cip <- ifelse(total_FA_per_snail$treatment %in% c("Cip", "Mix"),
                                1, 0)

total_FA_per_snail$Cip <- factor(total_FA_per_snail$Cip, levels = c(0, 1))

total_FA_per_snail$Pro <- ifelse(total_FA_per_snail$treatment %in% c("Pro", "Mix"),
                                1, 0)

total_FA_per_snail$Pro <- factor(total_FA_per_snail$Pro, levels = c(0, 1))

# sanity check
table(total_FA_per_snail$treatment,
      total_FA_per_snail$Cip,
      total_FA_per_snail$Pro)

# interaction effect model
model_mix_FA_snails <- lm(sum_FA ~ Cip * Pro, data = total_FA_per_snail)
par(mfrow = c(2, 2))
plot(model_mix_FA_snails) 
qreference(residuals(model_mix_FA_snails)) # normality acceptable
drop1(model_mix_FA_snails, test = 'F')
# Cip:Pro interaction not significant

# look deeper into interaction:
# df.ref only control values
df.ref.FA_snails <- total_FA_per_snail |>
  dplyr::filter(Cip == 0, Pro == 0) 

# remove grouping
df.ref.FA_snails <- df.ref.FA_snails %>%
  dplyr::ungroup()

nd_FA_snails <- nd.generator(df.ref.FA_snails, stressor1 = "Cip", stressor2 = "Pro", diff = NULL)

#sanity check  
table(nd_FA_snails$dataset, nd_FA_snails$Cip)
table(nd_FA_snails$dataset, nd_FA_snails$Pro)

#check for AD, MU (all values >0) and OD
hyp_all_FA_snails <- predictions(
  model_mix_FA_snails,
  newdata = nd_FA_snails,
  hypothesis = function(x) {
    hypothesis.function(
      x,
      stressor1 = "Cip",
      stressor2 = "Pro",
      null = "all",        # <- AD, MU, DO together
      strategy = "average"
    )
  }
) |>
  inferences(method = "simulation")

hyp_all_FA_snails

hyp_all_FA_snails$term <- dplyr::recode(hyp_all_FA_snails$term,
                                    "ad" = "AD",
                                    "do" = "DO",
                                    "mu" = "MU")

ggplot(hyp_all_FA_snails, aes(x = term, y = estimate)) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.2) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  labs(x = "Null model", y = "Interaction estimate")

```
Post-estimation inference following Pimentel et al. (2025) indicated that the additive, dominance, and multiplicative null models could not be rejected, as the confidence intervals of the interaction estimates overlapped zero. This suggests that the combined effect of ciprofloxacin and propyzamide did not deviate significantly from expectations based on independent single-stressor effects.

## Histogram FA snails

```{r}
hist(FA_Snails_absolute $ fa_conc) 
skewness(FA_Snails_absolute $ fa_conc, na.rm = TRUE) 
shapiro.test(FA_Snails_absolute $ fa_conc)
```

Fatty acid concentration not normal distributed and right skewed.

## NMDS FA snails

Change table to wide format:

```{r}
# change table to wide format (sqrt for square root transformation in next step)
FA_Snails_absolute_wide_sqrt <-
  pivot_wider(FA_Snails_absolute,
              names_from = fa,
              values_from = fa_conc)

# square root transformation, to reduce discriminatory power of dominant FA groups
FA_Snails_absolute_wide_sqrt[, 3:36] <- sqrt(FA_Snails_absolute_wide_sqrt[, 3:36])

# Only keep columns/FAs with more than 5 non-zero values (fatty acids that have just been found in 5 replicates will be removed)
FA_Snails_absolute_wide_sqrt <-
  FA_Snails_absolute_wide_sqrt %>%
  dplyr::select(
    1:2,
    where(~sum(. !=0, na.rm = TRUE)>5)
  )
```

Perform metaMDS:

```{r}
# Remove first and second column for metaMDS
FA_Snails_absolute_wide_sqrt <- FA_Snails_absolute_wide_sqrt[, -c(1, 2)]

# Rename (- C, - () ) fatty acids for later plot
FA_Snails_absolute_wide_sqrt <- FA_Snails_absolute_wide_sqrt %>%
  rename(
    `12:0`     = `C12:0`,   
    `13:0`     = `C13:0`,
    `14:0`     = `C14:0`,
    `15:0`     = `C15:0`,
    `16:0`     = `C16:0`,
    `16:1`     = `C16:1`,
    `18:0`     = `C18:0`,
    `18:1`     = `C18:1`,
    `18:2n-6c` = `C18:2(n-6c)`,
    `18:3n-6`  = `C18:3(n-6)`,
    `18:3n-3`  = `C18:3(n-3)`,
    `18:4n-3`  = `C18:4(n-3)`,
    `20:0`     = `C20:0`,
    `20:1`     = `C20:1`,
    `20:2n-6`  = `C20:2(n-6)`,
    `20:3n-6`  = `C20:3(n-6)`,
    `20:4n-6`  = `C20:4(n-6)`,
    `20:5n-3`  = `C20:5(n-3)`,
    `22:1`     = `C22:1`,
    `22:5n-3`  = `C22:5(n-3)`,
    `22:6n-3`  = `C22:6(n-3)`
)


NMDS_Snails_absolute <-
  metaMDS(
    FA_Snails_absolute_wide_sqrt,
    autotransform = FALSE,
    k = 2 ,
    distance = "bray",
    na.rm = TRUE
  )

NMDS_Snails_absolute$stress
```

Create fa vectors:

```{r, warning=FALSE}
# separation in SAFA; MUFA & PUFA to build vectors in NMDS
# "sa" = snail absolute
FA_variables.SAFAs_sa <-
  envfit(
    NMDS_Snails_absolute,
    subset(
      FA_Snails_absolute_wide_sqrt,
      select = c(
        "12:0",
        "13:0",
        "14:0",
        "15:0",
        "16:0",
        "18:0",
        "20:0"
      )
    )
    ,
    permutations = 999,
    na.rm = TRUE
  )

FA_variables.MUFAs_sa <-
  envfit(
    NMDS_Snails_absolute,
    subset(
      FA_Snails_absolute_wide_sqrt,
      select = c(
        "16:1",
        "18:1",
        "20:1",
        "22:1"
      )
    )
    ,
    permutations = 999,
    na.rm = TRUE
  )

FA_variables.PUFAs_sa <-
  envfit(
    NMDS_Snails_absolute,
    subset(
      FA_Snails_absolute_wide_sqrt,
      select = c(
        "18:2n-6c",
        "18:3n-6",
        "18:3n-3",
        "18:4n-3",
        "20:2n-6",
        "20:3n-6",
        "20:4n-6",
        "20:5n-3",
        "22:5n-3",
        "22:6n-3"
      )
    )
    ,
    permutations = 999,
    na.rm = TRUE
  )

# Extract NMDS scores 
data.scores_snail_absolute = as.data.frame(scores(NMDS_Snails_absolute)$sites)

Treatments <-
  FA_Snails_absolute $ treatment[FA_Snails_absolute $ fa == "C16:0"]

# add treatment column to NMDS scores
data.scores_snail_absolute$Treatment = Treatments

# vectors for SAFA, MUFA and PUFA
SAFAs_coord_cont_sa = as.data.frame(scores(FA_variables.SAFAs_sa, "vectors")) * ordiArrowMul(FA_variables.SAFAs_sa)
MUFAs_coord_cont_sa = as.data.frame(scores(FA_variables.MUFAs_sa, "vectors")) * ordiArrowMul(FA_variables.MUFAs_sa)
PUFAs_coord_cont_sa = as.data.frame(scores(FA_variables.PUFAs_sa, "vectors")) * ordiArrowMul(FA_variables.PUFAs_sa)
```

Plot NMDS:

```{r, warning=FALSE}
# For the plot: reorder treatment variable in data.scores_snail_absolute
data.scores_snail_absolute$Treatment <- factor(
  data.scores_snail_absolute$Treatment,
  levels = c("NC", "Pro", "Cip", "Mix"),
  labels  = c("Control", "Propyzamide", "Ciprofloxacin", "Mixture")
)

# Compute convex hull for each Treatment
polygon_data_snails <- data.scores_snail_absolute %>%
  group_by(Treatment) %>%
  summarise(hull_points = list(data.frame(
    NMDS1 = NMDS1[chull(NMDS1, NMDS2)],
    NMDS2 = NMDS2[chull(NMDS1, NMDS2)]
  ))) %>%
  unnest(hull_points)

# Function to rescale coordinates
rescale_vectors <- function(df, scale_factor = 1) {
  df$length <- sqrt(df$NMDS1^2 + df$NMDS2^2)
  max_length <- max(df$length)
  df$NMDS1 <- df$NMDS1 / max_length * scale_factor
  df$NMDS2 <- df$NMDS2 / max_length * scale_factor
  df$length <- NULL
  return(df)
}

# Apply to each group 
PUFAs_coord_cont_sa <- rescale_vectors(PUFAs_coord_cont_sa, 0.4)
MUFAs_coord_cont_sa <- rescale_vectors(MUFAs_coord_cont_sa, 0.4)
SAFAs_coord_cont_sa <- rescale_vectors(SAFAs_coord_cont_sa, 0.4)

# plot
absol_snail_with_vectors <- ggplot(data = data.scores_snail_absolute, aes(x = NMDS1, y = NMDS2)) +
    geom_polygon(
    data = polygon_data_snails,
    aes(x = NMDS1, y = NMDS2, group = Treatment, fill = Treatment),
    alpha = 0.2,
    show.legend = FALSE
    ) +
  geom_point(
    aes(shape = Treatment, color = Treatment),
    size = 5,
    alpha = 1,
    show.legend = TRUE
  ) +
  scale_shape_manual(values = c("Control" = 16, "Propyzamide" = 17, "Ciprofloxacin" = 18, "Mixture" = 15)) +
  scale_color_manual(values = c("Control" = "#009E73", "Propyzamide" = "#E69F00", "Ciprofloxacin" = "#56B4E9", "Mixture" = "#999999")) +
  scale_fill_manual(values = c("Control" = "#009E73", "Propyzamide" = "#E69F00", "Ciprofloxacin" = "#56B4E9", "Mixture" = "#999999")) + 
  geom_segment(aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2), 
               data = PUFAs_coord_cont_sa, linewidth = 0.5, alpha = 0.2, colour = "black") +
  geom_segment(aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2), 
               data = MUFAs_coord_cont_sa, linewidth = 0.5, alpha = 0.2, colour = "black") +
  geom_segment(aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2), 
               data = SAFAs_coord_cont_sa, linewidth = 0.5, alpha = 0.2, colour = "black") +
  geom_text_repel(
    data = PUFAs_coord_cont_sa, 
    aes(x = NMDS1, y = NMDS2, label = row.names(PUFAs_coord_cont_sa)), 
    colour = "black",
    size = 2.5) +
  geom_text_repel(
    data = MUFAs_coord_cont_sa, 
    aes(x = NMDS1, y = NMDS2, label = row.names(MUFAs_coord_cont_sa)), 
    colour = "#996600",
    size = 2.5) +
  geom_text_repel(
    data = SAFAs_coord_cont_sa, 
    aes(x = NMDS1, y = NMDS2, label = row.names(SAFAs_coord_cont_sa)), 
    colour = "#0000FF",
    size = 2.5) +
  theme_bw() +
  theme(
    axis.title = element_text(
      size = 13,
      face = "bold",
      colour = "grey30"
    ),
    panel.background = element_blank(),
    panel.border = element_rect(fill = NA, colour = "grey30"),
    legend.key = element_blank(),
    legend.title = element_text(
      size = 11,
      face = "bold",
      colour = "grey30"
    ),
    legend.text = element_text(size = 10, colour = "grey30"),
    panel.grid.major = element_line(colour = "grey95", size = 0.4),
    panel.grid.minor = element_line(colour = "grey95", size = 0.2)
  ) 

absol_snail_with_vectors
# position of FA names were manually adjusted for publication in vector graphic in the software Inkscape
```

Save the plot as .png:
```{r, warning=FALSE}
ggsave(
  "output/NMDS_snails_absolute_vectors.png",
  width = 16.5,
  height = 12,
  units = "cm",
  dpi = 600
)
```

Save the plot as vector graphic (minor adaptions for better readability with Inkscape):
```{r, warning=FALSE}
ggsave(
  "output/NMDS_snails_absolute_vectors.svg",
  device = "svg",
  width = 16.5,
  height = 16,
  units = "cm",
  dpi = 600
)
```

Now plot NMDS without fatty acid vectors:
```{r}
# plot
absol_snail <- ggplot(data = data.scores_snail_absolute, aes(x = NMDS1, y = NMDS2)) +
    geom_polygon(
    data = polygon_data_snails,
    aes(x = NMDS1, y = NMDS2, group = Treatment, fill = Treatment),
    alpha = 0.2,
    show.legend = FALSE
  ) +
  geom_point(
    aes(shape = Treatment, color = Treatment),
    size = 5,
    alpha = 1,
    show.legend = TRUE
  ) +
  scale_shape_manual(values = c("Control" = 16, "Propyzamide" = 17, "Ciprofloxacin" = 18, "Mixture" = 15)) +
  scale_color_manual(values = c("Control" = "#009E73", "Propyzamide" = "#E69F00", "Ciprofloxacin" = "#56B4E9", "Mixture" = "#999999")) +
  scale_color_manual(values = c("Control" = "#009E73", "Propyzamide" = "#E69F00", "Ciprofloxacin" = "#56B4E9", "Mixture" = "#999999")) +  
  scale_fill_manual(values = c("Control" = "#009E73", "Propyzamide" = "#E69F00", "Ciprofloxacin" = "#56B4E9", "Mixture" = "#999999")) + 
  theme_bw() +
  theme(
    axis.title = element_text(
      size = 11,
      face = "bold",
      colour = "grey30"
    ),
    panel.background = element_blank(),
    panel.border = element_rect(fill = NA, colour = "grey30"),
    legend.key = element_blank(),
    legend.title = element_text(
      size = 11,
      face = "bold",
      colour = "grey30"
    ),
    legend.text = element_text(size = 10, colour = "grey30"),
    panel.grid.major = element_line(colour = "grey95", size = 0.4),
    panel.grid.minor = element_line(colour = "grey95", size = 0.2)
  ) 

absol_snail
```

Save the plot as .png:
```{r}
ggsave(
  "output/NMDS_snails_absolute.png",
  width = 16.5,
  height = 12,
  units = "cm",
  dpi = 600
)
```

Save the plot as vector graphic:
```{r}
ggsave(
  "output/NMDS_snails_absolute.svg",
  device = "svg",
  width = 16.5,
  height = 16,
  units = "cm",
  dpi = 600
)
```

## PERMANOVA FA snails

Create distance matrix with fatty acids:

```{r}
FA_Snails_dist <- vegdist(FA_Snails_absolute_wide_sqrt, method = 'bray')
```

Separate explanatory variables from fatty acid information:
```{r}
# change initial table to wide format again
FA_Snails_absolute_wide_permanova <-
  pivot_wider(FA_Snails_absolute,
              names_from = fa,
              values_from = fa_conc)

# square root transformation again, to reduce discriminatory power of dominant fa groups
FA_Snails_absolute_wide_permanova[, 3:28] <- sqrt(FA_Snails_absolute_wide_permanova[, 3:28])

# Only keep columns/FAs with more than 5 non-zero values (fatty acids that have just been found in 5 replicates will be removed)
FA_Snails_absolute_wide_permanova <-
  FA_Snails_absolute_wide_permanova %>%
  dplyr::select(
    1:2,
    where(~sum(. !=0, na.rm = TRUE)>5)
  )

# delete fa columns
FA_Snails_absolute_wide_permanova <- FA_Snails_absolute_wide_permanova[, c(1, 2)]
FA_Snails_absolute_wide_permanova $ treatment <- as.factor(FA_Snails_absolute_wide_permanova $ treatment)
```

Perform PERMANOVA:

```{r}
perm_FA_Snails <- adonis2(FA_Snails_dist ~ treatment, data = FA_Snails_absolute_wide_permanova, permutations = 999)
print(perm_FA_Snails)
```

Pairwise comparisons (by: Martinez Arbizu, P. (2020). pairwiseAdonis: Pairwise multilevel comparison using adonis. R package version 0.4):

```{r}
pairwise_results <- pairwise.adonis2(
  FA_Snails_dist ~ treatment,
  data = FA_Snails_absolute_wide_permanova,
  permutations = 999
)

# View the results
print(pairwise_results)

# Cip*Pro interaction?
# add presence / absence columns
FA_Snails_absolute_wide_permanova$Cip <- ifelse(FA_Snails_absolute_wide_permanova$treatment %in% c("Cip", "Mix"),
                                                 1,
                                                 0)

FA_Snails_absolute_wide_permanova$Cip <- 
  factor(FA_Snails_absolute_wide_permanova$Cip, levels = c(0, 1))

FA_Snails_absolute_wide_permanova$Pro <- ifelse(FA_Snails_absolute_wide_permanova$treatment %in% c("Pro", "Mix"),
                                                 1,
                                                 0)

FA_Snails_absolute_wide_permanova$Pro <- 
  factor(FA_Snails_absolute_wide_permanova$Pro, levels = c(0, 1))

# sanity check
table(FA_Snails_absolute_wide_permanova$treatment,
      FA_Snails_absolute_wide_permanova$Cip)

table(FA_Snails_absolute_wide_permanova$treatment,
      FA_Snails_absolute_wide_permanova$Pro)

# permanova
perm_FA_snails_Cip_Pro <- adonis2(FA_Snails_dist ~ Cip * Pro,
        data = FA_Snails_absolute_wide_permanova,
        permutations = 999)

print(perm_FA_snails_Cip_Pro)
# no interaction
```

All treatments except of Pro_vs_Mix significant differences. No Cip * Pro interaction.

## SIMPER FA snails

```{r}
# Perform SIMPER
simper_FA_Snails <-
  with(
    FA_Snails_absolute_wide_permanova,
    simper(FA_Snails_absolute_wide_sqrt, treatment,  permutations = 999)
  )

summary(simper_FA_Snails)

str(simper_FA_Snails)

# save tsv tables of FAs which contribute > 70% to differences
simper_sum_snails <- summary(simper_FA_Snails)

out_dir <- "output"
dir.create(out_dir, showWarnings = FALSE)

# Safe filename helper
safe_name <- function(x) gsub("[^A-Za-z0-9_]+", "_", x)

iwalk(simper_sum_snails, function(x, nm) {
  if (is.null(x)) return(invisible(NULL))

  # SIMPER summary as a data.frame:
  # columns: average, sd, ratio, ava, avb, cumsum, p
  # rownames are the fatty acid IDs
  df <- as.data.frame(x)

  # keep rows up to the first that reaches/exceeds 70%
  idx <- which(df$cumsum >= 0.70)[1]
  if (is.na(idx)) idx <- nrow(df)
  out <- df[seq_len(idx), , drop = FALSE]

  # just keep 3 decimels
  out_write <- out
  num_cols <- vapply(out_write, is.numeric, logical(1))
  out_write[, num_cols] <- lapply(out_write[, num_cols, drop = FALSE],
                                  function(col) formatC(col, format = "f", digits = 3))

  file <- file.path(out_dir, paste0(safe_name(nm), "_top70_snails.tsv"))
  # write row names so the TSV mirrors the SIMPER structure
  write.table(out_write, file, sep = "\t", quote = FALSE, col.names = NA)
})

# Combine results into a data frame
simper_results_snails <- do.call(rbind, lapply(simper_FA_Snails, function(x) {
  if (!is.null(x$p)) {
    data.frame(
      fatty_acid = names(x$p),  
      p = as.numeric(x$p),      
      average = as.numeric(x$average),  
      sd = as.numeric(x$sd),
      ratio = as.numeric(x$ratio),
      cusum = as.numeric(x$cusum),
      stringsAsFactors = FALSE
    )
  } else {
    NULL  
  }
}))

# Show only significant results
simper_results_snails %>%
  dplyr::filter(p <= 0.05) %>%
  dplyr::select(fatty_acid, p, average, sd, ratio, cusum)  

# How often do they occur
simper_results_snails %>%
  dplyr::filter(p <= 0.05) %>%
  dplyr::count(fatty_acid) %>%
  dplyr::arrange(desc(n)) %>%
  print()
```

# Temperature and light conditions

## Biofilm colonisation and contamination

```{r, warning=FALSE}
# insert data
colonisation_temp_light <-
  read.table(file.path("data\\colonisation_temp_light.txt"),
             header = TRUE, sep = "\t")

View(colonisation_temp_light)

# average temperature
mean(colonisation_temp_light $ temperature_colonisation)
# 18.28751 °C

# mean light intensity
# filter only daytime
average_light_intensity_colonisation <- colonisation_temp_light %>%
  dplyr::filter(light_intensity_colonisation > 0) %>%
  dplyr::summarize(average_light_intensity_colonisation = mean(light_intensity_colonisation))

print(average_light_intensity_colonisation)
# 14019.26 lux

# temperature plot
# add a small amount of random noise to the y-values
colonisation_temp_light $ temperature_colonisation <- jitter(colonisation_temp_light $ temperature_colonisation)

t2 <- ggplot(
  data = colonisation_temp_light,
  mapping = aes( x = day, y = temperature_colonisation, color = time_of_day)) +
  geom_point(alpha = 0.5) +  # Individual measurements with transparency
  geom_smooth() +
  labs(
    x = "Day", y = "Temperature (°C)"
  ) +
  theme_bw() +
  scale_x_continuous(breaks = seq(0, 77, by = 5)) +
  scale_color_continuous(
    name = "Time of Day",  
    breaks = c(0, 5, 10, 15, 20),
    labels = c("08:00 pm", "03:00 pm", "10:00 am", "05:00 am", "0:00 am"),
    limits = c(0, 22),
  )

t2 

# save the plot
ggsave(
  "output/temperature_colonisation.png",
  width = 16.5,
  height = 12,
  units = "cm",
  dpi = 600
)

# light intensity plot
# Filter the data to include only rows where light_intensity > 0
filtered_data2 <- colonisation_temp_light %>% 
  filter(light_intensity_colonisation > 0)

# add a small amount of random noise to the y-values
filtered_data2 $ light_intensity_colonisation <- jitter(filtered_data2 $ light_intensity_colonisation)

l2 <- ggplot(
  data = filtered_data2,
  mapping = aes( x = day, y = light_intensity_colonisation)) +
  geom_point(size = 1) +  # Individual measurements with transparency
  geom_smooth(color = "#E69F00", linewidth = 1) +
  labs(
    x = "Day", y = "Light intensity (lux)"
  ) +
  theme_bw() +
  scale_x_continuous(breaks = seq(0, 77, by = 5)) 

l2 

# save the plot
ggsave(
  "output/light_intensity_colonisation.png",
  width = 16.5,
  height = 12,
  units = "cm",
  dpi = 600
)
```

## Feeding assay

```{r, warning = FALSE}
# insert data
feeding_assay_temp_light <-
  read.table(file.path("data\\feeding_assay_temp_light.txt"),
             header = TRUE, 
             sep = "\t"
  )
             
View(feeding_assay_temp_light)

# average temperature
mean(feeding_assay_temp_light $ temperature)
# 13.27927 °C

# mean light intensity
# filter only daytime
average_light_intensity_feeding <- feeding_assay_temp_light %>%
  dplyr::filter(light_intensity > 0) %>%
  dplyr::summarize(average_light_intensity_feeding = mean(light_intensity))

print(average_light_intensity_feeding)
# 330.3552 lux

# temperature plot
# add a small amount of random noise to the y-values
feeding_assay_temp_light $ temperature <- jitter(feeding_assay_temp_light $ temperature)

t1 <- ggplot(
  data = feeding_assay_temp_light,
  mapping = aes( x = day, y = temperature, color = time_of_day)) +
  geom_point(alpha = 0.5) +  # Individual measurements with transparency
  geom_smooth() +
  labs(
    x = "Day", y = "Temperature (°C)"
  ) +
  theme_bw() +
  scale_x_continuous(breaks = 0:21) +
  scale_color_continuous(
    name = "Time of Day",  
    breaks = c(0, 5, 10, 15, 20),
    labels = c("08:00 pm", "03:00 pm", "10:00 am", "05:00 am", "0:00 am"),
    limits = c(0, 22),
  )

t1 

# save the plot
ggsave(
  "output/temperature_feeding_assay.png",
  width = 16.5,
  height = 12,
  units = "cm",
  dpi = 600
)

# light intensity plot
# Filter the data to include only rows where light_intensity > 0
filtered_data1 <- feeding_assay_temp_light %>% 
  filter(light_intensity > 0)

# add a small amount of random noise to the y-values
filtered_data1 $ light_intensity <- jitter(filtered_data1 $ light_intensity)

l1 <- ggplot(
  data = filtered_data1,
  mapping = aes( x = day, y = light_intensity)) +
  geom_point(size = 1) +  # Individual measurements with transparency
  geom_smooth(color = "#E69F00", linewidth = 1) +
  labs(
    x = "Day", y = "Light intensity (lux)"
  ) +
  theme_bw() +
  scale_x_continuous(breaks = 0:21) 

l1 

# save the plot
ggsave(
  "output/light_intensity_feeding_assay.png",
  width = 16.5,
  height = 12,
  units = "cm",
  dpi = 600
)
```

# Chemical analyses
```{r, warning=FALSE}
# insert data
chemical_analyses <-
  read.table(file.path("data\\chemical_analyses.txt"),
             header = TRUE, sep = "\t")

View(chemical_analyses)

# LOQ to NA
chemical_analyses[chemical_analyses == "<LOQ"] <- NA
# concentration values as numeric
chemical_analyses$concentration_ug_L <- as.numeric(
  ifelse(chemical_analyses$concentration_ug_L == "<LOQ", NA, chemical_analyses$concentration_ug_L)
)

# calculate means, se, 95% CI
chemical_means <- chemical_analyses %>%
  filter(!is.na(concentration_ug_L)) %>% 
  dplyr::mutate(percent_deviation = 100 * (concentration_ug_L - nominal_conc) / nominal_conc) %>%
  dplyr::group_by(treatment, mixture) %>%
  dplyr::summarize(
    mean_concentration = mean(concentration_ug_L, na.rm = TRUE),
    n = sum(!is.na(concentration_ug_L)),
    se = sd(concentration_ug_L, na.rm = TRUE) / sqrt(n),
    t_crit = qt(0.975, df = n - 1),
    ci_lower = mean_concentration - t_crit * se,
    ci_upper = mean_concentration + t_crit * se,
    mean_percent_deviation = mean(percent_deviation, na.rm = TRUE),
    se_percent_deviation = sd(percent_deviation, na.rm = TRUE) / sqrt(sum(!is.na(percent_deviation))),
    .groups = "drop"
  ) %>%
  dplyr::select(treatment, mixture, mean_concentration, ci_lower, ci_upper, se, mean_percent_deviation, se_percent_deviation)

print(chemical_means)
```

