---
title: "Bottom-up effects of an herbicide and an antibiotic on *Potamopyrgus antipodarum*"
format: 
  html:
    self-contained: true
    toc: true
    style: "floating"
---

Sophie Oster, Verena C. Schreiner, Tobias Schmitt, Patrick Fink, Markus Weitere, Ibrahim Araoye, Ralf Schulz, Mirco Bundschuh

R script written by Sophie Oster\
checked by Verena C. Schreiner

\
RPTU Kaiserslautern-Landau\
Fortstrasse 7\
76829 Landau in der Pfalz\
GERMANY

Corresponding author email address:\
1) Mirco Bundschuh: mirco.bundschuh\@rptu.de

This study was conducted with four treatments and three time points explanatory variables depending on the endpoint:

-   treatment (categorical)
    -   uncontaminated control (negative control, NC)
    -   herbicide propyzamide (Pro)
    -   antibiotic ciprofloxacin (Cip)
    -   mixture of propyzamide and ciprofloxacin (Mix)
-   time.point (categorical)
    -   week 1
    -   week 2
    -   week 3

# 1 Packages in a reproducible environment

```{r, message = FALSE, warning = FALSE}
# renv::init() 
# renv::snapshot()

library(stats)
library(rstatix)
library(dplyr)
library(tidyverse)
library(tidyr)
library(broom)
library(dunn.test)
library(patchwork)
library(vegan)
library(data.table)
library(ggplot2)
library(ggrepel)
library(ggforce)
library(concaveman)
library(gridExtra)
library(stringr)
library(shades)
library(ggpubr)
library(base)
library(ggokabeito)
library(multcomp)
library(car)
library(asbio)
library(plotrix)
library(lmtest)
library(boot)
library(moments)
library(lme4)
library(DAAG)
library(performance)
library(emmeans)
library(purrr)
library(DT)
library(pairwiseAdonis)
```

# 2 Set working directory

```{r}
prj <- getwd()
setwd(prj)
```

# 3 Biofilm endpoints

## Chlorophyll and ash free dry weight of biofilm after colonisation + chronic contamination of 14 days

-   Unit chlorophyll a, b & c: µg chlorophyll / cm²

-   Unit afdw after colonisation: mg / cm²

```{r}
# insert data
biofilm_endpoints <-
  read.table(file.path(prj, "data\\biofilm_endpoints.txt"),
             header = TRUE, sep = "\t")

View(biofilm_endpoints)
```

## 3.1 Chlorophyll

#### Caclulate mean chla, chlb + chlc of each treatment, to compare in manuscript

```{r, message = FALSE, warning = FALSE}
#chl a
mean_chlorophyll_a <- biofilm_endpoints %>%
  group_by(week, treatment) %>%
  summarize(mean_chlorophyll_a = mean(chla))

print(mean_chlorophyll_a)

# chl b
mean_chlorophyll_b <- biofilm_endpoints %>%
  group_by(week, treatment) %>%
  summarize(mean_chlorophyll_b = mean(chlb))

print(mean_chlorophyll_b)

# chl c
mean_chlorophyll_c <- biofilm_endpoints %>%
  group_by(week, treatment) %>%
  summarize(mean_chlorophyll_c = mean(chlc))

print(mean_chlorophyll_c)
```

#### Histogram chlorophyll a, b + c

```{r}
# chl a
a <- hist(biofilm_endpoints $ chla) 
skewness(biofilm_endpoints $ chla, na.rm = TRUE) 

# chl b
b <- hist(biofilm_endpoints $ chlb) 
skewness(biofilm_endpoints $ chlb, na.rm = TRUE) 

# chl c
c <- hist(biofilm_endpoints $ chlc) 
skewness(biofilm_endpoints $ chlc, na.rm = TRUE) 
```

All not normal distributed and right skewed -\> log10 transformation in model

#### Model chla

Linear model with treatment and week and their interaction (week:treatment) as explanatory variables.

```{r}
# week and treatment as factor
biofilm_endpoints $ week <- factor(biofilm_endpoints $ week)
biofilm_endpoints $ treatment <- factor(biofilm_endpoints $ treatment)

model_chla <- lm(log10(chla) ~ week + treatment + week:treatment, data = biofilm_endpoints)

plot(model_chla) # point cloud looks evenly distributed --> normal distribution

qreference(residuals(model_chla)) # normal distribution assumed

drop1(model_chla, test = 'F')

```

Interaction week:treatment significant (p=0.02965). Now check differences between treatments within weeks.

```{r}
chla_inter <- lsmeans(model_chla, ~ treatment | week )
chla_inter_dunn1 <- contrast(chla_inter, 'pairwise')
(chla_inter_dunn_sum1 <- summary(chla_inter_dunn1, infer = TRUE, adjust = 'mvt'))
```

No significant differences of the treatments within week 1 and 3. Only NC-Mix significant in week 2 (p=0.003).

Now check for differences between weeks.

```{r}
anova(model_chla)
# week variable highly significant (p=7.689e-08)

chla_inter2 <- lsmeans(model_chla, ~ week |  treatment)
chla_inter_dunn2 <- contrast(chla_inter2, 'pairwise')
(chla_inter_dunn_sum2 <- summary(chla_inter_dunn2, infer = TRUE, adjust = 'mvt'))

```

-   NC\
    -   Week 1 & 2 similar\
    -   Week 2 & 3 similar\
    -   difference between week 1 & 3
-   Cip\
    -   Week 1 & 2 similar\
    -   Week 1 & 3 similar\
    -   difference between week 2 & 3
-   Pro\
    -   Week 1 & 2 similar\
    -   differences between week 1 & 3 and 2 & 3
-   Mix\
    -   all weeks different

Week 1 & 2 quite similar, expect of Mix treatment. Week 3 quite different to week 1 & 2 despite of Cip treatment.

#### Model chlb

```{r}
model_chlb <- lm(log10(chlb) ~ week + treatment + week:treatment, data = biofilm_endpoints)

plot(model_chlb) # point cloud looks evenly distributed --> normal distribution

qreference(residuals(model_chlb)) # normal distribution assumed

drop1(model_chlb, test = 'F')
```

No significant differences. Still check for differences between treatments.

```{r}
chlb_inter <- lsmeans(model_chlb, ~ treatment | week )
chlb_inter_dunn1 <- contrast(chlb_inter, 'pairwise')
(chlb_inter_dunn_sum1 <- summary(chlb_inter_dunn1, infer = TRUE, adjust = 'mvt'))
```

No differences between treatments for chlb.

Now check for differences between weeks.

```{r}
anova(model_chlb)
# week variable highly significant (p=5.259e-05)

chlb_inter2 <- lsmeans(model_chlb, ~ week |  treatment)
chlb_inter_dunn3 <- contrast(chlb_inter2, 'pairwise')
(chlb_inter_dunn_sum3 <- summary(chlb_inter_dunn3, infer = TRUE, adjust = 'mvt'))
```

Similar results/trends as for chla. Week 1 & 2 quite similar; differences in:

-   Pro treatment: week 2 & 3

-   Mix treatment: week 1 & 2; 2 & 3

#### Model chlc

```{r}
model_chlc <- lm(log10(chlc) ~ week + treatment + week:treatment, data = biofilm_endpoints)

plot(model_chlc) # point cloud looks evenly distributed --> normal distribution

qreference(residuals(model_chlc)) # normal distribution assumed

drop1(model_chlc, test = 'F')
```

No significant differences. Still check for differences between treatments.

```{r}
chlc_inter <- lsmeans(model_chlc, ~ treatment | week )
chlc_inter_dunn1 <- contrast(chlc_inter, 'pairwise')
(chlc_inter_dunn_sum1 <- summary(chlc_inter_dunn1, infer = TRUE, adjust = 'mvt'))
```

No differences between treatments for chlorophyll c.

#### Boxplots chlorophyll a, b + c

```{r}
# chlorophyll a
chla <- ggplot(data = biofilm_endpoints, 
               mapping = aes(x = factor(week), 
                             y = chla, 
                             fill = factor(treatment, levels = c("NC", "Cip", "Pro", "Mix")))) +
  geom_boxplot() +
  scale_fill_manual(values = c("NC" = "#009E73", "Cip" = "#56B4E9", "Pro" = "#E69F00", "Mix" = "#F0E442")) + 
  scale_color_okabe_ito() + 
  theme_bw() +
  labs(x = NULL , y = expression("Chlorophyll a (µg/cm"^"2"*")"), fill = "Treatment") +
  scale_x_discrete(labels = c("Week 1", "Week 2", "Week 3"))

chla 

# save the plot
ggsave(chla, 
       filename = "chla.png", 
       path = 'C:\\Users\\Oster Test\\OneDrive\\Desktop\\Promotion\\Bioassays\\Schneckenversuch 2023\\Bottom_up_effects_biofilm_grazer\\output')


# chlorophyll b
chlb <- ggplot(data = biofilm_endpoints, 
               mapping = aes(x = factor(week), 
                             y = chlb, 
                             fill = factor(treatment, levels = c("NC", "Cip", "Pro", "Mix")))) +
  geom_boxplot() +
  scale_fill_manual(values = c("NC" = "#009E73", "Cip" = "#56B4E9", "Pro" = "#E69F00", "Mix" = "#F0E442")) + 
  scale_color_okabe_ito() + 
  theme_bw() +
  labs(x = NULL , y = expression("Chlorophyll b (µg/cm"^"2"*")"), fill = "Treatment") +
  scale_x_discrete(labels = c("Week 1", "Week 2", "Week 3"))

chlb 

# save the plot
ggsave(chlb, 
       filename = "chlb.png", 
       path = 'C:\\Users\\Oster Test\\OneDrive\\Desktop\\Promotion\\Bioassays\\Schneckenversuch 2023\\Bottom_up_effects_biofilm_grazer\\output')


# chlorophyll c
chlc <- ggplot(data = biofilm_endpoints, 
               mapping = aes(x = factor(week), 
                             y = chlc, 
                             fill = factor(treatment, levels = c("NC", "Cip", "Pro", "Mix")))) +
  geom_boxplot() +
  scale_fill_manual(values = c("NC" = "#009E73", "Cip" = "#56B4E9", "Pro" = "#E69F00", "Mix" = "#F0E442")) + 
  scale_color_okabe_ito() + 
  theme_bw() +
  labs(x = NULL , y = expression("Chlorophyll c (µg/cm"^"2"*")"), fill = "Treatment") +
  scale_x_discrete(labels = c("Week 1", "Week 2", "Week 3"))

chlc 

# save the plot
ggsave(chlc, 
       filename = "chlc.png", 
       path = 'C:\\Users\\Oster Test\\OneDrive\\Desktop\\Promotion\\Bioassays\\Schneckenversuch 2023\\Bottom_up_effects_biofilm_grazer\\output')
```

## 3.2 AFDW after colonisation

#### Calculate mean AFDW of each treatment for manuscript.

```{r}
mean_afdw_after_colonisation <- biofilm_endpoints %>%
  group_by(week, treatment) %>%
  summarize(mean_afdw_after_colonisation = mean(afdw_after_colonisation))

print(mean_afdw_after_colonisation)
```

#### Histogram AFDW after colonisation

```{r}
hist(biofilm_endpoints $ afdw_after_colonisation) 
skewness(biofilm_endpoints $ afdw_after_colonisation, na.rm = TRUE) 
```

Not normal distributed and right skewed. Log10 transformation for model

#### Model AFDW after colonisation

Linear model with treatment and week and their interaction (week:treatment) as explanatory variables.

```{r}
model_afdw_after_colonisation <- lm(log10(afdw_after_colonisation) ~ week + treatment + week:treatment, data = biofilm_endpoints)

plot(model_afdw_after_colonisation) # not perfectly even distributed, but ok

qreference(residuals(model_afdw_after_colonisation)) # normal distribution assumed

drop1(model_afdw_after_colonisation, test = 'F')
```

Significant difference for week:treatment (p=0.009063). Check differences between treatments within weeks.

```{r}
afdw_after_colonisation_inter <- lsmeans(model_afdw_after_colonisation, ~ treatment | week )

afdw_after_colonisation_inter_dunn1 <- contrast(afdw_after_colonisation_inter, 'pairwise')

(afdw_after_colonisation_inter_dunn_sum1 <- summary(afdw_after_colonisation_inter_dunn1, infer = TRUE, adjust = 'mvt'))
```

Significant differences between week 2 Mix-NC and NC-Pro; and week 3 Cip-NC and Cip-Pro.

Now check for differences within weeks.

```{r}
anova(model_afdw_after_colonisation)
# week variable highly significant (p=2.541e-07)

afdw_after_colonisation_inter2 <- lsmeans(model_afdw_after_colonisation, ~ week |  treatment)

afdw_after_colonisation_inter_dunn2 <- contrast(afdw_after_colonisation_inter2, 'pairwise')

(afdw_after_colonisation_inter_dunn_sum2 <- summary(afdw_after_colonisation_inter_dunn2, infer = TRUE, adjust = 'mvt'))
```

-   Cip

    -   Difference between week 1 and 3

-   Mix

    -   Week 1 & 2 similar; week 1 & 3 and 2 & 3 different.

-   NC

    -   all weeks different from each other

-   Pro

    -   Week 1 & 2 similar; week 1 & 3 and 2 & 3 different.

#### Boxplot AFDW after colonisation

```{r}
afdw_after_colonisation_boxplot <- ggplot(data = biofilm_endpoints, 
               mapping = aes(x = factor(week), 
                             y = afdw_after_colonisation, 
                             fill = factor(treatment, levels = c("NC", "Cip", "Pro", "Mix")))) +
  geom_boxplot() +
  scale_fill_manual(values = c("NC" = "#009E73", "Cip" = "#56B4E9", "Pro" = "#E69F00", "Mix" = "#F0E442")) + 
  scale_color_okabe_ito() + 
  theme_bw() +
  labs(x = NULL , y = expression("Ash free dry weight (mg/cm"^"2"*")"), fill = "Treatment") +
  scale_x_discrete(labels = c("Week 1", "Week 2", "Week 3"))

afdw_after_colonisation_boxplot 

# @ VS: Hier auch log Skala, um die Daten besser zu sehen?

# save the plot
ggsave(afdw_after_colonisation_boxplot, 
       filename = "afdw_after_colonisation.png", 
       path = 'C:\\Users\\Oster Test\\OneDrive\\Desktop\\Promotion\\Bioassays\\Schneckenversuch 2023\\Bottom_up_effects_biofilm_grazer\\output')

```

# 4 Feeding rate snails and ash free dry weight after feeding

-   NA's due to missing snail and zero values of feeding rate due to no measurable feeding in respective week

-   Unit feeding rate: µg / day and unit ash free dry weight (afdw): mg / cm²

-   Column "used_for_FA_analysis" highlights the ten snails used for FA analyses for each treatment

```{r}
# insert data
afdw_feeding_rate <-
  read.table(file.path(prj, "data\\feeding_assay_afdw_feeding_rate.txt"),
             header = TRUE, sep = "\t")

View(afdw_feeding_rate)
```

## 4.1 Feeding rate snails

#### Calculate mean feeding rate for each week

```{r}
mean_feeding_rate <- afdw_feeding_rate %>%
  group_by(week, treatment) %>%
  summarize(mean_feeding_rate = mean(feeding_rate, na.rm = TRUE))

print(mean_feeding_rate)
```

#### Histogram feeding rate

```{r}
hist(afdw_feeding_rate $ feeding_rate) 
skewness(afdw_feeding_rate $ feeding_rate, na.rm = TRUE) 
```

Feeding rate variable not normal distributed and right skewed -\> log10 transformation for model

#### Model feeding rate

```{r}
# week, treatment and beaker as factor
afdw_feeding_rate $ week <- factor(afdw_feeding_rate $ week)
afdw_feeding_rate $ treatment <- factor(afdw_feeding_rate $ treatment)
afdw_feeding_rate $ beaker <- factor(afdw_feeding_rate $ beaker)

# log10 transformation; adding 0.001 to variable for samples with feeding rate = 0
afdw_feeding_rate <- afdw_feeding_rate %>%
  mutate(log_feeding_rate = log10(feeding_rate + 0.001))

# response variable: log_feeding_rate
# fixed effects: week + treatment + week:treatment
# random effect with constant intercept: (1|beaker)
# first remove NAs
afdw_feeding_rate <- na.omit(afdw_feeding_rate)

feeding_rate_lmer <- lmer(log_feeding_rate ~ week + treatment + week:treatment + (1|beaker), data = afdw_feeding_rate)
```

Warning "boundary (singular) fit" because random effects of beaker are very small.

```{r}
isSingular(feeding_rate_lmer)
# isSingular = TRUE

# look at fitted model
ranef(feeding_rate_lmer)
```

No systematic effect by "beaker", so simplify model:

```{r}
model_feeding_rate <- lm(log_feeding_rate ~ week + treatment + week:treatment, data = afdw_feeding_rate)

plot(model_feeding_rate) # point cloud looks evenly distributed --> normal distribution

qreference(residuals(model_feeding_rate)) # normality not perfect, but ok

drop1(model_feeding_rate, test = 'Chisq')
# interaction of timepoint (week) and treatment highly significant
# now check at which timepoints is a difference between treatments
anova(model_feeding_rate)

# conduct Tukey HSD-test for pairwise comparisons
feeding_rate_tukey <- lsmeans(model_feeding_rate, ~ treatment | week)
feeding_rate_tukey

feeding_rate_tukey1 <- contrast(feeding_rate_tukey, 'pairwise')
feeding_rate_tukey1

feeding_rate_summary <- summary(feeding_rate_tukey1, infer = TRUE, adjust = 'mvt')
feeding_rate_summary

```

Treatments within week 2 & 3 similar. Within week 2 differences between Cip-Pro and Mix-Pro. Now check differences between weeks.

```{r}
feeding_rate_inter1 <- lsmeans(model_feeding_rate, ~ week |  treatment)

feeding_rate_inter_dunn1 <- contrast(feeding_rate_inter1, 'pairwise')

(feeding_rate_inter_dunn_sum1 <- summary(feeding_rate_inter_dunn1, infer = TRUE, adjust = 'mvt'))
```

Weeks for all treatments similiar, only difference for treatment Pro week 1 & 2 and 1 & 3.

#### Boxplot feeding rate after 7/14/21 days

```{r}
feeding_rate_boxplot <- ggplot(data = afdw_feeding_rate, 
       mapping = aes(x = factor(week), 
                     y = log_feeding_rate, 
                     fill = factor(treatment, levels = c("NC", "Cip", "Pro", "Mix")))) +
  geom_boxplot() +
  scale_fill_manual(values = c("NC" = "#009E73", "Cip" = "#56B4E9", "Pro" = "#E69F00", "Mix" = "#F0E442")) + 
  scale_color_okabe_ito() + 
  theme_bw() +
  labs(x = NULL , fill = "Treatment") +
  scale_x_discrete(labels = c("Week 1", "Week 2", "Week 3")) +
  scale_y_continuous(trans = "log10") +
  labs(y = expression("Feeding rate (µg/day)" ~ log[10]))

feeding_rate_boxplot

# save the plot
ggsave(feeding_rate_boxplot, 
       filename = "feeding_rate_boxplot.png", 
       path = 'C:\\Users\\Oster Test\\OneDrive\\Desktop\\Promotion\\Bioassays\\Schneckenversuch 2023\\Bottom_up_effects_biofilm_grazer\\output')


```

## 4.2 AFDW after feeding

#### Calculate mean of afdw after feeding for each week

```{r}
mean_afdw_after_feeding <- afdw_feeding_rate %>%
  group_by(week, treatment) %>%
  summarize(mean_afdw_after_feeding = mean(afdw))

print(mean_afdw_after_feeding)
```

#### Histogram afdw after feeding

```{r}
hist(afdw_feeding_rate $ afdw) 
skewness(afdw_feeding_rate $ afdw, na.rm = TRUE) 
shapiro.test(afdw_feeding_rate $ afdw)
```

AFDW after feeding not normal distributed. Log10 transformation for model.

#### Model afdw after feeding

```{r}
# log10 transformation; adding 0.001 
afdw_feeding_rate <- afdw_feeding_rate %>%
  mutate(log_afdw = log10(afdw + 0.001))

# response variable: log_afdw
# fixed effects: week + treatment + week:treatment
# random effect with constant intercept: (1|beaker)
# remove NAs
afdw_feeding_rate <- na.omit(afdw_feeding_rate)

afdw_after_feeding_lmer <- lmer(log_afdw ~ week + treatment + week:treatment + (1|beaker), data = afdw_feeding_rate)

#----- new try without random effect
model_afdw2 <- lm(log_afdw ~ week + treatment + week:treatment, data = afdw_feeding_rate)

drop1(model_afdw2, test = "Chisq")

anova(model_afdw2)
#----------
plot(afdw_after_feeding_lmer) # point cloud looks evenly distributed --> normal distribution

qreference(residuals(afdw_after_feeding_lmer)) # normality

drop1(afdw_after_feeding_lmer, test = 'Chisq')
# interaction of timepoint (week) and treatment highly significant
# now check at which timepoints is a difference between treatments
anova(afdw_after_feeding_lmer)

# conduct Tukey HSD-test for pairwise comparisons
afdw_tukey <- lsmeans(afdw_after_feeding_lmer, ~ treatment | week)
afdw_tukey

afdw_tukey1 <- contrast(afdw_tukey, 'pairwise')
afdw_tukey

afdw_summary <- summary(afdw_tukey1, infer = TRUE, adjust = 'mvt')
afdw_summary
```

Here treatments quite different within weeks. Compare weeks:

```{r}
afdw_after_feeding_inter1 <- lsmeans(afdw_after_feeding_lmer, ~ week |  treatment)

afdw_after_feeding_dunn1 <- contrast(afdw_after_feeding_inter1, 'pairwise')

(afdw_after_feeding_dunn_sum1 <- summary(afdw_after_feeding_dunn1, infer = TRUE, adjust = 'mvt'))
```

Week 1 & 2 for treatments Mix & Pro similar. Everything else shows differences.

#### Boxplot afdw after feeding of 7/14/21 days

```{r}
afdw_after_feeding_boxplot <- ggplot(data = afdw_feeding_rate, 
       mapping = aes(x = factor(week), 
                     y = afdw, 
                     fill = factor(treatment, levels = c("NC", "Cip", "Pro", "Mix")))) +
  geom_boxplot() +
  scale_fill_manual(values = c("NC" = "#009E73", "Cip" = "#56B4E9", "Pro" = "#E69F00", "Mix" = "#F0E442")) + 
  scale_color_okabe_ito() + 
  theme_bw() +
  labs(x = NULL , fill = "Treatment") +
  scale_x_discrete(labels = c("Week 1", "Week 2", "Week 3")) +
  labs(y = expression("AFDW (mg/cm"^"2"*")"))

afdw_after_feeding_boxplot

# save the plot
ggsave(afdw_after_feeding_boxplot, 
       filename = "afdw_after_feeding_boxplot.png", 
       path = 'C:\\Users\\Oster Test\\OneDrive\\Desktop\\Promotion\\Bioassays\\Schneckenversuch 2023\\Bottom_up_effects_biofilm_grazer\\output')
```

# 5 Dryweight and growth snails

```{r}
# insert data
dryweight_snails <-
  read.table(file.path(prj, "data\\dryweight_snails.txt"),
             header = TRUE, sep = "\t")

View(dryweight_snails)
# 3 snails from the 20 baseline (culture) snails were lost during the analyses.
```

#### Calculate mean dryweight of all treatments and the culture snails

```{r}
mean_dryweight_snails <- dryweight_snails %>%
              group_by(treatment) %>%
              summarize(mean_dryweight_snails = mean(dryweight_mg))

print(mean_dryweight_snails)
```

Snails from feeding assay (fed with biofilm) heavier than snails from culture (fed with Tetraphyll flakes).

Calculate growth:

```{r}
growth_snails <- dryweight_snails %>%
  dplyr::mutate(growth_µg_dw_per_day = ((dryweight_mg - median(dryweight_mg[treatment == "Culture"])) / 7) * 1000) 
# multiply by 1000 to obtain a growth unit of µg dryweight per day

# calculate the growth in percentage in comparison to culture snails
percental_growth <- growth_snails %>%
  group_by(treatment) %>%
  summarize(mean_dryweight = mean(dryweight_mg)) %>%
  mutate(percental_growth = (mean_dryweight / mean_dryweight[treatment == "Culture"]) * 100 - 100)

percental_growth

# calculate effects size compared to control snails
effect_size_growth <- growth_snails %>%
  group_by(treatment) %>%
  summarize(mean_dryweight = mean(dryweight_mg)) %>%
  mutate(percental_growth = (mean_dryweight / mean_dryweight[treatment == "NC"]) * 100 - 100)

effect_size_growth

```

#### Histogram snail growth

```{r}
# filter culture snails
filtered_growth <- growth_snails %>%
  filter(!treatment %in% c("Culture"))

hist(filtered_growth $ growth_µg_dw_per_day) 
skewness(filtered_growth $ growth_µg_dw_per_day, na.rm = TRUE) 
shapiro.test(filtered_growth $ growth_µg_dw_per_day)
```

Growth normal distributed.

#### Model growth snails

```{r}
# explanatory variable treatment as factor
filtered_growth $ treatment <- factor(filtered_growth $ treatment)
  
  
model_growth_snails <- lm(growth_µg_dw_per_day ~ treatment, data = filtered_growth)

plot(model_growth_snails) # point cloud looks evenly distributed --> normal distribution

qreference(residuals(model_growth_snails)) # normal distribution

anova(model_growth_snails)
```

Explanatory variable "treatment" highly significant. Now check differences between treatments:

```{r}
growth_snails_inter <- lsmeans(model_growth_snails, ~ treatment)

growth_snails_inter_dunn1 <- contrast(growth_snails_inter, 'pairwise')

growth_snails_inter_dunn_sum1 <- summary(growth_snails_inter_dunn1, infer = TRUE, adjust = 'mvt')

growth_snails_inter_dunn1

```

-   Significant difference between growth of Cip and Mix snails compared to NC; almost significant difference (p=0.0564) between Pro treatment and NC

#### Boxplot growth snails

```{r}
# plot
growth_snails_boxplot <- ggplot(data = filtered_growth, 
                                mapping = aes(x = factor(treatment, levels = c("NC", "Cip", "Pro", "Mix")), 
                                              y = growth_µg_dw_per_day, 
                                              fill = factor(treatment, levels = c("NC", "Cip", "Pro", "Mix")))) +
  geom_boxplot() +
  scale_fill_manual(values = c("NC" = "#009E73", "Cip" = "#56B4E9", "Pro" = "#E69F00", "Mix" = "#F0E442"),
                    breaks = c("NC", "Cip", "Pro", "Mix")) +
  geom_point() +
  scale_color_okabe_ito() + 
  theme_bw() +
  labs(x = NULL , y = "Growth (µg dryweight / day)", fill = "Treatment") +
  ylim(-10, 90) +
  scale_x_discrete(labels = c("NC", "Cip", "Pro", "Mix")) +
  labs(title = expression(paste("Growth of ", italic("Potamopyrgus antipodarum"), " after 21-day feeding assay"))) +
  annotate("text", x = 1.5, y = max(filtered_growth$growth_µg_dw_per_day) + 5, 
           label = "p=0.022*", size = 3) +
  geom_segment(aes(x = 1, xend = 2, y = max(growth_µg_dw_per_day) + 2, yend = max(growth_µg_dw_per_day) + 2),
               lineend = "round", linewidth = 0.7) +
  annotate("text", x = 2.5, y = max(filtered_growth$growth_µg_dw_per_day) + 21, 
           label = "p=0.005**", size = 3) +
  geom_segment(aes(x = 1, xend = 4, y = max(growth_µg_dw_per_day) + 18, yend = max(growth_µg_dw_per_day) + 18),
               lineend = "round", linewidth = 0.7) +
    annotate("text", x = 2, y = max(filtered_growth$growth_µg_dw_per_day) + 13, 
           label = "p=0.056", size = 3) +
    geom_segment(aes(x = 1, xend = 3, y = max(growth_µg_dw_per_day) + 10, yend = max(growth_µg_dw_per_day) + 10),
               lineend = "round", linewidth = 0.7) 


growth_snails_boxplot

# save the plot
ggsave(growth_snails_boxplot, 
       filename = "growth_snails_boxplot.png", 
       path = 'C:\\Users\\Oster Test\\OneDrive\\Desktop\\Promotion\\Bioassays\\Schneckenversuch 2023\\Bottom_up_effects_biofilm_grazer\\output')
```

# 6 Fatty acid composition snails

Statistics for fatty acid composition of snails in absolute values (µg fatty acid / mg dry weight snail).

```{r}
# insert data
FA_Snails_absolute <-
  read.table(file.path(prj, "data\\FA_Snails_absolute.txt"),
             header = TRUE, sep = "\t")

View(FA_Snails_absolute)

any(is.na(FA_Snails_absolute))
```

Following fatty acids were not considered for FA content in snails due to too less values (less than 3 findings): C10:0, C11:0, C14:1, C15:1, C17:1, C20:0 and C22:0.

Therefore, delete rows with respective fatty acids:

```{r}
FA_Snails_absolute <- FA_Snails_absolute %>%
  filter(!fa %in% c("C10:0", "C11:0", "C14:1", "C15:1", "C17:1", "C20:0", "C22:0"))
```

#### Total SAFA, MUFA & PUFA content

1\) Total SAFA content

```{r}
# total SAFA, MUFA, PUFA content; sa = "snail absolute" 
# total SAFA content 
# filter saturated fatty acids in new table 
SAFA_sa <- FA_Snails_absolute %>%
filter(fa %in% c("C08:0", "C12:0", "C13:0", "C14:0", "C15:0", "C16:0", "C17:0", "C18:0", "C21:0", "C24:0")) 

# calculate the total SAFA content per snail 
total_SAFA_per_snail <- SAFA_sa %>% 
  group_by(replicate, treatment) %>% 
  summarize(sum_SAFA = sum(fa_conc)) 
print(total_SAFA_per_snail) 

# mean 
mean_SAFA_sa <- total_SAFA_per_snail %>% 
  group_by(treatment) %>% 
  summarize(mean_SAFA = mean(sum_SAFA, na.rm = TRUE)) 
print(mean_SAFA_sa)
```

2\) Total MUFA content

```{r}
# filter mono-unsaturated fatty acids in new table
MUFA_sa <- FA_Snails_absolute %>%
  filter(fa %in% c("C16:1", "C18:1", "C20:1", "C22:1", "C24:1"))

# total MUFA content per snail
total_MUFA_per_snail <- MUFA_sa %>%
  group_by(replicate, treatment) %>%
  summarize(sum_MUFA = sum(fa_conc))
print(total_MUFA_per_snail)

# mean
mean_MUFA_sa <- total_MUFA_per_snail %>%
  group_by(treatment) %>%
  summarize(mean_MUFA = mean(sum_MUFA, na.rm = TRUE))
print(mean_MUFA_sa)
```

3\) Total PUFA content

```{r}
# filter poly-unsaturated fatty acids in new table
PUFA_sa <- FA_Snails_absolute %>%
  filter(fa %in% c("C18:2(n-6c)", "C18:2(n-6t)", "C18:3(n-3)", "C18:3(n-6)", "C18:4(n-3)", "C20:2(n-6)", "C20:3(n-6)", "C20:4(n-6)", "C20:5(n-3)", "C22:2", "C22:5(n-3)", "C22:6(n-3)"))
  
# calculate total PUFA content per snail
total_PUFA_per_snail <- PUFA_sa %>%
  group_by(replicate, treatment) %>%
  summarize(sum_PUFA = sum(fa_conc))
print(total_PUFA_per_snail)

# mean
mean_PUFA_sa <- total_PUFA_per_snail %>%
  group_by(treatment) %>%
  summarize(mean_PUFA = mean(sum_PUFA, na.rm = TRUE))
print(mean_PUFA_sa)
```

4\) Total FA per snail

```{r}
# total FA 
total_FA_snails <- bind_rows(total_SAFA_per_snail, total_MUFA_per_snail, total_PUFA_per_snail)

total_FA_per_snail <- total_FA_snails %>%
  group_by(replicate, treatment) %>%
  summarize(sum_FA = sum(sum_SAFA, sum_MUFA, sum_PUFA, na.rm = TRUE))
print(total_FA_per_snail)

# mean
mean_total_FA_sa <- total_FA_per_snail %>%
  group_by(treatment) %>%
  summarize(mean_FA = mean(sum_FA, na.rm = TRUE))
print(mean_total_FA_sa)
```

#### Histogram FA snails

```{r}
hist(FA_Snails_absolute $ fa_conc) 
skewness(FA_Snails_absolute $ fa_conc, na.rm = TRUE) 
shapiro.test(FA_Snails_absolute $ fa_conc)
```

Fatty acid concentration not normal distributed and right skewed.

#### NMDS FA snails

Change table to wide format:

```{r}
# change table to wide format (sqrt for square root transformation in next step)
FA_Snails_absolute_wide_sqrt <-
  pivot_wider(FA_Snails_absolute,
              names_from = fa,
              values_from = fa_conc)

# square root transformation, to reduce discriminatory power of dominant fa groups
FA_Snails_absolute_wide_sqrt[, 3:28] <- sqrt(FA_Snails_absolute_wide_sqrt[, 3:28])

any(is.na(FA_Snails_absolute_wide_sqrt))
```
Check for outlier:

```{r}
#Run function one time, after that you can use the shortcut outlierKD (your subset, variable)

outlierKD_multiple <- function(dt, id_col = "replicate", start_col = 3, end_col = 28) {
  # Initialize an empty list to store results for each column
  outlier_results <- list()
  
  # Loop over each column from 3 to 28 (or specify your range)
  for (i in start_col:end_col) {
    col_name <- colnames(dt)[i]  # Get the column name
    
    # Extract the column values
    var_name <- dt[[col_name]]
    
    # Check if the column is numeric
    if (is.numeric(var_name)) {
      na1 <- sum(is.na(var_name))
      m1 <- mean(var_name, na.rm = TRUE)
      
      # Identify outliers using boxplot stats
      outlier <- boxplot.stats(var_name)$out  
      
      # Find rows with outliers
      outlier_rows <- which(var_name %in% outlier)
      
      # If no outliers, skip the rest of the loop
      if (length(outlier_rows) == 0) next
      
      # Get replicate values for the outliers
      id_values <- dt[[id_col]][outlier_rows]
      
      # Mean of outliers
      mo <- mean(outlier)
      
      # Collect the outlier information for this column
      outlier_results[[col_name]] <- data.frame(
        Column = col_name,
        Row = outlier_rows,
        Replicate = id_values,
        Outlier = outlier
      )
      
      cat("\nOutliers for column:", col_name, "\n")
      cat("Outliers identified:", length(outlier), "\n")
      cat("Proportion (%) of outliers:", 
          round(length(outlier) / sum(!is.na(var_name)) * 100, 1), "\n")
      cat("Mean of the outliers:", round(mo, 2), "\n")
      cat("Mean with outliers:", round(m1, 2), "\n")
    }
  }
  
  # Combine all outlier information into one data frame
  if (length(outlier_results) > 0) {
    combined_outliers <- do.call(rbind, outlier_results)
  } else {
    combined_outliers <- data.frame()  # If no outliers were found, return an empty data frame
  }
  
  # Return combined data frame with outliers information
  return(combined_outliers)
}



# Call the function and store the results
outlier_info_combined <- outlierKD_multiple(FA_Snails_absolute_wide_sqrt, id_col = "replicate", start_col = 3, end_col = 28)


print(outlier_info_combined)



outlier_info_combined <- outlier_info_combined  

# Count the number of outliers for each replicate
outlier_counts <- table(outlier_info_combined$Replicate)

# Convert to a data frame for better readability
outlier_counts_df <- as.data.frame(outlier_counts)
colnames(outlier_counts_df) <- c("Replicate", "Outlier_Count")

# Sort the results by the number of outliers (descending)
outlier_counts_df_sorted <- outlier_counts_df[order(-outlier_counts_df$Outlier_Count), ]


print(outlier_counts_df_sorted)

```

As the replicates S_1, S_13, S_25 and S_30 show more than 10 outliers within 26 fatty acids concentrations, these replicates will be removed.

```{r}
FA_Snails_absolute_wide_sqrt <- FA_Snails_absolute_wide_sqrt %>%
  filter(!(replicate %in% c("S_1", "S_13", "S_25", "S_30")))
```

Perform metaMDS:

```{r}
# Remove first and second column for metaMDS
FA_Snails_absolute_wide_sqrt <- FA_Snails_absolute_wide_sqrt[, -c(1, 2)]

# Rename (- C, - () ) fatty acids for later plot
FA_Snails_absolute_wide_sqrt <- FA_Snails_absolute_wide_sqrt %>%
  rename(
    `8:0`      = `C08:0`,   
    `12:0`     = `C12:0`,   
    `13:0`     = `C13:0`,
    `14:0`     = `C14:0`,
    `15:0`     = `C15:0`,
    `16:0`     = `C16:0`,
    `16:1`     = `C16:1`,
    `18:0`     = `C18:0`,
    `18:1`     = `C18:1`,
    `18:2n-6t` = `C18:2(n-6t)`,
    `18:2n-6c` = `C18:2(n-6c)`,
    `18:3n-6`  = `C18:3(n-6)`,
    `18:3n-3`  = `C18:3(n-3)`,
    `18:4n-3`  = `C18:4(n-3)`,
    `20:1`     = `C20:1`,
    `20:2n-6`  = `C20:2(n-6)`,
    `20:3n-6`  = `C20:3(n-6)`,
    `20:4n-6`  = `C20:4(n-6)`,
    `21:0`     = `C21:0`,
    `20:5n-3`  = `C20:5(n-3)`,
    `22:1`     = `C22:1`,
    `22:2`     = `C22:2`,
    `22:5n-3`  = `C22:5(n-3)`,
    `22:6n-3`  = `C22:6(n-3)`,
    `24:0`     = `C24:0`,
    `24:1`     = `C24:1`
  )



NMDS_Snails_absolute <-
  metaMDS(
    FA_Snails_absolute_wide_sqrt,
    autotransform = FALSE,
    k = 2 ,
    distance = "bray",
    na.rm = TRUE
  )

NMDS_Snails_absolute$stress
```

Create fa vectors

```{r}
# separation in SAFA; MUFA & PUFA to build vectors in NMDS
# "sa" = snail absolute
FA_variables.SAFAs_sa <-
  envfit(
    NMDS_Snails_absolute,
    subset(
      FA_Snails_absolute_wide_sqrt,
      select = c(
        "8:0",
        "12:0",
        "13:0",
        "14:0",
        "15:0",
        "16:0",
        "18:0",
        "21:0",
        "24:0"
      )
    )
    ,
    permutations = 999,
    na.rm = TRUE
  )

FA_variables.MUFAs_sa <-
  envfit(
    NMDS_Snails_absolute,
    subset(
      FA_Snails_absolute_wide_sqrt,
      select = c(
        "16:1",
        "18:1",
        "20:1",
        "22:1",
        "24:1"
      )
    )
    ,
    permutations = 999,
    na.rm = TRUE
  )

FA_variables.PUFAs_sa <-
  envfit(
    NMDS_Snails_absolute,
    subset(
      FA_Snails_absolute_wide_sqrt,
      select = c(
        "18:2n-6t",
        "18:2n-6c",
        "18:3n-6",
        "18:3n-3",
        "18:4n-3",
        "20:2n-6",
        "20:3n-6",
        "20:4n-6",
        "20:5n-3",
        "22:2",
        "22:5n-3",
        "22:6n-3"
      )
    )
    ,
    permutations = 999,
    na.rm = TRUE
  )

# Extract NMDS scores 
data.scores_snail_absolute = as.data.frame(scores(NMDS_Snails_absolute)$sites)

# Remove S_1, S_13, S_25 and S_30 from first data frame "FA_Snails_absolute"
FA_Snails_absolute <- FA_Snails_absolute %>%
  filter(!(replicate %in% c("S_1", "S_13", "S_25", "S_30")))

Treatments <-
  FA_Snails_absolute $ treatment[FA_Snails_absolute $ fa == "C16:0"]

# add treatment column to NMDS scores
data.scores_snail_absolute$Treatment = Treatments

# vectors for SAFA, MUFA and PUFA
SAFAs_coord_cont_sa = as.data.frame(scores(FA_variables.SAFAs_sa, "vectors")) * ordiArrowMul(FA_variables.SAFAs_sa)
MUFAs_coord_cont_sa = as.data.frame(scores(FA_variables.MUFAs_sa, "vectors")) * ordiArrowMul(FA_variables.MUFAs_sa)
PUFAs_coord_cont_sa = as.data.frame(scores(FA_variables.PUFAs_sa, "vectors")) * ordiArrowMul(FA_variables.PUFAs_sa)
```

Plot NMDS:

```{r}
# plot
absol_Snail <- ggplot(data = data.scores_snail_absolute, aes(x = NMDS1, y = NMDS2)) +
  geom_point(
    aes(shape = Treatment, color = Treatment),
    size = 5,
    alpha = 1,
    show.legend = TRUE
  ) +
  scale_shape_manual(values = c("NC" = 16, "Pro" = 17, "Cip" = 18, "Mix" = 15)) +
  scale_color_manual(values = c("NC" = "#009E73", "Pro" = "#E69F00", "Cip" = "#56B4E9", "Mix" = "#F0E442")) +
  stat_conf_ellipse(
    aes(colour = Treatment, fill = Treatment),
    alpha = 0.3,
    geom = "polygon",
    show.legend = FALSE
  ) +
  scale_color_manual(values = c("NC" = "#009E73", "Pro" = "#E69F00", "Cip" = "#56B4E9", "Mix" = "#F0E442")) +  
  scale_fill_manual(values = c("NC" = "#009E73", "Pro" = "#E69F00", "Cip" = "#56B4E9", "Mix" = "#F0E442")) + 
  geom_segment(aes(x = 0, y = 0, xend = NMDS1 * 0.2, yend = NMDS2 * 0.2), 
               data = PUFAs_coord_cont_sa, linewidth = 0.5, alpha = 0.2, colour = "black") +
    geom_segment(aes(x = 0, y = 0, xend = NMDS1 * 0.2, yend = NMDS2 * 0.2), 
               data = MUFAs_coord_cont_sa, linewidth = 0.5, alpha = 0.2, colour = "black") +
      geom_segment(aes(x = 0, y = 0, xend = NMDS1 * 0.1, yend = NMDS2 * 0.1), 
               data = SAFAs_coord_cont_sa, linewidth = 0.5, alpha = 0.2, colour = "black") +
  geom_text_repel(
    data = PUFAs_coord_cont_sa, 
    aes(x = NMDS1 * 0.2, y = NMDS2 * 0.2, label = row.names(PUFAs_coord_cont_sa)), 
    colour = "black",
    size = 2.5) +
  geom_text_repel(
    data = MUFAs_coord_cont_sa, 
    aes(x = NMDS1 * 0.2, y = NMDS2 * 0.2, label = row.names(MUFAs_coord_cont_sa)), 
    colour = "#996600",
    size = 2.5) +
  geom_text_repel(
    data = SAFAs_coord_cont_sa, 
    aes(x = NMDS1 * 0.1, y = NMDS2 * 0.1, label = row.names(SAFAs_coord_cont_sa)), 
    colour = "#0000FF",
    size = 2.5) +
  theme_bw() +
  theme(
    axis.title = element_text(
      size = 11,
      face = "bold",
      colour = "grey30"
    ),
    panel.background = element_blank(),
    panel.border = element_rect(fill = NA, colour = "grey30"),
    legend.key = element_blank(),
    legend.title = element_text(
      size = 11,
      face = "bold",
      colour = "grey30"
    ),
    legend.text = element_text(size = 10, colour = "grey30"),
    #plot.margin = margin(
     # t = 0.2,
    #  b = 0.2,
    #  l = 0.2,
    #  r = 0.2,
    #  unit = "cm"
  #  ),
    panel.grid.major = element_line(colour = "grey95", size = 0.4),
    panel.grid.minor = element_line(colour = "grey95", size = 0.2)
  ) +
  annotate(
    "text",
    x = 0.35,
    y = -0.19,
    label = paste("stress:", NMDS_Snails_absolute$stress %>% round(2)),
    size = 3.5
  ) 


absol_Snail
# position of FA names manually adjusted in vector graphic
```

Save the plot as .png:

```{r}
ggsave(absol_Snail, 
       filename = "NMDS_snails_absolute.png", 
       path = 'C:\\Users\\Oster Test\\OneDrive\\Desktop\\Promotion\\Bioassays\\Schneckenversuch 2023\\Bottom_up_effects_biofilm_grazer\\output')
```

Save the plot as vector graphic:

```{r}
ggsave(plot = absol_Snail,
       "NMDS_snails_absolute.svg", 
       device = "svg",
       path = 'C:\\Users\\Oster Test\\OneDrive\\Desktop\\Promotion\\Bioassays\\Schneckenversuch 2023\\Bottom_up_effects_biofilm_grazer\\output')
```

#### PERMANOVA FA snails

Create distance matrix with fatty acids:

```{r}
FA_Snails_dist <- vegdist(FA_Snails_absolute_wide_sqrt, method = 'bray')
```

Separate explanatory variables from fatty acid information:

```{r}
# change initial table to wide format again
FA_Snails_absolute_wide_permanova <-
  pivot_wider(FA_Snails_absolute,
              names_from = fa,
              values_from = fa_conc)

# square root transformation again, to reduce discriminatory power of dominant fa groups
FA_Snails_absolute_wide_permanova[, 3:28] <- sqrt(FA_Snails_absolute_wide_permanova[, 3:28])

# delete fa columns
FA_Snails_absolute_wide_permanova <- FA_Snails_absolute_wide_permanova[, c(1, 2)]
FA_Snails_absolute_wide_permanova $ treatment <- as.factor(FA_Snails_absolute_wide_permanova $ treatment)
```

Perform PERMANOVA:

```{r}
perm_FA_Snails <- adonis2(FA_Snails_dist ~ treatment, data = FA_Snails_absolute_wide_permanova, permutations = 999)
print(perm_FA_Snails)
```

Pairwise comparisons (by: Martinez Arbizu, P. (2020). pairwiseAdonis: Pairwise multilevel comparison using adonis. R package version 0.4):

```{r}
pairwise_results <- pairwise.adonis2(
  FA_Snails_dist ~ treatment,
  data = FA_Snails_absolute_wide_permanova,
  permutations = 999
)

# View the results
print(pairwise_results)

```

All treatments except of Pro_vs_Mix significant differences.

#### SIMPER FA snails

```{r}
# Upload extra table with treatments
snail_treatments_table <-
  read.table(file.path(prj, "data\\snail_treatments_table.txt"),
             header = TRUE, sep = "\t")

View(snail_treatments_table)

# Perform SIMPER
simper_FA_Snails <- with(snail_treatments_table, simper(FA_Snails_absolute_wide_sqrt, treatment,  permutations = 999))

summary(simper_FA_Snails)


str(simper_FA_Snails)

# Combine results into a data frame
simper_results_snails <- do.call(rbind, lapply(simper_FA_Snails, function(x) {
  # Check if the element contains valid data
  if (!is.null(x$p)) {
    data.frame(
      fatty_acid = names(x$p),  
      p = as.numeric(x$p),      
      average = as.numeric(x$average),  
      sd = as.numeric(x$sd),
      ratio = as.numeric(x$ratio),
      cusum = as.numeric(x$cusum),
      stringsAsFactors = FALSE
    )
  } else {
    NULL  # Skip if no p-values
  }
}))


# Show only significant results
simper_results_snails %>%
  dplyr::filter(p <= 0.05) %>%
  dplyr::select(fatty_acid, p, average, sd, ratio, cusum)  

# How often do they occur
simper_results_snails %>%
  dplyr::filter(p <= 0.05) %>%
  dplyr::count(fatty_acid) %>%
  dplyr::arrange(desc(n)) %>%
  print()

```

# 7 Fatty acid composition biofilm

Statistics for fatty acid composition of the biofilm in absolute values (µg fatty acid / mg dry weight biofilm).

```{r}
# insert data
FA_Biofilm_absolute <-
  read.table(file.path(prj, "data\\FA_Biofilm_absolute.txt"),
             header = TRUE, sep = "\t")

View(FA_Biofilm_absolute)

any(is.na(FA_Biofilm_absolute))
```

Following fatty acids were not considered for biofilm analysis due to too less values: C10:0, C11:0, C14:1.

Therefore, delete rows with respective fatty acids:

```{r}
FA_Biofilm_absolute <- FA_Biofilm_absolute %>%
  filter(!fa %in% c("C10:0", "C11:0", "C14:1"))
```



#### Total SAFA, MUFA & PUFA content

1\) Total SAFA content

```{r}
# total SAFA, MUFA, PUFA content; ba = "biofilm absolute"
# total SAFA content
# filter saturated fatty acids in new table
SAFA_ba <- FA_Biofilm_absolute %>%
    filter(fa %in% c("C08:0", "C10:0", "C12:0", "C13:0", "C14:0", "C15:0", "C16:0", "C17:0", "C18:0", "C20:0", "C22:0", "C24:0"))

# calculate the total SAFA content per replicate
total_SAFA_per_replicate <- SAFA_ba %>%
  group_by(replicate, treatment) %>%
  summarize(sum_SAFA = sum(fa_conc))
print(total_SAFA_per_replicate)

# mean
mean_SAFA_ba <- total_SAFA_per_replicate %>%
  group_by(treatment) %>%
  summarize(mean_SAFA = mean(sum_SAFA, na.rm = TRUE))
print(mean_SAFA_ba)
```

2\) Total MUFA content

```{r}
# filter mono-unsaturated fatty acids in new table
MUFA_ba <- FA_Biofilm_absolute %>%
  filter(fa %in% c("C16:1", "C17:1", "C18:1", "C20:1"))

# total MUFA content per replicate
total_MUFA_per_replicate <- MUFA_ba %>%
  group_by(replicate, treatment) %>%
  summarize(sum_MUFA = sum(fa_conc))
print(total_MUFA_per_replicate)

# mean
mean_MUFA_ba <- total_MUFA_per_replicate %>%
  group_by(treatment) %>%
  summarize(mean_MUFA = mean(sum_MUFA, na.rm = TRUE))
print(mean_MUFA_ba)
```

3\) Total PUFA content

```{r}
# filter poly-unsaturated fatty acids in new table
PUFA_ba <- FA_Biofilm_absolute %>%
  filter(fa %in% c("C18:2(n-6c)", "C18:2(n-6t)", "C18:3(n-3)", "C18:3(n-6)", "C18:4(n-3)", "C20:2(n-6)", "C20:3(n-6)", "C20:3(n-3)", "C20:4(n-6)", "C20:5(n-3)", "C22:2", "C22:5(n-3)", "C22:6(n-3)"))
  
# calculate total PUFA content per replicate
total_PUFA_per_replicate <- PUFA_ba %>%
  group_by(replicate, treatment) %>%
  summarize(sum_PUFA = sum(fa_conc))
print(total_PUFA_per_replicate)

# mean
mean_PUFA_ba <- total_PUFA_per_replicate %>%
  group_by(treatment) %>%
  summarize(mean_PUFA = mean(sum_PUFA, na.rm = TRUE))
print(mean_PUFA_ba)
```

4\) Total FA content

```{r}
# total FA 
total_FA_biofilm <- bind_rows(total_SAFA_per_replicate, total_MUFA_per_replicate, total_PUFA_per_replicate)

total_FA_per_replicate <- total_FA_biofilm %>%
  group_by(replicate, treatment) %>%
  summarize(sum_FA = sum(sum_SAFA, sum_MUFA, sum_PUFA, na.rm = TRUE))
print(total_FA_per_replicate)

# mean
mean_total_FA_ba <- total_FA_per_replicate %>%
  group_by(treatment) %>%
  summarize(mean_FA = mean(sum_FA, na.rm = TRUE))
print(mean_total_FA_ba)
```

#### Histogram FA biofilm

```{r}
hist(FA_Biofilm_absolute $ fa_conc) 
skewness(FA_Biofilm_absolute $ fa_conc, na.rm = TRUE) 
shapiro.test(FA_Biofilm_absolute $ fa_conc)
```
Biofilm fatty acid concentration not normal distributed and right skewed.


#### NMDS FA biofilm

Change table to wide format:

```{r}
# add prefix (week) 1,2 or 3 to treatments for later subgrouping
FA_Biofilm_absolute <- FA_Biofilm_absolute %>%
  mutate(treatment = if_else(week == "1", paste("1_", treatment, sep = ""), treatment))

FA_Biofilm_absolute <- FA_Biofilm_absolute %>%
  mutate(treatment = if_else(week == "2", paste("2_", treatment, sep = ""), treatment))

FA_Biofilm_absolute <- FA_Biofilm_absolute %>%
  mutate(treatment = if_else(week == "3", paste("3_", treatment, sep = ""), treatment))

# change table to wide format
FA_Biofilm_absolute_wide <-
  pivot_wider(FA_Biofilm_absolute,
              names_from = fa,
              values_from = fa_conc)

# square root transformation, to reduce discriminatory power of dominant fa groups
FA_Biofilm_absolute_wide_sqrt <- sqrt(FA_Biofilm_absolute_wide[, -c(1, 2, 3)])
View(FA_Biofilm_absolute_wide_sqrt)

any(is.na(FA_Biofilm_absolute_wide_sqrt))

```
Check for outliers:
```{r}
# noch machen
```


Perform metaMDS:

```{r}
# Rename and ease col-names for later NMDS plot
FA_Biofilm_absolute_wide_sqrt <- FA_Biofilm_absolute_wide_sqrt %>%
  rename(
    `8:0`      = `C08:0`,   
    `12:0`     = `C12:0`,   
    `13:0`     = `C13:0`,
    `14:0`     = `C14:0`,
    `15:0`     = `C15:0`,
    `15:1`     = `C15:1`,
    `16:0`     = `C16:0`,
    `16:1`     = `C16:1`,
    `17:1`     = `C17:1`,
    `18:0`     = `C18:0`,
    `18:1`     = `C18:1`,
    `18:2n-6t` = `C18:2(n-6t)`,
    `18:2n-6c` = `C18:2(n-6c)`,
    `18:3n-6`  = `C18:3(n-6)`,
    `18:3n-3`  = `C18:3(n-3)`,
    `18:4n-3`  = `C18:4(n-3)`,
    `20:0`     = `C20:0`,
    `20:1`     = `C20:1`,
    `20:2n-6`  = `C20:2(n-6)`,
    `20:3n-6`  = `C20:3(n-6)`,
    `20:4n-6`  = `C20:4(n-6)`,
    `20:3n-3`  = `C20:3(n-3)`,
    `21:0`     = `C21:0`,
    `20:5n-3`  = `C20:5(n-3)`,
    `22:0`     = `C22:0`,
    `22:1`     = `C22:1`,
    `22:2`     = `C22:2`,
    `22:5n-3`  = `C22:5(n-3)`,
    `22:6n-3`  = `C22:6(n-3)`,
    `24:0`     = `C24:0`,
    `24:1`     = `C24:1`
  )



NMDS_Biofilm_absolute <-
  metaMDS(
    FA_Biofilm_absolute_wide_sqrt,
    autotransform = FALSE,
    k = 2 ,
    distance = "bray",
    na.rm = TRUE
  )

NMDS_Biofilm_absolute$stress
```

Create fa vectors

```{r}
# separation in SAFA; MUFA & PUFA to build vectors in NMDS
# "ba" = biofilm absolute
FA_variables.SAFAs_ba <-
  envfit(
    NMDS_Biofilm_absolute,
    subset(
      FA_Biofilm_absolute_wide_sqrt,
      select = c(
        "8:0",
        "12:0",
        "13:0",
        "14:0",
        "15:0",
        "16:0",
        "18:0",
        "20:0",
        "21:0",
        "22:0",
        "24:0"
      )
    )
    ,
    permutations = 999,
    na.rm = TRUE
  )

FA_variables.MUFAs_ba <-
  envfit(
    NMDS_Biofilm_absolute,
    subset(
      FA_Biofilm_absolute_wide_sqrt,
      select = c(
        "15:1",
        "16:1",
        "17:1",
        "18:1",
        "20:1",
        "22:1",
        "24:1"
      )
    )
    ,
    permutations = 999,
    na.rm = TRUE
  )

FA_variables.PUFAs_ba <-
  envfit(
    NMDS_Biofilm_absolute,
    subset(
      FA_Biofilm_absolute_wide_sqrt,
      select = c(
        "18:2n-6t",
        "18:2n-6c",
        "18:3n-6",
        "18:3n-3",
        "18:4n-3",
        "20:2n-6",
        "20:3n-6",
        "20:4n-6",
        "20:3n-3",
        "20:5n-3",
        "22:2",
        "22:5n-3",
        "22:6n-3"
      )
    )
    ,
    permutations = 999,
    na.rm = TRUE
  )

# Extract NMDS scores
data.scores_biofilm_absolute = as.data.frame(scores(NMDS_Biofilm_absolute)$sites)

Treatments <-
  FA_Biofilm_absolute$treatment[FA_Biofilm_absolute$fa == "C16:0"]

# add treatment column to NMDS scores
data.scores_biofilm_absolute$Treatment = Treatments

# vectors for SAFA, MUFA and PUFA
SAFAs_coord_cont_ba = as.data.frame(scores(FA_variables.SAFAs_ba, "vectors")) * ordiArrowMul(FA_variables.SAFAs_ba)
MUFAs_coord_cont_ba = as.data.frame(scores(FA_variables.MUFAs_ba, "vectors")) * ordiArrowMul(FA_variables.MUFAs_ba)
PUFAs_coord_cont_ba = as.data.frame(scores(FA_variables.PUFAs_ba, "vectors")) * ordiArrowMul(FA_variables.PUFAs_ba)

# create weeks as subgroups for plotting
data.scores_biofilm_absolute <- data.scores_biofilm_absolute %>%
  mutate(time.point = case_when(
    str_detect(Treatment, "^1_") ~ "Week1",
    str_detect(Treatment, "^2_") ~ "Week2",
    str_detect(Treatment, "^3_") ~ "Week3",
    TRUE ~ "Other"
  ))

```

```{r}
# first delete 1_, 2_ & 3_ prefix in treatments
data.scores_biofilm_absolute $ Treatment <- sub("1_", "", data.scores_biofilm_absolute $ Treatment)
data.scores_biofilm_absolute $ Treatment <- sub("2_", "", data.scores_biofilm_absolute $ Treatment)
data.scores_biofilm_absolute $ Treatment <- sub("3_", "", data.scores_biofilm_absolute $ Treatment)
```

Plot NMDS:

```{r}
absol_Biofilm = ggplot(data = data.scores_biofilm_absolute, aes(x = NMDS1, y = NMDS2)) +
  geom_point( aes(shape = time.point, color = Treatment),
    size = 5,
    alpha = 1
  ) +
  scale_shape_manual(values = c(15, 16, 17)) +
  scale_color_manual(values = c("NC" = "#009E73", "Pro" = "#E69F00", "Cip" = "#56B4E9", "Mix" = "#F0E442")) +
  labs(shape = "Week", color = "Treatment") +
  geom_segment(data = PUFAs_coord_cont_ba, aes(x = 0, y = 0, xend = NMDS1 * 0.3, yend = NMDS2 * 0.3), linewidth = 0.5, alpha = 0.2, colour = "black") +
  geom_segment(data = MUFAs_coord_cont_ba, aes(x = 0, y = 0, xend = NMDS1 * 0.2, yend = NMDS2 * 0.2), linewidth = 0.5, alpha = 0.2, colour = "black") +
  geom_segment(data = SAFAs_coord_cont_ba, aes(x = 0, y = 0, xend = NMDS1 * 0.3, yend = NMDS2 * 0.3), linewidth = 0.5, alpha = 0.2, colour = "black") +
  geom_text_repel(
    data = PUFAs_coord_cont_ba, 
    aes(x = NMDS1 * 0.3, y = NMDS2 * 0.3, label = row.names(PUFAs_coord_cont_ba)), 
    colour = "grey30",
    size = 3) +
  geom_text_repel(
    data = MUFAs_coord_cont_ba, 
    aes(x = NMDS1 * 0.2, y = NMDS2 * 0.2, label = row.names(MUFAs_coord_cont_ba)), 
    colour = "grey30",
    size = 3) +
  geom_text_repel(
    data = SAFAs_coord_cont_ba, 
    aes(x = NMDS1 * 0.3, y = NMDS2 * 0.3, label = row.names(SAFAs_coord_cont_ba)), 
    colour = "grey30",
    size = 3) +
  theme_bw() +
  theme(
    axis.title = element_text(
      size = 11,
      face = "bold",
      colour = "grey30"
    ),
    panel.background = element_blank(),
    panel.border = element_rect(fill = NA, colour = "grey30"),
    #axis.ticks = element_blank(),
    #axis.text = element_blank(),
    legend.key = element_blank(),
    legend.title = element_text(
      size = 11,
      face = "bold",
      colour = "grey30"
    ),
    legend.text = element_text(size = 10, colour = "grey30"),
    plot.margin = margin(
      t = 0.2,
      b = 0.2,
      l = 0.2,
      r = 0.2,
      unit = "cm"
    ),
    panel.grid.major = element_line(colour = "grey95", size = 0.4),
    panel.grid.minor = element_line(colour = "grey95", size = 0.2)
  ) +
  annotate(
    "text",
    x = 0.32,
    y = -0.3,
    label = paste("stress:", NMDS_Biofilm_absolute$stress %>% round(2)),
    size = 3
  )
  
  
absol_Biofilm

```

Save the plot
```{r}
ggsave(absol_Biofilm, 
       filename = "NMDS_Biofilm_absolute.png", 
       path = 'C:\\Users\\Oster Test\\OneDrive\\Desktop\\Promotion\\Bioassays\\Schneckenversuch 2023\\Bottom_up_effects_biofilm_grazer\\output')
```




#### PERMANOVA FA biofilm

Create distance matrix with fatty acids:
```{r}
#FA_Snails_dist <- vegdist(FA_Snails_absolute_wide_sqrt, method = 'bray')
```




#### SIMPER FA biofilm

```{r}
# try out Simper
#biofilm_treatments_table <-
 # read.table(file.path(prj, "data\\biofilm_treatments_table.txt"),
            # header = TRUE, sep = "\t")

#View(biofilm_treatments_table)

# drei erste Spalten weg vor Simper (generell wide format muss chronologisch hier darüber)
#FA_Biofilm_absolute_wide <- FA_Biofilm_absolute_wide[, -c(1, 2, 3)]

#simper_FA_Snails <- with(biofilm_treatments_table, simper(FA_Biofilm_absolute_wide, treatment,  permutations = 999))

#summary(simper_FA_Snails)
```


# 8 Temperature and light conditions

## 8.1 Biofilm colonisation and contamination

```{r}
# insert data
colonisation_temp_light <-
  read.table(file.path(prj, "data\\colonisation_temp_light.txt"),
             header = TRUE, sep = "\t")

View(colonisation_temp_light)

# average temperature
mean(colonisation_temp_light $ temperature_colonisation)
# 18.28751 °C

# mean light intensity
# filter only daytime
average_light_intensity_colonisation <- colonisation_temp_light %>%
  dplyr::filter(light_intensity_colonisation > 0) %>%
  summarize(average_light_intensity_colonisation = mean(light_intensity_colonisation))

print(average_light_intensity_colonisation)
# 14019.26 lux

# temperature plot
# add a small amount of random noise to the y-values
colonisation_temp_light $ temperature_colonisation <- jitter(colonisation_temp_light $ temperature_colonisation)

t2 <- ggplot(
  data = colonisation_temp_light,
  mapping = aes( x = day, y = temperature_colonisation, color = time_of_day)) +
  geom_point(alpha = 0.5) +  # Individual measurements with transparency
  geom_smooth() +
  labs(
    title = "Temperature profile during biofilm colonisation",
    x = "Day", y = "Temperature (°C)"
  ) +
  theme_bw() +
  scale_x_continuous(breaks = seq(0, 77, by = 5)) +
  scale_color_continuous(
    name = "Time of Day",  
    breaks = c(0, 5, 10, 15, 20),
    labels = c("0:00 am", "05:00 am", "10:00 am", "03:00 pm", "08:00 pm"),
    limits = c(0, 22),
  )

t2 

# save the plot
ggsave(t2, 
       filename = "temperature_colonisation.png", 
       path = 'C:\\Users\\Oster Test\\OneDrive\\Desktop\\Promotion\\Bioassays\\Schneckenversuch 2023\\Bottom_up_effects_biofilm_grazer\\output')

# light intensity plot
# Filter the data to include only rows where light_intensity > 0
filtered_data2 <- colonisation_temp_light %>% 
  filter(light_intensity_colonisation > 0)

# add a small amount of random noise to the y-values
filtered_data2 $ light_intensity_colonisation <- jitter(filtered_data2 $ light_intensity_colonisation)

l2 <- ggplot(
  data = filtered_data2,
  mapping = aes( x = day, y = light_intensity_colonisation)) +
  geom_point(size = 1) +  # Individual measurements with transparency
  geom_smooth(color = "#E69F00", linewidth = 1) +
  labs(
    title = "Light intensity profile during biofilm colonisation",
    x = "Day", y = "Light intensity (lux)"
  ) +
  theme_bw() +
  scale_x_continuous(breaks = seq(0, 77, by = 5)) 


l2 

# save the plot
ggsave(l2, 
       filename = "light_intensity_colonisation.png", 
       path = 'C:\\Users\\Oster Test\\OneDrive\\Desktop\\Promotion\\Bioassays\\Schneckenversuch 2023\\Bottom_up_effects_biofilm_grazer\\output')


```

## 8.2 Feeding assay

```{r}
# insert data
feeding_assay_temp_light <-
  read.table(file.path(prj, "data\\feeding_assay_temp_light.txt"),
             header = TRUE, 
             sep = "\t"
  )
             
View(feeding_assay_temp_light)

# average temperature
mean(feeding_assay_temp_light $ temperature)
# 13.27927 °C

# mean light intensity
# filter only daytime
average_light_intensity_feeding <- feeding_assay_temp_light %>%
  dplyr::filter(light_intensity > 0) %>%
  summarize(average_light_intensity_feeding = mean(light_intensity))

print(average_light_intensity_feeding)
# 330.3552 lux

# temperature plot
# add a small amount of random noise to the y-values
feeding_assay_temp_light $ temperature <- jitter(feeding_assay_temp_light $ temperature)

t1 <- ggplot(
  data = feeding_assay_temp_light,
  mapping = aes( x = day, y = temperature, color = time_of_day)) +
  geom_point(alpha = 0.5) +  # Individual measurements with transparency
  geom_smooth() +
  labs(
    title = "Temperature profile during 21-day feeding assay",
    x = "Day", y = "Temperature (°C)"
  ) +
  theme_bw() +
  scale_x_continuous(breaks = 0:21) +
  scale_color_continuous(
    name = "Time of Day",  
    breaks = c(0, 5, 10, 15, 20),
    labels = c("0:00 am", "05:00 am", "10:00 am", "03:00 pm", "08:00 pm"),
    limits = c(0, 22),
  )

t1 

# save the plot
ggsave(t1, 
       filename = "temperature_feeding_assay.png", 
       path = 'C:\\Users\\Oster Test\\OneDrive\\Desktop\\Promotion\\Bioassays\\Schneckenversuch 2023\\Bottom_up_effects_biofilm_grazer\\output')


# light intensity plot
# Filter the data to include only rows where light_intensity > 0
filtered_data1 <- feeding_assay_temp_light %>% 
  filter(light_intensity > 0)

# add a small amount of random noise to the y-values
filtered_data1 $ light_intensity <- jitter(filtered_data1 $ light_intensity)

l1 <- ggplot(
  data = filtered_data1,
  mapping = aes( x = day, y = light_intensity)) +
  geom_point(size = 1) +  # Individual measurements with transparency
  geom_smooth(color = "#E69F00", linewidth = 1) +
  labs(
    title = "Light intensity profile during 21-day feeding assay",
    x = "Day", y = "Light intensity (lux)"
  ) +
  theme_bw() +
  scale_x_continuous(breaks = 0:21) 


l1 

# save the plot
ggsave(l1, 
       filename = "light_intensity_feeding_assay.png", 
       path = 'C:\\Users\\Oster Test\\OneDrive\\Desktop\\Promotion\\Bioassays\\Schneckenversuch 2023\\Bottom_up_effects_biofilm_grazer\\output')


```
