---
title: "Old R code snail paper"
format: html
editor: visual
---

# Relative values of FA snails

Calculate the relative amount of one fatty acid in a snail in relation to the total amount of fatty acids in this one snail.

```{r}
FA_Snails_relative <- FA_Snails_absolute_filtered %>%
  group_by(replicate) %>%
  mutate(fa_conc_rel = (fa_conc / sum(fa_conc)) * 100)
print(FA_Snails_relative)
```

Histogram FA Snails (relative)

```{r}
hist(FA_Snails_relative $ fa_conc_rel)
skewness(FA_Snails_relative $ fa_conc_rel, na.rm = TRUE)
shapiro.test(FA_Snails_relative $ fa_conc_rel)
```

Not normal distributed and right skewed.

Model FA Snails (relative)

```{r}
# Log10 transformation; adding 0.001
FA_Snails_relative <- FA_Snails_relative %>%
  mutate(log_fa_conc_rel = log10(fa_conc_rel + 0.001))
# response variable: log_fa_conc
# fixed effects: treatment + fa + treatment:fa
# random effect with constant intercept: (1|replicate)
FA_Snails_relative_lmer <- lmer(log_fa_conc_rel ~ treatment + fa + treatment:fa + (1|replicate), data = FA_Snails_relative) 

plot(FA_Snails_relative_lmer) # @ VS: siehe oben bei FA Snails, wegen den 0-Werten sieht das komisch aus, oder ist das als Normalverteilung noch ok?
qreference(residuals(FA_Snails_relative_lmer)) 

anova(FA_Snails_relative_lmer)
```

```{r}
FA_snails_relative_inter <- lsmeans(FA_Snails_relative_lmer, ~ treatment | fa)
FA_snails_relative_inter


FA_snails_relative_inter_dunn1 <- contrast(FA_snails_relative_inter, 'pairwise')

FA_snails_relative_inter_dunn_sum1 <- summary(FA_snails_relative_inter_dunn1, infer = TRUE, adjust = 'mvt')

FA_snails_relative_inter_dunn_sum1
```

Results quite similar to absolute values. Except the following marks, fatty acid concentration of the 4 treatments quite similar:

C12:0 : Ciprofloxacin treatment to Mix and NC (and almost to Pro) significantly different

C13:0 : Ciprofloxacin to all other treatments significantly different. Mix and Pro also different

Mix to all treatments different

C18:3(n-6) : Cip-NC almost different

C22:1 : NC different to all chemical treatments. Mix and Pro also different.

Total SAFA, MUFA & PUFA content (relative)

1\) Total SAFA content

```{r}
# total SAFA, MUFA, PUFA content; sr = "snail relative"
# total SAFA content
# filter saturated fatty acids in new table
SAFA_sr <- FA_Snails_relative %>%
  filter(fa %in% c("C08:0", "C12:0", "C13:0", "C14:0", "C15:0", "C16:0", "C17:0", "C18:0", "C20:0", "C24:0"))

# calculate the total SAFA content per snail
total_SAFA_per_snail_relative <- SAFA_sr %>%
  group_by(replicate, treatment) %>%
  summarize(sum_SAFA = sum(fa_conc_rel))

print(total_SAFA_per_snail_relative)

# mean
mean_SAFA_sr <- total_SAFA_per_snail_relative %>%
  group_by(treatment) %>%
  summarize(mean_SAFA = mean(sum_SAFA, na.rm = TRUE))
print(mean_SAFA_sr)
```

2\) Total MUFA content

```{r}
# filter mono-unsaturated fatty acids in new table
MUFA_sr <- FA_Snails_relative %>%
  filter(fa %in% c("C16:1", "C18:1", "C20:1", "C22:1"))

# total MUFA content per snail
total_MUFA_per_snail_relative <- MUFA_sr %>%
  group_by(replicate, treatment) %>%
  summarize(sum_MUFA = sum(fa_conc_rel))

print(total_MUFA_per_snail_relative)

# mean
mean_MUFA_sr <- total_MUFA_per_snail_relative %>%
  group_by(treatment) %>%
  summarize(mean_MUFA = mean(sum_MUFA, na.rm = TRUE))
print(mean_MUFA_sr)
```

3\) PUFA content

```{r}
# filter poly-unsaturated fatty acids in new table
PUFA_sr <- FA_Snails_relative %>%
  filter(fa %in% c("C18:2(n-6c)", "C18:2(n-6t)", "C18:3(n-3)", "C18:3(n-6)", "C18:4(n-3)", "C20:2(n-6)", "C20:3(n-6)", "C20:4(n-6)", "C20:5(n-3)", "C22:2", "C22:5(n-3)", "C22:6(n-3)"))

# calculate total PUFA content per snail
total_PUFA_per_snail_relative <- PUFA_sr %>%
  group_by(replicate, treatment) %>%
  summarize(sum_PUFA = sum(fa_conc_rel))

print(total_PUFA_per_snail_relative)

# mean
mean_PUFA_sr <- total_PUFA_per_snail_relative %>%
  group_by(treatment) %>%
  summarize(mean_PUFA = mean(sum_PUFA, na.rm = TRUE))
print(mean_PUFA_sr)
```

4\) Total FA (relative) per snail = 100 %.

# Relative values for biofilm FA

```{r}
FA_Biofilm_relative <- FA_Biofilm_absolute %>%
  group_by(replicate) %>%
  mutate(fa_conc_rel = (fa_conc / sum(fa_conc)) * 100)
print(FA_Biofilm_relative)
```

```{r}
# Remove the "log_fa_conc" column
FA_Biofilm_relative <- FA_Biofilm_relative[, !names(FA_Biofilm_relative) %in% "log_fa_conc"]
```

#### **Histogram FA Biofilm (relative)**

```{r}
hist(FA_Biofilm_relative $ fa_conc_rel)
skewness(FA_Biofilm_relative $ fa_conc_rel, na.rm = TRUE)
shapiro.test(FA_Biofilm_relative $ fa_conc_rel)
```

Not normal distributed and right skewed.

#### **Model FA Biofilm (relative)**

Hier weitermachen, wie Verena für Model FA Biofilm vorschlägt.

```{r}
# log10 transformation; adding 0.001 
# FA_Biofilm_relative <- FA_Biofilm_relative %>%
  #mutate(log_fa_conc_rel = log10(fa_conc_rel + 0.001))

# response variable: log_fa_conc
# fixed effects: treatment + fa +  + week + treatment:fa + week:fa + treatment:week
# random effect with constant intercept: (1|replicate)
#FA_Biofilm_relative_lmer <- lmer(log_fa_conc_rel ~ treatment + fa + week + week:fa + treatment:week + treatment:fa + (1|replicate), data = FA_Biofilm_relative)
# @ VS: ist das richtig so? alle 3 explanatory variables und jeweils alle interaktionen zu testen?

#plot(FA_Biofilm_relative_lmer) # @ VS: siehe oben, ist das normal verteilt?
#qreference(residuals(FA_Biofilm_relative_lmer)) 

#anova(FA_Biofilm_relative_lmer)
```

#### **Total SAFA, MUFA & PUFA content (relative)**

1\) SAFA content

```{r}
# total SAFA, MUFA, PUFA content; br = "biofilm relative"
# total SAFA content
# filter saturated fatty acids in new table
SAFA_br <- FA_Biofilm_relative %>%
    filter(fa %in% c("C08:0", "C10:0", "C12:0", "C13:0", "C14:0", "C15:0", "C16:0", "C17:0", "C18:0", "C20:0", "C22:0", "C24:0"))

# calculate the total SAFA content per replicate
total_SAFA_per_replicate_relative <- SAFA_br %>%
  group_by(replicate, treatment) %>%
  summarize(sum_SAFA = sum(fa_conc_rel))

print(total_SAFA_per_replicate_relative)

# mean
mean_SAFA_br <- total_SAFA_per_replicate_relative %>%
  group_by(treatment) %>%
  summarize(mean_SAFA = mean(sum_SAFA, na.rm = TRUE))
print(mean_SAFA_br)
```

2\) Total MUFA content

```{r}
# filter mono-unsaturated fatty acids in new table
MUFA_br <- FA_Biofilm_relative %>%
  filter(fa %in% c("C16:1", "C17:1", "C18:1", "C20:1"))

# total MUFA content per replicate
total_MUFA_per_replicate_relative <- MUFA_br %>%
  group_by(replicate, treatment) %>%
  summarize(sum_MUFA = sum(fa_conc_rel))

print(total_MUFA_per_replicate_relative)

# mean
mean_MUFA_br <- total_MUFA_per_replicate_relative %>%
  group_by(treatment) %>%
  summarize(mean_MUFA = mean(sum_MUFA, na.rm = TRUE))
print(mean_MUFA_br)
```

3\) Total PUFA content

```{r}
# filter poly-unsaturated fatty acids in new table
PUFA_br <- FA_Biofilm_relative %>%
  filter(fa %in% c("C18:2(n-6c)", "C18:2(n-6t)", "C18:3(n-3)", "C18:3(n-6)", "C18:4(n-3)", "C20:2(n-6)", "C20:3(n-6)", "C20:3(n-3)", "C20:4(n-6)", "C20:5(n-3)", "C22:2", "C22:5(n-3)", "C22:6(n-3)"))
  
# calculate total PUFA content per replicate
total_PUFA_per_replicate_relative <- PUFA_br %>%
  group_by(replicate, treatment) %>%
  summarize(sum_PUFA = sum(fa_conc_rel))

print(total_PUFA_per_replicate_relative)

# mean
mean_PUFA_br <- total_PUFA_per_replicate_relative %>%
  group_by(treatment) %>%
  summarize(mean_PUFA = mean(sum_PUFA, na.rm = TRUE))
print(mean_PUFA_br)
```

# 16.11.2024

\#### NMDS FA snails

Standardization to the maximum of each individual FA or FA group (i.e. saturated FA, SAFA; monounsaturated FA, MUFA, etc.) to reduce discriminatory power of dominant FA or FA groups.

```{r}
# SAFA
SAFA_sa <- SAFA_sa %>%
  mutate(standardized_FA_conc = fa_conc /  total_SAFA_per_snail$sum_SAFA)

# MUFA
MUFA_sa <- MUFA_sa %>%
  mutate(standardized_FA_conc = fa_conc / total_MUFA_per_snail$sum_MUFA)

# PUFA
PUFA_sa <- PUFA_sa %>%
  mutate(standardized_FA_conc = fa_conc / total_PUFA_per_snail$sum_PUFA)


# Combine data sets and add source:
SAFA_sa <- SAFA_sa %>% mutate(source = "SAFA")
MUFA_sa <- MUFA_sa %>% mutate(source = "MUFA")
PUFA_sa <- PUFA_sa %>% mutate(source = "PUFA")

# Now combine with source identifiers
combined_data_sa <- bind_rows(SAFA_sa, MUFA_sa, PUFA_sa) %>%
  arrange(replicate)

# deselect columns which are unneccessary for NMDS: fa_conc, treatment & source
combined_data_sa_subset <- combined_data_sa[, -c(3, 4, 6)]
```

Change table to wide format:

```{r}
FA_Snails_absolute_wide <-
  pivot_wider(combined_data_sa_subset,
              names_from = fa,
              values_from = standardized_FA_conc)


any(is.na(FA_Snails_absolute_wide))

# set NA to 0 
# FA_Snails_absolute_wide[is.na(FA_Snails_absolute_wide)] <- 0
# View(FA_Snails_absolute_wide)
```

Perform metaMDS:

```{r}
# Remove outlier replicates S_1, S_13, S_25 and S_30
FA_Snails_absolute_wide <- FA_Snails_absolute_wide %>%
  filter(!(replicate %in% c("S_1", "S_13", "S_25", "S_30")))


# Remove replicate column for metaMDS
FA_Snails_absolute_wide <- FA_Snails_absolute_wide[, -c(1)]

# Rename (- C, - () ) fatty acids for later plot
FA_Snails_absolute_wide <- FA_Snails_absolute_wide %>%
  rename(
    `8:0`      = `C08:0`,   
    `12:0`     = `C12:0`,   
    `13:0`     = `C13:0`,
    `14:0`     = `C14:0`,
    `15:0`     = `C15:0`,
    `16:0`     = `C16:0`,
    `16:1`     = `C16:1`,
    `18:0`     = `C18:0`,
    `18:1`     = `C18:1`,
    `18:2n-6t` = `C18:2(n-6t)`,
    `18:2n-6c` = `C18:2(n-6c)`,
    `18:3n-6`  = `C18:3(n-6)`,
    `18:3n-3`  = `C18:3(n-3)`,
    `18:4n-3`  = `C18:4(n-3)`,
    `20:1`     = `C20:1`,
    `20:2n-6`  = `C20:2(n-6)`,
    `20:3n-6`  = `C20:3(n-6)`,
    `20:4n-6`  = `C20:4(n-6)`,
    `21:0`     = `C21:0`,
    `20:5n-3`  = `C20:5(n-3)`,
    `22:1`     = `C22:1`,
    `22:2`     = `C22:2`,
    `22:5n-3`  = `C22:5(n-3)`,
    `22:6n-3`  = `C22:6(n-3)`,
    `24:0`     = `C24:0`,
    `24:1`     = `C24:1`
  )



NMDS_Snails_absolute <-
  metaMDS(
    FA_Snails_absolute_wide,
    autotransform = FALSE,
    k = 2 ,
    distance = "bray",
    na.rm = TRUE
  )

NMDS_Snails_absolute$stress 
```

Create fa vectors

```{r}
# separation in SAFA; MUFA & PUFA to build vectors in NMDS
# "sa" = snail absolute
FA_variables.SAFAs_sa <-
  envfit(
    NMDS_Snails_absolute,
    subset(
      FA_Snails_absolute_wide,
      select = c(
        "8:0",
        "12:0",
        "13:0",
        "14:0",
        "15:0",
        "16:0",
        "18:0",
        "24:0"
      )
    )
    ,
    permutations = 999,
    na.rm = TRUE
  )

FA_variables.MUFAs_sa <-
  envfit(
    NMDS_Snails_absolute,
    subset(
      FA_Snails_absolute_wide,
      select = c(
        "16:1",
        "18:1",
        "20:1",
        "22:1",
        "24:1"
      )
    )
    ,
    permutations = 999,
    na.rm = TRUE
  )

FA_variables.PUFAs_sa <-
  envfit(
    NMDS_Snails_absolute,
    subset(
      FA_Snails_absolute_wide,
      select = c(
        "18:2n-6t",
        "18:2n-6c",
        "18:3n-6",
        "18:3n-3",
        "18:4n-3",
        "20:2n-6",
        "20:3n-6",
        "20:4n-6",
        "20:5n-3",
        "22:5n-3",
        "22:6n-3"
      )
    )
    ,
    permutations = 999,
    na.rm = TRUE
  )

# Extract NMDS scores 
data.scores_snail_absolute = as.data.frame(scores(NMDS_Snails_absolute)$sites)

# Remove S_1, S_13, S_25 and S_30 from first data frame "FA_Snails_absolute"
FA_Snails_absolute <- FA_Snails_absolute %>%
  filter(!(replicate %in% c("S_1", "S_13", "S_25", "S_30")))


Treatments <-
  FA_Snails_absolute $ treatment[FA_Snails_absolute $ fa == "C16:0"]

# add treatment column to NMDS scores
data.scores_snail_absolute$Treatment = Treatments

# vectors for SAFA, MUFA and PUFA
SAFAs_coord_cont_sa = as.data.frame(scores(FA_variables.SAFAs_sa, "vectors")) * ordiArrowMul(FA_variables.SAFAs_sa)
MUFAs_coord_cont_sa = as.data.frame(scores(FA_variables.MUFAs_sa, "vectors")) * ordiArrowMul(FA_variables.MUFAs_sa)
PUFAs_coord_cont_sa = as.data.frame(scores(FA_variables.PUFAs_sa, "vectors")) * ordiArrowMul(FA_variables.PUFAs_sa)
```

Plot NMDS:

```{r}
# plot
absol_Snail <- ggplot(data = data.scores_snail_absolute, aes(x = NMDS1, y = NMDS2)) +
  geom_point(
    aes(shape = Treatment, color = Treatment),
    size = 5,
    alpha = 1,
    show.legend = TRUE
  ) +
  scale_shape_manual(values = c("NC" = 16, "Pro" = 17, "Cip" = 18, "Mix" = 15)) +
  scale_color_manual(values = c("NC" = "#009E73", "Pro" = "#E69F00", "Cip" = "#56B4E9", "Mix" = "#F0E442")) +
  stat_conf_ellipse(
    aes(colour = Treatment, fill = Treatment),
    alpha = 0.3,
    geom = "polygon",
    show.legend = FALSE
  ) +
  scale_color_manual(values = c("NC" = "#009E73", "Pro" = "#E69F00", "Cip" = "#56B4E9", "Mix" = "#F0E442")) +  
  scale_fill_manual(values = c("NC" = "#009E73", "Pro" = "#E69F00", "Cip" = "#56B4E9", "Mix" = "#F0E442")) + 
  geom_segment(aes(x = 0, y = 0, xend = NMDS1 * 0.2, yend = NMDS2 * 0.2), 
               data = PUFAs_coord_cont_sa, linewidth = 0.5, alpha = 0.2, colour = "black") +
    geom_segment(aes(x = 0, y = 0, xend = NMDS1 * 0.2, yend = NMDS2 * 0.2), 
               data = MUFAs_coord_cont_sa, linewidth = 0.5, alpha = 0.2, colour = "black") +
      geom_segment(aes(x = 0, y = 0, xend = NMDS1 * 0.1, yend = NMDS2 * 0.1), 
               data = SAFAs_coord_cont_sa, linewidth = 0.5, alpha = 0.2, colour = "black") +
  geom_text_repel(
    data = PUFAs_coord_cont_sa, 
    aes(x = NMDS1 * 0.2, y = NMDS2 * 0.2, label = row.names(PUFAs_coord_cont_sa)), 
    colour = "grey30",
    size = 2) +
  geom_text_repel(
    data = MUFAs_coord_cont_sa, 
    aes(x = NMDS1 * 0.2, y = NMDS2 * 0.2, label = row.names(MUFAs_coord_cont_sa)), 
    colour = "grey30",
    size = 2) +
  geom_text_repel(
    data = SAFAs_coord_cont_sa, 
    aes(x = NMDS1 * 0.1, y = NMDS2 * 0.1, label = row.names(SAFAs_coord_cont_sa)), 
    colour = "grey30",
    size = 2) +
  theme_bw() +
  theme(
    axis.title = element_text(
      size = 11,
      face = "bold",
      colour = "grey30"
    ),
    panel.background = element_blank(),
    panel.border = element_rect(fill = NA, colour = "grey30"),
    legend.key = element_blank(),
    legend.title = element_text(
      size = 11,
      face = "bold",
      colour = "grey30"
    ),
    legend.text = element_text(size = 10, colour = "grey30"),
    #plot.margin = margin(
     # t = 0.2,
    #  b = 0.2,
    #  l = 0.2,
    #  r = 0.2,
    #  unit = "cm"
  #  ),
    panel.grid.major = element_line(colour = "grey95", size = 0.4),
    panel.grid.minor = element_line(colour = "grey95", size = 0.2)
  ) +
  annotate(
    "text",
    x = 0.35,
    y = -0.19,
    label = paste("stress:", NMDS_Snails_absolute$stress %>% round(2)),
    size = 3.5
  ) 


absol_Snail

# position of FA names manually adjusted in vector graphic
```

Model FA Snails

```{r}
# log10 transformation; adding "0.1" to the lowest value
# which is lowest value?
FA_Snails_no0 <- subset(FA_Snails_absolute, fa_conc != 0)

lowest_value <- min(FA_Snails_no0 $ fa_conc , na.rm = TRUE)
print(lowest_value)

# lowest value in dataset is 0.00776, use 0.10776 for transformation

FA_Snails_absolute <- FA_Snails_absolute %>%
  mutate(log_fa_conc = log10(fa_conc + 0.10776))

# response variable: log_fa_conc
# fixed effects: treatment + fa + treatment:fa
# random effect with constant intercept: (1|replicate); first change to factor
FA_Snails_absolute $ replicate <- as.factor(FA_Snails_absolute $ replicate)
FA_Snails_absolute $ treatment <- as.factor(FA_Snails_absolute $ treatment)
FA_Snails_absolute $ fa <- as.factor(FA_Snails_absolute $ fa)

FA_snails_lmer <- lmer(log_fa_conc ~ treatment + fa + treatment:fa + (1|replicate), data = FA_Snails_absolute)
```

```{r}
plot(FA_snails_lmer) 
# @ VS: das sieht nicht wirklich gleichmäßig verteilt aus, oder? Was würdest du jetzt machen?

qreference(residuals(FA_snails_lmer)) 
# @ VS: das sieht schon wieder okay aus 

anova(FA_snails_lmer)
```

```{r}
FA_snails_inter <- lsmeans(FA_snails_lmer, ~ treatment | fa)
FA_snails_inter

FA_snails_inter_dunn1 <- contrast(FA_snails_inter, 'pairwise')

FA_snails_inter_dunn_sum1 <- summary(FA_snails_inter_dunn1, infer = TRUE, adjust = 'mvt')

FA_snails_inter_dunn_sum1
```

Except the following remarks, fatty acid concentration of the 4 treatments quite similar.

\- C12:0 : Ciprofloxacin treatment to Mix and NC (and almost to Pro) significantly different

\- C13:0 : Ciprofloxacin to all other treatments significantly different. Mix and Pro also different

\- C14:0 : Cip-Mix and Mix-NC different

\- C18:3(n-6) : Cip-NC and almost Mix-NC different

\- C22:1 : NC different to all chemical treatments. Mix and Pro also different.

# Model biofilm FA

```{r}
# log10 transformation; adding 0.001 
FA_Biofilm_absolute <- FA_Biofilm_absolute %>%
  mutate(log_fa_conc = log10(fa_conc + 0.001))

# response variable: log_fa_conc
# fixed effects: treatment + fa +  + week + treatment:fa + week:fa + treatment:week
# random effect with constant intercept: (1|replicate); first change to factor
FA_Biofilm_absolute $ replicate <- as.factor(FA_Biofilm_absolute $ replicate)
FA_Biofilm_absolute $ week <- as.factor(FA_Biofilm_absolute $ week)
FA_Biofilm_absolute $ treatment <- as.factor(FA_Biofilm_absolute $ treatment)
FA_Biofilm_absolute $ fa <- as.factor(FA_Biofilm_absolute $ fa)

FA_Biofilm_lmer <- lmer(log_fa_conc ~ treatment + fa + week + week:fa + treatment:week + treatment:fa + (1|replicate), data = FA_Biofilm_absolute)
# @ VS: ist das richtig so? alle 3 explanatory variables und jeweils alle interaktionen zu testen?
```

```{r}
plot(FA_Biofilm_lmer) 
# @ VS: das sieht nicht wirklich gleichmäßig verteilt aus, oder? Was würdest du jetzt machen?

qreference(residuals(FA_Biofilm_lmer)) 
# @ VS: das sieht schon wieder okay aus 

anova(FA_Biofilm_lmer)

drop1(FA_Biofilm_lmer, test = 'Chisq')
```

```{r}
FA_Biofilm_inter <- lsmeans(FA_Biofilm_lmer, ~ treatment | fa)
FA_Biofilm_inter

FA_Biofilm_inter_dunn1 <- contrast(FA_Biofilm_inter, 'pairwise')

FA_Biofilm_inter_dunn_sum1 <- summary(FA_Biofilm_inter_dunn1, infer = TRUE, adjust = 'mvt')

FA_Biofilm_inter_dunn_sum1
# @ VS hier kommen bei mir "Inf" Werte raus, irgendwie klappt das so nicht, ich habe nicht herausgefunden, was ich ändern muss
```
