---
title: "Bottom-up effects of an herbicide and an antibiotic on *Potamopyrgus antipodarum*"
format: 
  html:
    self-contained: true
    toc: true
    toc-depth: 2
    number-sections: true
    style: "floating"
editor_options: 
  chunk_output_type: console
---

Sophie Oster, Patrick Fink, Tobias Schmitt, Verena C. Schreiner & Mirco Bundschuh

R script written by Sophie Oster\
checked by Verena C. Schreiner

\
RPTU Kaiserslautern-Landau\
Fortstrasse 7\
76829 Landau in der Pfalz\
GERMANY

This study was conducted with four chemical treatments, as explanatory variables:

-   treatment (categorical)
    -   uncontaminated control (negative control, NC)
    -   herbicide propyzamide 10 ug / L (Pro)
    -   antibiotic ciprofloxacin 100 ug / L (Cip)
    -   mixture of 10 ug / L  propyzamide and 100 ug / L ciprofloxacin (Mix)


# R version 4.3.2

R version used for this script: 4.3.2, use the following code chunk to download the correct version.

```{r}
# Check the current R version
current_r_version <- as.character(getRversion())

# Specify the required R version
required_r_version <- "4.3.2"

if (current_r_version != required_r_version) {
  # Install 'rig' if not already installed
  if (!requireNamespace("rig", quietly = TRUE)) {
    install.packages("rig")
  }
  
  # Load the 'rig' package
  library(rig)
  
  # Check if the required R version is installed
  installed_versions <- rig::rig_ls()
  if (!(required_r_version %in% installed_versions$version)) {
    # Install the required R version using 'rig'
    rig::rig_add(required_r_version)
  }
  
  # Set the required R version as default
  rig::rig_default(required_r_version)
  
  # Notify the user about the version change
  stop(paste("Please restart R using version", required_r_version, "and rerun the script."))
} else {
  message("You are already using the correct R version:", required_r_version)
}

```

# Packages in a reproducible environment

```{r, message = FALSE, warning = FALSE}
# Restore the environment:
# renv::restore()

library(stats)
library(rstatix)
library(dplyr)
library(tidyverse)
library(tidyr)
library(broom)
library(dunn.test)
library(patchwork)
library(vegan)
library(data.table)
library(ggplot2)
library(ggrepel)
library(ggforce)
library(concaveman)
library(gridExtra)
library(stringr)
library(shades)
library(ggpubr)
library(base)
library(ggokabeito)
library(multcomp)
library(car)
library(asbio)
library(plotrix)
library(lmtest)
library(boot)
library(moments)
library(lme4)
library(DAAG)
library(performance)
library(emmeans)
library(purrr)
library(DT)
library(Hmisc)

if (!requireNamespace("devtools", quietly = TRUE)) {
  install.packages("devtools")
}

library(devtools)

if (!requireNamespace("pairwiseAdonis", quietly = TRUE)) {
  devtools::install_github("pmartinezarbizu/pairwiseAdonis/pairwiseAdonis")
}

library(pairwiseAdonis)

# renv::snapshot()
```

# Biofilm endpoints

Chlorophyll concentrations and ash free dry weigh (AFDW) of biofilm

-   Unit chlorophyll a, b & c: µg chlorophyll / cm²

-   Unit afdw: mg / cm²

```{r}
# insert data
biofilm_endpoints <-
  read.table(file.path("data\\biofilm_endpoints.txt"),
             header = TRUE, sep = "\t")

View(biofilm_endpoints)
```

##  Chlorophyll a, b and c concentrations

Caclulate mean, sd and 95% CI chla, chlb + chlc of each treatment:

```{r}
# chla
mean_chla <- biofilm_endpoints %>%
  group_by(treatment) %>%
  dplyr::summarize(
    mean_chla = mean(chla, na.rm = TRUE),
    n = sum(!is.na(chla)),
    se = sd(chla, na.rm = TRUE) / sqrt(n),
    t_crit = qt(0.975, df = n - 1),
    ci_lower = mean_chla - t_crit * se,
    ci_upper = mean_chla + t_crit * se,
    .groups = "drop"
  ) %>%
  dplyr::select(treatment, mean_chla, ci_lower, ci_upper, se)

print(mean_chla)

#chlb
mean_chlb <- biofilm_endpoints %>%
  group_by(treatment) %>%
  dplyr::summarize(
    mean_chlb = mean(chlb, na.rm = TRUE),
    n = sum(!is.na(chlb)),
    se = sd(chlb, na.rm = TRUE) / sqrt(n),
    t_crit = qt(0.975, df = n - 1),
    ci_lower = mean_chlb - t_crit * se,
    ci_upper = mean_chlb + t_crit * se,
    .groups = "drop"
  ) %>%
  dplyr::select(treatment, mean_chlb, ci_lower, ci_upper, se)

print(mean_chlb)

#chlc
mean_chlc <- biofilm_endpoints %>%
  group_by(treatment) %>%
  dplyr::summarize(
    mean_chlc = mean(chlc, na.rm = TRUE),
    n = sum(!is.na(chlc)),
    se = sd(chlc, na.rm = TRUE) / sqrt(n),
    t_crit = qt(0.975, df = n - 1),
    ci_lower = mean_chlc - t_crit * se,
    ci_upper = mean_chlc + t_crit * se,
    .groups = "drop"
  ) %>%
  dplyr::select(treatment, mean_chlc, ci_lower, ci_upper, se)

print(mean_chlc)
```


### Histogram chlorophyll a, b + c

```{r}
# chl a
hist(biofilm_endpoints $ chla) 
skewness(biofilm_endpoints $ chla, na.rm = TRUE)
shapiro.test(biofilm_endpoints $ chla)

# chl b
hist(biofilm_endpoints $ chlb) 
skewness(biofilm_endpoints $ chlb, na.rm = TRUE) 
shapiro.test(biofilm_endpoints $ chlb)

# chl c
hist(biofilm_endpoints $ chlc) 
skewness(biofilm_endpoints $ chlc, na.rm = TRUE) 
shapiro.test(biofilm_endpoints $ chlc)
```

All not normal distributed, right skewed but skewness \< 1.5 therefore no data transformation for model

### Model chla

Linear model with treatment as explanatory variable.

```{r}
# treatment as factor
biofilm_endpoints $ treatment <- factor(biofilm_endpoints $ treatment)

model_chla <- lm(chla ~ treatment, data = biofilm_endpoints)

par(mfrow = c(2, 2))

plot(model_chla) # point cloud looks evenly distributed --> normal distribution

qreference(residuals(model_chla)) 

drop1(model_chla, test = 'F')

anova(model_chla)

chla_inter2 <- lsmeans(model_chla, ~ treatment)
chla_inter_dunn2 <- contrast(chla_inter2, 'pairwise')
(chla_inter_dunn_sum2 <- summary(chla_inter_dunn2, infer = TRUE, adjust = 'mvt'))
```
No significant differences between chemical treatments in chlorophyll a concentrations.

### Model chlb

```{r}
model_chlb <- lm(chlb ~ treatment, data = biofilm_endpoints)

par(mfrow = c(2, 2))

plot(model_chlb) # point cloud looks evenly distributed --> normal distribution

qreference(residuals(model_chlb)) # normal distribution assumed

drop1(model_chlb, test = 'F')

anova(model_chlb)

chlb_inter <- lsmeans(model_chlb, ~ treatment)
chlb_inter_dunn1 <- contrast(chlb_inter, 'pairwise')
(chlb_inter_dunn_sum1 <- summary(chlb_inter_dunn1, infer = TRUE, adjust = 'mvt'))
```

No significant differences between chemical treatments in chlorophyll b concentrations.

### Model chlc

```{r}
model_chlc <- lm(chlc ~ treatment, data = biofilm_endpoints)

par(mfrow = c(2, 2))

plot(model_chlc) # point cloud looks evenly distributed --> normal distribution

qreference(residuals(model_chlc)) # normal distribution assumed

drop1(model_chlc, test = 'F')

anova(model_chlc)

chlc_inter <- lsmeans(model_chlc, ~ treatment)
chlc_inter_dunn1 <- contrast(chlc_inter, 'pairwise')
(chlc_inter_dunn_sum1 <- summary(chlc_inter_dunn1, infer = TRUE, adjust = 'mvt'))
```
No significant differences between chemical treatments in chlorophyll c concentrations.

### Graphs chlorophyll a, b + c

```{r}
# chl a
chla_plot <-
  ggplot(data = biofilm_endpoints,
         mapping = aes(x = factor(
           treatment, levels = c("NC", "Pro", "Cip", "Mix")
         ), y = chla)) +
  # Mean
  stat_summary(
    fun = mean,
    geom = "point",
    aes(fill = treatment),
    shape = 21,
    size = 5,
    color = "grey20",
    stroke = 0.5
  ) +
  
  # 95% CI
  # Black outline (thicker, underneath)
  stat_summary(
    fun.data = mean_cl_normal,
    geom = "errorbar",
    aes(group = treatment),
    width = 0.2,
    linewidth = 1,
    color = "grey30"
  ) +
  
  # Colored CI on top
  stat_summary(
    fun.data = mean_cl_normal,
    geom = "errorbar",
    aes(color = treatment),
    width = 0.2,
    linewidth = 0.8
  ) +
  
  # Jittered
  geom_point(
    aes(color = treatment),
    position = position_jitter(width = 0.25, height = 0),
    alpha = 0.8,
    size = 2
  ) +
  
  # Color scales
  scale_fill_manual(values = c(
    "NC" = "#009E73",
    "Pro" = "#E69F00",
    "Cip" = "#56B4E9",
    "Mix" = "#999999"
  )) +
  scale_color_manual(values = c(
    "NC" = "#009E73",
    "Pro" = "#E69F00",
    "Cip" = "#56B4E9",
    "Mix" = "#999999"
  )) +
  guides(fill = "none") +
  
  # Theme and labels
  labs(x = NULL,
       y = expression(bold("Chl a (µg/cm" ^ "2" * ")")),
       color = "") +
  scale_x_discrete(labels = NULL) +
  theme_minimal(base_size = 11) +
  theme(
    legend.position = "none",
    panel.background = element_rect(fill = "grey100"),
    axis.title.y = element_text(
      size = 13,
      face = "bold",
      colour = "grey30"
  ),
    axis.text.x = element_text(size = 12),  
    axis.text.y = element_text(size = 12),
    panel.border = element_rect(fill = NA, colour = "grey30"),
    panel.grid.major = element_line(colour = "grey85", size = 0.4),
    panel.grid.minor = element_line(colour = "grey95", size = 0.2)
  ) 

# Chlorophyll b
chlb_plot <-
  ggplot(data = biofilm_endpoints,
         mapping = aes(x = factor(
           treatment, levels = c("NC", "Pro", "Cip", "Mix")
         ), y = chlb)) +
  # Mean
  stat_summary(
    fun = mean,
    geom = "point",
    aes(fill = treatment),
    shape = 21,
    size = 5,
    color = "grey20",
    stroke = 0.5
  ) +
  
  # 95% CI
  # Black outline (thicker, underneath)
  stat_summary(
    fun.data = mean_cl_normal,
    geom = "errorbar",
    aes(group = treatment),
    width = 0.2,
    linewidth = 1,
    color = "grey30"
  ) +
  
  # Colored CI on top
  stat_summary(
    fun.data = mean_cl_normal,
    geom = "errorbar",
    aes(color = treatment),
    width = 0.2,
    linewidth = 0.8
  ) +
  
  # Jittered
  geom_point(
    aes(color = treatment),
    position = position_jitter(width = 0.25, height = 0),
    alpha = 0.8,
    size = 2
  ) +
  
  # Color scales
  scale_fill_manual(values = c(
    "NC" = "#009E73",
    "Pro" = "#E69F00",
    "Cip" = "#56B4E9",
    "Mix" = "#999999"
  )) +
  scale_color_manual(values = c(
    "NC" = "#009E73",
    "Pro" = "#E69F00",
    "Cip" = "#56B4E9",
    "Mix" = "#999999"
  )) +
  guides(fill = "none") +
  
  # Theme and labels
  labs(x = NULL,
       y = expression(bold("Chl b (µg/cm" ^ "2" * ")")),
       color = "") +
  scale_x_discrete(labels = NULL) +
  theme_minimal(base_size = 11) +
  theme(
    legend.position = "none",
    panel.background = element_rect(fill = "grey100"),
    axis.title = element_text(
      size = 13,
      face = "bold",
      colour = "grey30"
    ),
    axis.text.x = element_text(size = 12),  
    axis.text.y = element_text(size = 12),
    panel.border = element_rect(fill = NA, colour = "grey30"),
    panel.grid.major = element_line(colour = "grey85", size = 0.4),
    panel.grid.minor = element_line(colour = "grey95", size = 0.2)
  ) 

# Chlorophyll c
chlc_plot <-
  ggplot(data = biofilm_endpoints,
         mapping = aes(x = factor(
           treatment, levels = c("NC", "Pro", "Cip", "Mix")
         ), y = chlc)) +
  # Mean
  stat_summary(
    fun = mean,
    geom = "point",
    aes(fill = treatment),
    shape = 21,
    size = 5,
    color = "grey20",
    stroke = 0.5
  ) +
  
  # 95% CI
  # Black outline (thicker, underneath)
  stat_summary(
    fun.data = mean_cl_normal,
    geom = "errorbar",
    aes(group = treatment),
    width = 0.2,
    linewidth = 1,
    color = "grey30"
  ) +
  
  # Colored CI on top
  stat_summary(
    fun.data = mean_cl_normal,
    geom = "errorbar",
    aes(color = treatment),
    width = 0.2,
    linewidth = 0.8
  ) +
  
  # Jittered
  geom_point(
    aes(color = treatment),
    position = position_jitter(width = 0.25, height = 0),
    alpha = 0.8,
    size = 2
  ) +
  
  # Color scales
  scale_fill_manual(values = c(
    "NC" = "#009E73",
    "Pro" = "#E69F00",
    "Cip" = "#56B4E9",
    "Mix" = "#999999"
  )) +
  scale_color_manual(values = c(
    "NC" = "#009E73",
    "Pro" = "#E69F00",
    "Cip" = "#56B4E9",
    "Mix" = "#999999"
  )) +
  guides(fill = "none") +
  
  # Theme and labels
  labs(x = "Treatment",
       y = expression(bold("Chl c (µg/cm" ^ "2" * ")")),
       color = "") +
  scale_x_discrete(labels = c("Control", "Propyzamide", "Ciprofloxacin", "Mixture")) +
  theme_minimal(base_size = 11) +
  theme(
    legend.position = "none",
    panel.background = element_rect(fill = "grey100"),
    axis.title = element_text(
      size = 13,
      face = "bold",
      colour = "grey30"
    ),
    axis.text.x = element_text(size = 11),  
    axis.text.y = element_text(size = 12),
    panel.border = element_rect(fill = NA, colour = "grey30"),
    panel.grid.major = element_line(colour = "grey85", size = 0.4),
    panel.grid.minor = element_line(colour = "grey95", size = 0.2)
  ) 

# combine and save with AFDW plot later
```

## AFDW 

Calculate mean, sd and 95% CI AFDW of each treatment.

```{r}
mean_afdw <- biofilm_endpoints %>%
  group_by(treatment) %>%
  dplyr::summarize(
    mean_afdw = mean(afdw, na.rm = TRUE),
    n = sum(!is.na(afdw)),
    se = sd(afdw, na.rm = TRUE) / sqrt(n),
    t_crit = qt(0.975, df = n - 1),
    ci_lower = mean_afdw - t_crit * se,
    ci_upper = mean_afdw + t_crit * se,
    .groups = "drop"
  ) %>%
  dplyr::select(treatment, mean_afdw, ci_lower, ci_upper, se)

print(mean_afdw)
```

### Histogram AFDW after colonisation

```{r}
hist(biofilm_endpoints $ afdw) 
skewness(biofilm_endpoints $ afdw, na.rm = TRUE) 
shapiro.test(biofilm_endpoints $ afdw)
```

Not normal distributed, right skewed but skewness \<1.5 therefore no data transformation for model.

### Model AFDW 

```{r}
model_afdw <- lm(afdw ~ treatment, data = biofilm_endpoints)

par(mfrow = c(2, 2))

plot(model_afdw) # not perfectly even distributed, but ok

qreference(residuals(model_afdw)) # normal distribution assumed

drop1(model_afdw, test = 'F')

anova(model_afdw)

afdw_inter <- lsmeans(model_afdw, ~ treatment)

afdw_inter_dunn1 <- contrast(afdw_inter, 'pairwise')

(afdw_inter_dunn_sum1 <- summary(afdw_inter_dunn1, infer = TRUE, adjust = 'mvt'))
```

No significant differences for treatment variable of total biofilm biomass.

### Boxplot AFDW 

```{r}
afdw_boxplot <-
  ggplot(data = biofilm_endpoints,
         mapping = aes(x = factor(
           treatment, levels = c("NC", "Pro", "Cip", "Mix")
         ), y = afdw)) +
  # Mean
  stat_summary(
    fun = mean,
    geom = "point",
    aes(fill = treatment),
    shape = 21,
    size = 5,
    color = "grey20",
    stroke = 0.5
  ) +
  
  # 95% CI
  # Black outline (thicker, underneath)
  stat_summary(
    fun.data = mean_cl_normal,
    geom = "errorbar",
    aes(group = treatment),
    width = 0.2,
    linewidth = 1,
    color = "grey30"
  ) +
  
  # Colored CI on top
  stat_summary(
    fun.data = mean_cl_normal,
    geom = "errorbar",
    aes(color = treatment),
    width = 0.2,
    linewidth = 0.8
  ) +
  
  # Jittered
  geom_point(
    aes(color = treatment),
    position = position_jitter(width = 0.25, height = 0),
    alpha = 0.8,
    size = 2
  ) +
  
  # Color scales
  scale_fill_manual(values = c(
    "NC" = "#009E73",
    "Pro" = "#E69F00",
    "Cip" = "#56B4E9",
    "Mix" = "#999999"
  )) +
  scale_color_manual(values = c(
    "NC" = "#009E73",
    "Pro" = "#E69F00",
    "Cip" = "#56B4E9",
    "Mix" = "#999999"
  )) +
  guides(fill = "none") +
  
  # Theme and labels
  labs(x = "Treatment",
       y = expression(bold("AFDW (mg/cm"^"2"*")")),
       color = "") +
  scale_x_discrete(labels = c("Control", "Propyzamide", "Ciprofloxacin", "Mixture")) +
  theme_minimal(base_size = 11) +
  theme(
    legend.position = "none",
    panel.background = element_rect(fill = "grey100"),
    axis.title = element_text(
      size = 13,
      face = "bold",
      colour = "grey30"
    ),
    axis.text.x = element_text(size = 11),  
    axis.text.y = element_text(size = 12),
    panel.border = element_rect(fill = NA, colour = "grey30"),
    panel.grid.major = element_line(colour = "grey85", size = 0.4),
    panel.grid.minor = element_line(colour = "grey95", size = 0.2)
  ) 


afdw_boxplot 

# Now combine with chlorophyll plots
combined_plot <- chla_plot / chlb_plot / chlc_plot / afdw_boxplot + 
  plot_layout(ncol = 2) 

# save the plot
ggsave(
  "output/biofilm_endpoints.png",
  plot = combined_plot,
  width = 25,
  height = 30,
  units = "cm",
  dpi = 600
)

```

# Fatty acid composition biofilm

Statistics for fatty acid composition of the biofilm in absolute values (µg fatty acid / mg dry weight biofilm).

```{r}
# insert data
FA_Biofilm_absolute <-
  read.table(file.path("data\\FA_Biofilm_absolute.txt"),
             header = TRUE,
             sep = "\t")

View(FA_Biofilm_absolute)
```

## Total SAFA, MUFA & PUFA content

1\) Total SAFA content

```{r}
# total SAFA, MUFA, PUFA content; ba = "biofilm absolute"
# total SAFA content
# filter saturated fatty acids in new table
SAFA_ba <- FA_Biofilm_absolute %>%
    filter(fa %in% c("C12:0", "C13:0", "C14:0", "C15:0", "C16:0", "C18:0", "C20:0", "C22:0", "C24:0"))

# Sum SAFA per replicate and treatment
total_SAFA_per_replicate <- SAFA_ba %>%
  dplyr::group_by(replicate, treatment) %>%
  dplyr::summarize(sum_SAFA = sum(fa_conc), .groups = "drop")
print(total_SAFA_per_replicate)

# Statistical differences
hist(total_SAFA_per_replicate $ sum_SAFA) # normal distributed
shapiro.test(total_SAFA_per_replicate $ sum_SAFA) # normal distributed
leveneTest(sum_SAFA ~ treatment, data = total_SAFA_per_replicate) # homogenous
anova_SAFA <- aov(sum_SAFA ~ treatment, data = total_SAFA_per_replicate)
summary(anova_SAFA)
TukeyHSD(anova_SAFA)

# Calculate mean and 95% CI per treatment
mean_SAFA_ba <- total_SAFA_per_replicate %>%
  dplyr::group_by(treatment) %>%
  dplyr::summarize(
    mean_SAFA = mean(sum_SAFA, na.rm = TRUE),
    n = sum(!is.na(sum_SAFA)),
    se = sd(sum_SAFA, na.rm = TRUE) / sqrt(n),
    t_crit = qt(0.975, df = n - 1),
    ci_lower = mean_SAFA - t_crit * se,
    ci_upper = mean_SAFA + t_crit * se,
    .groups = "drop"
  ) %>%
  dplyr::select(treatment, mean_SAFA, ci_lower, ci_upper, se)

# Extract mean of control group for percental difference
control_mean_SAFA_ba <- mean_SAFA_ba %>%
  filter(treatment == "NC") %>%
  pull(mean_SAFA)

# Calculate percental difference from control
mean_SAFA_ba <- mean_SAFA_ba %>%
  mutate(percent_diff_from_NC = 100 * (mean_SAFA - control_mean_SAFA_ba) / control_mean_SAFA_ba)

print(mean_SAFA_ba)

```

2\) Total MUFA content

```{r}
# filter mono-unsaturated fatty acids in new table
MUFA_ba <- FA_Biofilm_absolute %>%
  filter(fa %in% c("C16:1", "C17:1", "C18:1", "C20:1","C22:1"))

# Sum MUFA per replicate and treatment
total_MUFA_per_replicate <- MUFA_ba %>%
  dplyr::group_by(replicate, treatment) %>%
  dplyr::summarize(sum_MUFA = sum(fa_conc), .groups = "drop")
print(total_MUFA_per_replicate)

# Statistical differences
hist(total_MUFA_per_replicate $ sum_MUFA) # normal distributed
shapiro.test(total_MUFA_per_replicate $ sum_MUFA) # normal distributed
leveneTest(sum_MUFA ~ treatment, data = total_MUFA_per_replicate) # homogenous
anova_MUFA <- aov(sum_MUFA ~ treatment, data = total_MUFA_per_replicate)
summary(anova_MUFA)
TukeyHSD(anova_MUFA)

# Calculate mean and 95% CI per treatment
mean_MUFA_ba <- total_MUFA_per_replicate %>%
  dplyr::group_by(treatment) %>%
  dplyr::summarize(
    mean_MUFA = mean(sum_MUFA, na.rm = TRUE),
    n = sum(!is.na(sum_MUFA)),
    se = sd(sum_MUFA, na.rm = TRUE) / sqrt(n),
    t_crit = qt(0.975, df = n - 1),
    ci_lower = mean_MUFA - t_crit * se,
    ci_upper = mean_MUFA + t_crit * se,
    .groups = "drop"
  ) %>%
  dplyr::select(treatment, mean_MUFA, ci_lower, ci_upper, se)

# Extract mean of control group for percental difference
control_mean_MUFA_ba <- mean_MUFA_ba %>%
  filter(treatment == "NC") %>%
  pull(mean_MUFA)

# Calculate percental difference from control
mean_MUFA_ba <- mean_MUFA_ba %>%
  mutate(percent_diff_from_NC = 100 * (mean_MUFA - control_mean_MUFA_ba) / control_mean_MUFA_ba)

print(mean_MUFA_ba)
```

3\) Total PUFA content

```{r}
# filter poly-unsaturated fatty acids in new table
PUFA_ba <- FA_Biofilm_absolute %>%
  filter(fa %in% c("C18:2(n-6c)", "C18:3(n-3)", "C18:3(n-6)", "C18:4(n-3)", "C20:2(n-6)", "C20:3(n-6)", "C20:4(n-6)", "C20:5(n-3)", "C22:6(n-3)"))
  
# Sum PUFA per replicate and treatment
total_PUFA_per_replicate <- PUFA_ba %>%
  dplyr::group_by(replicate, treatment) %>%
  dplyr::summarize(sum_PUFA = sum(fa_conc), .groups = "drop")
print(total_PUFA_per_replicate)

# Statistical differences
hist(total_PUFA_per_replicate $ sum_PUFA) # normal distributed
shapiro.test(total_PUFA_per_replicate $ sum_PUFA) # normal distributed
leveneTest(sum_PUFA ~ treatment, data = total_PUFA_per_replicate) # homogenous
anova_PUFA <- aov(sum_PUFA ~ treatment, data = total_PUFA_per_replicate)
summary(anova_PUFA)
TukeyHSD(anova_PUFA)

# Calculate mean and 95% CI per treatment
mean_PUFA_ba <- total_PUFA_per_replicate %>%
  dplyr::group_by(treatment) %>%
  dplyr::summarize(
    mean_PUFA = mean(sum_PUFA, na.rm = TRUE),
    n = sum(!is.na(sum_PUFA)),
    se = sd(sum_PUFA, na.rm = TRUE) / sqrt(n),
    t_crit = qt(0.975, df = n - 1),
    ci_lower = mean_PUFA - t_crit * se,
    ci_upper = mean_PUFA + t_crit * se,
    .groups = "drop"
  ) %>%
  dplyr::select(treatment, mean_PUFA, ci_lower, ci_upper, se)

# Extract mean of control group for percental difference
control_mean_PUFA_ba <- mean_PUFA_ba %>%
  filter(treatment == "NC") %>%
  pull(mean_PUFA)

# Calculate percental difference from control
mean_PUFA_ba <- mean_PUFA_ba %>%
  mutate(percent_diff_from_NC = 100 * (mean_PUFA - control_mean_PUFA_ba) / control_mean_PUFA_ba)

print(mean_PUFA_ba)
```

4\) Total FA content

```{r}
# total FA 
total_FA_biofilm <- bind_rows(total_SAFA_per_replicate, total_MUFA_per_replicate, total_PUFA_per_replicate)

total_FA_per_replicate <- total_FA_biofilm %>%
  dplyr::group_by(replicate, treatment) %>%
  dplyr::summarize(sum_FA = sum(sum_SAFA, sum_MUFA, sum_PUFA, na.rm = TRUE))
print(total_FA_per_replicate)

# statistical differences will be calculated by PERMANOVA

# Calculate mean and 95% CI per treatment
mean_FA_ba <- total_FA_per_replicate %>%
  dplyr::group_by(treatment) %>%
  dplyr::summarize(
    mean_FA = mean(sum_FA, na.rm = TRUE),
    n = sum(!is.na(sum_FA)),
    se = sd(sum_FA, na.rm = TRUE) / sqrt(n),
    t_crit = qt(0.975, df = n - 1),
    ci_lower = mean_FA - t_crit * se,
    ci_upper = mean_FA + t_crit * se,
    .groups = "drop"
  ) %>%
  dplyr::select(treatment, mean_FA, ci_lower, ci_upper, se)

# Extract mean of control group for percental difference
control_mean_FA_ba <- mean_FA_ba %>%
  filter(treatment == "NC") %>%
  pull(mean_FA)

# Calculate percental difference from control
mean_FA_ba <- mean_FA_ba %>%
  mutate(percent_diff_from_NC = 100 * (mean_FA - control_mean_FA_ba) / control_mean_FA_ba)

print(mean_FA_ba)
```

## Histogram FA biofilm

```{r}
hist(FA_Biofilm_absolute $ fa_conc) 
skewness(FA_Biofilm_absolute $ fa_conc, na.rm = TRUE) 
shapiro.test(FA_Biofilm_absolute $ fa_conc)
```

Biofilm fatty acid concentration not normal distributed and right skewed.

## NMDS FA biofilm

Change table to wide format:

```{r}
# change table to wide format
FA_Biofilm_absolute_wide_sqrt <-
  pivot_wider(FA_Biofilm_absolute,
              names_from = fa,
              values_from = fa_conc)

# square root transformation, to reduce discriminatory power of dominant fa groups
FA_Biofilm_absolute_wide_sqrt[, 4:37] <- sqrt(FA_Biofilm_absolute_wide_sqrt[, 4:37])

View(FA_Biofilm_absolute_wide_sqrt)

# Only keep columns/FAs with more than 5 non-zero values (fatty acids that have just been found in 5 replicates will be removed)
FA_Biofilm_absolute_wide_sqrt <-
  FA_Biofilm_absolute_wide_sqrt %>%
  dplyr::select(
    1:3,
    where(~sum(. !=0, na.rm = TRUE)>5)
  )
```

Perform metaMDS:

```{r}
# Rename and ease col-names for later NMDS plot
FA_Biofilm_absolute_wide_sqrt <- FA_Biofilm_absolute_wide_sqrt %>%
  rename(
    `12:0`     = `C12:0`,   
    `13:0`     = `C13:0`,
    `14:0`     = `C14:0`,
    `15:0`     = `C15:0`,
    `16:0`     = `C16:0`,
    `16:1`     = `C16:1`,
    `17:1`     = `C17:1`,
    `18:0`     = `C18:0`,
    `18:1`     = `C18:1`,
    `18:2n-6c` = `C18:2(n-6c)`,
    `18:3n-6`  = `C18:3(n-6)`,
    `18:3n-3`  = `C18:3(n-3)`,
    `18:4n-3`  = `C18:4(n-3)`,
    `20:0`     = `C20:0`,
    `20:1`     = `C20:1`,
    `20:2n-6`  = `C20:2(n-6)`,
    `20:4n-6`  = `C20:4(n-6)`,
    `20:5n-3`  = `C20:5(n-3)`,
    `22:0`     = `C22:0`,
    `22:1`     = `C22:1`,
    `22:6n-3`  = `C22:6(n-3)`,
    `24:0`     = `C24:0`
  )

# Remove columns 1:3 for metaMDS
FA_Biofilm_absolute_wide_sqrt <- FA_Biofilm_absolute_wide_sqrt[, -c(1, 2, 3)]

NMDS_Biofilm_absolute <-
  metaMDS(
    FA_Biofilm_absolute_wide_sqrt,
    autotransform = FALSE,
    k = 2 ,
    distance = "bray",
    na.rm = TRUE
  )

NMDS_Biofilm_absolute$stress
```

Create fa vectors:

```{r}
# separation in SAFA; MUFA & PUFA to build vectors in NMDS
# "ba" = biofilm absolute
FA_variables.SAFAs_ba <-
  envfit(
    NMDS_Biofilm_absolute,
    subset(
      FA_Biofilm_absolute_wide_sqrt,
      select = c(
        "12:0",
        "13:0",
        "14:0",
        "15:0",
        "16:0",
        "18:0",
        "20:0",
        "22:0",
        "24:0"
      )
    )
    ,
    permutations = 999,
    na.rm = TRUE
  )

FA_variables.MUFAs_ba <-
  envfit(
    NMDS_Biofilm_absolute,
    subset(
      FA_Biofilm_absolute_wide_sqrt,
      select = c(
        "16:1",
        "17:1",
        "18:1",
        "20:1",
        "22:1"
      )
    )
    ,
    permutations = 999,
    na.rm = TRUE
  )

FA_variables.PUFAs_ba <-
  envfit(
    NMDS_Biofilm_absolute,
    subset(
      FA_Biofilm_absolute_wide_sqrt,
      select = c(
        "18:2n-6c",
        "18:3n-6",
        "18:3n-3",
        "18:4n-3",
        "20:2n-6",
        "20:4n-6",
        "20:5n-3",
        "22:6n-3"
      )
    )
    ,
    permutations = 999,
    na.rm = TRUE
  )

# Extract NMDS scores
data.scores_biofilm_absolute = as.data.frame(scores(NMDS_Biofilm_absolute)$sites)

Treatments <-
  FA_Biofilm_absolute$treatment[FA_Biofilm_absolute$fa == "C16:0"]

# add treatment column to NMDS scores
data.scores_biofilm_absolute$Treatment = Treatments

# vectors for SAFA, MUFA and PUFA
SAFAs_coord_cont_ba = as.data.frame(scores(FA_variables.SAFAs_ba, "vectors")) * ordiArrowMul(FA_variables.SAFAs_ba)
MUFAs_coord_cont_ba = as.data.frame(scores(FA_variables.MUFAs_ba, "vectors")) * ordiArrowMul(FA_variables.MUFAs_ba)
PUFAs_coord_cont_ba = as.data.frame(scores(FA_variables.PUFAs_ba, "vectors")) * ordiArrowMul(FA_variables.PUFAs_ba)

```

Plot NMDS:

```{r}
# Reorder treatment variable for the plot
data.scores_biofilm_absolute$Treatment <- factor(
  data.scores_biofilm_absolute$Treatment,
  levels = c("NC", "Pro", "Cip", "Mix")
)

# Compute convex hull for each Treatment
polygon_data_biofilm_treatment <- data.scores_biofilm_absolute %>%
  group_by(Treatment) %>%
  summarise(hull_points = list(data.frame(
    NMDS1 = NMDS1[chull(NMDS1, NMDS2)],
    NMDS2 = NMDS2[chull(NMDS1, NMDS2)]
  ))) %>%
  unnest(hull_points)

# plot
absol_biofilm_with_vectors = ggplot(data = data.scores_biofilm_absolute, aes(x = NMDS1, y = NMDS2)) +
  geom_polygon(
    data = polygon_data_biofilm_treatment,
    aes(
      x = NMDS1,
      y = NMDS2,
      group = Treatment,
      fill = Treatment
    ),
    alpha = 0.2,
    show.legend = FALSE
  ) +
  geom_point(aes(shape = Treatment, color = Treatment),
             size = 5,
             alpha = 1,
             show.legend = TRUE) +
  scale_shape_manual(values = c(
    "NC" = 16,
    "Pro" = 17,
    "Cip" = 18,
    "Mix" = 15
  )) +
  scale_color_manual(values = c(
    "NC" = "#009E73",
    "Pro" = "#E69F00",
    "Cip" = "#56B4E9",
    "Mix" = "#999999"
  )) +
  scale_fill_manual(values = c(
    "NC" = "#009E73",
    "Pro" = "#E69F00",
    "Cip" = "#56B4E9",
    "Mix" = "#999999"
  ))  +
  labs(color = "Treatment") +
  geom_segment(
    data = PUFAs_coord_cont_ba,
    aes(
      x = 0,
      y = 0,
      xend = NMDS1,
      yend = NMDS2
    ),
    linewidth = 0.5,
    alpha = 0.2,
    colour = "black"
  ) +
  geom_segment(
    data = MUFAs_coord_cont_ba,
    aes(
      x = 0,
      y = 0,
      xend = NMDS1,
      yend = NMDS2
    ),
    linewidth = 0.5,
    alpha = 0.2,
    colour = "black"
  ) +
  geom_segment(
    data = SAFAs_coord_cont_ba,
    aes(
      x = 0,
      y = 0,
      xend = NMDS1,
      yend = NMDS2
    ),
    linewidth = 0.5,
    alpha = 0.2,
    colour = "black"
  ) +
  geom_text_repel(
    data = PUFAs_coord_cont_ba,
    aes(
      x = NMDS1,
      y = NMDS2,
      label = row.names(PUFAs_coord_cont_ba)
    ),
    colour = "black",
    size = 2.5
  ) +
  geom_text_repel(
    data = MUFAs_coord_cont_ba,
    aes(
      x = NMDS1,
      y = NMDS2,
      label = row.names(MUFAs_coord_cont_ba)
    ),
    colour = "#996600",
    size = 2.5
  ) +
  geom_text_repel(
    data = SAFAs_coord_cont_ba,
    aes(
      x = NMDS1,
      y = NMDS2,
      label = row.names(SAFAs_coord_cont_ba)
    ),
    colour = "#0000FF",
    size = 2.5
  ) +
  theme_bw() +
  theme(
    axis.title = element_text(
      size = 13,
      face = "bold",
      colour = "grey30"
    ),
    panel.background = element_blank(),
    panel.border = element_rect(fill = NA, colour = "grey30"),
    legend.key = element_blank(),
    legend.title = element_text(
      size = 11,
      face = "bold",
      colour = "grey30"
    ),
    legend.text = element_text(size = 10, colour = "grey30"),
    plot.margin = margin(
      t = 0.2,
      b = 0.2,
      l = 0.2,
      r = 0.2,
      unit = "cm"
    ),
    panel.grid.major = element_line(colour = "grey95", size = 0.4),
    panel.grid.minor = element_line(colour = "grey95", size = 0.2)
  )

absol_biofilm_with_vectors
# position of FA names manually adjusted in vector graphic with Inkscape
```

Save the plot as .png:

```{r}
ggsave(
  "output/NMDS_Biofilm_absolute.png",
  width = 16.5,
  height = 12,
  units = "cm",
  dpi = 600
)
```

Save the plot as vector graphic:

```{r}
ggsave(
  "output/NMDS_Biofilm_absolute.svg",
  device = "svg",
  width = 16.5,
  height = 16,
  units = "cm",
  dpi = 600
)
```

## PERMANOVA FA biofilm

Create distance matrix with fatty acids:

```{r}
FA_biofilm_dist <- vegdist(FA_Biofilm_absolute_wide_sqrt, method = 'bray')
```

Separate explanatory variables from fatty acid information:

```{r}
FA_biofilm_absolute_wide_permanova <-
  pivot_wider(FA_Biofilm_absolute,
              names_from = fa,
              values_from = fa_conc)

# square root transformation again, to reduce discriminatory power of dominant fa groups
FA_biofilm_absolute_wide_permanova[, 4:37] <- sqrt(FA_biofilm_absolute_wide_permanova[, 4:37])

# delete fa columns
FA_biofilm_absolute_wide_permanova <-
  FA_biofilm_absolute_wide_permanova[, c(1, 2, 3)]

FA_biofilm_absolute_wide_permanova$treatment <-
  as.factor(FA_biofilm_absolute_wide_permanova$treatment)
```

Perform PERMANOVA for treatments:

```{r}
# perform PERMANOVA for treatments
perm_FA_biofilm_treatment <- adonis2(FA_biofilm_dist ~ treatment, data = FA_biofilm_absolute_wide_permanova, permutations = 999)
print(perm_FA_biofilm_treatment)

```

No significant differences between biofilm treatments. Anyway pairwise comparisons (by: Martinez Arbizu, P. (2020). pairwiseAdonis: Pairwise multilevel comparison using adonis. R package version 0.4):

```{r}
pairwise_results_biofilm_treatment <- pairwise.adonis2(
  FA_biofilm_dist ~ treatment,
  data = FA_biofilm_absolute_wide_permanova,
  permutations = 99
)

# View the results
print(pairwise_results_biofilm_treatment)
```

## SIMPER FA biofilm

```{r}
# create extra table with treatments
biofilm_treatments_table <- FA_biofilm_absolute_wide_permanova

View(biofilm_treatments_table)

# Perform SIMPER
simper_FA_biofilm_treatments <-
  with(
    biofilm_treatments_table,
    simper(FA_Biofilm_absolute_wide_sqrt, treatment,  permutations = 999)
  )

summary(simper_FA_biofilm_treatments)
```

```{r}
simper_sum <- summary(simper_FA_biofilm_treatments)

# save results
out_dir <- "output"
dir.create(out_dir, showWarnings = FALSE)

safe_name <- function(x) gsub("[^A-Za-z0-9_]+", "_", x)

iwalk(simper_sum, function(x, nm) {
  if (is.null(x)) return(invisible(NULL))

  # SIMPER summary as a data.frame:
  # columns: average, sd, ratio, ava, avb, cumsum, p
  # rownames are the fatty acid IDs
  df <- as.data.frame(x)

  # keep rows up to the first that reaches/exceeds 70%
  idx <- which(df$cumsum >= 0.70)[1]
  if (is.na(idx)) idx <- nrow(df)
  out <- df[seq_len(idx), , drop = FALSE]

  # just keep 3 decimals
  out_write <- out
  num_cols <- vapply(out_write, is.numeric, logical(1))
  out_write[, num_cols] <- lapply(out_write[, num_cols, drop = FALSE],
                                  function(col) formatC(col, format = "f", digits = 3))

  file <- file.path(out_dir, paste0(safe_name(nm), "_top70.tsv"))
  # write row names so the TSV mirrors the SIMPER structure
  write.table(out_write, file, sep = "\t", quote = FALSE, col.names = NA)
})
```

```{r}
# Combine results into a data frame
simper_results_biofilm_treatments <- do.call(rbind, lapply(simper_FA_biofilm_treatments, function(x) {
  # Check if the element contains valid data
  if (!is.null(x$p)) {
    data.frame(
      fatty_acid = names(x$p),  
      p = as.numeric(x$p),      
      average = as.numeric(x$average),  
      sd = as.numeric(x$sd),
      ratio = as.numeric(x$ratio),
      cusum = as.numeric(x$cusum),
      stringsAsFactors = FALSE
    )
  } else {
    NULL  # Skip if no p-values
  }
}))

# Show only significant results
simper_results_biofilm_treatments %>%
  dplyr::filter(p <= 0.05) %>%
  dplyr::select(fatty_acid, p, average, sd, ratio, cusum)  
```

```{r}
# How often do they occur
simper_results_biofilm_treatments %>%
  dplyr::filter(p <= 0.05) %>%
  dplyr::count(fatty_acid) %>%
  dplyr::arrange(desc(n)) %>%
  print()
```

# Dryweight snails

Unit of dry weight = mg

```{r}
# insert data (Culture snails from in-house breeding)
dryweight_snails <-
  read.table(file.path("data\\dryweight_snails.txt"),
             header = TRUE, sep = "\t")

View(dryweight_snails)
# 3 snails from the 20 baseline (culture) snails were lost during the analyses.
```

## Calculate mean dryweight, sd and 95% CI of all treatment snails and the culture snails

```{r}
mean_dryweight_snails <- dryweight_snails %>%
  group_by(treatment) %>%
  dplyr::summarize(
    mean_dryweight_snails = mean(dryweight_mg, na.rm = TRUE),
    n = sum(!is.na(dryweight_mg)),
    se = sd(dryweight_mg, na.rm = TRUE) / sqrt(n),
    t_crit = qt(0.975, df = n - 1),
    ci_lower = mean_dryweight_snails - t_crit * se,
    ci_upper = mean_dryweight_snails + t_crit * se,
    .groups = "drop"
  ) %>%
  dplyr::select(treatment, mean_dryweight_snails, ci_lower, ci_upper, se)
  
print(mean_dryweight_snails)
```

Snails from feeding assay (fed with biofilm) on average heavier than snails from culture (fed with Tetraphyll flakes).

```{r}
#culture/baseline snails
culture_snails <- dryweight_snails %>%
  dplyr::filter(c(treatment == "Culture"))

hist(culture_snails $ dryweight_mg) 
skewness(culture_snails $ dryweight_mg, na.rm = TRUE) 
shapiro.test(culture_snails $ dryweight_mg)
# culture / baseline snails normal distributed

# calculate differences in final dry weight in comparison to culture snails
percentage_differences <- dryweight_snails %>%
  group_by(treatment) %>%
  dplyr::summarize(mean_dryweight = mean(dryweight_mg)) %>%
  mutate(percentage_differences = (mean_dryweight / mean_dryweight[treatment == "Culture"]) * 100 - 100)

percentage_differences
```

## Histograms

```{r}
# filter culture snails
filtered_dry_weight <- dryweight_snails %>%
  filter(!treatment %in% c("Culture"))

hist(filtered_dry_weight $ dryweight_mg) 
skewness(filtered_dry_weight $ dryweight_mg, na.rm = TRUE) 
shapiro.test(filtered_dry_weight $ dryweight_mg)
```

Soft tissue dry weights normal distributed.

## Model dry weight snails

```{r}
# explanatory variable treatment as factor
filtered_dry_weight $ treatment <- factor(filtered_dry_weight $ treatment)
  
model_snails <- lm(dryweight_mg ~ treatment, data = filtered_dry_weight)

plot(model_snails) # point cloud looks evenly distributed --> normal distribution

qreference(residuals(model_snails)) # normal distribution

anova(model_snails)
```

Explanatory variable "treatment" highly significant. 

```{r}
dw_snails_inter <- lsmeans(model_snails, ~ treatment)

dw_snails_inter_dunn1 <- contrast(dw_snails_inter, 'pairwise')

dw_snails_inter_dunn_sum1 <- summary(dw_snails_inter_dunn1, infer = TRUE, adjust = 'mvt')

dw_snails_inter_dunn1
```

Significant difference between dry weight of Cip and Mix snails compared to NC; almost significant difference (p=0.0564) between Pro treatment and NC.

## Boxplot dry weight snails

```{r, warning = FALSE}
# Plot
snails_boxplot <-
  ggplot(data = filtered_dry_weight,
         mapping = aes(x = factor(
           treatment, levels = c("NC", "Pro", "Cip", "Mix")
         ), y = dryweight_mg)) +
  # Mean
  stat_summary(
    fun = mean,
    geom = "point",
    aes(fill = treatment),
    shape = 21,
    size = 5,
    color = "grey20",
    stroke = 0.5
  ) +
  
  # 95% CI
  # Black outline (thicker, underneath)
  stat_summary(
    fun.data = mean_cl_normal,
    geom = "errorbar",
    aes(group = treatment),
    width = 0.2,
    linewidth = 1,
    color = "grey30"
  ) +
  
  # Colored CI on top
  stat_summary(
    fun.data = mean_cl_normal,
    geom = "errorbar",
    aes(color = treatment),
    width = 0.2,
    linewidth = 0.8
  ) +
  
  # Jittered
geom_point(
  aes(color = treatment),
  position = position_jitter(width = 0.25, height = 0),
  alpha = 0.8,
  size = 2
) +
  
  # Color scales
  scale_fill_manual(values = c(
    "NC" = "#009E73",
    "Pro" = "#E69F00",
    "Cip" = "#56B4E9",
    "Mix" = "#999999"
  )) +
  scale_color_manual(values = c(
    "NC" = "#009E73",
    "Pro" = "#E69F00",
    "Cip" = "#56B4E9",
    "Mix" = "#999999"
  )) +
  guides(fill = "none") +
  
  # Theme and labels
  labs(x = "Treatment", y = "Grazer Dry Weight (mg)", color = "") +
 # scale_y_continuous(limits = c(0, 1.1)) +
  scale_x_discrete(labels = c("Control", "Propyzamide", "Ciprofloxacin", "Mixture")) +
  theme_minimal(base_size = 11) +
  theme(
    legend.position = "none",
    panel.background = element_rect(fill = "grey100"),
    axis.title = element_text(
      size = 13,
      face = "bold",
      colour = "grey30"
    ),
    axis.text.x = element_text(size = 12),  
    axis.text.y = element_text(size = 12),
    panel.border = element_rect(fill = NA, colour = "grey30"),
    panel.grid.major = element_line(colour = "grey85", size = 0.4),
    panel.grid.minor = element_line(colour = "grey95", size = 0.2)
  ) +
  # Significance annotations
  annotate(
    "text",
    x = 1.5,
    y = max(filtered_dry_weight$dryweight_mg) + 0.07,
    label = "p=0.056",
    size = 4,
    color = "grey30"
  ) +
  geom_segment(
    aes(
      x = 1,
      xend = 2,
      y = max(dryweight_mg) + 0.05,
      yend = max(dryweight_mg) + 0.05
    ),
    lineend = "round",
    linewidth = 0.5,
    color = "grey30"
  ) +
  # Vertical tick on the left (x = 1)
  geom_segment(
    aes(
      x = 1,
      xend = 1,
      y = max(dryweight_mg) + 0.04,
      yend = max(dryweight_mg) + 0.05
    ),
    color = "grey30",
    linewidth = 0.5
  ) +
  # # Vertical tick on the right (x = 4)
  geom_segment(
    aes(
      x = 2,
      xend = 2,
      y = max(dryweight_mg) + 0.04,
      yend = max(dryweight_mg) + 0.05
    ),
    color = "grey30",
    linewidth = 0.5
  ) +
  annotate(
    "text",
    x = 2.5,
    y = max(filtered_dry_weight$dryweight_mg) + 0.17,
    label = "p=0.005**",
    size = 4,
    color = "grey30"
  ) +
  geom_segment(
    aes(
      x = 1,
      xend = 4,
      y = max(dryweight_mg) + 0.15,
      yend = max(dryweight_mg) + 0.15
    ),
    lineend = "round",
    linewidth = 0.5,
    color = "grey30"
  ) +
  geom_segment(
    aes(
      x = 1,
      xend = 1,
      y = max(dryweight_mg) + 0.14,
      yend = max(dryweight_mg) + 0.15
    ),
    color = "grey30",
    linewidth = 0.5
  ) +
  geom_segment(
    aes(
      x = 4,
      xend = 4,
      y = max(dryweight_mg) + 0.14,
      yend = max(dryweight_mg) + 0.15
    ),
    color = "grey30",
    linewidth = 0.5
  ) +
  annotate(
    "text",
    x = 2,
    y = max(filtered_dry_weight$dryweight_mg) + 0.12,
    label = "p=0.022*",
    size = 4,
    color = "grey30"
  ) +
  geom_segment(
    aes(
      x = 1,
      xend = 3,
      y = max(dryweight_mg) + 0.1,
      yend = max(dryweight_mg) + 0.1
    ),
    lineend = "round",
    linewidth = 0.5,
    color = "grey30"
  ) +
  geom_segment(
    aes(
      x = 1,
      xend = 1,
      y = max(dryweight_mg) + 0.09,
      yend = max(dryweight_mg) + 0.1
    ),
    color = "grey30",
    linewidth = 0.5
  ) +
  geom_segment(
    aes(
      x = 3,
      xend = 3,
      y = max(dryweight_mg) + 0.09,
      yend = max(dryweight_mg) + 0.1
    ),
    color = "grey30",
    linewidth = 0.5
  )


snails_boxplot
```

```{r, warning = FALSE}
# save the plot
ggsave(
  "output/snails_boxplot.png",
  width = 16.5,
  height = 12,
  units = "cm",
  dpi = 600
)
```

# Fatty acid composition snails

Statistics for fatty acid composition of snails in absolute values (µg fatty acid / mg dry weight snail).

```{r}
# insert data
FA_Snails_absolute <-
  read.table(file.path("data\\FA_Snails_absolute.txt"),
             header = TRUE, sep = "\t")

View(FA_Snails_absolute)
# Mix S_29 no data, fatty acid extraction did not work
```

## Total SAFA, MUFA & PUFA content

1\) Total SAFA content

```{r}
# total SAFA, MUFA, PUFA content; sa = "snail absolute" 
# total SAFA content 
# filter saturated fatty acids in new table 
SAFA_sa <- FA_Snails_absolute %>%
filter(fa %in% c("C12:0", "C13:0", "C14:0", "C15:0", "C16:0", "C18:0", "C20:0"))

# Sum SAFA per replicate and treatment
total_SAFA_per_snail <- SAFA_sa %>% 
  dplyr::group_by(replicate, treatment) %>% 
  dplyr::summarize(sum_SAFA = sum(fa_conc), .groups = "drop") 
print(total_SAFA_per_snail) 

# Statistical differences
hist(total_SAFA_per_snail $ sum_SAFA) # normal distributed
shapiro.test(total_SAFA_per_snail $ sum_SAFA) # normal distributed
leveneTest(sum_SAFA ~ treatment, data = total_SAFA_per_snail) # homogenous
anova_SAFA <- aov(sum_SAFA ~ treatment, data = total_SAFA_per_snail)
summary(anova_SAFA)
TukeyHSD(anova_SAFA)

# Calculate mean and 95% CI per treatment
mean_SAFA_sa <- total_SAFA_per_snail %>%
  dplyr::group_by(treatment) %>%
  dplyr::summarize(
    mean_SAFA = mean(sum_SAFA, na.rm = TRUE),
    n = sum(!is.na(sum_SAFA)),
    se = sd(sum_SAFA, na.rm = TRUE) / sqrt(n),
    t_crit = qt(0.975, df = n - 1),
    ci_lower = mean_SAFA - t_crit * se,
    ci_upper = mean_SAFA + t_crit * se,
    .groups = "drop"
  ) %>%
  dplyr::select(treatment, mean_SAFA, ci_lower, ci_upper, se)

# Extract mean of control group for percental difference
control_mean_SAFA_sa <- mean_SAFA_sa %>%
  filter(treatment == "NC") %>%
  pull(mean_SAFA)

# Calculate percental difference from control
mean_SAFA_sa <- mean_SAFA_sa %>%
  mutate(percent_diff_from_NC = 100 * (mean_SAFA - control_mean_SAFA_sa) / control_mean_SAFA_sa)

print(mean_SAFA_sa)
```

2\) Total MUFA content

```{r}
# filter mono-unsaturated fatty acids in new table
MUFA_sa <- FA_Snails_absolute %>%
  filter(fa %in% c("C16:1", "C18:1", "C20:1", "C22:1"))

# Sum MUFA per replicate and treatment
total_MUFA_per_snail <- MUFA_sa %>% 
  dplyr::group_by(replicate, treatment) %>% 
  dplyr::summarize(sum_MUFA = sum(fa_conc), .groups = "drop") 
print(total_MUFA_per_snail) 

# Statistical differences
hist(total_MUFA_per_snail $ sum_MUFA) 
shapiro.test(total_MUFA_per_snail $ sum_MUFA) 
leveneTest(sum_MUFA ~ treatment, data = total_MUFA_per_snail)
pairwise.wilcox.test(total_MUFA_per_snail$sum_MUFA,
                     total_MUFA_per_snail$treatment,
                     p.adjust.method = "BH")

# Calculate mean and 95% CI per treatment
mean_MUFA_sa <- total_MUFA_per_snail %>%
  dplyr::group_by(treatment) %>%
  dplyr::summarize(
    mean_MUFA = mean(sum_MUFA, na.rm = TRUE),
    n = sum(!is.na(sum_MUFA)),
    se = sd(sum_MUFA, na.rm = TRUE) / sqrt(n),
    t_crit = qt(0.975, df = n - 1),
    ci_lower = mean_MUFA - t_crit * se,
    ci_upper = mean_MUFA + t_crit * se,
    .groups = "drop"
  ) %>%
  dplyr::select(treatment, mean_MUFA, ci_lower, ci_upper, se)

# Extract mean of control group for percental difference
control_mean_MUFA_sa <- mean_MUFA_sa %>%
  filter(treatment == "NC") %>%
  pull(mean_MUFA)

# Calculate percental difference from control
mean_MUFA_sa <- mean_MUFA_sa %>%
  mutate(percent_diff_from_NC = 100 * (mean_MUFA - control_mean_MUFA_sa) / control_mean_MUFA_sa)

print(mean_MUFA_sa)
```

3\) Total PUFA content

```{r}
# filter poly-unsaturated fatty acids in new table
PUFA_sa <- FA_Snails_absolute %>%
  filter(fa %in% c("C18:2(n-6c)", "C18:3(n-3)", "C18:3(n-6)", "C18:4(n-3)", "C20:2(n-6)", "C20:3(n-6)", "C20:4(n-6)", "C20:5(n-3)", "C22:5(n-3)", "C22:6(n-3)"))
  
# Sum PUFA per replicate and treatment
total_PUFA_per_snail <- PUFA_sa %>% 
  dplyr::group_by(replicate, treatment) %>% 
  dplyr::summarize(sum_PUFA = sum(fa_conc), .groups = "drop") 
print(total_PUFA_per_snail) 

# statistics
hist(total_PUFA_per_snail $ sum_PUFA)
shapiro.test(total_PUFA_per_snail $ sum_PUFA) 
leveneTest(sum_PUFA ~ treatment, data = total_PUFA_per_snail)
anova_PUFA <- aov(sum_PUFA ~ treatment, data = total_PUFA_per_snail)
summary(anova_PUFA)
TukeyHSD(anova_PUFA)

# Calculate mean and 95% CI per treatment
mean_PUFA_sa <- total_PUFA_per_snail %>%
  dplyr::group_by(treatment) %>%
  dplyr::summarize(
    mean_PUFA = mean(sum_PUFA, na.rm = TRUE),
    n = sum(!is.na(sum_PUFA)),
    se = sd(sum_PUFA, na.rm = TRUE) / sqrt(n),
    t_crit = qt(0.975, df = n - 1),
    ci_lower = mean_PUFA - t_crit * se,
    ci_upper = mean_PUFA + t_crit * se,
    .groups = "drop"
  ) %>%
  dplyr::select(treatment, mean_PUFA, ci_lower, ci_upper, se)

# Extract mean of control group for percental difference
control_mean_PUFA_sa <- mean_PUFA_sa %>%
  filter(treatment == "NC") %>%
  pull(mean_PUFA)

# Calculate percental difference from control
mean_PUFA_sa <- mean_PUFA_sa %>%
  mutate(percent_diff_from_NC = 100 * (mean_PUFA - control_mean_PUFA_sa) / control_mean_PUFA_sa)

print(mean_PUFA_sa)
```

4\) Total FA per snail

```{r}
# total FA 
total_FA_snails <- bind_rows(total_SAFA_per_snail, total_MUFA_per_snail, total_PUFA_per_snail)

total_FA_per_snail <- total_FA_snails %>%
  dplyr::group_by(replicate, treatment) %>%
  dplyr::summarize(sum_FA = sum(sum_SAFA, sum_MUFA, sum_PUFA, na.rm = TRUE))
print(total_FA_per_snail)

# statistical differences will be calculated by PERMANOVA

# Calculate mean and 95% CI per treatment
mean_FA_sa <- total_FA_per_snail %>%
  dplyr::group_by(treatment) %>%
  dplyr::summarize(
    mean_FA = mean(sum_FA, na.rm = TRUE),
    n = sum(!is.na(sum_FA)),
    se = sd(sum_FA, na.rm = TRUE) / sqrt(n),
    t_crit = qt(0.975, df = n - 1),
    ci_lower = mean_FA - t_crit * se,
    ci_upper = mean_FA + t_crit * se,
    .groups = "drop"
  ) %>%
  dplyr::select(treatment, mean_FA, ci_lower, ci_upper, se)

# Extract mean of control group for percental difference
control_mean_FA_sa <- mean_FA_sa %>%
  filter(treatment == "NC") %>%
  pull(mean_FA)

# Calculate percental difference from control
mean_FA_sa <- mean_FA_sa %>%
  mutate(percent_diff_from_NC = 100 * (mean_FA - control_mean_FA_sa) / control_mean_FA_sa)

print(mean_FA_sa)
```

## Histogram FA snails

```{r}
hist(FA_Snails_absolute $ fa_conc) 
skewness(FA_Snails_absolute $ fa_conc, na.rm = TRUE) 
shapiro.test(FA_Snails_absolute $ fa_conc)
```

Fatty acid concentration not normal distributed and right skewed.

## NMDS FA snails

Change table to wide format:

```{r}
# change table to wide format (sqrt for square root transformation in next step)
FA_Snails_absolute_wide_sqrt <-
  pivot_wider(FA_Snails_absolute,
              names_from = fa,
              values_from = fa_conc)

# square root transformation, to reduce discriminatory power of dominant FA groups
FA_Snails_absolute_wide_sqrt[, 3:23] <- sqrt(FA_Snails_absolute_wide_sqrt[, 3:23])

# Only keep columns/FAs with more than 5 non-zero values (fatty acids that have just been found in 5 replicates will be removed)
FA_Snails_absolute_wide_sqrt <-
  FA_Snails_absolute_wide_sqrt %>%
  dplyr::select(
    1:2,
    where(~sum(. !=0, na.rm = TRUE)>5)
  )
```

Perform metaMDS:

```{r}
# Remove first and second column for metaMDS
FA_Snails_absolute_wide_sqrt <- FA_Snails_absolute_wide_sqrt[, -c(1, 2)]

# Rename (- C, - () ) fatty acids for later plot
FA_Snails_absolute_wide_sqrt <- FA_Snails_absolute_wide_sqrt %>%
  rename(
    `12:0`     = `C12:0`,   
    `13:0`     = `C13:0`,
    `14:0`     = `C14:0`,
    `15:0`     = `C15:0`,
    `16:0`     = `C16:0`,
    `16:1`     = `C16:1`,
    `18:0`     = `C18:0`,
    `18:1`     = `C18:1`,
    `18:2n-6c` = `C18:2(n-6c)`,
    `18:3n-6`  = `C18:3(n-6)`,
    `18:3n-3`  = `C18:3(n-3)`,
    `18:4n-3`  = `C18:4(n-3)`,
    `20:0`     = `C20:0`,
    `20:1`     = `C20:1`,
    `20:2n-6`  = `C20:2(n-6)`,
    `20:3n-6`  = `C20:3(n-6)`,
    `20:4n-6`  = `C20:4(n-6)`,
    `20:5n-3`  = `C20:5(n-3)`,
    `22:1`     = `C22:1`,
    `22:5n-3`  = `C22:5(n-3)`,
    `22:6n-3`  = `C22:6(n-3)`
)


NMDS_Snails_absolute <-
  metaMDS(
    FA_Snails_absolute_wide_sqrt,
    autotransform = FALSE,
    k = 2 ,
    distance = "bray",
    na.rm = TRUE
  )

NMDS_Snails_absolute$stress
```

Create fa vectors:

```{r}
# separation in SAFA; MUFA & PUFA to build vectors in NMDS
# "sa" = snail absolute
FA_variables.SAFAs_sa <-
  envfit(
    NMDS_Snails_absolute,
    subset(
      FA_Snails_absolute_wide_sqrt,
      select = c(
        "12:0",
        "13:0",
        "14:0",
        "15:0",
        "16:0",
        "18:0",
        "20:0"
      )
    )
    ,
    permutations = 999,
    na.rm = TRUE
  )

FA_variables.MUFAs_sa <-
  envfit(
    NMDS_Snails_absolute,
    subset(
      FA_Snails_absolute_wide_sqrt,
      select = c(
        "16:1",
        "18:1",
        "20:1",
        "22:1"
      )
    )
    ,
    permutations = 999,
    na.rm = TRUE
  )

FA_variables.PUFAs_sa <-
  envfit(
    NMDS_Snails_absolute,
    subset(
      FA_Snails_absolute_wide_sqrt,
      select = c(
        "18:2n-6c",
        "18:3n-6",
        "18:3n-3",
        "18:4n-3",
        "20:2n-6",
        "20:3n-6",
        "20:4n-6",
        "20:5n-3",
        "22:5n-3",
        "22:6n-3"
      )
    )
    ,
    permutations = 999,
    na.rm = TRUE
  )

# Extract NMDS scores 
data.scores_snail_absolute = as.data.frame(scores(NMDS_Snails_absolute)$sites)

Treatments <-
  FA_Snails_absolute $ treatment[FA_Snails_absolute $ fa == "C16:0"]

# add treatment column to NMDS scores
data.scores_snail_absolute$Treatment = Treatments

# vectors for SAFA, MUFA and PUFA
SAFAs_coord_cont_sa = as.data.frame(scores(FA_variables.SAFAs_sa, "vectors")) * ordiArrowMul(FA_variables.SAFAs_sa)
MUFAs_coord_cont_sa = as.data.frame(scores(FA_variables.MUFAs_sa, "vectors")) * ordiArrowMul(FA_variables.MUFAs_sa)
PUFAs_coord_cont_sa = as.data.frame(scores(FA_variables.PUFAs_sa, "vectors")) * ordiArrowMul(FA_variables.PUFAs_sa)
```

Plot NMDS:

```{r, warning=FALSE}
# For the plot: reorder treatment variable in data.scores_snail_absolute
data.scores_snail_absolute$Treatment <- factor(
  data.scores_snail_absolute$Treatment,
  levels = c("NC", "Pro", "Cip", "Mix")
)

# Compute convex hull for each Treatment
polygon_data_snails <- data.scores_snail_absolute %>%
  group_by(Treatment) %>%
  summarise(hull_points = list(data.frame(
    NMDS1 = NMDS1[chull(NMDS1, NMDS2)],
    NMDS2 = NMDS2[chull(NMDS1, NMDS2)]
  ))) %>%
  unnest(hull_points)

# Function to rescale coordinates
rescale_vectors <- function(df, scale_factor = 1) {
  df$length <- sqrt(df$NMDS1^2 + df$NMDS2^2)
  max_length <- max(df$length)
  df$NMDS1 <- df$NMDS1 / max_length * scale_factor
  df$NMDS2 <- df$NMDS2 / max_length * scale_factor
  df$length <- NULL
  return(df)
}

# Apply to each group with same scale_factor (e.g. 0.2)
PUFAs_coord_cont_sa <- rescale_vectors(PUFAs_coord_cont_sa, 0.4)
MUFAs_coord_cont_sa <- rescale_vectors(MUFAs_coord_cont_sa, 0.4)
SAFAs_coord_cont_sa <- rescale_vectors(SAFAs_coord_cont_sa, 0.4)

# plot
absol_snail_with_vectors <- ggplot(data = data.scores_snail_absolute, aes(x = NMDS1, y = NMDS2)) +
    geom_polygon(
    data = polygon_data_snails,
    aes(x = NMDS1, y = NMDS2, group = Treatment, fill = Treatment),
    alpha = 0.2,
    show.legend = FALSE
    ) +
  geom_point(
    aes(shape = Treatment, color = Treatment),
    size = 5,
    alpha = 1,
    show.legend = TRUE
  ) +
  scale_shape_manual(values = c("NC" = 16, "Pro" = 17, "Cip" = 18, "Mix" = 15)) +
  scale_color_manual(values = c("NC" = "#009E73", "Pro" = "#E69F00", "Cip" = "#56B4E9", "Mix" = "#999999")) +
  scale_fill_manual(values = c("NC" = "#009E73", "Pro" = "#E69F00", "Cip" = "#56B4E9", "Mix" = "#999999")) + 
  geom_segment(aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2), 
               data = PUFAs_coord_cont_sa, linewidth = 0.5, alpha = 0.2, colour = "black") +
  geom_segment(aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2), 
               data = MUFAs_coord_cont_sa, linewidth = 0.5, alpha = 0.2, colour = "black") +
  geom_segment(aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2), 
               data = SAFAs_coord_cont_sa, linewidth = 0.5, alpha = 0.2, colour = "black") +
  geom_text_repel(
    data = PUFAs_coord_cont_sa, 
    aes(x = NMDS1, y = NMDS2, label = row.names(PUFAs_coord_cont_sa)), 
    colour = "black",
    size = 2.5) +
  geom_text_repel(
    data = MUFAs_coord_cont_sa, 
    aes(x = NMDS1, y = NMDS2, label = row.names(MUFAs_coord_cont_sa)), 
    colour = "#996600",
    size = 2.5) +
  geom_text_repel(
    data = SAFAs_coord_cont_sa, 
    aes(x = NMDS1, y = NMDS2, label = row.names(SAFAs_coord_cont_sa)), 
    colour = "#0000FF",
    size = 2.5) +
  theme_bw() +
  theme(
    axis.title = element_text(
      size = 13,
      face = "bold",
      colour = "grey30"
    ),
    panel.background = element_blank(),
    panel.border = element_rect(fill = NA, colour = "grey30"),
    legend.key = element_blank(),
    legend.title = element_text(
      size = 11,
      face = "bold",
      colour = "grey30"
    ),
    legend.text = element_text(size = 10, colour = "grey30"),
    panel.grid.major = element_line(colour = "grey95", size = 0.4),
    panel.grid.minor = element_line(colour = "grey95", size = 0.2)
  ) 

absol_snail_with_vectors
# position of FA names manually adjusted in vector graphic
```

Save the plot as .png:

```{r}
ggsave(
  "output/NMDS_snails_absolute_vectors.png",
  width = 16.5,
  height = 12,
  units = "cm",
  dpi = 600
)
```

Save the plot as vector graphic (minor adaptions for better readability with Inkscape):

```{r}
ggsave(
  "output/NMDS_snails_absolute_vectors.svg",
  device = "svg",
  width = 16.5,
  height = 16,
  units = "cm",
  dpi = 600
)
```

Now plot NMDS without fatty acid vectors:

```{r}
# plot
absol_snail <- ggplot(data = data.scores_snail_absolute, aes(x = NMDS1, y = NMDS2)) +
    geom_polygon(
    data = polygon_data_snails,
    aes(x = NMDS1, y = NMDS2, group = Treatment, fill = Treatment),
    alpha = 0.2,
    show.legend = FALSE
  ) +
  geom_point(
    aes(shape = Treatment, color = Treatment),
    size = 5,
    alpha = 1,
    show.legend = TRUE
  ) +
  scale_shape_manual(values = c("NC" = 16, "Pro" = 17, "Cip" = 18, "Mix" = 15)) +
  scale_color_manual(values = c("NC" = "#009E73", "Pro" = "#E69F00", "Cip" = "#56B4E9", "Mix" = "#999999")) +
  scale_color_manual(values = c("NC" = "#009E73", "Pro" = "#E69F00", "Cip" = "#56B4E9", "Mix" = "#999999")) +  
  scale_fill_manual(values = c("NC" = "#009E73", "Pro" = "#E69F00", "Cip" = "#56B4E9", "Mix" = "#999999")) + 
  theme_bw() +
  theme(
    axis.title = element_text(
      size = 11,
      face = "bold",
      colour = "grey30"
    ),
    panel.background = element_blank(),
    panel.border = element_rect(fill = NA, colour = "grey30"),
    legend.key = element_blank(),
    legend.title = element_text(
      size = 11,
      face = "bold",
      colour = "grey30"
    ),
    legend.text = element_text(size = 10, colour = "grey30"),
    panel.grid.major = element_line(colour = "grey95", size = 0.4),
    panel.grid.minor = element_line(colour = "grey95", size = 0.2)
  ) 

absol_snail
```

Save the plot as .png:

```{r}
ggsave(
  "output/NMDS_snails_absolute.png",
  width = 16.5,
  height = 12,
  units = "cm",
  dpi = 600
)
```

Save the plot as vector graphic:

```{r}
ggsave(
  "output/NMDS_snails_absolute.svg",
  device = "svg",
  width = 16.5,
  height = 16,
  units = "cm",
  dpi = 600
)
```

## PERMANOVA FA snails

Create distance matrix with fatty acids:

```{r}
FA_Snails_dist <- vegdist(FA_Snails_absolute_wide_sqrt, method = 'bray')
```

Separate explanatory variables from fatty acid information:

```{r}
# change initial table to wide format again
FA_Snails_absolute_wide_permanova <-
  pivot_wider(FA_Snails_absolute,
              names_from = fa,
              values_from = fa_conc)

# square root transformation again, to reduce discriminatory power of dominant fa groups
FA_Snails_absolute_wide_permanova[, 3:28] <- sqrt(FA_Snails_absolute_wide_permanova[, 3:28])

# Only keep columns/FAs with more than 5 non-zero values (fatty acids that have just been found in 5 replicates will be removed)
FA_Snails_absolute_wide_permanova <-
  FA_Snails_absolute_wide_permanova %>%
  dplyr::select(
    1:2,
    where(~sum(. !=0, na.rm = TRUE)>5)
  )

# delete fa columns
FA_Snails_absolute_wide_permanova <- FA_Snails_absolute_wide_permanova[, c(1, 2)]
FA_Snails_absolute_wide_permanova $ treatment <- as.factor(FA_Snails_absolute_wide_permanova $ treatment)
```

Perform PERMANOVA:

```{r}
perm_FA_Snails <- adonis2(FA_Snails_dist ~ treatment, data = FA_Snails_absolute_wide_permanova, permutations = 999)
print(perm_FA_Snails)
```

Pairwise comparisons (by: Martinez Arbizu, P. (2020). pairwiseAdonis: Pairwise multilevel comparison using adonis. R package version 0.4):

```{r}
pairwise_results <- pairwise.adonis2(
  FA_Snails_dist ~ treatment,
  data = FA_Snails_absolute_wide_permanova,
  permutations = 999
)

# View the results
print(pairwise_results)
```

All treatments except of Pro_vs_Mix significant differences.

## SIMPER FA snails

```{r}
# Perform SIMPER
simper_FA_Snails <-
  with(
    FA_Snails_absolute_wide_permanova,
    simper(FA_Snails_absolute_wide_sqrt, treatment,  permutations = 999)
  )

summary(simper_FA_Snails)

str(simper_FA_Snails)

# save tsv tables of FAs which contribute > 70% to differences
simper_sum_snails <- summary(simper_FA_Snails)

out_dir <- "output"
dir.create(out_dir, showWarnings = FALSE)

# Safe filename helper
safe_name <- function(x) gsub("[^A-Za-z0-9_]+", "_", x)

iwalk(simper_sum_snails, function(x, nm) {
  if (is.null(x)) return(invisible(NULL))

  # SIMPER summary as a data.frame:
  # columns: average, sd, ratio, ava, avb, cumsum, p
  # rownames are the fatty acid IDs
  df <- as.data.frame(x)

  # keep rows up to the first that reaches/exceeds 70%
  idx <- which(df$cumsum >= 0.70)[1]
  if (is.na(idx)) idx <- nrow(df)
  out <- df[seq_len(idx), , drop = FALSE]

  # just keep 3 decimels
  out_write <- out
  num_cols <- vapply(out_write, is.numeric, logical(1))
  out_write[, num_cols] <- lapply(out_write[, num_cols, drop = FALSE],
                                  function(col) formatC(col, format = "f", digits = 3))

  file <- file.path(out_dir, paste0(safe_name(nm), "_top70_snails.tsv"))
  # write row names so the TSV mirrors the SIMPER structure
  write.table(out_write, file, sep = "\t", quote = FALSE, col.names = NA)
})

# Combine results into a data frame
simper_results_snails <- do.call(rbind, lapply(simper_FA_Snails, function(x) {
  if (!is.null(x$p)) {
    data.frame(
      fatty_acid = names(x$p),  
      p = as.numeric(x$p),      
      average = as.numeric(x$average),  
      sd = as.numeric(x$sd),
      ratio = as.numeric(x$ratio),
      cusum = as.numeric(x$cusum),
      stringsAsFactors = FALSE
    )
  } else {
    NULL  
  }
}))

# Show only significant results
simper_results_snails %>%
  dplyr::filter(p <= 0.05) %>%
  dplyr::select(fatty_acid, p, average, sd, ratio, cusum)  

# How often do they occur
simper_results_snails %>%
  dplyr::filter(p <= 0.05) %>%
  dplyr::count(fatty_acid) %>%
  dplyr::arrange(desc(n)) %>%
  print()
```


# Temperature and light conditions

## Biofilm colonisation and contamination

```{r, warning=FALSE}
# insert data
colonisation_temp_light <-
  read.table(file.path("data\\colonisation_temp_light.txt"),
             header = TRUE, sep = "\t")

View(colonisation_temp_light)

# average temperature
mean(colonisation_temp_light $ temperature_colonisation)
# 18.28751 °C

# mean light intensity
# filter only daytime
average_light_intensity_colonisation <- colonisation_temp_light %>%
  dplyr::filter(light_intensity_colonisation > 0) %>%
  dplyr::summarize(average_light_intensity_colonisation = mean(light_intensity_colonisation))

print(average_light_intensity_colonisation)
# 14019.26 lux

# temperature plot
# add a small amount of random noise to the y-values
colonisation_temp_light $ temperature_colonisation <- jitter(colonisation_temp_light $ temperature_colonisation)

t2 <- ggplot(
  data = colonisation_temp_light,
  mapping = aes( x = day, y = temperature_colonisation, color = time_of_day)) +
  geom_point(alpha = 0.5) +  # Individual measurements with transparency
  geom_smooth() +
  labs(
    x = "Day", y = "Temperature (°C)"
  ) +
  theme_bw() +
  scale_x_continuous(breaks = seq(0, 77, by = 5)) +
  scale_color_continuous(
    name = "Time of Day",  
    breaks = c(0, 5, 10, 15, 20),
    labels = c("08:00 pm", "03:00 pm", "10:00 am", "05:00 am", "0:00 am"),
    limits = c(0, 22),
  )

t2 

# save the plot
ggsave(
  "output/temperature_colonisation.png",
  width = 16.5,
  height = 12,
  units = "cm",
  dpi = 600
)

# light intensity plot
# Filter the data to include only rows where light_intensity > 0
filtered_data2 <- colonisation_temp_light %>% 
  filter(light_intensity_colonisation > 0)

# add a small amount of random noise to the y-values
filtered_data2 $ light_intensity_colonisation <- jitter(filtered_data2 $ light_intensity_colonisation)

l2 <- ggplot(
  data = filtered_data2,
  mapping = aes( x = day, y = light_intensity_colonisation)) +
  geom_point(size = 1) +  # Individual measurements with transparency
  geom_smooth(color = "#E69F00", linewidth = 1) +
  labs(
    x = "Day", y = "Light intensity (lux)"
  ) +
  theme_bw() +
  scale_x_continuous(breaks = seq(0, 77, by = 5)) 

l2 

# save the plot
ggsave(
  "output/light_intensity_colonisation.png",
  width = 16.5,
  height = 12,
  units = "cm",
  dpi = 600
)
```

## Feeding assay

```{r, warning = FALSE}
# insert data
feeding_assay_temp_light <-
  read.table(file.path("data\\feeding_assay_temp_light.txt"),
             header = TRUE, 
             sep = "\t"
  )
             
View(feeding_assay_temp_light)

# average temperature
mean(feeding_assay_temp_light $ temperature)
# 13.27927 °C

# mean light intensity
# filter only daytime
average_light_intensity_feeding <- feeding_assay_temp_light %>%
  dplyr::filter(light_intensity > 0) %>%
  dplyr::summarize(average_light_intensity_feeding = mean(light_intensity))

print(average_light_intensity_feeding)
# 330.3552 lux

# temperature plot
# add a small amount of random noise to the y-values
feeding_assay_temp_light $ temperature <- jitter(feeding_assay_temp_light $ temperature)

t1 <- ggplot(
  data = feeding_assay_temp_light,
  mapping = aes( x = day, y = temperature, color = time_of_day)) +
  geom_point(alpha = 0.5) +  # Individual measurements with transparency
  geom_smooth() +
  labs(
    x = "Day", y = "Temperature (°C)"
  ) +
  theme_bw() +
  scale_x_continuous(breaks = 0:21) +
  scale_color_continuous(
    name = "Time of Day",  
    breaks = c(0, 5, 10, 15, 20),
    labels = c("08:00 pm", "03:00 pm", "10:00 am", "05:00 am", "0:00 am"),
    limits = c(0, 22),
  )

t1 

# save the plot
ggsave(
  "output/temperature_feeding_assay.png",
  width = 16.5,
  height = 12,
  units = "cm",
  dpi = 600
)

# light intensity plot
# Filter the data to include only rows where light_intensity > 0
filtered_data1 <- feeding_assay_temp_light %>% 
  filter(light_intensity > 0)

# add a small amount of random noise to the y-values
filtered_data1 $ light_intensity <- jitter(filtered_data1 $ light_intensity)

l1 <- ggplot(
  data = filtered_data1,
  mapping = aes( x = day, y = light_intensity)) +
  geom_point(size = 1) +  # Individual measurements with transparency
  geom_smooth(color = "#E69F00", linewidth = 1) +
  labs(
    x = "Day", y = "Light intensity (lux)"
  ) +
  theme_bw() +
  scale_x_continuous(breaks = 0:21) 

l1 

# save the plot
ggsave(
  "output/light_intensity_feeding_assay.png",
  width = 16.5,
  height = 12,
  units = "cm",
  dpi = 600
)
```
